### 2019-01-03 Paper1/ProofOneShotIPM.tex
proof one shot IPM started

\documentclass[10pt, a4paper, titlepage, bibliography=totocnumbered]{article}

\renewcommand*\rmdefault{ppl}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}                 % Anpassung f\"ur deutsche Sprache
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{algorithm,algorithmic}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\usepackage{amssymb}

\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\usepackage{bm}
\usepackage{color}

\newcommand{\qed}{\hfill \ensuremath{\Box}}

\usepackage{geometry}
%\geometry{verbose,a4paper,tmargin=10mm,bmargin=15mm,lmargin=12mm,rmargin=12mm}

\begin{document}

\section*{Proof One-Shot IPM}
For classical IPM, the iteation scheme for the momen vectors is
\begin{align}\label{eq:momentIteration}
\bm{u}_j^{n+1} = \bm{c}\left(\bm{\lambda}(\bm{u}_{j-1}^n),\bm{\lambda}(\bm{u}_{j}^n),\bm{\lambda}(\bm{u}_{j+1}^n)\right),
\end{align}
where $\bm{c}:\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$ is given by
\begin{align*}
\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle u(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}\rangle - \frac{\Delta t}{\Delta x}\left(\langle g(u(\bm{\lambda}_c^T\bm{\varphi}),u(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}\rangle-\langle g(u(\bm{\lambda}_{\ell}^T\bm{\varphi}),u(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}\rangle\right).
\end{align*}
The map from the moment vector to the dual variables is given by the exact fix point of $\bm{d}:\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$,
\begin{align*}
\bm{d}(\bm{u},\bm{\lambda}):= \bm{\lambda}-\bm{B}\cdot \left(\langle u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{u}\right),
\end{align*}
where $\bm{B}$ is a preconditioner, which is used in the numerical scheme to ensure convergence of the fix point iteration (which we call dual iteration in the following)
\begin{align}\label{eq:dualIteration}
\bm{\lambda}_j^{m+1} = \bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^m).
\end{align}
Note that here, $\bm{u}_j^{n}$ is a fixed parameter. By letting the dual iteration converge, i.e. $m\rightarrow\infty$, we obtain $\lim_{m\rightarrow\infty}\bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^m) =:\bm{\lambda}_j^{n} \equiv \bm\lambda(\bm{u}_j^{n})$, which are called the dual variables of $\bm{u}_j^{n}$. Hence, the numerical scheme for IPM converges the dual iteration \eqref{eq:dualIteration} to compute the mapping from the moment vectors to their dual variables and then performs one iteration of \eqref{eq:momentIteration} to obtain the time update of the moment vector. Note that this can become extremely expensive, since in each iteration of the moment vectors, we fully converge the dual iteration for all spatial cells.

To simplify notation, we leave out the dependency on $\bm{\lambda}$ when writing the moment scheme in the following: Hence with 
\begin{align*}
\bm{\tilde c}\left(\bm{u}_{\ell}^n,\bm{u}_{c}^n,\bm{u}_{r}^n\right):=\bm{c}\left(\bm{\lambda}(\bm{u}_{\ell}^n),\bm{\lambda}(\bm{u}_{c}^n),\bm{\lambda}(\bm{u}_{r}^n)\right)
\end{align*}
we can rewrite the moment iteration as
\begin{align}
\bm{u}_j^{n+1} = \bm{\tilde c}\left(\bm{u}_{j-1}^n,\bm{u}_{j}^n,\bm{u}_{j+1}^n\right).
\end{align}

For steady problems, we assume that the IPM scheme converges to a fix point, i.e. we must have that $\rho (\bm{\tilde c}_{\bm{u}})<1$. The main idea of \textit{One-Shot IPM} is to not fully converge the dual iteration, since the moment vectors are not yet converged to the exact steady solution. So if we successively perform one update of the moment iteration and one update of the dual iteration, we obtain 
\begin{subequations}\label{eq:oneshotIPM}
\begin{align}
&\bm{\lambda}_{j}^{n} =  \bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^{n-1}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^n,\bm{\lambda}_{j}^n,\bm{\lambda}_{j+1}^n\right)\label{eq:oneshotIPMmoment}.
\end{align}
\end{subequations}
\textcolor{blue}{[first moments, the duals?]}
In the following, we prove convergence of this scheme:
\begin{theorem}
The \textit{One-Shot IPM} scheme \eqref{eq:oneshotIPM} converges with the same rate as the moment iteration \eqref{eq:momentIteration} if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n-1})\bm{\varphi}\bm{\varphi}^T\rangle$ is used.
\end{theorem}
\begin{proof}
We first point out that since we assume convergence of the classical IPM scheme, we have $\rho (\bm{\tilde c}_{\bm{u}})<1$ with $\bm{\tilde c}_{\bm{u}}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by\textcolor{blue}{[write for all j]}
\begin{align*}
\bm{\tilde c}_{\bm{u}} = 
\begin{pmatrix} 
    \partial_{\bm{u}_c}\bm{\tilde c}_{1} & \partial_{\bm{u}_r}\bm{\tilde c}_{1}& 0 & 0 & \dots \\
    \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{2} & \partial_{\bm{u}_c}\bm{\tilde c}_{2} & \partial_{\bm{u}_r}\bm{\tilde c}_{2}& 0 & \dots \\
    0 & \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{3} & \partial_{\bm{u}_c}\bm{\tilde c}_{3} & \partial_{\bm{u}_r}\bm{\tilde c}_{3}\\
    \vdots & & & \ddots & \\
    0 &\cdots &  0 & \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{N_x} & \partial_{\bm{u}_c}\bm{\tilde c}_{N_x}
    \end{pmatrix},
\end{align*}
where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{u}_{j-1}^n,\bm{u}_{j}^n,\bm{u}_{j+1}^n\right)$ for all $j$. Now we have for each term inside the matrix $\bm{\tilde c}_{\bm{u}}$
\begin{align*}
\partial_{\bm{u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{\lambda}(\bm{u}_{j-1})}{\partial \bm{u}},\enskip\partial_{\bm{u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_c}\frac{\partial \bm{\lambda}(\bm{u}_j)}{\partial \bm{u}},\enskip\partial_{\bm{u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_r}\frac{\partial \bm{\lambda}(\bm{u}_{j+1})}{\partial \bm{u}}.
\end{align*}
We now wish to understand the structure of the terms $\partial_{\bm{u}} \bm{\lambda}(\bm{u})$. For this, we note that the exact dual state fulfills
\begin{align}\label{eq:ulambda}
\bm{u} = \langle u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle =: \bm{h}(\bm{\lambda}),
\end{align}
which is why we have the mapping $\bm{u}:\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$, $\bm{u}(\bm{\lambda}) = \bm{h}(\bm{\lambda})$. Since the solution of the dual problem for a given moment vector is unique, this mapping is bijective and therefore we have an inverse function
\begin{align}\label{eq:lambdau}
\bm{\lambda} = \bm{h}^{-1}(\bm{u}(\bm{\lambda}))
\end{align}
Now we differentiate both sides by $\bm{\lambda}$ to get
\begin{align*}
\bm{I}_{d} = \frac{\partial \bm{h}^{-1}(\bm{u}(\bm{\lambda}))}{\partial \bm{u}}\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}
\end{align*}
We multiply with the matrix inverse of $\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}$ and rewrite $\bm{h}^{-1}(\bm{u}(\bm{\lambda}))$ as $\bm{\lambda}(\bm{u})$ to get
\begin{align*}
\left(\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}\right)^{-1} = \bm{h}^{-1}(\bm{u}(\bm{\lambda})).
\end{align*}
Note that on the left-hand-side we have the inverse of a matrix and on the right-hand-side, we have the inverse of a multi-dimensional function. By rewriting $\bm{h}^{-1}(\bm{u}(\bm{\lambda}))$ as $\bm{\lambda}(\bm{u})$ and simply computing the term $\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}$ by differentiating \eqref{eq:ulambda} w.r.t. $\bm{\lambda}$, one obtains
\begin{align*}
\partial_{\bm{u}} \bm{\lambda}(\bm{u}) = \langle u'(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle.
\end{align*}
Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}. To obtain a single iteration function, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
\begin{align*}
\bm{u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{u}_{j-1}^{n},\bm{\lambda}_{j-1}^{n-1}),\bm{d}(\bm{u}_{j}^{n},\bm{\lambda}_{j}^{n-1}),\bm{d}(\bm{u}_{j+1}^{n},\bm{\lambda}_{j+1}^{n-1})\right)
\end{align*}
\end{proof}

\end{document}

### 2019-01-03 Paper1/ProofOneShotIPM.tex
switched one-shot IPM scheme

@@
 \begin{align}
-&\bm{\lambda}_{j}^{n} =  \bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^{n-1}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
-&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^n,\bm{\lambda}_{j}^n,\bm{\lambda}_{j+1}^n\right)\label{eq:oneshotIPMmoment}.
+&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^n,\bm{\lambda}_{j}^n,\bm{\lambda}_{j+1}^n\right)\label{eq:oneshotIPMmoment}\\
+&\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{u}_j^{n+1},\bm{\lambda}_j^{n}). \label{eq:oneshotIPMdual}
 \end{align}
@@
 Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
+To obtain a single iteration function, we plug the moment iteration step \eqref{eq:oneshotIPMmoment} into the dual iteration \eqref{eq:oneshotIPMdual} to obtain
+\begin{align*}
+\bm{\lambda}_j^{n+1} =  \bm{d}\left( \bm{c}\left(\bm{\lambda}_{j-1}^n,\bm{\lambda}_{j}^n,\bm{\lambda}_{j+1}^n\right),\bm{\lambda}_j^n \right)
+\end{align*}
+The 
+\qed
+\end{proof}
+
+We can first update the dual and then the moment vector:
+\begin{subequations}\label{eq:oneshotIPMvv}
+\begin{align}
+&\bm{\lambda}_{j}^{n} =  \bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^{n-1}) \enskip \text{ for all j} \label{eq:oneshotIPMdualvv}\\
+&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^n,\bm{\lambda}_{j}^n,\bm{\lambda}_{j+1}^n\right)\label{eq:oneshotIPMmomentvv}.
+\end{align}
+\end{subequations}
+In the following, we prove convergence of this scheme:
+\begin{theorem}
+The \textit{One-Shot IPM} scheme \eqref{eq:oneshotIPM} converges with the same rate as the moment iteration \eqref{eq:momentIteration} if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n-1})\bm{\varphi}\bm{\varphi}^T\rangle$ is used.
+\end{theorem}
+\begin{proof}
+Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
 To obtain a single iteration function, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
@@
 \end{align*}
+The 
 \end{proof}
-
 \end{document}

### 2019-01-03 Paper1/ProofOneShotIPM.tex
Proof done? There are some issues
- clean up notation
- can we fix B? This is the strategy in common One Shot

--- 2019-01-03 Paper1/ProofOneShotIPM.tex
@@
 \begin{align*}
-\bm{d}(\bm{u},\bm{\lambda}):= \bm{\lambda}-\bm{B}\cdot \left(\langle u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{u}\right),
+\bm{d}(\bm{\lambda},\bm{u}):= \bm{\lambda}-\bm{B}\cdot \left(\langle u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{u}\right),
 \end{align*}
@@
 \begin{align}\label{eq:dualIteration}
-\bm{\lambda}_j^{m+1} = \bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^m).
+\bm{\lambda}_j^{m+1} = \bm{d}(\bm{\lambda}_j^m,\bm{u}_j^{n}).
 \end{align}
@@
 By letting the dual iteration converge, i.e.
-$m\rightarrow\infty$, we obtain $\lim_{m\rightarrow\infty}\bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^m) =:\bm{\lambda}_j^{n} \equiv \bm\lambda(\bm{u}_j^{n})$, which are called the dual variables of $\bm{u}_j^{n}$.
+$m\rightarrow\infty$, we obtain $\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^m,\bm{u}_j^{n}) =:\bm{\lambda}_j^{n} \equiv \bm\lambda(\bm{u}_j^{n})$, which are called the dual variables of $\bm{u}_j^{n}$.
 Hence, the numerical scheme for IPM converges the dual iteration \eqref{eq:dualIteration} to compute the mapping from the moment vectors to their dual variables and then performs one iteration of \eqref{eq:momentIteration} to obtain the time update of the moment vector.
@@
 \begin{align}
-&\bm{\lambda}_{j}^{n} =  \bm{d}(\bm{u}_j^{n},\bm{\lambda}_j^{n-1}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
-&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^n,\bm{\lambda}_{j}^n,\bm{\lambda}_{j+1}^n\right)\label{eq:oneshotIPMmoment}.
+&\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
+&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_{j}^{n+1},\bm{\lambda}_{j+1}^{n+1}\right)\label{eq:oneshotIPMmoment}.
 \end{align}
 \end{subequations}
-\textcolor{blue}{[first moments, the duals?]}
 In the following, we prove convergence of this scheme:
 \begin{theorem}
-The \textit{One-Shot IPM} scheme \eqref{eq:oneshotIPM} converges with the same rate as the moment iteration \eqref{eq:momentIteration} if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n-1})\bm{\varphi}\bm{\varphi}^T\rangle$ is used.
+The \textit{One-Shot IPM} scheme \eqref{eq:oneshotIPM} converges with the same rate as the moment iteration \eqref{eq:momentIteration} if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$ is used.
 \end{theorem}
 \begin{proof}
-We first point out that since we assume convergence of the classical IPM scheme, we have $\rho (\bm{\tilde c}_{\bm{u}})<1$ with $\bm{\tilde c}_{\bm{u}}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by\textcolor{blue}{[write for all j]}
+We first point out that since we assume convergence of the classical IPM scheme, we have $\rho (\bm{\tilde c}_{\bm{u}})<1$ with $\bm{\tilde c}_{\bm{u}}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by
 \begin{align*}
@@
 where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{u}_{j-1}^n,\bm{u}_{j}^n,\bm{u}_{j+1}^n\right)$ for all $j$. Now we have for each term inside the matrix $\bm{\tilde c}_{\bm{u}}$
-\begin{align*}
+\begin{align}\label{eq:cTildeDer}
 \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{\lambda}(\bm{u}_{j-1})}{\partial \bm{u}},\enskip\partial_{\bm{u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_c}\frac{\partial \bm{\lambda}(\bm{u}_j)}{\partial \bm{u}},\enskip\partial_{\bm{u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_r}\frac{\partial \bm{\lambda}(\bm{u}_{j+1})}{\partial \bm{u}}.
-\end{align*}
+\end{align}
 We now wish to understand the structure of the terms $\partial_{\bm{u}} \bm{\lambda}(\bm{u})$. For this, we note that the exact dual state fulfills
@@
 \end{align}
-Now we differentiate both sides by $\bm{\lambda}$ to get
+Now we differentiate both sides w.r.t. $\bm{\lambda}$ to get
 \begin{align*}
@@
 $\bm{\lambda}$, one obtains
+\begin{align}\label{eq:dudlambdaex}
+\partial_{\bm{u}} \bm{\lambda}(\bm{u}) = \langle u'(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle.
+\end{align}
+Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
+Note that this iteration is not really a fix point iteration, since it uses the time updated dual variables in \eqref{eq:oneshotIPMmoment}.
+To obtain a fix point iteration, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
 \begin{align*}
-\partial_{\bm{u}} \bm{\lambda}(\bm{u}) = \langle u'(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle.
+&\bm{\lambda}_j^{n+1} = \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all j} \\
+&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{u}_{j+1}^{n})\right)
 \end{align*}
-Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
-To obtain a single iteration function, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
+The Jacobian $\bm{J}\in\mathbb{R}^{2(N+1)\cdot N_x \times 2(N+1)\cdot N_x}$ has the form
 \begin{align*}
-\bm{u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{u}_{j-1}^{n},\bm{\lambda}_{j-1}^{n-1}),\bm{d}(\bm{u}_{j}^{n},\bm{\lambda}_{j}^{n-1}),\bm{d}(\bm{u}_{j+1}^{n},\bm{\lambda}_{j+1}^{n-1})\right)
+\bm{J} = 
+\begin{pmatrix}
+ \partial_{\bm{\lambda}} \bm{d} & \partial_{\bm{u}} \bm{d}  \\
+\partial_{\bm{\lambda}} \bm{c} & \partial_{\bm{u}} \bm{c}
+\end{pmatrix},
 \end{align*}
+where each block has entries for all spatial cells. We start by looking at $\partial_{\bm{\lambda}} \bm{d}$. For columns $(j-1)\cdot(N+1)$ to $j\cdot(N+1)$, we have
+\begin{align*}
+\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{u}_j^n) = \bm{I}_d - \bm{B} \cdot \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle.
+\end{align*}
+Recall the choice $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-1}$, meaning that $\partial_{\bm{\lambda}} \bm{d} = \bm{0}$.
+For the block $\partial_{\bm{u}} \bm{d}$, we get
+\begin{align*}
+\partial_{\bm{u}} \bm{d}(\bm{\lambda}_j^n,\bm{u}_j^n) = \bm{B},
+\end{align*}
+hence $\partial_{\bm{u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$:
+\begin{align*}
+\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{u}_{j-1}^{n})}{\partial \bm{\lambda}} = \bm{0},
+\end{align*}
+since we already showed that by the choice of $\bm{B}$ the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
+We can show the same result for all spatial cells and all inputs of $\bm{c}$ analogously, hence $\partial_{\bm{\lambda}} \bm{c} = 0$.
+For the last block, we have that
+\begin{align*}
+\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{u}_{j-1}^{n})}{\partial \bm{u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{B} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle u'(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-1} = \partial_{\bm{u}_{\ell}}\bm{\tilde c}_j
+\end{align*}
+by the choice of $\bm{B}$ as well as \eqref{eq:cTildeDer} and \eqref{eq:dudlambdaex}.
+Hence, we have that $\partial_{\bm{u}} \bm{c} = \bm{\tilde c}_{\bm{u}}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM converges.
+Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM converges.\qed
 \end{proof}

### 2019-01-06 Paper1/ProofOneShotIPM.tex
small fixes, B is still constant

@@
 \section*{Proof One-Shot IPM}
-For classical IPM, the iteation scheme for the momen vectors is
+For classical IPM, the iteration scheme for the moment vectors is
 \begin{align}\label{eq:momentIteration}
@@
 \end{align}
-We now wish to understand the structure of the terms $\partial_{\bm{u}} \bm{\lambda}(\bm{u})$. For this, we note that the exact dual state fulfills
+We first wish to understand the structure of the terms $\partial_{\bm{u}} \bm{\lambda}(\bm{u})$. For this, we note that the exact dual state fulfills
 \begin{align}\label{eq:ulambda}
@@
 \begin{align*}
-\bm{I}_{d} = \frac{\partial \bm{h}^{-1}(\bm{u}(\bm{\lambda}))}{\partial \bm{u}}\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}
+\bm{I}_{d} = \frac{\partial \bm{h}^{-1}(\bm{u}(\bm{\lambda}))}{\partial \bm{u}}\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}.
 \end{align*}
-We multiply with the matrix inverse of $\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}$ and rewrite $\bm{h}^{-1}(\bm{u}(\bm{\lambda}))$ as $\bm{\lambda}(\bm{u})$ to get
+We multiply with the matrix inverse of $\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}$ to get
 \begin{align*}
-\left(\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}\right)^{-1} = \bm{h}^{-1}(\bm{u}(\bm{\lambda})).
+\left(\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}\right)^{-1} = \frac{\partial \bm{h}^{-1}(\bm{u}(\bm{\lambda}))}{\partial \bm{u}}.
 \end{align*}
@@
 \begin{align}\label{eq:dudlambdaex}
-\partial_{\bm{u}} \bm{\lambda}(\bm{u}) = \langle u'(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle.
+\partial_{\bm{u}} \bm{\lambda}(\bm{u}) = \langle u'(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}.
 \end{align}
 Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
-Note that this iteration is not really a fix point iteration, since it uses the time updated dual variables in \eqref{eq:oneshotIPMmoment}.
+Note that in its current form this iteration is not really a fix point iteration, since it uses the time updated dual variables in \eqref{eq:oneshotIPMmoment}.
 To obtain a fix point iteration, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain

### 2019-01-07 Paper1/ProofOneShotIPM.tex
proof local convergence

@@
 \newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{remark}[theorem]{Remark}
 
@@
 \newenvironment{example}[1][Example]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{remark}[1][Remark]{\begin{trivlist}
 \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
@@
 
-\section*{Proof One-Shot IPM}
+\section*{One-Shot IPM}
 For classical IPM, the iteration scheme for the moment vectors is
@@
 we can rewrite the moment iteration as
-\begin{align}
+\begin{align}\label{eq:shortIPMIt}
 \bm{u}_j^{n+1} = \bm{\tilde c}\left(\bm{u}_{j-1}^n,\bm{u}_{j}^n,\bm{u}_{j+1}^n\right).
@@
 \end{subequations}
-In the following, we prove convergence of this scheme:
-\begin{theorem}
-The \textit{One-Shot IPM} scheme \eqref{eq:oneshotIPM} converges with the same rate as the moment iteration \eqref{eq:momentIteration} if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$ is used.
+In the following, we study convergence of this scheme.
+For this we take an approach commonly chosen to prove the local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fix point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence.
+\begin{theorem}\label{th:Contractive}
+Assume that the classical IPM iteration \eqref{eq:shortIPMIt} is contractive at its fix point $\bm{u}^*$.
+Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fix point $(\bm{\lambda}^*,\bm{u}^*)$ if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$, i.e.
+the Hessian of the dual problem is used.
 \end{theorem}
 \begin{proof}
-We first point out that since we assume convergence of the classical IPM scheme, we have $\rho (\bm{\tilde c}_{\bm{u}})<1$ with $\bm{\tilde c}_{\bm{u}}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by
+We first point out that since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\bm{\tilde c}_{\bm{u}}(\bm{u}^*))<1$ with $\bm{\tilde c}_{\bm{u}}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by
 \begin{align*}
@@
 \end{align*}
-where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{u}_{j-1}^n,\bm{u}_{j}^n,\bm{u}_{j+1}^n\right)$ for all $j$. Now we have for each term inside the matrix $\bm{\tilde c}_{\bm{u}}$
+where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{u}_{j-1}^*,\bm{u}_{j}^*,\bm{u}_{j+1}^*\right)$ for all $j$. Now we have for each term inside the matrix $\bm{\tilde c}_{\bm{u}}$
 \begin{align}\label{eq:cTildeDer}
-\partial_{\bm{u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{\lambda}(\bm{u}_{j-1})}{\partial \bm{u}},\enskip\partial_{\bm{u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_c}\frac{\partial \bm{\lambda}(\bm{u}_j)}{\partial \bm{u}},\enskip\partial_{\bm{u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_r}\frac{\partial \bm{\lambda}(\bm{u}_{j+1})}{\partial \bm{u}}.
+\partial_{\bm{u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{\lambda}(\bm{u}_{j-1}^*)}{\partial \bm{u}},\enskip\partial_{\bm{u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_c}\frac{\partial \bm{\lambda}(\bm{u}_j^*)}{\partial \bm{u}},\enskip\partial_{\bm{u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_r}\frac{\partial \bm{\lambda}(\bm{u}_{j+1}^*)}{\partial \bm{u}}.
 \end{align}
@@
 The Jacobian $\bm{J}\in\mathbb{R}^{2(N+1)\cdot N_x \times 2(N+1)\cdot N_x}$ has the form
-\begin{align*}
+\begin{align}\label{eq:Jacobian}
 \bm{J} = 
@@
 \end{pmatrix},
+\end{align}
+where each block has entries for all spatial cells. We start by looking at $\partial_{\bm{\lambda}} \bm{d}$. For the columns belonging to cell $j$, we have
+\begin{align*}
+\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{u}_j^n) &= \bm{I}_d - \bm{B} \cdot \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle - \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle u(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi} \rangle - \bm{u}\right) \\
+&=- \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle u(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi} \rangle - \bm{u}\right),
 \end{align*}
-where each block has entries for all spatial cells. We start by looking at $\partial_{\bm{\lambda}} \bm{d}$. For columns $(j-1)\cdot(N+1)$ to $j\cdot(N+1)$, we have
-\begin{align*}
-\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{u}_j^n) = \bm{I}_d - \bm{B} \cdot \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle.
-\end{align*}
-Recall the choice $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-1}$, meaning that $\partial_{\bm{\lambda}} \bm{d} = \bm{0}$.
+where we used $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-1}$.
+Recall that at the fix point $(\bm{\lambda}^*,\bm{u}^*)$, we have $\langle u(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi} \rangle = \bm{u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
 For the block $\partial_{\bm{u}} \bm{d}$, we get
@@
 since we already showed that by the choice of $\bm{B}$ the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
-We can show the same result for all spatial cells and all inputs of $\bm{c}$ analogously, hence $\partial_{\bm{\lambda}} \bm{c} = 0$.
+We can show the same result for all spatial cells and all inputs of $\bm{c}$ analogously, hence $\partial_{\bm{\lambda}} \bm{c} = \bm{0}$.
 For the last block, we have that
@@
 by the choice of $\bm{B}$ as well as \eqref{eq:cTildeDer} and \eqref{eq:dudlambdaex}.
-Hence, we have that $\partial_{\bm{u}} \bm{c} = \bm{\tilde c}_{\bm{u}}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM converges.
-Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM converges.\qed
+Hence, we have that $\partial_{\bm{u}} \bm{c} = \bm{\tilde c}_{\bm{u}}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM iteration is contractive.
+Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fix point.\qed
 \end{proof}
+\begin{theorem}\label{th:localConvergence}
+With the assumptions from Theorem \ref{th:Contractive}, the One-Shot IPM converges locally, i.e.
+there exists a $\delta>0$ s.t.
+for all starting points $(\bm{\lambda}^0,\bm{u}^0)\in B_{\delta}(\bm{\lambda}^*,\bm{u}^*)$ we have
+\begin{align*}
+\Vert (\bm{\lambda}^n,\bm{u}^n) - (\bm{\lambda}^*,\bm{u}^*)\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
+\end{align*}
+\end{theorem}
+\begin{proof}
+By Theorem \ref{th:Contractive}, the One-Shot scheme is contractive at its fix point.
+Since we assumed convergence of the classical IPM scheme, we can conclude that all entries in the Jacobian $\bm{J}$ are continuous functions.
+Furthermore, the determinant of $\bm{\tilde{J}}:=\bm{J}-\lambda \bm{I}_d$ is a polynomial of continuous functions, since
+\begin{align*}
+\text{det}(\bm{\tilde J}) = \sum_{\sigma} \text{sgn}(\sigma)\prod_{i = 1}^{2 N_x (N+1)} \tilde J_{\sigma(i),i}.
+\end{align*}
+Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{u})$.
+Hence there exists an open ball with radius $\delta$ around the fix point in which the eigenvalues remain in the interval $(-1,1)$.\qed
+\end{proof}
+\begin{remark}
+Theorem \ref{th:localConvergence} gives some insights in how to set up the One-Shot IPM iteration: If we look at the Jacobian \eqref{eq:Jacobian} at an arbitrary point, we obtain
+
+Let us assume that the classical IPM scheme converges globally.
+
+\begin{itemize}
+\item Choose a moment vector $\bm{u}^0$ and compute the corresponding exact dual variables $\bm{\lambda}^0:=\bm{\lambda}(\bm{u}^0)$ as starting vector.
+In this case, the term $\partial_{\lambda}\bm{d}$ is equal to zero for the starting conditions.
+\end{itemize}
+\end{remark}
 

### 2019-02-19 code/src/euler2d.cpp
added paper tex style; fixed wrong sigma association

@@ -183,13 +183,13 @@ Vector Euler2D::IC( const Vector& x, const Vector& xi ) {
         double T     = 273.15;
         double p     = 101325.0;
         double Ma    = 0.8;
-        if( xi.size() == 1 ) {
-            Ma = Ma + xi[1] * _sigma[0];
+        if( xi.size() == 2 ) {
+            Ma = Ma + xi[1] * _sigma[1];
         }
         double a = sqrt( gamma * R * T );
 
         double uMax  = Ma * a;
-        double angle = ( 1.25 + 0.0 * _sigma[0] * xi[0] ) * ( 2.0 * M_PI ) / 360.0;
+        double angle = ( 1.25 + _sigma[0] * xi[0] ) * ( 2.0 * M_PI ) / 360.0;
         double uF    = uMax * cos( angle );
         double vF    = uMax * sin( angle );
 

### 2019-02-19 code/src/euler.cpp
first changes

@@ -20,7 +20,7 @@ Euler::~Euler() {}
 void Euler::Solve() {}
 
 Vector Euler::G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n ) {
-    double rhoInv = 1.0 / u[0];    //
+    double rhoInv = 1.0 / u[0];
     double vU     = u[1] * rhoInv;
     double p      = ( _gamma - 1.0 ) * ( u[2] - 0.5 * u[0] * pow( vU, 2 ) );
     double aU     = sqrt( _gamma * p * rhoInv );

### 2019-02-19 code/src/expliciteuler.cpp

@@ -4,7 +4,6 @@ ExplicitEuler::ExplicitEuler( Settings* settings, Mesh* mesh ) : TimeSolver( set
 
 void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, const Matrix&, const Vector&, const Vector& )> const& fluxFunc,
                              MatVec& uNew,
-                             MatVec& u,
                              MatVec& uQ,
                              double dt ) {
     auto numCells = _mesh->GetNumCells();
@@ -17,7 +16,7 @@ void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, const M
         auto neighbors = cell->GetNeighborIDs();
         if( cell->IsBoundaryCell() ) {
             if( cell->GetBoundaryType() == BoundaryType::DIRICHLET ) {
-                uNew[j] = u[j];
+                uNew[j] = uQ[j];
                 continue;
             }
             else if( cell->GetBoundaryType() == BoundaryType::NOSLIP ) {
@@ -56,7 +55,7 @@ void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, const M
             }
         }
 
-        Matrix rhs( u[0].rows(), u[0].columns(), 0.0 );
+        Matrix rhs( uQ[0].rows(), uQ[0].columns(), 0.0 );
         for( unsigned l = 0; l < neighbors.size(); ++l ) {
             if( ( _mesh->GetBoundaryType( j ) == BoundaryType::NOSLIP || _mesh->GetBoundaryType( j ) == BoundaryType::SWWALL ) &&
                 neighbors[l] == numCells ) {
@@ -66,6 +65,6 @@ void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, const M
                 fluxFunc( rhs, uQ[j], uQ[neighbors[l]], cell->GetUnitNormal( l ), cell->GetNormal( l ) );
             }
         }
-        uNew[j] = u[j] - ( dt / cell->GetArea() ) * rhs;
+        uNew[j] = uQ[j] - ( dt / cell->GetArea() ) * rhs;
     }
 }

### 2019-02-19 code/src/expliciteuler.h

@@ -13,7 +13,6 @@ class ExplicitEuler : public TimeSolver
     ExplicitEuler( Settings* settings, Mesh* mesh );
     virtual void Advance( std::function<void( Matrix&, const Matrix&, const Matrix&, const Vector&, const Vector& )> const& fluxFunc,
                           MatVec& uNew,
-                          MatVec& u,
                           MatVec& uQ,
                           double dt );
 };

### 2019-02-19 code/src/momentsolver.cpp

@@ -29,20 +29,19 @@ void MomentSolver::Solve() {
 
     std::chrono::steady_clock::time_point tic = std::chrono::steady_clock::now();
 
+    MatVec uQ = MatVec( _nCells + 1, Matrix( _nStates, _settings->GetNQTotal() ) );
+
     // create solution fields
     MatVec u( _nCells, Matrix( _nStates, _nTotal ) );
     if( _settings->HasRestartFile() ) {
         this->ImportTime();
-        u       = this->ImportMoments();
-        _lambda = this->ImportDuals();
+        uQ = this->ImportMoments();
     }
     else {
-        u       = SetupIC();
-        _lambda = MatVec( _nCells + 1, Matrix( _nStates, _nTotal ) );
+        uQ = SetupIC();
     }
-    MatVec uNew = u;
-    MatVec uOld = u;
-    MatVec uQ   = MatVec( _nCells + 1, Matrix( _nStates, _settings->GetNqPE() ) );
+    MatVec uNew = uQ;
+    MatVec uOld = uNew;
 
     Vector ds( _nStates );
     Vector u0( _nStates );
@@ -52,36 +51,6 @@ void MomentSolver::Solve() {
     std::vector<unsigned> cellIndexPE = _settings->GetCellIndexPE();
     std::vector<int> PEforCell        = _settings->GetPEforCell();
 
-    log->info( "PE {0}: kStart {1}, kEnd {2}", _settings->GetMyPE(), _settings->GetKStart(), _settings->GetKEnd() );
-
-    for( unsigned j = 0; j < _nCells; ++j ) {
-        for( unsigned l = 0; l < _nStates; ++l ) {
-            u0[l] = u[j]( l, 0 );
-        }
-        if( !_settings->HasRestartFile() ) {
-            _closure->DS( ds, u0 );
-            for( unsigned l = 0; l < _nStates; ++l ) {
-                _lambda[j]( l, 0 ) = ds[l];
-            }
-        }
-    }
-
-    // Converge initial condition entropy variables for One Shot IPM
-    if( _settings->GetMaxIterations() == 1 ) {
-
-        _settings->SetMaxIterations( 10000 );
-        for( unsigned j = 0; j < _nCells; ++j ) {
-            _closure->SolveClosure( _lambda[j], u[j] );
-        }
-        _settings->SetMaxIterations( 1 );
-
-        // recompute moments with inexact lambda
[…]

### 2019-02-19 code/src/momentsolver.h

@@ -35,7 +35,6 @@ class MomentSolver
     void numFlux( Matrix& out, const Matrix& u1, const Matrix& u2, const Vector& nUnit, const Vector& n );
     void SetupIC( MatVec& out );
     MatVec SetupIC();
-    void CalculateMoments( MatVec& out, const MatVec& lambda );
     Vector EvalLambda( const Vector& lambda, const Vector& xi );
     void Plot( double time, unsigned nSteps );
     void Export( const MatVec& u, const MatVec& lambda ) const;

### 2019-02-19 code/src/timesolver.h

@@ -30,7 +30,6 @@ class TimeSolver
     static TimeSolver* Create( Settings* settings, Mesh* mesh );
     virtual void Advance( std::function<void( Matrix&, const Matrix&, const Matrix&, const Vector&, const Vector& )> const& fluxFunc,
                           MatVec& uNew,
-                          MatVec& u,
                           MatVec& uQ,
                           double dt ) = 0;
     double GetTimeStepSize();

### 2019-02-19 code/src/expliciteuler.cpp
Collocation done

@@ -22,39 +22,40 @@ void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, const M
             else if( cell->GetBoundaryType() == BoundaryType::NOSLIP ) {
                 ghostCell = uQ[j];
                 Vector v( 2, 0.0 );
-                for( unsigned k = 0; k < uQ[numCells].columns(); ++k ) {
-                    v.reset();
-                    v[0]              = ghostCell( 1, k ) / ghostCell( 0, k );
-                    v[1]              = ghostCell( 2, k ) / ghostCell( 0, k );
-                    Vector n          = cell->GetBoundaryUnitNormal();
-                    double vn         = dot( n, v );
-                    Vector Vn         = vn * n;
-                    Vector Vb         = -Vn + v;
-                    double velMagB    = Vb[0] * Vb[0] + Vb[1] * Vb[1];
-                    double velMag     = v[0] * v[0] + v[1] * v[1];
-                    double rho        = ghostCell( 0, k );
-                    ghostCell( 1, k ) = rho * ( Vb[0] );
-                    ghostCell( 2, k ) = rho * ( Vb[1] );
-                    ghostCell( 3, k ) += rho * 0.5 * ( velMagB - velMag );
-                }
+                v.reset();
+                v[0]              = ghostCell( 1, 0 ) / ghostCell( 0, 0 );
+                v[1]              = ghostCell( 2, 0 ) / ghostCell( 0, 0 );
+                Vector n          = cell->GetBoundaryUnitNormal();
+                double vn         = dot( n, v );
+                Vector Vn         = vn * n;
+                Vector Vb         = -Vn + v;
+                double velMagB    = Vb[0] * Vb[0] + Vb[1] * Vb[1];
+                double velMag     = v[0] * v[0] + v[1] * v[1];
+                double rho        = ghostCell( 0, 0 );
+                ghostCell( 1, 0 ) = rho * ( Vb[0] );
+                ghostCell( 2, 0 ) = rho * ( Vb[1] );
+                ghostCell( 3, 0 ) += rho * 0.5 * ( velMagB - velMag );
             }
             else if( cell->GetBoundaryType() == BoundaryType::SWWALL ) {
                 ghostCell = uQ[j];
                 Vector v( 2, 0.0 );
-                for( unsigned k = 0; k < uQ[numCells].columns(); ++k ) {
[…]

### 2019-02-19 code/src/momentsolver.cpp

@@ -9,7 +9,7 @@ MomentSolver::MomentSolver( Settings* settings, Mesh* mesh, Problem* problem ) :
     _tEnd        = _settings->GetTEnd();
     _nStates     = _settings->GetNStates();
     _nQuadPoints = _settings->GetNQuadPoints();
-    _nQTotal     = _settings->GetNQTotal();
+    _nQTotal     = 1;
     _nTotal      = _settings->GetNTotal();
 
     _closure = Closure::Create( _settings );
@@ -24,32 +24,23 @@ MomentSolver::~MomentSolver() {
     delete _time;
 }
 
-void MomentSolver::Solve() {
+MatVec MomentSolver::Solve( Vector xi ) {
     auto log = spdlog::get( "event" );
 
     std::chrono::steady_clock::time_point tic = std::chrono::steady_clock::now();
 
-    MatVec uQ = MatVec( _nCells + 1, Matrix( _nStates, _settings->GetNQTotal() ) );
+    MatVec uQ = MatVec( _nCells + 1, Matrix( _nStates, 1 ) );
 
     // create solution fields
-    MatVec u( _nCells, Matrix( _nStates, _nTotal ) );
+    MatVec u( _nCells, Matrix( _nStates, 1 ) );
     if( _settings->HasRestartFile() ) {
         this->ImportTime();
         uQ = this->ImportMoments();
     }
     else {
-        uQ = SetupIC();
+        uQ = SetupIC( xi );
     }
     MatVec uNew = uQ;
-    MatVec uOld = uNew;
-
-    Vector ds( _nStates );
-    Vector u0( _nStates );
-
-    double residualFull;
-
-    std::vector<unsigned> cellIndexPE = _settings->GetCellIndexPE();
-    std::vector<int> PEforCell        = _settings->GetPEforCell();
 
     auto numFluxPtr = std::bind( &MomentSolver::numFlux,
                                  this,
@@ -58,8 +49,6 @@ void MomentSolver::Solve() {
                                  std::placeholders::_3,
                                  std::placeholders::_4,
                                  std::placeholders::_5 );
-
-    std::cout << uQ[2]( 0, 0 ) << std::endl;
 
     log->info( "{:10}   {:10}", "t", "residual" );
     // Begin time loop
@@ -81,8 +70,8 @@ void MomentSolver::Solve() {
         for( unsigned j = 0; j < _nCells; ++j ) {
[…]

### 2019-02-19 Paper1/main.tex
first small changes in Paper

\documentclass[3p]{elsarticle}

% Packages and macros go here
\usepackage{bm}
\usepackage[parfill]{parskip}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{color}
\usepackage{bm}
\usepackage{graphicx}
%\usepackage{subfig}
\usepackage{subcaption}
%\usepackage[]{algorithm2e}
%\usepackage{algorithmic}
%\usepackage{algorithmicx}
\usepackage[linesnumbered,ruled]{algorithm2e}

\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{algorithmic}
\ifpdf
  \DeclareGraphicsExtensions{.eps,.pdf,.png,.jpg}
\else
  \DeclareGraphicsExtensions{.eps}
\fi

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\DeclareMathOperator*{\argmin}{arg\,min}

\makeatletter
\newsavebox\myboxA
\newsavebox\myboxB
\newlength\mylenA

\newcommand*\xoverline[2][0.75]{%
    \sbox{\myboxA}{$\m@th#2$}%
    \setbox\myboxB\null% Phantom box
    \ht\myboxB=\ht\myboxA%
    \dp\myboxB=\dp\myboxA%
    \wd\myboxB=#1\wd\myboxA% Scale phantom
    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
    \addtolength\mylenA{-\the\wd\myboxB}%
    \ifdim\wd\myboxB<\wd\myboxA%
       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
    \else
        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
    \fi}
\makeatother

\journal{arXiv.org}

%strongly recommended
%\numberwithin{theorem}{section}

% Declare title and authors, without \thanks
\newcommand{\TheTitle}{A semi-intrusive Code Framework for Uncertainty Quantification}

\date{\today}

% Sets running headers as well as PDF title and authors
%\headers{\RunningTitle}{\TheAuthors}

\usepackage{amsopn}
\DeclareMathOperator{\diag}{diag}
%\DeclareMathOperator{\argmin}{arg\min}

% FundRef data to be entered by SIAM
%<funding-group>
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\def\lambdabar{\xoverline{\bm{\lambda}}}
\def\lambdabarjn{\lambdabar_j^n}

\newcommand{\commentRyan}[1]{{\Large{\textcolor{red}{\newline[RYAN: #1]\newline}}}}
\newcommand{\commentJonas}[1]{{\Large{\textcolor{blue}{\newline[JONAS: #1]\newline}}}}

\begin{document}

\begin{frontmatter}

\title{\TheTitle}

\author[adressJonas]{Jonas Kusch}
\author[adressJannick]{Jannick Wolters}

\address[adressJonas]{Karlsruhe Institute of Technology, Karlsruhe,
    jonas.kusch@kit.edu}
\address[adressJannick]{Karlsruhe Institute of Technology, Karlsruhe, jannick.wolters@kit.edu}

\begin{abstract}
Standard methods for quantifying the effects of uncertainties in hyperbolic problems can yield oscillatory solution approximations. ...
\end{abstract}

\begin{keyword}
conservation laws, hyperbolic, intrusive, UQ, IPM, SG, Collocation
\end{keyword}

\end{frontmatter}
\section{Introduction}

\section{Methods of Uncertainty Quantification}
\subsection{Collocation}
\subsection{Stochastic Galerkin}
\subsection{Intrusive Polynomial Moment Method}
\section{Method discussion}
\begin{itemize}
	\item kleinere Fehler [11]
	\item Programmieraufwand [31]
	\item zweiabhängig
	\item adaptiv (breack up black box approach )
	\item grobe Auflösung für Momente 
\end{itemize}
\section{UQCreator}
\subsection{Adaptivity}
\subsection{restart IPM} %relay(-strategy) IPM?
\subsection{one-shot IPM}
\subsection{adaptive IPM}
\section{Results}
\subsection{2D Euler flow over NACA0012}
\begin{itemize}
	\item AoA study
	\item AoA + Ma number + mehr ?
\end{itemize}
\subsection{2D shallow water dam break}

\bibliographystyle{unsrt}  
\bibliography{references}
\end{document}

### 2019-02-19 code/src/sparsegrid.cpp
added cc sparse grids

@@ -1,0 +1,414 @@
+#include "sparsegrid.h"
+
+SparseGrid::SparseGrid( std::string type, unsigned pDim, unsigned pLevel ) : _dim( pDim ), _level( pLevel ), _nodeCount( 0 ), _unionCount( 0 ) {
+    if( !( type.length() != 0 && pDim > 0 ) ) {
+        std::cerr << "\n[sparseGrid]: Invalid input parameter" << std::endl;
+        exit( EXIT_FAILURE );
+    }
+    if( _level == 0 ) {
+        _Rl = 1;
+    }
+    else {
+        _Rl = static_cast<unsigned>( std::pow( 2, _level ) + 1 );
+    }
+
+    if( !std::strcmp( type.c_str(), "CC" ) ) {
+        createCCgrid();
+    }
+    else {
+        std::cerr << "\n[sparseGrid]: Invalid node type" << std::endl;
+        exit( EXIT_FAILURE );
+    }
+}
+
+SparseGrid::~SparseGrid() {
+    for( unsigned i = 0; i < _dim; i++ ) delete[] _nodes[i];
+    delete[] _nodes;
+
+    delete[] _weights;
+
+    for( unsigned i = 0; i < _dim; i++ ) delete[] _unionIndex[i];
+    delete[] _unionIndex;
+
+    for( unsigned i = 0; i < _dim; i++ ) delete[] _nodeOrder[i];
+    delete[] _nodeOrder;
+}
+
+void SparseGrid::createCCgrid() {
+    computeUnionIndex();
+    computeNodeOrder();
+
+    _nodes = new double*[_dim];
+    for( unsigned i = 0; i < _dim; i++ ) {
+        _nodes[i] = new double[_nodeCount];
+    }
+    for( unsigned i = 0; i < _dim; i++ )
+        for( unsigned j = 0; j < _nodeCount; j++ ) _nodes[i][j] = 0.0;
+    computeCCNodes();
+
+    _weights = new double[_nodeCount];
+    for( unsigned i = 0; i < _nodeCount; i++ ) _weights[i] = 0.0;
+    computeCCWeights();
+}
+
+void SparseGrid::computeCCNodes() {
+    for( unsigned i = 0; i < _dim; i++ ) {
+        for( unsigned j = 0; j < _nodeCount; j++ ) {
+            unsigned n = _nodeOrder[i][j] + 1;
+            if( n < 1 || _Rl < n ) {
[…]

### 2019-02-19 code/src/sparsegrid.h

@@ -1,0 +1,53 @@
+#ifndef SPARSE_GRID_H
+#define SPARSE_GRID_H
+
+#include <assert.h>
+#include <cmath>
+#include <cstring>
+#include <functional>
+#include <iostream>
+#include <limits>
+#include <numeric>
+#include <vector>
+
+class SparseGrid
+{
+  private:
+    std::string _nodeType;
+    unsigned _dim;
+    unsigned _level;
+    unsigned _Rl;
+    long unsigned _nodeCount;
+    long unsigned _unionCount;
+
+    unsigned** _unionIndex;
+    unsigned** _nodeOrder;
+    std::vector<std::vector<std::vector<unsigned>>> _gridIndices;
+
+    double** _nodes;
+    double* _weights;
+
+    void computeUnionIndex();
+    void computeNodeOrder();
+    void createCCgrid();
+    void computeCCNodes();
+    void computeCCWeights();
+    std::vector<double> CCweights1D( unsigned order );
+    std::vector<double>
+    CCweightsND( unsigned dimIndex, unsigned order, unsigned orderND, std::vector<double> weights1D, std::vector<double> weightsND );
+    void colexOrdering( std::vector<unsigned> base, std::vector<unsigned>& mat, bool& run );
+    template <class T> int isEven( T i );
+    template <class T> double choose( T n, T k );
+
+    SparseGrid() = delete;
+
+  public:
+    SparseGrid( std::string type, unsigned pDim, unsigned pLevel );
+    ~SparseGrid();
+    std::vector<std::pair<std::vector<double>, double>> GetGrid() const;
+    std::vector<std::vector<double>> GetNodes() const;
+    std::vector<double> GetWeights() const;
+    unsigned long GetNodeCount() const;
+};
+
+#endif

### 2019-02-19 Paper1/main.tex
paper continued

@@
 \begin{abstract}
-Standard methods for quantifying the effects of uncertainties in hyperbolic problems can yield oscillatory solution approximations. ...
+Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
+Intrusive methods require the implementation of a new code and yield
+
 \end{abstract}
@@
 \section{Introduction}
+Hyperbolic equations play an important role in various research areas such as fluid dynamics or plasma physics.
+Efficient numerical methods combined with robust implementations are available for these problems, however they do not account for uncertainties which can arise in measurement data or modeling assumptions.
+Including the effects of uncertainties in differential equations has become an important topic in the last decades.
+%Examples include computational fluid dynamics \cite{bijl2013uncertainty}.
+
+A general hyperbolic set of equations with random initial data can be written as
+\begin{subequations}\label{eq:hyperbolicProblem}
+\begin{align}
+\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{F}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0}, \\ \label{eq:ic}
+\bm{u}(t=0,\bm{x},&\bm{\xi}) = \bm{u}_{\text{IC}}(\bm{x},\bm{\xi}),
+\end{align}
+\end{subequations}
+where the solution $u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subset \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subset\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
+The physical flux is given by $\bm{F}:\mathbb{R}^p\rightarrow\mathbb{R}^p$.
+To simplify notation, we assume that the initial condition is random, i.e.
+$\bm{\xi}$ enters through the definition of $\bm{u}_{IC}$.
+Equations \eqref{eq:hyperbolicProblem} are usually supplemented with boundary conditions, which we will specify later for the individual problems.
+
+Due to the randomness of the solution, one is interested in determining the expectation value or the variance of the solution, i.e.
+\begin{align*}
+\text{E}[\bm{u}] = \langle \bm{u} \rangle,\qquad \text{Var}[\bm{u}] = \langle \left( \bm{u}-\text{E}[\bm{u}]\right)^2\rangle,
+\end{align*}
+where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot \prod_{i=1}^s f_{\Xi,i}(\xi_i)d\xi_1 \cdots d\xi_s$.
+More generally, one is interested in determining the moments of the solution for a given set of orthonormal basis functions $\varphi_{i}$ for $i = 0,\cdots,N$.
+Here, $i$ is a multi-index and we define basis functions such that the total degree is smaller or equal to $N$.
+The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$.\\
+
+Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods.
+The main idea of intrusive methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i:\Theta\rightarrow\mathbb{R}$ for $|i|\leq M$ yields
+\begin{subequations}\label{eq:nonClosedMomentSystem}
+\begin{align}
+\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{F}(\bm{u}(t,\bm{x},\cdot)) \varphi_i\rangle = \bm{0}, \\
+\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
+\end{align}
+\end{subequations}
+To obtain a closed set of equations, one needs to derive a closure $\mathcal{U}$ such that 
+\begin{align*}
+\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N).
+\end{align*}
+A commonly used closure is the stochastic-Galerkin, which represents the solution by a polynomial:
+\begin{align*}
+\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N):= \sum_{i=0}^N \hat{u}_i\varphi_i.
+\end{align*}
+
+, which both come with certain advantages and disadvantages.
+
+When investigating hyperbolic problems, standard methods tend to suffer from non-physical oscillations, which yield
 

### 2019-02-19 Paper1/main.tex
continued

@@
 where the solution $u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subset \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subset\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
-The physical flux is given by $\bm{F}:\mathbb{R}^p\rightarrow\mathbb{R}^p$.
+The physical flux is given by $\bm{F}:\mathbb{R}^p\to\mathbb{R}^p$.
 To simplify notation, we assume that the initial condition is random, i.e.
@@
 where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot \prod_{i=1}^s f_{\Xi,i}(\xi_i)d\xi_1 \cdots d\xi_s$.
-More generally, one is interested in determining the moments of the solution for a given set of orthonormal basis functions $\varphi_{i}$ for $i = 0,\cdots,N$.
-Here, $i$ is a multi-index and we define basis functions such that the total degree is smaller or equal to $N$.
+More generally, one is interested in determining the moments of the solution for a given set of orthonormal basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq N$.
 The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$.\\
@@
 Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods.
-The main idea of intrusive methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i:\Theta\rightarrow\mathbb{R}$ for $|i|\leq M$ yields
+The main idea of intrusive methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
 \begin{subequations}\label{eq:nonClosedMomentSystem}
@@
 \begin{align*}
-\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N).
+\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi).
 \end{align*}
@@
 \begin{align*}
-\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N):= \sum_{i=0}^N \hat{u}_i\varphi_i.
+\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}).
 \end{align*}
-
-, which both come with certain advantages and disadvantages.
+When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic, meaning that it cannot be solved with standard methods.
+An approach to derive a closure, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure, which has been presented in \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.
+For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is
+\begin{align}\label{eq:primalProblem}
+\mathcal{U}_{ME}(\bm{u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
+\end{align}
+This problem can be rewritten as
+\begin{align}\label{eq:dualProblem}
+ \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{s\cdot(N + 1)}}
+  \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}\right\},
+\end{align}
+where $s_*:\mathbb{R}\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}$ are called the dual variables. The solution to \eqref{eq:primalProblem} is then given by
+\begin{align}\label{eq:ansatz}
+ \mathcal{U}_{ME}(\bm{u}) = \left( s_{\bm{u}} \right)^{-1}(\bm{\hat{\lambda}}(\bm{u})^T \bm{\varphi}).
+\end{align}
+The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin method, i.e.
+IPM generalizes different intrusive method.
+Furthermore, at least for scalar problems, IPM is less oscillatory compared to SG \cite{kusch2017maximum}.
 

### 2019-02-20 Paper1/colloc_ipm.tex
added file structure

\section{Collocation accelerated IPM for steady state problems}

### 2019-02-20 Paper1/introduction.tex
added file structure

--- 2019-02-19 Paper1/main.tex
@@
-\documentclass[3p]{elsarticle}
-
-% Packages and macros go here
-\usepackage{bm}
-\usepackage[parfill]{parskip}
-\usepackage[utf8]{inputenc}
-\usepackage{url}
-\usepackage{amsmath,amssymb,amsfonts,amsthm}
-\usepackage{color}
-\usepackage{bm}
-\usepackage{graphicx}
-%\usepackage{subfig}
-\usepackage{subcaption}
-%\usepackage[]{algorithm2e}
-%\usepackage{algorithmic}
-%\usepackage{algorithmicx}
-\usepackage[linesnumbered,ruled]{algorithm2e}
-
-\usepackage{lipsum}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{epstopdf}
-\usepackage{algorithmic}
-\ifpdf
-  \DeclareGraphicsExtensions{.eps,.pdf,.png,.jpg}
-\else
-  \DeclareGraphicsExtensions{.eps}
-\fi
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{proposition}{Proposition}
-
-\DeclareMathOperator*{\argmin}{arg\,min}
-
-\makeatletter
-\newsavebox\myboxA
-\newsavebox\myboxB
-\newlength\mylenA
-
-\newcommand*\xoverline[2][0.75]{%
-    \sbox{\myboxA}{$\m@th#2$}%
-    \setbox\myboxB\null% Phantom box
-    \ht\myboxB=\ht\myboxA%
-    \dp\myboxB=\dp\myboxA%
-    \wd\myboxB=#1\wd\myboxA% Scale phantom
-    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
-    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
-    \addtolength\mylenA{-\the\wd\myboxB}%
-    \ifdim\wd\myboxB<\wd\myboxA%
-       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
-    \else
-        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
-    \fi}
-\makeatother
-
-\journal{arXiv.org}
-
-%strongly recommended
-%\numberwithin{theorem}{section}
-
-% Declare title and authors, without \thanks
-\newcommand{\TheTitle}{A semi-intrusive Code Framework for Uncertainty Quantification}
-
-\date{\today}
-
-% Sets running headers as well as PDF title and authors
-%\headers{\RunningTitle}{\TheAuthors}
-
-\usepackage{amsopn}
-\DeclareMathOperator{\diag}{diag}
-%\DeclareMathOperator{\argmin}{arg\min}
-
-% FundRef data to be entered by SIAM
-%<funding-group>
-%<award-group>
-%<funding-source>
-%<named-content content-type="funder-name"> 
-%</named-content> 
-%<named-content content-type="funder-identifier"> 
-%</named-content>
-%</funding-source>
-%<award-id> </award-id>
-%</award-group>
-%</funding-group>
-
-\def\lambdabar{\xoverline{\bm{\lambda}}}
-\def\lambdabarjn{\lambdabar_j^n}
-
-\newcommand{\commentRyan}[1]{{\Large{\textcolor{red}{\newline[RYAN: #1]\newline}}}}
-\newcommand{\commentJonas}[1]{{\Large{\textcolor{blue}{\newline[JONAS: #1]\newline}}}}
-
-\begin{document}
-
-\begin{frontmatter}
-
-\title{\TheTitle}
-
-\author[adressJonas]{Jonas Kusch}
-\author[adressJannick]{Jannick Wolters}
-
-\address[adressJonas]{Karlsruhe Institute of Technology, Karlsruhe,
-    jonas.kusch@kit.edu}
-\address[adressJannick]{Karlsruhe Institute of Technology, Karlsruhe, jannick.wolters@kit.edu}
-
-\begin{abstract}
-Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
-Intrusive methods require the implementation of a new code and yield
-
-\end{abstract}
-
-\begin{keyword}
-conservation laws, hyperbolic, intrusive, UQ, IPM, SG, Collocation
-\end{keyword}
-
-\end{frontmatter}
 \section{Introduction}
@@
 When investigating hyperbolic problems, standard methods tend to suffer from non-physical oscillations, which yield
-
-\section{Methods of Uncertainty Quantification}
-\subsection{Collocation}
-\subsection{Stochastic Galerkin}
-\subsection{Intrusive Polynomial Moment Method}
-\section{Method discussion}
-\begin{itemize}
-	\item kleinere Fehler [11]
-	\item Programmieraufwand [31]
-	\item zweiabhängig
-	\item adaptiv (breack up black box approach )
-	\item grobe Auflösung für Momente 
-\end{itemize}
-\section{UQCreator}
-\subsection{Adaptivity}
-\subsection{restart IPM} %relay(-strategy) IPM?
-\subsection{one-shot IPM}
-\subsection{adaptive IPM}
-\section{Results}
-\subsection{2D Euler flow over NACA0012}
-\begin{itemize}
-	\item AoA study
-	\item AoA + Ma number + mehr ?
-\end{itemize}
-\subsection{2D shallow water dam break}
-
-\bibliographystyle{unsrt}  
-\bibliography{references}
-\end{document}

### 2019-02-20 code/src/momentsolver.cpp
Allreduce for dt

@@ -121,12 +121,12 @@ void MomentSolver::Solve() {
             dtCurrent = _problem->ComputeDt( uQ[j], _mesh->GetMaxEdge( j ) / _mesh->GetArea( j ) );
             if( dtCurrent < dtMinOnPE ) dtMinOnPE = dtCurrent;
         }
-        MPI_Reduce( &dtMinOnPE, &dt, 1, MPI_DOUBLE, MPI_MIN, 0, MPI_COMM_WORLD );
+        MPI_Allreduce( &dtMinOnPE, &dt, 1, MPI_DOUBLE, MPI_MIN, MPI_COMM_WORLD );
         t += dt;
 
         _time->Advance( numFluxPtr, uNew, u, uQ, dt );
 
-        // perform reduction to obtain full moments on all PEs
+        // perform reduction to obtain full moments on PE PEforCell[j], which is the PE that solves the dual problem for cell j
         for( unsigned j = 0; j < _nCells; ++j ) {
             MPI_Reduce( uNew[j].GetPointer(), u[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, MPI_SUM, PEforCell[j], MPI_COMM_WORLD );
         }

### 2019-02-20 Paper1/introduction.tex

@@
  \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{s\cdot(N + 1)}}
-  \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}\right\},
+  \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
 \end{align}
-where $s_*:\mathbb{R}\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}$ are called the dual variables. The solution to \eqref{eq:primalProblem} is then given by
+where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}$ are called the dual variables. The solution to \eqref{eq:primalProblem} is then given by
 \begin{align}\label{eq:ansatz}
- \mathcal{U}_{ME}(\bm{u}) = \left( s_{\bm{u}} \right)^{-1}(\bm{\hat{\lambda}}(\bm{u})^T \bm{\varphi}).
+ \mathcal{U}_{ME}(\bm{\hat u}) = \left( s_{\bm{\hat u}} \right)^{-1}(\bm{\hat{\lambda}}(\bm{u})^T \bm{\varphi}).
 \end{align}
@@
 IPM generalizes different intrusive method.
-Furthermore, at least for scalar problems, IPM is less oscillatory compared to SG \cite{kusch2017maximum}.
+Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
+Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the entropy
+\begin{align*}
+S(\bm{\hat u}) := \langle s( \mathcal{U}_{ME}(\bm{\hat u}))\rangle,
+\end{align*}
+i.e.
+the IPM method yields a physically correct entropy solution.
+This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution, which can then lead to a failure of the method.
+The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves the computation of the optimization problem \eqref{eq:dualProblem}.
 

### 2019-02-20 Paper1/main.tex

--- 2019-02-19 Paper1/main.tex
@@
 \author[adressJannick]{Jannick Wolters}
+\author[adressMartin]{Martin Frank}
 
@@
 \address[adressJannick]{Karlsruhe Institute of Technology, Karlsruhe, jannick.wolters@kit.edu}
+\address[adressMartin]{Karlsruhe Institute of Technology, Karlsruhe, martin.frank@kit.edu}
 
 \begin{abstract}
-Standard methods for quantifying the effects of uncertainties in hyperbolic problems can yield oscillatory solution approximations. ...
+Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
+Intrusive methods require the implementation of a new code and yield
+
 \end{abstract}
@@
 \end{frontmatter}
-\section{Introduction}
 
-\section{Methods of Uncertainty Quantification}
-\subsection{Collocation}
-\subsection{Stochastic Galerkin}
-\subsection{Intrusive Polynomial Moment Method}
-\section{Method discussion}
-\begin{itemize}
-	\item kleinere Fehler [11]
-	\item Programmieraufwand [31]
-	\item zweiabhängig
-	\item adaptiv (breack up black box approach )
-	\item grobe Auflösung für Momente 
-\end{itemize}
-\section{UQCreator}
-\subsection{Adaptivity}
-\subsection{restart IPM} %relay(-strategy) IPM?
-\subsection{one-shot IPM}
-\subsection{adaptive IPM}
-\section{Results}
-\subsection{2D Euler flow over NACA0012}
-\begin{itemize}
-	\item AoA study
-	\item AoA + Ma number + mehr ?
-\end{itemize}
-\subsection{2D shallow water dam break}
+\input{introduction}
+\input{framework}
+\input{colloc_ipm}
+\input{one_shot_ipm}
+\input{results}
+
+%\section{Methods of Uncertainty Quantification}
+%\subsection{Collocation}
+%\subsection{Stochastic Galerkin}
+%\subsection{Intrusive Polynomial Moment Method}
+%\section{Method discussion}
+%\begin{itemize}
+%	\item kleinere Fehler [11]
+%	\item Programmieraufwand [31]
+%	\item zweiabhängig
+%	\item adaptiv (breack up black box approach )
+%	\item grobe Auflösung für Momente 
+%\end{itemize}
+%\section{UQCreator}
+%\subsection{Adaptivity}
+%\subsection{restart IPM} %relay(-strategy) IPM?
+%\subsection{one-shot IPM}
+%\subsection{adaptive IPM}
+%\section{Results}
+%\subsection{2D Euler flow over NACA0012}
+%\begin{itemize}
+%	\item AoA study
+%	\item AoA + Ma number + mehr ?
+%\end{itemize}
+%\subsection{2D shallow water dam break}
 

### 2019-02-20 code/src/momentsolver.cpp
added option to not load dual state on restart (not loading is the default)

@@ -33,8 +33,10 @@ void MomentSolver::Solve() {
     MatVec u( _nCells, Matrix( _nStates, _nTotal ) );
     if( _settings->HasRestartFile() ) {
         this->ImportTime();
-        u       = this->ImportMoments();
-        _lambda = this->ImportDuals();
+        u = this->ImportMoments();
+        if( _settings->LoadLambda() ) {
+            _lambda = this->ImportDuals();
+        }
     }
     else {
         u       = SetupIC();
@@ -58,7 +60,7 @@ void MomentSolver::Solve() {
         for( unsigned l = 0; l < _nStates; ++l ) {
             u0[l] = u[j]( l, 0 );
         }
-        if( !_settings->HasRestartFile() ) {
+        if( !_settings->LoadLambda() ) {
             _closure->DS( ds, u0 );
             for( unsigned l = 0; l < _nStates; ++l ) {
                 _lambda[j]( l, 0 ) = ds[l];

### 2019-02-20 code/src/settings.cpp

@@ -56,6 +56,7 @@ Settings::Settings( std::string inputFile ) : _inputFile( inputFile ), _hasExact
         auto restartFile = general->get_as<std::string>( "restartFile" );
         if( restartFile ) {
             _restartFile = _inputDir.string() + "/" + *restartFile;
+            _loadLambda  = general->get_as<bool>( "importDualState" ).value_or( false );
         }
 
         auto icFile = general->get_as<std::string>( "icFile" );
@@ -281,6 +282,7 @@ bool Settings::HasICFile() const { return !_icFile.empty(); }
 std::string Settings::GetICFile() const { return _icFile; }
 bool Settings::HasRestartFile() const { return !_restartFile.empty(); }
 std::string Settings::GetRestartFile() const { return _restartFile; }
+bool Settings::LoadLambda() const { return _loadLambda; }
 
 // problem
 TimesteppingType Settings::GetTimesteppingType() const { return _timesteppingType; }

### 2019-02-20 code/src/settings.h

@@ -25,6 +25,7 @@ class Settings
     std::filesystem::path _outputFile;
     std::filesystem::path _icFile;
     std::filesystem::path _restartFile;
+    bool _loadLambda;
 
     // requied settings
     unsigned _meshDimension;
@@ -92,6 +93,7 @@ class Settings
     std::string GetICFile() const;
     bool HasRestartFile() const;
     std::string GetRestartFile() const;
+    bool LoadLambda() const;
 
     // problem
     TimesteppingType GetTimesteppingType() const;

### 2019-02-20 Paper1/introduction.tex
introduction fast fertig

@@
 Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods.
-The main idea of intrusive methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
+A popular non-intrusive method is the stochastic-Collocation (SC) method, which computes the moments with the help of a numerical quadrature rule: For a given set of $N_q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
+\begin{align*}
+\bm{\hat u_i} = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{N_q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi_k})f_{\Xi}(\bm{\xi}_k).
+\end{align*} 
+Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a big implementation effort.
+Furthermore, it is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
+A downside of collocation methods are aliasing effects, which stem from the inexact approximation of the integral.
+Furthermore, collocation methods require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation}.
+Despite their easy implementation, collocation methods can become cumbersome for unsteady problems, since in this case, the solution needs to written out at all quadrature points in every time step to compute the time evolution of the moments.
+This problem does not occur in intrusive methods, since the computation is carried out on the moments, i.e.
+the time evolution can be recorded during the computation.
+Also for steady problems, the fact that the computation is carried out on the moments provided the ability to compute the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
+However intrusive methods are in general more difficult to implement and come with higher numerical costs.
+The main idea of these methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
 \begin{subequations}\label{eq:nonClosedMomentSystem}
@@
 \begin{align}\label{eq:dualProblem}
- \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{s\cdot(N + 1)}}
+ \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}methods^{s\cdot(N + 1)}}
   \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
@@
 \end{align}
+Plugging the derived closure into the moment system \eqref{eq:nonClosedMomentSystem}, one obtains the IPM moment system
+\begin{subequations}\label{eq:IPMmomentSystem}
+\begin{align}
+\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{F}(\mathcal{U}_{ME}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
+\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
+\end{align}
+\end{subequations}
 The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin method, i.e.
@@
 the IPM method yields a physically correct entropy solution.
-This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution, which can then lead to a failure of the method.
+This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution, which can then cause a failure of the method.
 The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves the computation of the optimization problem \eqref{eq:dualProblem}.
+Hence, the desirable costs of IPM (and intrusive methods in general) require a significantly increased run time.
+A further advantages which we will not use in this work is the possibility of using adaptivity \cite{kroker2012finite}, i.e.
+refining the stochastic space at certain spatial position and time steps in which complex structures such as discontinuities occur.
+The main task here is to find an adequate refinement indicator \cite{kroker2012finite,giesselmann2017posteriori}.\\
 
-When investigating hyperbolic problems, standard methods tend to suffer from non-physical oscillations, which yield
+In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods.
+The framework only requires the numerical flux of the deterministic problem.
+If IPM is used, the entropy of the deterministic problem needs to be provided.
+We thereby provide the ability to recycle existing implementations of deterministic solvers.
+Furthermore, we investigate intrusive methods for steady problems and compare them to collocation methods.
+The steady setting provides different opportunities to take advantage of features of intrusive methods:
+\begin{itemize}
+\item We propose to apply intrusive methods as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method, which we then again converge to steady state.
+\item Since the moments during the iteration process are inaccurate, i.e.
+they are not the steady state solution, we propose not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
+\end{itemize}
+
+The paper is structured as follows: After

### 2019-02-21 code/src/momentsolver.cpp
broadcast Moments for restart calculation

@@ -142,6 +142,12 @@ void MomentSolver::Solve() {
             log->info( "{:03.8f}   {:01.5e}", t, residualFull );
         }
     }
+
+    // MPI Broadcast final moment vectors to all PEs
+    for( unsigned j = 0; j < _nCells; ++j ) {
+        MPI_Bcast( uNew[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, PEforCell[j], MPI_COMM_WORLD );
+    }
+
     if( _settings->GetMyPE() != 0 ) return;
 
     // save final moments on u

### 2019-02-21 Paper1/framework.tex
continued

\section{Discretization and code framework}
\label{sec:framework}

### 2019-02-21 Paper1/introduction.tex
continued

@@
 \begin{align}\label{eq:primalProblem}
-\mathcal{U}_{ME}(\bm{u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
+\mathcal{U}_{ME}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
 \end{align}
@@
 \begin{itemize}
-\item We propose to apply intrusive methods as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method, which we then again converge to steady state.
+\item Apply intrusive methods as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method, which we then again converge to steady state.
 \item Since the moments during the iteration process are inaccurate, i.e.
-they are not the steady state solution, we propose not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
+they are not the steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
 \end{itemize}
 
-The paper is structured as follows: After
+The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive framework in section \ref{sec:framework}.
+Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}.
+A comparison of results computed with the presented methods is given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.

### 2019-02-21 Paper1/one_shot_ipm.tex
continued

--- 2019-01-07 Paper1/ProofOneShotIPM.tex
@@
-\documentclass[10pt, a4paper, titlepage, bibliography=totocnumbered]{article}
-
-\renewcommand*\rmdefault{ppl}
-
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage[english]{babel}                 % Anpassung f\"ur deutsche Sprache
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-\usepackage{algorithm,algorithmic}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{remark}[theorem]{Remark}
-
-\usepackage{amssymb}
-
-\newenvironment{proof}[1][Proof]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{definition}[1][Definition]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{example}[1][Example]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-
-\usepackage{bm}
-\usepackage{color}
-
-\newcommand{\qed}{\hfill \ensuremath{\Box}}
-
-\usepackage{geometry}
-%\geometry{verbose,a4paper,tmargin=10mm,bmargin=15mm,lmargin=12mm,rmargin=12mm}
-
-\begin{document}
-
-\section*{One-Shot IPM}
+\section{One-shot IPM}
+\label{sec:OneShotIPM}
 For classical IPM, the iteration scheme for the moment vectors is
@@
 \end{remark}
-
-\end{document}

### 2019-02-21 Paper1/summary_outlook.tex
continued

\section{Summary and outlook}
\label{sec:summary_outlook}

### 2019-02-21 Paper1/framework.tex
moment solver restart bug fixed + Paper continued

@@
 \label{sec:framework}
+\subsection{Discretization}
+Omitting initial conditions, we can write the IPM moment system \eqref{eq:IPMmomentSystem} as
+\begin{align*}
+\partial_t \bm{u}+\partial_x \bm{F}(\bm{u}) = \bm{0}
+\end{align*}
+with the flux $\bm{F}:\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$, $\bm{F}(\bm{u})=\langle f(u_{ME}(\hat \Lambda(\bm{u})))\bm{\varphi} \rangle$ depending on the dual state 
+\begin{align*}
+\hat \Lambda(\bm{u}) = \bm{\hat{\lambda}}(\bm{u})^T\bm{\varphi}.
+\end{align*}
+For efficiency of exposition, we sometimes omit the dependence on $\bm{u}$.
+The IPM system is hyperbolic, so it is naturally solved by a finite-volume method.
+First we discretize the spatial domain into cells.
+The discrete unknowns are chosen to be the spatial averages over each cell at time $t_n$, given by
+\begin{align*}
+u_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}u_i(t_n,x) dx.
+\end{align*}
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{u}_j^n = (u_{0j}^n,\cdots,u_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$ as
+\begin{align}\label{eq:IPMDiscretization}
+\bm{u}_{j}^{n+1} = \bm{u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{u}_j^n,\bm{u}_{j+1}^n)- \bm{G}(\bm{u}_{j-1}^n,\bm{u}_{j}^n)\right)
+\end{align}
+for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
+%The discretization of the dual state $\Lambda$ at spatial cell $j$ and time step $n$ is given by
+%\begin{align*}
+%\hat{\Lambda}_j^n = \bm{\hat{\lambda}}(\bm{u}_j^n)^T\bm{\varphi}.
+%\end{align*}
+The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\bm{u},\bm{u})=\bm{F}(\bm{u})$.
+To ensure stability, a CFL condition has to be derived by investigating the eigenvalues of $\nabla \bm{F}$.
+
+When a consistent numerical flux $g:\mathbb{R}\times\mathbb{R}\to\mathbb{R}$, $g = g(u_\ell, u_r)$ is available for the deterministic problem \eqref{eq:origProblem}, then for the IPM system we can simply take
+\begin{align*}
+ \bm{G}(\bm{u}_{j}^n,\bm{u}_{j+1}^n) = \langle g(u_{ME}(\hat\Lambda(\bm{u}_{j}^n)),u_{ME}(\hat\Lambda(\bm{u}_{j+1}^n)))\bm{\varphi}\rangle.
+\end{align*}
+This choice of the numerical flux is a common choice in kinetic theory and is called kinetic flux.
+The time update of the moment vector now becomes
+\begin{align}\label{eq:exactUpdate}
+\bm{u}_{j}^{n+1} = \bm{u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\hat \Lambda_j^n),u_{ME}(\hat \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\hat \Lambda_{j-1}^n),u_{ME}(\hat \Lambda_{j}^n))\bm{\varphi} \rangle\right),
+\end{align}
+where $\hat\Lambda_{j}^n :=\hat\Lambda(\bm{u}_{j}^n)$ for all $j$.
+Note that the computation of $\hat\Lambda_{j}^n$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{u}_{j}^n$.
+
+Unfortunately \eqref{eq:exactUpdate} cannot be implemented because the dual problem cannot be solved exactly.%
+\footnote{
+Equation \eqref{eq:exactUpdate} also includes integral evaluations which cannot be computed in closed form.
+Their approximation by numerical quadrature, however, does not play a role in the realizability problems we discuss below.
+}
+Instead, it must be solved numerically, for example with Newton's method.
+The stopping criterion for the numerical optimizer ensures that the approximate multiplier vector it returns, which we denote $\xoverline{\bm{\lambda}}_j^n\in\mathbb{R}^{N+1}$ for the moment vector $\bm{u}_j^n$, satisfies the stopping criterion
+\begin{align}\label{eq:tauCrit}
+\left\Vert \bm{u}_j^n-\left\langle u_{ME}\left(\left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau.
+\end{align}
+This is derived from the first-order necessary conditions for the dual problem.
+Once the numerical optimizer finds such a $\xoverline{\bm{\lambda}}_j^n$, the corresponding dual state $\xoverline{\Lambda}_j^n := \left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi} \in \mathbb{P}(\Theta)$ can be used in \eqref{eq:exactUpdate} for the unknown $\hat \Lambda_j^n$.
+This gives Algorithm \ref{alg:seq}.
+
+\begin{algorithm}[H]
+\begin{algorithmic}[1]
+\FOR{$j=0$ to $N_x+1$}
+\STATE $\bm{u}_j^0 = \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_0(x, \cdot) \bm{\varphi} \rangle dx$
+\ENDFOR
+\FOR{$n=0$ to $N_t$}
+\FOR{$j=0$ to $N_x+1$}
+\STATE $\xoverline{\bm{\lambda}}_j^n \approx \argmin_{\bm{\lambda}}  \left( \langle s_*(\bm{\lambda}^T \bm{\varphi})\rangle - \bm{\lambda}^T \bm{u}_j^n \right)$
+\hfill such that \eqref{eq:tauCrit} holds
+\STATE $\xoverline \Lambda_j^n = \left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}$
+\ENDFOR
+\FOR{$j=1$ to $N_x$}
+\STATE $\bm{u}_{j}^{n+1} = \bm{u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\xoverline \Lambda_j^n),u_{ME}(\xoverline \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\xoverline \Lambda_{j-1}^n),u_{ME}(\xoverline \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
+\ENDFOR
+\ENDFOR
+\end{algorithmic}
+\caption{IPM for Uncertainty Quantification}
+\label{alg:seq}
+\end{algorithm}

### 2019-02-21 code/src/momentsolver.cpp

@@ -37,6 +37,9 @@ void MomentSolver::Solve() {
         if( _settings->LoadLambda() ) {
             _lambda = this->ImportDuals();
         }
+        else {
+            _lambda = MatVec( _nCells + 1, Matrix( _nStates, _nTotal ) );
+        }
     }
     else {
         u       = SetupIC();

### 2019-02-21 Paper1/framework.tex
introduction and framework continued

--- 2019-02-21 Paper1/framework.tex
@@
 \label{sec:framework}
+\subsection{Discretization}
+In the following, we discretize the moment system in space and time following according to \cite{kusch2017maximum}.
+Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:IPMmomentSystem}.
+Omitting initial conditions and assuming a one-dimensional spatial domain, we can write this system  as
+\begin{align*}
+\partial_t \bm{\hat u}+\partial_x \bm{F}(\bm{\hat u}) = \bm{0}
+\end{align*}
+with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{u})=\langle f(\mathcal{U}_{ME}(\bm{\hat u}))\bm{\varphi} \rangle$, where we omit the index ME in the following for efficiency of exposition.
+Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
+First we perform the discretization of the spatial domain: We choose the discrete unknowns to be the spatial averages over each cell at time $t_n$, given by
+\begin{align*}
+\bm{\hat u}_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}\bm{\hat u}_i(t_n,x) dx.
+\end{align*}
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$ as
+\begin{align}\label{eq:IPMDiscretization}
+\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{\hat u}_{j}^{n},\bm{\hat u}_{j+1}^{n})- \bm{G}(\bm{\hat u}_{j-1}^{n},\bm{\hat u}_{j}^{n})\right)
+\end{align}
+for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
+The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\bm{u},\bm{u})=\bm{F}(\bm{u})$.
+To ensure stability, a CFL condition has to be derived by investigating the eigenvalues of $\nabla \bm{F}$.
+
+When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the deterministic problem \eqref{eq:origProblem}, then for the IPM system we can simply take
+\begin{align*}
+ \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle g(u_{ME}(\hat\Lambda(\bm{\hat u}_{j}^{n})),u_{ME}(\hat\Lambda(\bm{\hat u}_{j+1}^{n})))\bm{\varphi}\rangle.
+\end{align*}
+This choice of the numerical flux is a common choice in kinetic theory and is called kinetic flux.
+The time update of the moment vector now becomes
+\begin{align}\label{eq:exactUpdate}
+\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\hat \Lambda_j^n),u_{ME}(\hat \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\hat \Lambda_{j-1}^n),u_{ME}(\hat \Lambda_{j}^n))\bm{\varphi} \rangle\right),
+\end{align}
+where $\hat\Lambda_{j}^n :=\hat\Lambda(\bm{\hat u}_{j}^{n})$ for all $j$.
+Note that the computation of $\hat\Lambda_{j}^n$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
+
+Unfortunately \eqref{eq:exactUpdate} cannot be implemented because the dual problem cannot be solved exactly.%
+\footnote{
+Equation \eqref{eq:exactUpdate} also includes integral evaluations which cannot be computed in closed form.
+Their approximation by numerical quadrature, however, does not play a role in the realizability problems we discuss below.
+}
+Instead, it must be solved numerically, for example with Newton's method.
+The stopping criterion for the numerical optimizer ensures that the approximate multiplier vector it returns, which we denote $\xoverline{\bm{\lambda}}_j^n\in\mathbb{R}^{N+1}$ for the moment vector $\bm{\hat u}_{j}^{n}$, satisfies the stopping criterion
+\begin{align}\label{eq:tauCrit}
+\left\Vert \bm{\hat u}_{j}^{n}-\left\langle u_{ME}\left(\left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau.
+\end{align}
+This is derived from the first-order necessary conditions for the dual problem.
+Once the numerical optimizer finds such a $\xoverline{\bm{\lambda}}_j^n$, the corresponding dual state $\xoverline{\Lambda}_j^n := \left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi} \in \mathbb{P}(\Theta)$ can be used in \eqref{eq:exactUpdate} for the unknown $\hat \Lambda_j^n$.
+This gives Algorithm \ref{alg:seq}.
+
+\begin{algorithm}[H]
+\begin{algorithmic}[1]
+\FOR{$j=0$ to $N_x+1$}
+\STATE $\bm{u}_j^0 = \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_0(x, \cdot) \bm{\varphi} \rangle dx$
+\ENDFOR
+\FOR{$n=0$ to $N_t$}
+\FOR{$j=0$ to $N_x+1$}
+\STATE $\xoverline{\bm{\lambda}}_j^n \approx \argmin_{\bm{\lambda}}  \left( \langle s_*(\bm{\lambda}^T \bm{\varphi})\rangle - \bm{\lambda}^T \bm{\hat u}_{j}^{n} \right)$
+\hfill such that \eqref{eq:tauCrit} holds
+\STATE $\xoverline \Lambda_j^n = \left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}$
+\ENDFOR
+\FOR{$j=1$ to $N_x$}
+\STATE $\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\xoverline \Lambda_j^n),u_{ME}(\xoverline \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\xoverline \Lambda_{j-1}^n),u_{ME}(\xoverline \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
+\ENDFOR
+\ENDFOR
+\end{algorithmic}
+\caption{IPM for Uncertainty Quantification}
+\label{alg:seq}
+\end{algorithm}

### 2019-02-21 Paper1/introduction.tex
introduction and framework continued

@@
 \end{align*}
-A commonly used closure is the stochastic-Galerkin, which represents the solution by a polynomial:
+A commonly used closure is given by stochastic-Galerkin, which represents the solution by a polynomial:
 \begin{align*}
-\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}).
+\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi},
 \end{align*}
+where we collect moments in $\hat{u}:=(\hat{u}_0,\cdots,\hat{u}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
 When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic, meaning that it cannot be solved with standard methods.
-An approach to derive a closure, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure, which has been presented in \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.
+A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.
 For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is
@@
 \begin{align}\label{eq:dualProblem}
- \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}methods^{s\cdot(N + 1)}}
-  \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
+ \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{(N+1) \times p}}
+  \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
 \end{align}
-where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}$ are called the dual variables. The solution to \eqref{eq:primalProblem} is then given by
+where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}\in \mathbb{R}^{(N+1) \times p}$ are called the dual variables.
+The solution to \eqref{eq:primalProblem} is then given by
 \begin{align}\label{eq:ansatz}

### 2019-02-22 code/src/momentsolver.cpp
added read-in of settings from a logfile

@@ -32,10 +32,10 @@ void MomentSolver::Solve() {
     // create solution fields
     MatVec u( _nCells, Matrix( _nStates, _nTotal ) );
     if( _settings->HasRestartFile() ) {
-        this->ImportTime();
-        u = this->ImportMoments();
+        this->ImportPrevSettings();
+        u = this->ImportPrevMoments();
         if( _settings->LoadLambda() ) {
-            _lambda = this->ImportDuals();
+            _lambda = this->ImportPrevDuals();
         }
         else {
             _lambda = MatVec( _nCells + 1, Matrix( _nStates, _nTotal ) );
@@ -299,21 +299,35 @@ void MomentSolver::Export( const MatVec& u, const MatVec& lambda ) const {
     dual_writer->flush();
 }
 
-void MomentSolver::ImportTime() {
+void MomentSolver::ImportPrevSettings() {
     auto file = std::ifstream( _settings->GetRestartFile() );
     std::string line;
+    std::stringstream prevSettingsStream;
+    bool configSection = false;
     while( std::getline( file, line ) ) {
-        if( line.find( "tEnd" ) != std::string::npos ) {
-            line.erase( 0, line.find_first_of( '=' ) + 1 );
-            line.erase( std::remove_if( line.begin(), line.end(), []( char c ) -> bool { return std::isspace<char>( c, std::locale::classic() ); } ),
-                        line.end() );
-            _tStart = std::stod( line );
+        if( line.find( "Config file" ) != std::string::npos ) {
+            configSection = true;
+            std::getline( file, line );
+            std::getline( file, line );
+        }
+        else if( configSection && line.find( "==================================" ) != std::string::npos ) {
             break;
         }
+        if( configSection ) {
+            line.erase( 0, line.find_first_of( '|' ) + 1 );
+            prevSettingsStream << line << std::endl;
+        }
+    }
+    std::istringstream inputStream( prevSettingsStream.str() );
+    Settings* prevSettings = new Settings( inputStream );
+    _tStart                = prevSettings->GetTEnd();
+    if( prevSettings->GetNMoments() != _settings->GetNMoments() ) {
+        Closure* prevClosure = Closure::Create( prevSettings );
+        // TODO
     }
 }
 
-MatVec MomentSolver::ImportMoments() {
+MatVec MomentSolver::ImportPrevMoments() {
[…]

### 2019-02-22 code/src/momentsolver.h

@@ -39,9 +39,9 @@ class MomentSolver
     Vector EvalLambda( const Vector& lambda, const Vector& xi );
     void Plot( double time, unsigned nSteps );
     void Export( const MatVec& u, const MatVec& lambda ) const;
-    void ImportTime();
-    MatVec ImportMoments();
-    MatVec ImportDuals();
+    void ImportPrevSettings();
+    MatVec ImportPrevMoments();
+    MatVec ImportPrevDuals();
 
   public:
     MomentSolver( Settings* settings, Mesh* mesh, Problem* problem );

### 2019-02-22 code/src/settings.cpp

@@ -2,6 +2,13 @@
 
 #include "settings.h"
 
+namespace cpptoml {
+inline std::shared_ptr<table> parse_file( const std::istringstream& content ) {
+    parser p{const_cast<std::istringstream&>( content )};
+    return p.parse();
+}
+}    // namespace cpptoml
+
 Settings::Settings( std::string inputFile ) : _inputFile( inputFile ), _hasExactSolution( false ) {
     auto log = spdlog::get( "event" );
 
@@ -13,6 +20,241 @@ Settings::Settings( std::string inputFile ) : _inputFile( inputFile ), _hasExact
         _inputDir = std::experimental::filesystem::path( _inputFile.parent_path() );
 
         auto file = cpptoml::parse_file( _inputFile );
+
+        // section general
+        auto general           = file->get_table( "general" );
+        auto problemTypeString = general->get_as<std::string>( "problem" );
+
+        if( problemTypeString ) {
+            if( problemTypeString->compare( "Burgers1D" ) == 0 ) {
+                _problemType = ProblemType::P_BURGERS_1D;
+            }
+            else if( problemTypeString->compare( "Euler1D" ) == 0 ) {
+                _problemType = ProblemType::P_EULER_1D;
+            }
+            else if( problemTypeString->compare( "Euler2D" ) == 0 ) {
+                _problemType = ProblemType::P_EULER_2D;
+            }
+            else if( problemTypeString->compare( "ShallowWater1D" ) == 0 ) {
+                _problemType = ProblemType::P_SHALLOWWATER_1D;
+            }
+            else if( problemTypeString->compare( "ShallowWater2D" ) == 0 ) {
+                _problemType = ProblemType::P_SHALLOWWATER_2D;
+            }
+            else {
+                log->error( "[inputfile] [general] 'problem' is invalid!\nPlease set one of the following types: Burgers1D, Euler1D, "
+                            "Euler2D,ShallowWater1D,ShallowWater2D" );
+                validConfig = false;
+            }
+        }
+        else {
+            log->error( "[inputfile] [general] 'problem' not set!\nPlease set one of the following types: Burgers, Euler, Euler2D" );
+            validConfig = false;
+        }
+        auto outputDir = general->get_as<std::string>( "outputDir" );
+        if( outputDir ) {
+            _outputDir = *outputDir;
+        }
+        else {
+            log->error( "[inputfile] [general] 'outputDir' not set!" );
[…]

### 2019-02-22 code/src/settings.h

@@ -71,10 +71,11 @@ class Settings
     // problem specific settings
     double _gamma;
 
-    Settings() {}
+    Settings() = delete;
 
   public:
     Settings( std::string inputFile );
+    Settings( const std::istringstream& inputStream );
     ~Settings();
 
     ProblemType GetProblemType() const;

### 2019-02-22 Paper1/framework.tex

@@
 \end{align*}
-with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{u})=\langle f(\mathcal{U}_{ME}(\bm{\hat u}))\bm{\varphi} \rangle$, where we omit the index ME in the following for efficiency of exposition.
+with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle f(\mathcal{U}_{ME}(\bm{\hat u}))\bm{\varphi} \rangle$, where we omit the index ME in the following for efficiency of exposition.
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
@@
 
-When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the deterministic problem \eqref{eq:origProblem}, then for the IPM system we can simply take
+When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the deterministic problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take
 \begin{align*}
- \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle g(u_{ME}(\hat\Lambda(\bm{\hat u}_{j}^{n})),u_{ME}(\hat\Lambda(\bm{\hat u}_{j+1}^{n})))\bm{\varphi}\rangle.
+ \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}_{ME}(\bm{\hat u}_j^n),\mathcal{U}_{ME}(\bm{\hat u}_{j+1}^n))\bm{\varphi}\rangle.
 \end{align*}
 This choice of the numerical flux is a common choice in kinetic theory and is called kinetic flux.
-The time update of the moment vector now becomes
-\begin{align}\label{eq:exactUpdate}
-\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\hat \Lambda_j^n),u_{ME}(\hat \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\hat \Lambda_{j-1}^n),u_{ME}(\hat \Lambda_{j}^n))\bm{\varphi} \rangle\right),
-\end{align}
-where $\hat\Lambda_{j}^n :=\hat\Lambda(\bm{\hat u}_{j}^{n})$ for all $j$.
-Note that the computation of $\hat\Lambda_{j}^n$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
+
+The numerical flux requires evaluating the closure $\mathcal{U}_{ME}(\bm{\hat u}_j^n)$. To simplify notation, we define
+\begin{align*}
+\bm{u}_{ME}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
+\end{align*}
+meaning that we can write the closure as
+\begin{align*}
+\mathcal{U}_{ME}(\bm{\hat u}_j^n) = \bm{u}_{ME}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
+\end{align*}
+The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
+Commonly, this optimization problem is solved with Newton's method:
 

### 2019-02-22 Paper1/main.tex
continued

@@
 \newtheorem{proposition}{Proposition}
+\newtheorem{remark}[theorem]{Remark}
 
@@
 Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
-Intrusive methods require the implementation of a new code and yield
-
+A main drawbacks of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver.
+Furthermore, intrusive methods come with increased computational costs, making it hard to compete with non-intrusive methods.
+In this work, we present a code framework, which facilitates the implementation of intrusive methods by recycling parts of deterministic codes.
+Additionally, we introduce methods to decrease numerical costs of intrusive methods for steady problems.
+We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as the shallow water equations.
 \end{abstract}
@@
 \begin{keyword}
-conservation laws, hyperbolic, intrusive, UQ, IPM, SG, Collocation
+uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation
 \end{keyword}
@@
 \input{results}
+\input{summary_outlook}
 

### 2019-02-22 code/src/closure.cpp
restart with higher order

@@ -10,7 +10,8 @@
 #include "stochasticgalerkin.h"
 
 Closure::Closure( Settings* settings )
-    : _settings( settings ), _nMoments( _settings->GetNMoments() ), _nQuadPoints( _settings->GetNQuadPoints() ), _nStates( _settings->GetNStates() ) {
+    : _settings( settings ), _nMoments( _settings->GetNMoments() ), _nQuadPoints( _settings->GetNQuadPoints() ), _nStates( _settings->GetNStates() ),
+      _maxIterations( _settings->GetMaxIterations() ) {
     _log = spdlog::get( "event" );
     // initialize classes: Basis Functions and Quadrature rules are defined for Legendre and Hermite
     _basis.resize( 2 );
@@ -144,23 +145,22 @@ Closure* Closure::Create( Settings* settings ) {
 void Closure::SolveClosure( Matrix& lambda, const Matrix& u ) {
     int maxRefinements = 1000;
 
-    Matrix H( _nStates * _nTotal, _nStates * _nTotal );
     Vector g( _nStates * _nTotal );
-    Vector dlambdaNew( _nStates * _nTotal );
 
     // check if initial guess is good enough
     Gradient( g, lambda, u );
-    // std::cout << "Gradient = " << g << std::endl;
     if( CalcNorm( g ) < _settings->GetEpsilon() ) {
         return;
     }
+    Matrix H( _nStates * _nTotal, _nStates * _nTotal );
+    Vector dlambdaNew( _nStates * _nTotal );
     // std::cout << "before first Hessian inversion..." << std::endl;
     // calculate initial Hessian and gradient
     Vector dlambda = -g;
     // std::cout << g << std::endl;
     Hessian( H, lambda );
     posv( H, g );
-    if( _settings->GetMaxIterations() == 1 ) {
+    if( _maxIterations == 1 ) {
         AddMatrixVectorToMatrix( lambda, -_alpha * g, lambda );
         return;
     }
@@ -168,7 +168,7 @@ void Closure::SolveClosure( Matrix& lambda, const Matrix& u ) {
     AddMatrixVectorToMatrix( lambda, -_alpha * g, lambdaNew );
     Gradient( dlambdaNew, lambdaNew, u );
     // perform Newton iterations
-    for( unsigned l = 0; l < _settings->GetMaxIterations(); ++l ) {
+    for( unsigned l = 0; l < _maxIterations; ++l ) {
[…]

### 2019-02-22 code/src/closure.h

@@ -29,11 +29,11 @@ class Closure
     unsigned _numDimXi;
     unsigned _nQTotal;
     unsigned _nTotal;
+    unsigned _maxIterations;
     void Hessian( Matrix& H, const Matrix& lambda );
     void Gradient( Vector& g, const Matrix& lambda, const Matrix& u );
     std::shared_ptr<spdlog::logger> _log;
     Matrix _dUdLambda;    // preallocated memory dor computation of Hessian
-
   public:
     /**
      * constructor of class Closure
@@ -94,6 +94,18 @@ class Closure
      * @param new step size
      */
     void SetAlpha( double alpha );
+
+    /**
+     * reset max iteration number for Newton
+     * @param new max iteration number
+     */
+    void SetMaxIterations( unsigned maxIterations );
+
+    /**
+     * get max iteration number for Newton
+     * @return max iteration number
+     */
+    unsigned GetMaxIterations() const;
 
     std::vector<Polynomial*> GetBasis();
     std::vector<Polynomial*> GetQuadrature();

### 2019-02-22 code/src/momentsolver.cpp

@@ -31,28 +31,38 @@ MomentSolver::~MomentSolver() {
 
 void MomentSolver::Solve() {
     auto log = spdlog::get( "event" );
+    Closure* prevClosure;
+    Settings* prevSettings;
 
     std::chrono::steady_clock::time_point tic = std::chrono::steady_clock::now();
 
     // create solution fields
     MatVec u( _nCells, Matrix( _nStates, _nTotal ) );
     if( _settings->HasRestartFile() ) {
-        this->ImportPrevSettings();
-        u = this->ImportPrevMoments();
-        if( _settings->LoadLambda() ) {
-            _lambda = this->ImportPrevDuals();
+        prevSettings = this->ImportPrevSettings();
+        prevSettings->SetNStates( _settings->GetNStates() );
+        prevSettings->SetGamma( _settings->GetGamma() );
+        if( prevSettings->GetNMoments() != _settings->GetNMoments() ) {
+            prevClosure = Closure::Create( prevSettings );
         }
         else {
-            _lambda = MatVec( _nCells + 1, Matrix( _nStates, _nTotal ) );
+            prevClosure = _closure;
+        }
+
+        u = this->ImportPrevMoments( prevSettings->GetNTotal() );
+        if( _settings->LoadLambda() ) {
+            _lambda = this->ImportPrevDuals( prevSettings->GetNTotal() );
+        }
+        else {
+            _lambda = MatVec( _nCells + 1, Matrix( _nStates, prevSettings->GetNTotal() ) );
         }
     }
     else {
-        u       = SetupIC();
-        _lambda = MatVec( _nCells + 1, Matrix( _nStates, _nTotal ) );
+        u           = SetupIC();
+        _lambda     = MatVec( _nCells + 1, Matrix( _nStates, _nTotal ) );
+        prevClosure = _closure;
     }
-    MatVec uNew = u;
-    MatVec uOld = u;
-    MatVec uQ   = MatVec( _nCells + 1, Matrix( _nStates, _settings->GetNqPE() ) );
+    MatVec uQ = MatVec( _nCells + 1, Matrix( _nStates, _settings->GetNqPE() ) );
 
     Vector ds( _nStates );
     Vector u0( _nStates );
@@ -77,20 +87,49 @@ void MomentSolver::Solve() {
     }
 
     // Converge initial condition entropy variables for One Shot IPM
-    if( _settings->GetMaxIterations() == 1 ) {
-
-        _settings->SetMaxIterations( 10000 );
[…]

### 2019-02-22 code/src/momentsolver.h

@@ -40,10 +40,9 @@ class MomentSolver
     Vector EvalLambda( const Vector& lambda, const Vector& xi );
     void Plot( double time, unsigned nSteps );
     void Export( const MatVec& u, const MatVec& lambda ) const;
-    void ImportPrevSettings();
-    MatVec ImportPrevMoments();
-    MatVec ImportPrevDuals();
-    Vector CalculateErrorVar( Matrix solution, unsigned LNorm );
+    Settings* ImportPrevSettings();
+    MatVec ImportPrevMoments( unsigned nPrevTotal );
+    MatVec ImportPrevDuals( unsigned nPrevTotal );
 
   public:
     MomentSolver( Settings* settings, Mesh* mesh, Problem* problem );

### 2019-02-22 code/src/settings.cpp

@@ -552,6 +552,7 @@ bool Settings::HasExactSolution() const { return _hasExactSolution; }
 ClosureType Settings::GetClosureType() const { return _closureType; }
 unsigned Settings::GetNMoments() const { return _nMoments; }
 unsigned Settings::GetNQuadPoints() const { return _nQuadPoints; }
+void Settings::SetNQuadPoints( unsigned nqNew ) { _nQuadPoints = nqNew; }
 unsigned Settings::GetNQTotal() const { return _nQTotal; }
 bool Settings::UsesMaxDegree() const { return _useMaxDegree; }
 LimiterType Settings::GetLimiterType() const { return _limiterType; }

### 2019-02-22 code/src/settings.h

@@ -119,6 +119,7 @@ class Settings
     unsigned GetNMoments() const;
     unsigned GetNTotal() const;
     unsigned GetNQuadPoints() const;
+    void SetNQuadPoints( unsigned nqNew );
     unsigned GetNQTotal() const;
     bool UsesMaxDegree() const;
     LimiterType GetLimiterType() const;

### 2019-02-22 code/src/momentsolver.cpp
git errors corrected

@@ -192,13 +192,13 @@ void MomentSolver::Solve() {
 
     // MPI Broadcast final moment vectors to all PEs
     for( unsigned j = 0; j < _nCells; ++j ) {
-        MPI_Bcast( uNew[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, PEforCell[j], MPI_COMM_WORLD );
+        MPI_Bcast( u[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, PEforCell[j], MPI_COMM_WORLD );
     }
 
     if( _settings->GetMyPE() != 0 ) return;
 
-    // save final moments on u
-    u = uNew;
+    // save final moments on uNew
+    uNew = u;
 
     std::chrono::steady_clock::time_point toc = std::chrono::steady_clock::now();
     log->info( "" );

### 2019-02-22 code/src/momentsolver.h

@@ -43,6 +43,7 @@ class MomentSolver
     Settings* ImportPrevSettings();
     MatVec ImportPrevMoments( unsigned nPrevTotal );
     MatVec ImportPrevDuals( unsigned nPrevTotal );
+    Vector CalculateErrorVar( Matrix solution, unsigned LNorm );
 
   public:
     MomentSolver( Settings* settings, Mesh* mesh, Problem* problem );

### 2019-02-22 code/src/main.cpp
bugfix moment export

@@ -193,6 +193,7 @@ int main( int argc, char* argv[] ) {
     // perform reduction to obtain full moments on all PEs
     std::vector<int> PEforCell = settings->GetPEforCell();
     for( unsigned j = 0; j < nCells; ++j ) {
+        u[j].reset();
         MPI_Reduce( uMoments[j].GetPointer(), u[j].GetPointer(), int( nStates * nTotal ), MPI_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD );
     }
 

### 2019-02-22 code/src/momentsolver.cpp

@@ -121,8 +121,8 @@ void MomentSolver::Export( const MatVec& u ) const {
     for( unsigned i = 0; i < _nCells; ++i ) {
         std::stringstream line;
         for( unsigned j = 0; j < _nStates; ++j ) {
-            for( unsigned k = 0; k < 1; ++k ) {
-                line << u[i]( j, 0 ) << ",";
+            for( unsigned k = 0; k < _nTotal; ++k ) {
+                line << u[i]( j, k ) << ",";
             }
         }
         moment_writer->info( line.str() );

### 2019-02-24 code/src/momentsolver.cpp
residual without dt added

@@ -182,12 +182,12 @@ void MomentSolver::Solve() {
         }
 
         for( unsigned j = 0; j < cellIndexPE.size(); ++j ) {
-            residual += std::fabs( u[cellIndexPE[j]]( 0, 0 ) - uOld[cellIndexPE[j]]( 0, 0 ) ) * _mesh->GetArea( cellIndexPE[j] ) / _dt;
+            residual += std::fabs( u[cellIndexPE[j]]( 0, 0 ) - uOld[cellIndexPE[j]]( 0, 0 ) ) * _mesh->GetArea( cellIndexPE[j] );
         }
 
         MPI_Reduce( &residual, &residualFull, 1, MPI_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD );
         if( _settings->GetMyPE() == 0 ) {
-            log->info( "{:03.8f}   {:01.5e}", t, residualFull );
+            log->info( "{:03.8f}   {:01.5e}   {:01.5e}", t, residualFull, residualFull / dt );
         }
     }
 

### 2019-02-24 Paper1/framework.tex
restart with higher order
residual without dt added
paper continued

@@
 \subsection{Discretization}
-In the following, we discretize the moment system in space and time following according to \cite{kusch2017maximum}.
+In the following, we discretize the moment system in space and time according to \cite{kusch2017maximum}.
 Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:IPMmomentSystem}.
@@
 \end{align*}
-with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle f(\mathcal{U}_{ME}(\bm{\hat u}))\bm{\varphi} \rangle$, where we omit the index ME in the following for efficiency of exposition.
+with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle f(\mathcal{U}(\bm{\hat u}))\bm{\varphi} \rangle$.
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
-First we perform the discretization of the spatial domain: We choose the discrete unknowns to be the spatial averages over each cell at time $t_n$, given by
+First we perform the discretization of the spatial domain: We choose the discrete unknowns which represent the solution to be the spatial averages over each cell at time $t_n$, given by
 \begin{align*}
@@
 for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
-The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\bm{u},\bm{u})=\bm{F}(\bm{u})$.
+The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\hat{\bm{u}},\hat{\bm{u}})=\bm{F}(\hat{\bm{u}})$.
 To ensure stability, a CFL condition has to be derived by investigating the eigenvalues of $\nabla \bm{F}$.
@@
 \begin{align*}
- \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}_{ME}(\bm{\hat u}_j^n),\mathcal{U}_{ME}(\bm{\hat u}_{j+1}^n))\bm{\varphi}\rangle.
+ \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}\rangle.
 \end{align*}
-This choice of the numerical flux is a common choice in kinetic theory and is called kinetic flux.
 
-The numerical flux requires evaluating the closure $\mathcal{U}_{ME}(\bm{\hat u}_j^n)$. To simplify notation, we define
+The numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define
 \begin{align*}
-\bm{u}_{ME}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
+\bm{u}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
 \end{align*}
@@
 \begin{align*}
-\mathcal{U}_{ME}(\bm{\hat u}_j^n) = \bm{u}_{ME}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
+\mathcal{U}(\bm{\hat u}_j^n) = \bm{u}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
 \end{align*}
 The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
-Commonly, this optimization problem is solved with Newton's method:
+We wish to minimize the cost function
+\begin{align*}
+L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i,
+\end{align*}
+i.e. we need to find the root of
+\begin{align*}
+\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla_{\bm{u}}s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi\rangle - \bm{\hat u}_i,
+\end{align*}
+which is done by an iterative method. Defining the dual iteration function
+\begin{align*}
+\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}\cdot \left(\langle \bm u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{\hat{u}}\right),
+\end{align*}
+the iteration process for spatial cell $j$ is given by
+\begin{align}\label{eq:dualIteration1}
+\bm{\lambda}^{m+1}_j = \bm{d}(\bm{\lambda}_j,\bm{\hat{u}_j})
+\end{align}
+and the exact dual state is then obtained by converging this iteration, i.e. $\bm{\hat\lambda}_j^n=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^m,\bm{\hat{u}_j^n})$.
 
@@
 \begin{align}\label{eq:tauCrit}
-\left\Vert \bm{\hat u}_{j}^{n}-\left\langle u_{ME}\left(\left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau.
+\left\Vert \bm{\hat u}_{j}^{n}-\left\langle u\left(\left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau.
 \end{align}
@@
 \FOR{$j=1$ to $N_x$}
-\STATE $\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\xoverline \Lambda_j^n),u_{ME}(\xoverline \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\xoverline \Lambda_{j-1}^n),u_{ME}(\xoverline \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
+\STATE $\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u(\xoverline \Lambda_j^n),u(\xoverline \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u(\xoverline \Lambda_{j-1}^n),u(\xoverline \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
 \ENDFOR

### 2019-02-24 Paper1/framework.tex
numerics continued

@@
 The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\hat{\bm{u}},\hat{\bm{u}})=\bm{F}(\hat{\bm{u}})$.
-To ensure stability, a CFL condition has to be derived by investigating the eigenvalues of $\nabla \bm{F}$.
 
@@
 \begin{align*}
-\bm{u}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
+\bm{u}_{ME}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
 \end{align*}
@@
 \begin{align*}
-\mathcal{U}(\bm{\hat u}_j^n) = \bm{u}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
+\mathcal{U}(\bm{\hat u}_j^n) = \bm{u}_{ME}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
 \end{align*}
 The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
-We wish to minimize the cost function
+To determine the dual variables for a given moment vector $\bm{\hat{u}}$, we must minimize the cost function
 \begin{align*}
@@
 \begin{align*}
-\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}\cdot \left(\langle \bm u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{\hat{u}}\right),
+\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{ME}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{\hat{u}}\right),
 \end{align*}
-the iteration process for spatial cell $j$ is given by
+with an adequate preconditioner $\bm{B}$, the iteration process for spatial cell $j$ is given by
 \begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{m+1}_j = \bm{d}(\bm{\lambda}_j,\bm{\hat{u}_j})
+\bm{\lambda}^{(m+1)}_j = \bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j}).
 \end{align}
-and the exact dual state is then obtained by converging this iteration, i.e. $\bm{\hat\lambda}_j^n=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^m,\bm{\hat{u}_j^n})$.
-
-Unfortunately \eqref{eq:exactUpdate} cannot be implemented because the dual problem cannot be solved exactly.%
-\footnote{
-Equation \eqref{eq:exactUpdate} also includes integral evaluations which cannot be computed in closed form.
-Their approximation by numerical quadrature, however, does not play a role in the realizability problems we discuss below.
-}
-Instead, it must be solved numerically, for example with Newton's method.
-The stopping criterion for the numerical optimizer ensures that the approximate multiplier vector it returns, which we denote $\xoverline{\bm{\lambda}}_j^n\in\mathbb{R}^{N+1}$ for the moment vector $\bm{\hat u}_{j}^{n}$, satisfies the stopping criterion
+The exact dual state is then obtained by converging this iteration, i.e.
+$\bm{\hat\lambda}_j^n=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
+A common preconditioner is the inverse Hessian of the dual problem, i.e.
+\begin{align*}
+\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}.
+\end{align*}
+To obtain a finite number of iterations, a stopping criterion 
 \begin{align}\label{eq:tauCrit}
-\left\Vert \bm{\hat u}_{j}^{n}-\left\langle u\left(\left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau.
+\left\Vert \bm{\hat u}_{j}^{n}-\left\langle \bm{u}_{ME}\left(\left(\hat{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau
 \end{align}
-This is derived from the first-order necessary conditions for the dual problem.
-Once the numerical optimizer finds such a $\xoverline{\bm{\lambda}}_j^n$, the corresponding dual state $\xoverline{\Lambda}_j^n := \left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi} \in \mathbb{P}(\Theta)$ can be used in \eqref{eq:exactUpdate} for the unknown $\hat \Lambda_j^n$.
-This gives Algorithm \ref{alg:seq}.
+is used. This gives Algorithm \ref{alg:seq}.
 
@@
 \FOR{$j=0$ to $N_x+1$}
-\STATE $\xoverline{\bm{\lambda}}_j^n \approx \argmin_{\bm{\lambda}}  \left( \langle s_*(\bm{\lambda}^T \bm{\varphi})\rangle - \bm{\lambda}^T \bm{\hat u}_{j}^{n} \right)$
+\STATE $\hat{\bm{\lambda}}_j^n \approx \argmin_{\bm{\lambda}}  \left( \langle s_*(\bm{\lambda}^T \bm{\varphi})\rangle - \bm{\lambda}^T \bm{\hat u}_{j}^{n} \right)$
 \hfill such that \eqref{eq:tauCrit} holds
-\STATE $\xoverline \Lambda_j^n = \left(\xoverline{\bm{\lambda}}_j^n\right)^T\bm{\varphi}$
+\STATE $\hat \Lambda_j^n = \left(\hat{\bm{\lambda}}_j^n\right)^T\bm{\varphi}$
 \ENDFOR
 \FOR{$j=1$ to $N_x$}
-\STATE $\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u(\xoverline \Lambda_j^n),u(\xoverline \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u(\xoverline \Lambda_{j-1}^n),u(\xoverline \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
+\STATE $\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(\bm{u}_{ME}(\hat \Lambda_j^n),\bm{u}_{ME}(\hat \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(\bm{u}_{ME}(\hat \Lambda_{j-1}^n),\bm{u}_{ME}(\hat \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
 \ENDFOR
@@
 \end{algorithm}
+
+\subsection{Code framework}

### 2019-02-26 Paper1/framework.tex
paper small changes

@@
 \end{align*}
-with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle f(\mathcal{U}(\bm{\hat u}))\bm{\varphi} \rangle$.
+with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
@@
 \begin{align*}
- \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}\rangle.
+ \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T.
 \end{align*}
-
 The numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define
@@
 \end{align*}
-meaning that we can write the closure as
+meaning that the closure \eqref{eq:ansatz} can be written as
 \begin{align*}
@@
 The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
-To determine the dual variables for a given moment vector $\bm{\hat{u}}$, we must minimize the cost function
+To determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
 \begin{align*}
-L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i,
+L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i
 \end{align*}
-i.e. we need to find the root of
+needs to be minimized, i.e. one needs to find the root of
 \begin{align*}
-\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla_{\bm{u}}s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi\rangle - \bm{\hat u}_i,
+\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla_{\bm{u}}s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle^T - \bm{\hat u}_i,
 \end{align*}
-which is done by an iterative method. Defining the dual iteration function
+which is usually done by an iterative method. Defining the dual iteration function
 \begin{align*}
-\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{ME}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{\hat{u}}\right),
+\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{ME}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T-\bm{\hat{u}}\right),
 \end{align*}

### 2019-02-26 Paper1/framework.tex
framework continued

@@
 \begin{align*}
-\bm{u}_{ME}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
+\bm{u}_{s}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
 \end{align*}
@@
 \begin{align*}
-\mathcal{U}(\bm{\hat u}_j^n) = \bm{u}_{ME}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
+\mathcal{U}(\bm{\hat u}_j^n) = \bm{u}_{s}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
 \end{align*}
@@
 \end{align*}
-which is usually done by an iterative method. Defining the dual iteration function
+which is usually done by an iterative method. Defining the dual iteration function $\bm{d}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$,
 \begin{align*}
-\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{ME}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T-\bm{\hat{u}}\right),
+\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T-\bm{\hat{u}}\right),
 \end{align*}
@@
 \end{align}
-The exact dual state is then obtained by converging this iteration, i.e.
+The exact dual state is then obtained by computing the fix point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
 $\bm{\hat\lambda}_j^n=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
@@
 \begin{align*}
-\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}.
+\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-T}.
 \end{align*}
-To obtain a finite number of iterations, a stopping criterion 
+To obtain a finite number of iterations for the iteration in cell $j$, a stopping criterion 
 \begin{align}\label{eq:tauCrit}
-\left\Vert \bm{\hat u}_{j}^{n}-\left\langle \bm{u}_{ME}\left(\left(\hat{\bm{\lambda}}_j^n\right)^T\bm{\varphi}\right)\bm{\varphi}\right\rangle \right\Vert < \tau
+\sum_{i=0}^p\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(m)};\bm{\hat{u}}_j^n) \right\Vert < \tau
 \end{align}
-is used. This gives Algorithm \ref{alg:seq}.
+is used.
+
+We now write down the entire scheme: To obtain a more compact notation, one defines
+\begin{align*}
+\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle \bm u_{s}(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}^T\rangle^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}^T\rangle^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{\ell}^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}^T\rangle^T\right).
+\end{align*}
+The moment iteration is then given by
+\begin{align}\label{eq:momentIteration}
+\bm{\hat u}_j^{n+1} = \bm{c}\left(\bm{\lambda}(\bm{\hat u}_{j-1}^n),\bm{\lambda}(\bm{\hat u}_{j}^n),\bm{\lambda}(\bm{\hat u}_{j+1}^n)\right),
+\end{align}
+where the map from the moment vector to the dual variables, i.e. $\bm{\lambda}(\bm{\hat u}_{j}^n)$, is obtained by iterating
+\begin{align}\label{eq:dualIteration}
+\bm{\lambda}_j^{(m+1)} = \bm{d}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n}).
+\end{align}
+until condition \eqref{eq:tauCrit} is fulfilled. This gives Algorithm \ref{alg:seq}.
 
@@
 \begin{algorithmic}[1]
-\FOR{$j=0$ to $N_x+1$}
-\STATE $\bm{u}_j^0 = \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_0(x, \cdot) \bm{\varphi} \rangle dx$
-\ENDFOR
-\FOR{$n=0$ to $N_t$}
-\FOR{$j=0$ to $N_x+1$}
-\STATE $\hat{\bm{\lambda}}_j^n \approx \argmin_{\bm{\lambda}}  \left( \langle s_*(\bm{\lambda}^T \bm{\varphi})\rangle - \bm{\lambda}^T \bm{\hat u}_{j}^{n} \right)$
-\hfill such that \eqref{eq:tauCrit} holds
-\STATE $\hat \Lambda_j^n = \left(\hat{\bm{\lambda}}_j^n\right)^T\bm{\varphi}$
-\ENDFOR
-\FOR{$j=1$ to $N_x$}
-\STATE $\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(\bm{u}_{ME}(\hat \Lambda_j^n),\bm{u}_{ME}(\hat \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(\bm{u}_{ME}(\hat \Lambda_{j-1}^n),\bm{u}_{ME}(\hat \Lambda_{j}^n))\bm{\varphi} \rangle\right)$
-\ENDFOR
-\ENDFOR
+\For{$j=0$ to $N_x+1$}
+\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle dx$
+\EndFor
+\For{$n=0$ to $N_t$}
+\For{$j=0$ to $N_x+1$}
+\State $\bm{\lambda}_j^{(0)} \leftarrow \bm{\hat \lambda}_j^{n}$
+\While{\eqref{eq:tauCrit} is violated}
+\State $\bm{\lambda}_j^{(m+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n})$
+\State $m \leftarrow m+1$
+\EndWhile
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(m)}$
+\EndFor
+\For{$j=1$ to $N_x$}
+\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\hat \lambda}_{j-1}^{n+1},\bm{\hat \lambda}_j^{n+1},\bm{\hat \lambda}_{j+1}^{n+1})$
+\EndFor
+\EndFor
 \end{algorithmic}
-\caption{IPM for Uncertainty Quantification}
+\caption{IPM implementation}
 \label{alg:seq}

### 2019-02-26 Paper1/introduction.tex
paper continued
framework continued

@@
 \begin{align}
-\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{F}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0}, \\ \label{eq:ic}
+\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D, \\ \label{eq:ic}
 \bm{u}(t=0,\bm{x},&\bm{\xi}) = \bm{u}_{\text{IC}}(\bm{x},\bm{\xi}),
@@
 \end{subequations}
-where the solution $u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subset \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subset\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
-The physical flux is given by $\bm{F}:\mathbb{R}^p\to\mathbb{R}^p$.
-To simplify notation, we assume that the initial condition is random, i.e.
+where the solution $\bm u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
+The physical flux is given by $\bm{f}:\mathbb{R}^p\to\mathbb{R}^{d\times p}$.
+To simplify notation, we assume that only the initial condition is random, i.e.
 $\bm{\xi}$ enters through the definition of $\bm{u}_{IC}$.
@@
 where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot \prod_{i=1}^s f_{\Xi,i}(\xi_i)d\xi_1 \cdots d\xi_s$.
-More generally, one is interested in determining the moments of the solution for a given set of orthonormal basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq N$.
+More generally, one is interested in determining the moments of the solution for a given set of basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq N$.
+Commonly one chooses orthonormal basis functions, i.e.
+$\langle \varphi_n \varphi_m \rangle = \delta_{nm}$.
 The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$.\\
@@
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a big implementation effort.
-Furthermore, it is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
-A downside of collocation methods are aliasing effects, which stem from the inexact approximation of the integral.
+Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
+A downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
 Furthermore, collocation methods require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation}.
-Despite their easy implementation, collocation methods can become cumbersome for unsteady problems, since in this case, the solution needs to written out at all quadrature points in every time step to compute the time evolution of the moments.
+Despite their easy implementation, collocation methods can become cumbersome for unsteady problems, since in this case, the solution needs to be written out, i.e.
+stored in an external file, at all quadrature points in every time step to compute the time evolution of the moments.
 This problem does not occur in intrusive methods, since the computation is carried out on the moments, i.e.
 the time evolution can be recorded during the computation.
-Also for steady problems, the fact that the computation is carried out on the moments provided the ability to compute the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
-However intrusive methods are in general more difficult to implement and come with higher numerical costs.
-The main idea of these methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
+Also for steady problems, the fact that the computation is carried out on the moments enables the computation of the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
+However intrusive methods are in general more difficult to implement and come along with higher numerical costs.
+The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
 \begin{subequations}\label{eq:nonClosedMomentSystem}
 \begin{align}
-\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{F}(\bm{u}(t,\bm{x},\cdot)) \varphi_i\rangle = \bm{0}, \\
+\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\bm{u}(t,\bm{x},\cdot)) \varphi_i\rangle = \bm{0}, \\
 \bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
@@
 \begin{align*}
-\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi},
+\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
 \end{align*}
-where we collect moments in $\hat{u}:=(\hat{u}_0,\cdots,\hat{u}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
+Here, we collect the moments in $\bm{\hat{u}}:=(\bm{\hat{u}}_0,\cdots,\bm{\hat{u}}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
 When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic, meaning that it cannot be solved with standard methods.
@@
 \begin{align}\label{eq:primalProblem}
-\mathcal{U}_{ME}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
+\mathcal{U}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
 \end{align}
@@
 \begin{align}\label{eq:ansatz}
- \mathcal{U}_{ME}(\bm{\hat u}) = \left( s_{\bm{\hat u}} \right)^{-1}(\bm{\hat{\lambda}}(\bm{u})^T \bm{\varphi}).
+ \mathcal{U}(\bm{\hat u}) = \left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm{\hat{\lambda}}(\bm{\hat u})^T \bm{\varphi}).
 \end{align}
@@
 \begin{align}
-\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{F}(\mathcal{U}_{ME}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
+\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
 \bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
@@
 The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin method, i.e.
-IPM generalizes different intrusive method.
+IPM generalizes different intrusive methods.
 Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
@@
 \begin{align*}
-S(\bm{\hat u}) := \langle s( \mathcal{U}_{ME}(\bm{\hat u}))\rangle,
+S(\bm{\hat u}) := \langle s( \mathcal{U}(\bm{\hat u}))\rangle,
 \end{align*}
@@
 the IPM method yields a physically correct entropy solution.
-This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution, which can then cause a failure of the method.
-The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves the computation of the optimization problem \eqref{eq:dualProblem}.
-Hence, the desirable costs of IPM (and intrusive methods in general) require a significantly increased run time.
-A further advantages which we will not use in this work is the possibility of using adaptivity \cite{kroker2012finite}, i.e.
+This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution values, which can then cause a failure of the method.
+The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
+Hence, the desirable properties of IPM come along with a significantly increased run time.
+A further advantage of intrusive methods in general is the possibility of using adaptivity \cite{kroker2012finite,kusch2018filtered}, i.e.
 refining the stochastic space at certain spatial position and time steps in which complex structures such as discontinuities occur.
-The main task here is to find an adequate refinement indicator \cite{kroker2012finite,giesselmann2017posteriori}.\\
+The main task here is to find an adequate refinement indicator \cite{kroker2012finite,giesselmann2017posteriori}.
+In this paper we compare methods without using refinement strategies.\\
 
 In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods.
-The framework only requires the numerical flux of the deterministic problem.
-If IPM is used, the entropy of the deterministic problem needs to be provided.
+The framework only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used).
 We thereby provide the ability to recycle existing implementations of deterministic solvers.
@@
 \begin{itemize}
-\item Apply intrusive methods as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method, which we then again converge to steady state.
+\item Accelerate convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
 \item Since the moments during the iteration process are inaccurate, i.e.
@@
 The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive framework in section \ref{sec:framework}.
-Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}.
+Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration.
 A comparison of results computed with the presented methods is given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.

### 2019-02-26 code/src/closure.cpp
some debugging stuff deleated and config changed

@@ -321,7 +321,7 @@ void Closure::SolveClosureSafe( Matrix& lambda, const Matrix& u ) {
         AddMatrixVectorToMatrix( lambda, -stepSize * _alpha * g, lambdaNew );
         Gradient( dlambdaNew, lambdaNew, u );
         int refinementCounter = 0;
-        // std::cout << "Res " << CalcNorm( dlambdaNew ) << std::endl;
+        std::cout << "Res " << CalcNorm( dlambdaNew ) << std::endl;
         while( CalcNorm( dlambda ) < CalcNorm( dlambdaNew ) || !std::isfinite( CalcNorm( dlambdaNew ) ) ) {
             stepSize *= 0.5;
             AddMatrixVectorToMatrix( lambda, -stepSize * _alpha * g, lambdaNew );

### 2019-02-26 code/src/momentsolver.cpp

@@ -88,14 +88,14 @@ void MomentSolver::Solve() {
     }
 
     // Converge initial condition entropy variables for One Shot IPM
-    if( _settings->GetMaxIterations() == 1 || prevSettings->GetNMoments() != _settings->GetNMoments() || true ) {
+    if( _settings->GetMaxIterations() == 1 || prevSettings->GetNMoments() != _settings->GetNMoments() ) {
         prevClosure->SetMaxIterations( 10000 );
         for( unsigned j = 0; j < _nCells; ++j ) {
             prevClosure->SolveClosureSafe( _lambda[j], u[j] );
         }
         prevClosure->SetMaxIterations( 1 );
     }
-
+    // std::cout << "PE " << _settings->GetMyPE() << ": First dual DONE." << std::endl;
     // for restart with lower order of moments reconstruct solution at finer quad points and compute moments at higher order
     if( prevSettings->GetNMoments() != _settings->GetNMoments() ) {
         MatVec uQFullProc      = MatVec( _nCells + 1, Matrix( _nStates, _settings->GetNQTotal() ) );
@@ -120,6 +120,8 @@ void MomentSolver::Solve() {
             }
             auto uQTest = _closure->U( _closure->EvaluateLambda( _lambda[j] ) );
             auto uTest  = uQTest * _closure->GetPhiTildeWf();
+            // std::cout << "----------------------------------" << std::endl;
+            // std::cout << "u = " << u[j] << std::endl;
             _closure->SolveClosureSafe( _lambda[j], u[j] );
         }
         _closure->SetMaxIterations( maxIterations );
@@ -128,6 +130,7 @@ void MomentSolver::Solve() {
         delete prevClosure;
         delete prevSettings;
     }
+    // std::cout << "PE " << _settings->GetMyPE() << ": Second dual DONE." << std::endl;
 
     MatVec uNew = u;
     MatVec uOld = u;

### 2019-02-26 Paper1/colloc_ipm.tex
continued

@@
 \section{Collocation accelerated IPM for steady state problems}
+\label{sec:collIPM}
+
+In the following, we look at steady state problems, i.e. we have
+\begin{align}\label{eq:hyperbolicProblemSteady}
+\nabla\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D
+\end{align}
+with adequate boundary conditions.
+A general strategy for computing the steady state solution to \eqref{eq:hyperbolicProblemSteady} is to introduce a pseudo-time and numerically treat \eqref{eq:hyperbolicProblemSteady} as an unsteady problem.
+A steady state solution is then obtained by iterating in pseudo-time untill the solution remains constant.
+When computing steady state moments for this problem, the run time of different UQ methods crucially depends on the chosen initial condition and its distance to the steady state solution.

### 2019-02-26 Paper1/one_shot_ipm.tex

@@
 \begin{align*}
-\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle u(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}\rangle - \frac{\Delta t}{\Delta x}\left(\langle g(u(\bm{\lambda}_c^T\bm{\varphi}),u(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}\rangle-\langle g(u(\bm{\lambda}_{\ell}^T\bm{\varphi}),u(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}\rangle\right).
+\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle u(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}\rangle - \frac{\Delta t}{\Delta x}\left(\langle g(u(\bm{\lambda}_c^T\bm{\varphi}),u(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}\rangle-\langle g(u(\bm{\lambda}_{\ell}^T\bm{\varphi}),u(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}\rangle\right).
 \end{align*}
@@
 Hence, we have that $\partial_{\bm{u}} \bm{c} = \bm{\tilde c}_{\bm{u}}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM iteration is contractive.
-Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fix point.\qed
+Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fix point.
 \end{proof}
@@
 Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{u})$.
-Hence there exists an open ball with radius $\delta$ around the fix point in which the eigenvalues remain in the interval $(-1,1)$.\qed
+Hence there exists an open ball with radius $\delta$ around the fix point in which the eigenvalues remain in the interval $(-1,1)$.
 \end{proof}

### 2019-02-28 Paper1/introduction.tex
Paper continued

@@
 
-Due to the randomness of the solution, one is interested in determining the expectation value or the variance of the solution, i.e.
+Due to the randomness of the solution, one is interested in determining the expectation value or the variance, i.e.
 \begin{align*}
@@
 where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot \prod_{i=1}^s f_{\Xi,i}(\xi_i)d\xi_1 \cdots d\xi_s$.
-More generally, one is interested in determining the moments of the solution for a given set of basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq N$.
-Commonly one chooses orthonormal basis functions, i.e.
+More generally, one is interested in determining the moments of the solution for a given set of basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
+Commonly one chooses orthonormal polynomials as basis functions \cite{xiu2002wiener}, i.e.
 $\langle \varphi_n \varphi_m \rangle = \delta_{nm}$.
-The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$.\\
+The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$ can for example be used to span the solution \cite{wiener1938homogeneous}.
+\\
 
 Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods.
-A popular non-intrusive method is the stochastic-Collocation (SC) method, which computes the moments with the help of a numerical quadrature rule: For a given set of $N_q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
+A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
+\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $N_q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
 \begin{align*}
@@
 stored in an external file, at all quadrature points in every time step to compute the time evolution of the moments.
-This problem does not occur in intrusive methods, since the computation is carried out on the moments, i.e.
-the time evolution can be recorded during the computation.
+
+Intrusive methods do not suffer from this problem, since the computation is directly carried out on the moments, i.e.
+their time evolution can be recorded during the computation.
 Also for steady problems, the fact that the computation is carried out on the moments enables the computation of the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
@@
 \end{align*}
-A commonly used closure is given by stochastic-Galerkin, which represents the solution by a polynomial:
+A commonly used closure is given by stochastic-Galerkin \cite{ghanem2003stochastic}, which represents the solution by a polynomial:
 \begin{align*}
@@
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.
-For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is
+For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is given by
 \begin{align}\label{eq:primalProblem}
@@
 \end{align}
-This problem can be rewritten as
+Rewritten in its dual form, the problem is given by
 \begin{align}\label{eq:dualProblem}
@@
 where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}\in \mathbb{R}^{(N+1) \times p}$ are called the dual variables.
-The solution to \eqref{eq:primalProblem} is then given by
+The solution to \eqref{eq:primalProblem} is then obtained by
 \begin{align}\label{eq:ansatz}
@@
 \end{subequations}
-The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin method, i.e.
+The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
 IPM generalizes different intrusive methods.

### 2019-02-28 Paper1/main.tex
paper continued
framework continued
Paper continued

@@
 \usepackage{graphicx}
-%\usepackage{subfig}
 \usepackage{subcaption}
-%\usepackage[]{algorithm2e}
-%\usepackage{algorithmic}
-%\usepackage{algorithmicx}
-\usepackage[linesnumbered,ruled]{algorithm2e}
-
 \usepackage{lipsum}
@@
 \usepackage{epstopdf}
-\usepackage{algorithmic}
+
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+
 \ifpdf
@@
 Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
-A main drawbacks of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver.
+A main drawback of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver.
 Furthermore, intrusive methods come with increased computational costs, making it hard to compete with non-intrusive methods.
@@
 \begin{keyword}
-uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation
+uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation, Intrusive Polynomial Moment Method
 \end{keyword}
@@
 
-%\section{Methods of Uncertainty Quantification}
-%\subsection{Collocation}
-%\subsection{Stochastic Galerkin}
-%\subsection{Intrusive Polynomial Moment Method}
-%\section{Method discussion}
-%\begin{itemize}
-%	\item kleinere Fehler [11]
-%	\item Programmieraufwand [31]
-%	\item zweiabhängig
-%	\item adaptiv (breack up black box approach )
-%	\item grobe Auflösung für Momente 
-%\end{itemize}
-%\section{UQCreator}
-%\subsection{Adaptivity}
-%\subsection{restart IPM} %relay(-strategy) IPM?
-%\subsection{one-shot IPM}
-%\subsection{adaptive IPM}
-%\section{Results}
-%\subsection{2D Euler flow over NACA0012}
-%\begin{itemize}
-%	\item AoA study
-%	\item AoA + Ma number + mehr ?
-%\end{itemize}
-%\subsection{2D shallow water dam break}
-
 \bibliographystyle{unsrt}  

### 2019-02-28 Paper1/framework.tex
Paper references added

@@
 
-When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the deterministic problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take
+When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the original problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take
 \begin{align*}
@@
 \end{align*}
-The numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define
+Note that the numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define
 \begin{align*}
-\bm{u}_{s}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
+\bm{u}_{s}(\bm\Lambda):=\left( \nabla_{\bm{u}} s \right)^{-1}(\bm\Lambda),
 \end{align*}
-meaning that the closure \eqref{eq:ansatz} can be written as
+meaning that the closure \eqref{eq:ansatz} in cell $j$ at timestep $n$ can be written as
 \begin{align*}
@@
 \begin{align*}
-\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla_{\bm{u}}s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle^T - \bm{\hat u}_i,
+\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle^T - \bm{\hat u}_i,
 \end{align*}

### 2019-02-28 Paper1/introduction.tex
Paper references added

@@
 their time evolution can be recorded during the computation.
-Also for steady problems, the fact that the computation is carried out on the moments enables the computation of the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
+Also for steady problems, the fact that the moments are known in each iteration enables the computation of the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
 However intrusive methods are in general more difficult to implement and come along with higher numerical costs.
@@
 \begin{align*}
-\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi).
+\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi).
 \end{align*}
@@
 \end{align*}
-Here, we collect the moments in $\bm{\hat{u}}:=(\bm{\hat{u}}_0,\cdots,\bm{\hat{u}}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
-When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic, meaning that it cannot be solved with standard methods.
+Here, we collect the moments for which $\vert i \vert \leq M$ holds in the moment matrix $\bm{\hat{u}}:=(\bm{\hat{u}}_0,\cdots,\bm{\hat{u}}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
+When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic and the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.
@@
 \end{align}
-where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}\in \mathbb{R}^{(N+1) \times p}$ are called the dual variables.
+where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_0,\cdots,\bm{\hat{\lambda}}_N)^T\in \mathbb{R}^{(N+1) \times p}$ are called the dual variables.
 The solution to \eqref{eq:primalProblem} is then obtained by
@@
 Hence, the desirable properties of IPM come along with a significantly increased run time.
-A further advantage of intrusive methods in general is the possibility of using adaptivity \cite{kroker2012finite,kusch2018filtered}, i.e.
-refining the stochastic space at certain spatial position and time steps in which complex structures such as discontinuities occur.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architecture, which can be used to significantly reduce runtime \cite{garrett2015optimization}.
+
+When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{pettersson2009numerical,offner2017stability}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non} tend to show incorrect discontinuities in certain regions of the physical space.
+These non-physical structures tend to dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added \cite{,kusch2018filtered}.
+Here, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies which locally increase the polynomial order \cite{kroker2012finite} or add artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur.
 The main task here is to find an adequate refinement indicator \cite{kroker2012finite,giesselmann2017posteriori}.
-In this paper we compare methods without using refinement strategies.\\
+In this paper we compare methods without using local refinement strategies.\\
 

### 2019-02-28 Paper1/colloc_ipm.tex
Paper continued

@@
-\section{Collocation accelerated IPM for steady state problems}
+\subsection{Collocation accelerated IPM}
 \label{sec:collIPM}
@@
 \begin{align}\label{eq:hyperbolicProblemSteady}
-\nabla\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D
+\nabla\cdot\bm{f}(\bm{u}(\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D
 \end{align}
@@
 A general strategy for computing the steady state solution to \eqref{eq:hyperbolicProblemSteady} is to introduce a pseudo-time and numerically treat \eqref{eq:hyperbolicProblemSteady} as an unsteady problem.
-A steady state solution is then obtained by iterating in pseudo-time untill the solution remains constant.
+A steady state solution is then obtained by iterating in pseudo-time until the solution remains constant.
+Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system 
+\begin{align}\label{eq:MomentSystemSteady}
+\nabla\cdot\langle\bm{f}(\bm{u}(\bm{x},\bm{\xi}))\bm{\varphi}^T\rangle^T = \bm{0} \enskip \text{ in } D
+\end{align}
+which is again needed for the construction of intrusive methdos.
+By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:IPMmomentSystem}, i.e.
+Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
+Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the residual
+\begin{align*}
+\sum_{j = 1}^{N_x} \Delta x_j \Vert \bm{\hat{u}}_j^n - \bm{\hat{u}}_j^{n-1} \Vert \leq \varepsilon
+\end{align*}
+is fulfilled. Note that since one is generally interested in low order moments such as the expectation value, this residual can be modified by only accounting for the zero order moments.
+
 When computing steady state moments for this problem, the run time of different UQ methods crucially depends on the chosen initial condition and its distance to the steady state solution.
+However, it is not always clear how to choose a satisfactory initial guess for the moments.
+We tackle this problem with the following method: For a bad initial guess, we iterate to the steady state solution with the help of a cheap method.
+Note that one expects to need a lot of iteration steps to arrive at a steady state solution, however the overall computational time is acceptable due to the choice of a cheap method.
+The obtained inaccurate moments are then used to start a more accurate, but at the same time more expensive method.
+Due to the choice of the initial condition, a smaller number of iterations is needed to arrive at the steady state solution of the expensive method, i.e.
+the numerical costs are again acceptable.
+
+I our case, the cheap method will be stochastic-Collocation with a small number of samples, whereas the expensive method will be IPM with a high number of moments.
+However different variants of the main idea can be applied:

### 2019-02-28 Paper1/introduction.tex
Paper continued

@@
 \begin{align}\label{eq:ansatz}
- \mathcal{U}(\bm{\hat u}) = \left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm{\hat{\lambda}}(\bm{\hat u})^T \bm{\varphi}).
+ \mathcal{U}(\bm{\hat u}) = \left( \nabla_{\bm{u}} s \right)^{-1}(\bm{\hat{\lambda}}(\bm{\hat u})^T \bm{\varphi}).
 \end{align}
@@
 Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
-Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the entropy
+Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the expectation value of the entropy, which is
 \begin{align*}
@@
 
-When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{pettersson2009numerical,offner2017stability}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non} tend to show incorrect discontinuities in certain regions of the physical space.
-These non-physical structures tend to dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added \cite{,kusch2018filtered}.
-Here, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies which locally increase the polynomial order \cite{kroker2012finite} or add artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur.
-The main task here is to find an adequate refinement indicator \cite{kroker2012finite,giesselmann2017posteriori}.
+When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
+These non-physical structures tend to dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}.
+Here, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies which locally increase the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or add artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur.
+The main task here is to find an adequate refinement indicator.
 In this paper we compare methods without using local refinement strategies.\\
@@
 \item Accelerate convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
-\item Since the moments during the iteration process are inaccurate, i.e.
-they are not the steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
+\item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e.
+they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
 \end{itemize}

### 2019-02-28 Paper1/introduction.tex
mini change introduction

@@
 $\langle \varphi_n \varphi_m \rangle = \delta_{nm}$.
-The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$ can for example be used to span the solution \cite{wiener1938homogeneous}.
+The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$.
+Besides facilitating the computation of the expectation value and variance, the moments can be used to span the solution in $\bm\xi$ \cite{wiener1938homogeneous}.
 \\

### 2019-02-28 scripts/convergence_plot/plot.py
convergence plot script is now working with the current code

@@ -2,65 +2,124 @@ import matplotlib.pyplot as plt
 import argparse
 import numpy as np
 import datetime as dt
-import os, time
-
-parser = argparse.ArgumentParser()
-parser.add_argument("--file", "-f", type=str, required=True)
-parser.add_argument("--compareTo", "-c", type=str, required=False)
-parser.add_argument("--live", "-l", action='store_true', required=False)
-args = parser.parse_args()
-
-def splitLine(line, splitchar):
-    list = line.split(splitchar)
-    list = [" ".join(x.split()) for x in list]
-    return list
+import os
+import pandas as pd
+from io import StringIO
 
 def time_to_float(time):
     t = dt.datetime.strptime(time, "%Y-%m-%d %H:%M:%S.%f")
     epoch = dt.datetime.utcfromtimestamp(0)
     return (t-epoch).total_seconds()
 
-def plot(x,y,xlabel,ylabel):
+def create_label(dir, file):    
+    section_ctr = 0
+    line_ctr = 0    
+    with open(dir+'/'+file, 'r') as content:
+        lines = content.readlines()        
+        closure = ''
+        nQuad = -1
+        nMoments = -1   
+        maxIter = -1
+        restartFile = False
+        while section_ctr<3:            
+            if "#" in lines[line_ctr]:
+                line_ctr = line_ctr + 1
+                continue
+            if "=====" in lines[line_ctr]:
+                section_ctr = section_ctr + 1
+            if "closure" in lines[line_ctr]:
+                closure = lines[line_ctr].split("=")[1].strip().strip('"').strip('\n')
+            elif "moments" in lines[line_ctr]:
+                nQuad = int(lines[line_ctr].split("=")[1])
+            elif "quadPoints" in lines[line_ctr]:
+                nMoments = int(lines[line_ctr].split("=")[1])
+            elif "maxIterations" in lines[line_ctr]:
+                maxIter = int(lines[line_ctr].split("=")[1])
+            elif "restartFile" in lines[line_ctr] and not "#restartFile" in lines[line_ctr]:
+                restartFile = True
+            line_ctr = line_ctr + 1        
+        if closure == 'StochasticGalerkin' and nQuad == nMoments:
+            return 'SC'
+        elif closure == 'StochasticGalerkin':
+            return 'SG'
+        elif closure == 'Euler2D' and restartFile:
+            return 'caos-IPM'
[…]

### 2019-02-28 code/src/main.cpp
each PE writes its own config file

@@ -76,7 +76,13 @@ void initLogger( spdlog::level::level_enum terminalLogLvl, spdlog::level::level_
         std::filesystem::create_directory( outputDir + "/logs" );
     }
 
-    auto fileSink = std::make_shared<spdlog::sinks::basic_file_sink_mt>( outputDir + "/logs/" + currentDateTime() );
+    int mype;
+    MPI_Comm_rank( MPI_COMM_WORLD, &mype );
+    std::ostringstream strs;
+    strs << mype;
+    std::string strPE = strs.str();
+
+    auto fileSink = std::make_shared<spdlog::sinks::basic_file_sink_mt>( outputDir + "/logs/" + currentDateTime() + "_PE" + strPE );
     fileSink->set_level( fileLogLvl );
     fileSink->set_pattern( "%Y-%m-%d %H:%M:%S.%f | %v" );
 
@@ -187,9 +193,6 @@ int main( int argc, char* argv[] ) {
 
     auto uMoments = u;
 
-    // if( settings->GetMyPE() == 0 )
-    // std::cout << "PE " << settings->GetMyPE() << ": u before reduce: " << u[0] << std::endl;
-
     // perform reduction to obtain full moments on all PEs
     std::vector<int> PEforCell = settings->GetPEforCell();
     for( unsigned j = 0; j < nCells; ++j ) {
@@ -218,8 +221,6 @@ int main( int argc, char* argv[] ) {
             mesh->ExportShallowWater( meanAndVar );
         else
             mesh->Export( meanAndVar );
-
-        // if( settings->GetMyPE() == 0 ) std::cout << "u after reduce: " << u[0] << std::endl;
     }
 
     delete solver;

### 2019-02-28 code/src/momentsolver.cpp

@@ -73,7 +73,7 @@ MatVec MomentSolver::Solve( Vector xi ) {
         for( unsigned j = 0; j < _nCells; ++j ) {
             residual += std::fabs( uNew[j]( 0, 0 ) - uQ[j]( 0, 0 ) ) * _mesh->GetArea( j );
         }
-        if( _settings->GetMyPE() == 0 ) log->info( "{:03.8f}   {:01.5e}   {:01.5e}", t, residual, residual / dt );
+        log->info( "{:03.8f}   {:01.5e}   {:01.5e}", t, residual, residual / dt );
         // std::cout << "dt = " << dt << std::endl;
 
         // update solution

### 2019-02-28 scripts/convergence_plot/plot.py
updated script

@@ -92,11 +92,11 @@ def create_runtime_residual_plots(data, labels, threshold=np.inf):
     plt.legend(labels)
     plt.savefig("convergence_runtime_residual.pdf")
 
-def create_time_residual_plots(data, labels, threshold=np.inf):   
+def create_time_residual_plots(data, labels, tEnd=np.inf):   
     plt.cla()
     plt.clf()
     for df in data:
-        df = df[df['residual_scaled'] < threshold]
+        df = df[df['t'] < threshold]
         plt.semilogy(df['t'],df['residual_scaled'])
     plt.xlabel('Time [s]')
     plt.ylabel('Residual')
@@ -107,6 +107,7 @@ if __name__ == "__main__":
     parser = argparse.ArgumentParser()
     parser.add_argument("--dir", "-d", type=str, required=True)
     parser.add_argument("--threshold", "-t", type=str, required=False)
+    parser.add_argument("--tEnd", "-e", type=str, required=False)
     args = parser.parse_args()
     data = []
     labels = []
@@ -119,7 +120,7 @@ if __name__ == "__main__":
         labels.append(create_label(args.dir, file))
     if(args.threshold):
         create_runtime_residual_plots(data, labels, args.threshold)
-        create_time_residual_plots(data, labels, args.threshold)
+        create_time_residual_plots(data, labels, args.tEnd)
     else:
         create_runtime_residual_plots(data, labels)
         create_time_residual_plots(data, labels)

### 2019-02-28 scripts/convergence_plot/plot.py
script hotfix

@@ -96,7 +96,7 @@ def create_time_residual_plots(data, labels, tEnd=np.inf):
     plt.cla()
     plt.clf()
     for df in data:
-        df = df[df['t'] < threshold]
+        df = df[df['t'] < tEnd]
         plt.semilogy(df['t'],df['residual_scaled'])
     plt.xlabel('Time [s]')
     plt.ylabel('Residual')

### 2019-03-08 Paper1/introduction.tex
vtk plot and paper continued

@@
 Here, we collect the moments for which $\vert i \vert \leq M$ holds in the moment matrix $\bm{\hat{u}}:=(\bm{\hat{u}}_0,\cdots,\bm{\hat{u}}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
-When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic and the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
+When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty} and the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.

### 2019-03-12 Paper1/colloc_ipm.tex
vtk plot and paper continued
paper continued + momentsolver minor changes

@@
-\subsection{Collocation accelerated IPM}
-\label{sec:collIPM}
-
 In the following, we look at steady state problems, i.e. we have
@@
 A steady state solution is then obtained by iterating in pseudo-time until the solution remains constant.
-Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system 
+Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system for \eqref{eq:hyperbolicProblemSteady} given by
 \begin{align}\label{eq:MomentSystemSteady}
@@
 
+\subsection{Collocation accelerated IPM}
+\label{sec:collIPM}
+
 When computing steady state moments for this problem, the run time of different UQ methods crucially depends on the chosen initial condition and its distance to the steady state solution.
@@
 I our case, the cheap method will be stochastic-Collocation with a small number of samples, whereas the expensive method will be IPM with a high number of moments.
-However different variants of the main idea can be applied:
+However different variants of the main idea can be applied: ...

### 2019-03-12 code/src/momentsolver.cpp

@@ -201,6 +201,8 @@ void MomentSolver::Solve() {
             log->info( "{:1d}       {:01.5e}   {:01.5e}", i, l1Error[i], sqrt( l2Error[i] ) );
         }
         meanAndVarErrors = this->CalculateErrorField( meanAndVar, 1 );
+        std::cout << std::setprecision( 9 ) << meanAndVar( 0, 19080 ) << " " << _referenceSolution[19080][0] << std::endl;
+        std::cout << "error is " << meanAndVarErrors( 0, 19080 ) << std::endl;
         _mesh->Export( meanAndVarErrors, "_errors" );
     }
 

### 2019-03-12 code/src/main.cpp
minor changes

@@ -155,7 +155,6 @@ int main( int argc, char* argv[] ) {
     std::vector<Polynomial*> quad = closure->GetQuadrature();
     unsigned nCells               = settings->GetNumCells();
     unsigned nStates              = settings->GetNStates();
-    unsigned nQTotal              = settings->GetNQTotal();
     unsigned nTotal               = settings->GetNTotal();
 
     for( unsigned k = 0; k < settings->GetNQTotal(); ++k ) {
@@ -208,8 +207,8 @@ int main( int argc, char* argv[] ) {
         Matrix meanAndVar( 2 * nStates, nCells, 0.0 );
         for( unsigned j = 0; j < nCells; ++j ) {
             // expected value
-            for( unsigned i = 0; i < nStates; ++i ) {
-                meanAndVar( i, j ) = u[j]( i, 0 );
+            for( unsigned s = 0; s < nStates; ++s ) {
+                meanAndVar( s, j ) = u[j]( s, 0 );
             }
             // variance
             for( unsigned s = 0; s < nStates; ++s ) {

### 2019-03-13 Paper1/results.tex

\section{Results}
\label{sec:results}

\subsection{2D Euler equations}
We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 profile computed with different methods. To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $50$ quadrature points and compare against Collocation with $5$ collocation points as well as SG and IPM with $5$ moments and $10$ quadrature points. Furthermore, we use convergence accelerated (caIPM) as well as convergence accelerated one-shot IPM (caosIPM), which we introduced in Sections \ref{sec:collIPM} and \ref{sec:OneShotIPM} with $10$ moments and $15$ quadrature points to converge the collocation solution with $5$ quadrature points to an entropy solution with increased accuracy.

\begin{figure}[h!]
\centering
		\centering
		\includegraphics[scale=0.5]{figs/{convergence_runtime_residual}.pdf}
		\label{fig:sub1}
	\caption{Convergence to steady state with 5 MPI and 2 OMP threads.}
	\label{fig:Obstacles2D}
\end{figure}

\newgeometry{top=1.5cm, left=1.0cm}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_sc_n5_s05_aoa}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa}.png}
		\label{fig:sub1}
	\end{subfigure}

	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_sc_n50_s05_aoa_legend}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\caption{E$[\rho]$ (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM, reference solution.}
	\label{fig:Obstacles2D}
\end{figure}

\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_sc_n5_s05_aoa_var}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa_var}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa_var}.png}
		\label{fig:sub1}
	\end{subfigure}

	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa_var}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa_var}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_sc_n50_s05_aoa_var_legend}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\caption{Var$[\rho]$ (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM, reference solution.}
	\label{fig:Obstacles2D}
\end{figure}

%%% error plots %%%
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_sc_n5_s05_aoa_error_legend}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa_error_legend}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa_error_legend}.png}
		\label{fig:sub1}
	\end{subfigure}

	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa_error_legend}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa_error_legend}.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ distance to reference solution (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM.}
	\label{fig:Obstacles2D}
\end{figure}

\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_sc_n5_s05_aoa_var_error_legend}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa_var_error_legend}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa_var_error_legend}.png}
		\label{fig:sub1}
	\end{subfigure}

	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa_var_error_legend}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa_var_error_legend}.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{Var$[\rho]$ distance to reference solution (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM.}
	\label{fig:Obstacles2D}
\end{figure}
\newgeometry{top=2.5cm, bottom=2.5cm}

### 2019-03-22 code/src/main.cpp
logfile now also contains mpi and openmp threads, but no comments from the inputfile

@@ -160,13 +160,17 @@ void PrintInit( std::string configFile ) {
     log->info( "================================================================" );
     log->info( "Git commit hash:\t{0}", GIT_HASH );
     log->info( "Config file:\t{0}", configFile );
+    int nprocs;
+    MPI_Comm_size( MPI_COMM_WORLD, &nprocs );
+    log->info( "MPI Threads:\t{0}", nprocs );
+    log->info( "OMP Threads:\t{0}", omp_get_max_threads() );
     log->info( "================================================================" );
     std::ifstream ifs( configFile );
     if( ifs.is_open() ) {
         std::string line;
         while( !ifs.eof() ) {
             std::getline( ifs, line );
-            log->info( " {0}", line );
+            if( line[0] != '#' ) log->info( " {0}", line );
         }
     }
     log->info( "================================================================" );

### 2019-03-27 Paper1/figs/errorRuntime/plot.py
Paper continued

@@ -1,0 +1,180 @@
+import matplotlib.pyplot as plt
+import argparse
+import numpy as np
+import datetime as dt
+import os
+import pandas as pd
+import re
+from io import StringIO
+
+def time_to_float(time):
+    t = dt.datetime.strptime(time, "%Y-%m-%d %H:%M:%S.%f")
+    epoch = dt.datetime.utcfromtimestamp(0)
+    return (t-epoch).total_seconds()
+
+def checkForRestartFile(dir, file):
+    section_ctr = 0
+    line_ctr = 0    
+    restartFile = False
+    restartFileName = ''
+    with open(dir+'/'+file, 'r') as content:
+        lines = content.readlines()        
+        while section_ctr<3 and line_ctr<len(lines):                     
+            if "=====" in lines[line_ctr]:
+                section_ctr = section_ctr + 1
+            elif "restartFile" in lines[line_ctr] and not "#" in lines[line_ctr]:
+                restartFile = True
+                restartFileName = lines[line_ctr].split('=')[-1].split("/")[-1][:-2]
+            line_ctr = line_ctr + 1 
+    return restartFile, restartFileName
+
+def create_label(dir, file):    
+    section_ctr = 0
+    line_ctr = 0    
+    with open(dir+'/'+file, 'r') as content:
+        lines = content.readlines()        
+        closure = ''
+        nQuad = -1
+        nMoments = -1   
+        maxIter = -1
+        restartFile = False
+        while section_ctr<3:            
+            if "#" in lines[line_ctr]:
+                line_ctr = line_ctr + 1
+                continue
+            if "=====" in lines[line_ctr]:
+                section_ctr = section_ctr + 1
+            if "closure" in lines[line_ctr]:
+                closure = lines[line_ctr].split("=")[1].strip().strip('"').strip('\n')
+            elif "moments" in lines[line_ctr]:
+                nQuad = int(lines[line_ctr].split("=")[1])
+            elif "quadPoints" in lines[line_ctr]:
+                nMoments = int(lines[line_ctr].split("=")[1])
+            elif "maxIterations" in lines[line_ctr]:
+                maxIter = int(lines[line_ctr].split("=")[1])
+            elif "restartFile" in lines[line_ctr] and not "#" in lines[line_ctr]:
+                restartFile = True                
[…]

### 2019-03-27 scripts/convergence_plot/plot.py
merged all plot scripts into one and improved plotting speed by a lot

@@ -1,188 +1,0 @@
-import matplotlib.pyplot as plt
-import argparse
-import numpy as np
-import datetime as dt
-import os
-import pandas as pd
-from io import StringIO
-
-def time_to_float(time):
-    t = dt.datetime.strptime(time, "%Y-%m-%d %H:%M:%S.%f")
-    epoch = dt.datetime.utcfromtimestamp(0)
-    return (t-epoch).total_seconds()
-
-def checkForRestartFile(dir, file):
-    section_ctr = 0
-    line_ctr = 0
-    restartFile = False
-    restartFileName = ''
-    with open(dir+'/'+file, 'r') as content:
-        lines = content.readlines()
-        while section_ctr<3:
-            if "=====" in lines[line_ctr]:
-                section_ctr = section_ctr + 1
-            elif "restartFile" in lines[line_ctr] and not "#" in lines[line_ctr]:
-                restartFile = True
-                restartFileName = lines[line_ctr].split('=')[-1].split("/")[-1][:-2]
-            line_ctr = line_ctr + 1
-    return restartFile, restartFileName
-
-def getTEnd(dir,file):
-    section_ctr = 0
-    line_ctr = 0
-    tEnd = 0.0
-    with open(dir+'/'+file, 'r') as content:
-        lines = content.readlines()
-        while section_ctr<3:
-            if "=====" in lines[line_ctr]:
-                section_ctr = section_ctr + 1
-            elif "tEnd" in lines[line_ctr] and not "#" in lines[line_ctr]:
-                tEnd = float(lines[line_ctr].split('=')[-1][:-1])
-            line_ctr = line_ctr + 1
-    return tEnd
-
-def create_label(dir, file):
-    section_ctr = 0
-    line_ctr = 0
-    with open(dir+'/'+file, 'r') as content:
-        lines = content.readlines()
-        closure = ''
-        nQuad = -1
-        nMoments = -1
-        maxIter = -1
-        restartFile = False
-        while section_ctr<3:
-            if "#" in lines[line_ctr]:
-                line_ctr = line_ctr + 1
-                continue
-            if "=====" in lines[line_ctr]:
-                section_ctr = section_ctr + 1
-            if "closure" in lines[line_ctr]:
-                closure = lines[line_ctr].split("=")[1].strip().strip('"').strip('\n')
-            elif "moments" in lines[line_ctr]:
-                nQuad = int(lines[line_ctr].split("=")[1])
-            elif "quadPoints" in lines[line_ctr]:
[…]

### 2019-03-27 scripts/error_plot/plot.py

@@ -1,180 +1,0 @@
-import matplotlib.pyplot as plt
-import argparse
-import numpy as np
-import datetime as dt
-import os
-import pandas as pd
-import re
-from io import StringIO
-
-def time_to_float(time):
-    t = dt.datetime.strptime(time, "%Y-%m-%d %H:%M:%S.%f")
-    epoch = dt.datetime.utcfromtimestamp(0)
-    return (t-epoch).total_seconds()
-
-def checkForRestartFile(dir, file):
-    section_ctr = 0
-    line_ctr = 0    
-    restartFile = False
-    restartFileName = ''
-    with open(dir+'/'+file, 'r') as content:
-        lines = content.readlines()        
-        while section_ctr<3 and line_ctr<len(lines):                     
-            if "=====" in lines[line_ctr]:
-                section_ctr = section_ctr + 1
-            elif "restartFile" in lines[line_ctr] and not "#" in lines[line_ctr]:
-                restartFile = True
-                restartFileName = lines[line_ctr].split('=')[-1].split("/")[-1][:-2]
-            line_ctr = line_ctr + 1 
-    return restartFile, restartFileName
-
-def create_label(dir, file):    
-    section_ctr = 0
-    line_ctr = 0    
-    with open(dir+'/'+file, 'r') as content:
-        lines = content.readlines()        
-        closure = ''
-        nQuad = -1
-        nMoments = -1   
-        maxIter = -1
-        restartFile = False
-        while section_ctr<3:            
-            if "#" in lines[line_ctr]:
-                line_ctr = line_ctr + 1
-                continue
-            if "=====" in lines[line_ctr]:
-                section_ctr = section_ctr + 1
-            if "closure" in lines[line_ctr]:
-                closure = lines[line_ctr].split("=")[1].strip().strip('"').strip('\n')
-            elif "moments" in lines[line_ctr]:
-                nQuad = int(lines[line_ctr].split("=")[1])
-            elif "quadPoints" in lines[line_ctr]:
-                nMoments = int(lines[line_ctr].split("=")[1])
-            elif "maxIterations" in lines[line_ctr]:
-                maxIter = int(lines[line_ctr].split("=")[1])
-            elif "restartFile" in lines[line_ctr] and not "#" in lines[line_ctr]:
-                restartFile = True                
[…]

### 2019-03-27 scripts/field_plot/plot.py

@@ -1,87 +1,0 @@
-import vtk
-import argparse
-import os
-
-parser = argparse.ArgumentParser()
-parser.add_argument("--dir", "-d", type=str, required=True)
-args = parser.parse_args()
-files = os.listdir(args.dir)
-fields = ["E(ρ)", "E(ρU)", "E(ρE)", "Var(ρ)", "Var(ρU)", "Var(ρE)"]
-for file in files:
-    if file.endswith('.vtk'):        
-        reader = vtk.vtkUnstructuredGridReader()
-        reader.SetFileName(args.dir+'/'+file)
-        reader.ReadAllScalarsOn()
-        reader.ReadAllVectorsOn()
-        reader.Update() 
-        output = reader.GetOutput()
-
-        numvals = 1024
-        ctf = vtk.vtkColorTransferFunction()
-        ctf.SetColorSpaceToRGB()
-        ctf.AddRGBPoint(-1, 1, 1, 0)
-        ctf.AddRGBPoint(0, 0, 0, 0)
-        ctf.AddRGBPoint(0.293069/1.72393, 0, 0, 1)
-        ctf.AddRGBPoint(0.586138/1.72393, 0, 1, 1)
-        ctf.AddRGBPoint(0.861967/1.72393, 0, 1, 0)
-        ctf.AddRGBPoint(1.155040/1.72393, 1, 1, 0)
-        ctf.AddRGBPoint(1.448100/1.72393, 1, 0, 0)
-        ctf.AddRGBPoint(1.723930/1.72393, 0.87843, 0, 1)
-        lut = vtk.vtkLookupTable()
-        lut.SetNumberOfTableValues(numvals)
-        lut.Build()
-
-        camera = vtk.vtkCamera()
-        camera.SetPosition(0.5, 0.3, 3);
-        camera.SetFocalPoint(0.5, 0.3, 0);
-
-        for i in range(0,numvals):
-            rgb = list(ctf.GetColor(float(i)/numvals))+[1]
-            lut.SetTableValue(i,rgb)
-        
-        for field_name in fields:
-            mapper = vtk.vtkDataSetMapper()
-            mapper.SetInputData(output)
-            mapper.ScalarVisibilityOn()
-            mapper.SetColorModeToMapScalars()
-            mapper.SetLookupTable(lut)
-            mapper.SetScalarModeToUsePointFieldData()
-            mapper.SelectColorArray(field_name)
[…]

### 2019-03-27 scripts/plot/plot.py

@@ -1,0 +1,359 @@
+import matplotlib.pyplot as plt
+import argparse
+import numpy as np
+import datetime as dt
+import os
+import pandas as pd
+import re
+from io import StringIO
+import vtk
+
+def extract_intervals(sequence, num):
+    length = float(len(sequence))
+    for i in range(num):
+        yield sequence[int(np.ceil(i * length / num))]
+
+def time_to_float(time):
+    t = dt.datetime.strptime(time, "%Y-%m-%d %H:%M:%S.%f")
+    epoch = dt.datetime.utcfromtimestamp(0)
+    return (t-epoch).total_seconds()
+
+def checkForRestartFile(dir, file):
+    section_ctr = 0
+    line_ctr = 0
+    restartFile = False
+    restartFileName = ''
+    with open(dir+'/'+file, 'r') as content:
+        lines = content.readlines()
+        while section_ctr<3 and line_ctr<len(lines):
+            current_line = lines[line_ctr]
+            if "=====" in current_line:
+                section_ctr = section_ctr + 1
+            elif "restartFile" in current_line:
+                restartFile = True
+                restartFileName = current_line.split('=')[-1].split("/")[-1][:-2]
+            line_ctr = line_ctr + 1
+    return restartFile, restartFileName
+
+def getTEnd(dir,file):
+    section_ctr = 0
+    line_ctr = 0
+    tEnd = 0.0
+    with open(dir+'/'+file, 'r') as content:
+        lines = content.readlines()
+        while section_ctr<3:
+            if "=====" in lines[line_ctr]:
+                section_ctr = section_ctr + 1
+            elif "tEnd" in lines[line_ctr] and not "#" in lines[line_ctr]:
+                tEnd = float(lines[line_ctr].split('=')[-1][:-1])
+            line_ctr = line_ctr + 1
+    return tEnd
+
+def create_label(dir, file):
+    section_ctr = 0
+    line_ctr = 0
+    with open(dir+'/'+file, 'r') as content:
+        lines = content.readlines()
+        closure = ''
+        nQuad = -1
+        nMoments = -1
+        maxIter = -1
+        restartFile = False
+        while section_ctr<3:
+            current_line = lines[line_ctr]
+            if "#" not in current_line:
+                if "=====" in current_line:
+                    section_ctr = section_ctr + 1
[…]

### 2019-04-08 Paper1/introduction.tex
paper continued

@@
 where the solution $\bm u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
+Hence, the probability density function of $\bm{\xi}$ is then given by $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
 The physical flux is given by $\bm{f}:\mathbb{R}^p\to\mathbb{R}^{d\times p}$.
@@
 \end{align*}
-where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot \prod_{i=1}^s f_{\Xi,i}(\xi_i)d\xi_1 \cdots d\xi_s$.
+where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_s$.
 More generally, one is interested in determining the moments of the solution for a given set of basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
@@
 stored in an external file, at all quadrature points in every time step to compute the time evolution of the moments.
+This contradicts modern HPC paradigms, which aim at reducing the amount of data produced by numerical methods.
 
@@
 \end{align}
-Rewritten in its dual form, the problem is given by
+Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstraint optimization problem, given by
 \begin{align}\label{eq:dualProblem}

### 2019-04-09 code/src/mesh.cpp
2d stuff added
plot script new colorbar, finer output

@@ -61,8 +61,10 @@ BoundaryType Mesh::GetBoundaryType( unsigned i ) const { return _boundaryType[i]
 
 void Mesh::PlotInXi( const Matrix& u, unsigned state ) const {
     std::ofstream out( "../results/plotInXi" );
-    for( unsigned k = 0; k < _settings->GetNQTotal(); ++k ) {
-        if( k != 0 && k % _settings->GetNQuadPoints() == 0 ) {
+    unsigned nQ   = u.columns();
+    unsigned nQ1D = unsigned( std::pow( nQ, 1.0 / _settings->GetNDimXi() ) );
+    for( unsigned k = 0; k < nQ; ++k ) {
+        if( k != 0 && k % nQ1D == 0 ) {
             out << std::endl;
         }
         out << u( k, state ) << " ";

### 2019-04-09 code/src/momentsolver.cpp

@@ -255,10 +255,17 @@ void MomentSolver::Solve() {
     }
 
     this->Export( uNew, _lambda );
-    /*
-        unsigned evalCell  = 300;    // 2404;
-        unsigned plotState = 0;
-        _mesh->PlotInXi( _closure->U( _closure->EvaluateLambda( _lambda[evalCell] ) ), plotState );*/
+
+    unsigned evalCell = 2404;
+    if( _settings->GetNumCells() > evalCell ) {
+        unsigned plotState  = 0;
+        unsigned nQFine     = 100;
+        unsigned nQOriginal = _settings->GetNQuadPoints();
+        _settings->SetNQuadPoints( nQFine );
+        Closure* closurePlot = Closure::Create( _settings );
+        _mesh->PlotInXi( closurePlot->U( closurePlot->EvaluateLambda( _lambda[evalCell] ) ), plotState );
+        _settings->SetNQuadPoints( nQOriginal );
+    }
 }
 
 MatVec MomentSolver::DetermineMoments( unsigned nTotal ) const {

### 2019-04-11 Paper1/colloc_ipm.tex

@@
 A steady state solution is then obtained by iterating in pseudo-time until the solution remains constant.
+It is important to point out that the time it takes to converge to a steady state solution is crucially affected by the chosen initial condition and its distance to the steady state solution.
 Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system for \eqref{eq:hyperbolicProblemSteady} given by
@@
 Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the residual
-\begin{align*}
+\begin{align}\label{eq:residualUnsteady}
 \sum_{j = 1}^{N_x} \Delta x_j \Vert \bm{\hat{u}}_j^n - \bm{\hat{u}}_j^{n-1} \Vert \leq \varepsilon
-\end{align*}
-is fulfilled. Note that since one is generally interested in low order moments such as the expectation value, this residual can be modified by only accounting for the zero order moments.
+\end{align}
+is fulfilled. Since one is generally interested in low order moments such as the expectation value, this residual can be modified by only accounting for the zero order moments.
 
@@
 \label{sec:collIPM}
+Commonly, a great amount of iterations in pseudo-time are needed to converge to a steady state solution.
+Consequently, the IPM method which requires solving the dual problem \eqref{eq:dualProblem} in every spacial cell in each iteration becomes prohibitively expensive.
+We tackle this problem by using IPM only as a postprocesssing step for the steady solution obtained by a cheap method.
+(Or vice-versa, we use a cheap method as a preprocessing step for IPM).
+In our case, we perform the preprocessing step with stochastic-Collocation, i.e.
+we converge the moments to a steady state solution by applying collocation steps.
+The obtained moments are then used as initial condition for the IPM moment system (for which the moments are no longer a steady state solution).
+After applying a significantly reduced number of IPM iterations, we obtain a steady state IPM solution.
+In our numerical experiments presented in section \ref{sec:results}, we can show that the overall costs are dominated by the large number of cheap collocation steps and not by the small number of expensive IPM steps, while the solution shows the expected desirable properties of the IPM solution.
 
-When computing steady state moments for this problem, the run time of different UQ methods crucially depends on the chosen initial condition and its distance to the steady state solution.
-However, it is not always clear how to choose a satisfactory initial guess for the moments.
-We tackle this problem with the following method: For a bad initial guess, we iterate to the steady state solution with the help of a cheap method.
-Note that one expects to need a lot of iteration steps to arrive at a steady state solution, however the overall computational time is acceptable due to the choice of a cheap method.
-The obtained inaccurate moments are then used to start a more accurate, but at the same time more expensive method.
-Due to the choice of the initial condition, a smaller number of iterations is needed to arrive at the steady state solution of the expensive method, i.e.
-the numerical costs are again acceptable.
-
-I our case, the cheap method will be stochastic-Collocation with a small number of samples, whereas the expensive method will be IPM with a high number of moments.
-However different variants of the main idea can be applied: ...
+Different variants of this method are possible:
+\begin{itemize}
+\item Since the IPM iterations will again modify the steady state Collocation solution, it is not necessary to converge Collocation to the exact steady state solution before starting IPM.
+Here, one needs to determine an indicator to choose at which residual the collocation iteration is sufficiently accurate and can therefore be switched to IPM.
+\item The main idea is to use a cheap but inexact method as a preconditioner for an expensive but accurate method.
+Here, one is not limited in choosing SC and IPM, but one can for example choose Monte Carlo methods as an accelerator or SC with an increased number of quadrature points as the expensive method.
+Note that in the latter case, a map from the moments to the solution at the increased quadrature set is required, for which the IPM reconstruction \eqref{eq:ansatz} would be a suitable choice.
+\end{itemize}

### 2019-04-11 Paper1/framework.tex
Paper continued

@@
 \end{align*}
-Note that the numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define
+Note that the numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^s \to \mathbb{R}^s$,
 \begin{align*}
@@
 \end{align*}
-meaning that the closure \eqref{eq:ansatz} in cell $j$ at timestep $n$ can be written as
+meaning that the closure \eqref{eq:ansatz} at cell $j$ in timestep $n$ can be written as
 \begin{align*}
@@
 The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
-To determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
+Hence, to determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
 \begin{align*}
@@
 The exact dual state is then obtained by computing the fix point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
-$\bm{\hat\lambda}_j^n=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
+$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
 A common preconditioner is the inverse Hessian of the dual problem, i.e.
 \begin{align*}
-\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-T}.
+\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-T}.
 \end{align*}
@@
 \begin{align}\label{eq:momentIteration}
-\bm{\hat u}_j^{n+1} = \bm{c}\left(\bm{\lambda}(\bm{\hat u}_{j-1}^n),\bm{\lambda}(\bm{\hat u}_{j}^n),\bm{\lambda}(\bm{\hat u}_{j+1}^n)\right),
+\bm{\hat u}_j^{n+1} = \bm{c}\left(\bm{\hat\lambda}(\bm{\hat u}_{j-1}^n),\bm{\hat\lambda}(\bm{\hat u}_{j}^n),\bm{\hat\lambda}(\bm{\hat u}_{j+1}^n)\right),
 \end{align}
@@
 \end{align}
-until condition \eqref{eq:tauCrit} is fulfilled. This gives Algorithm \ref{alg:seq}.
+until condition \eqref{eq:tauCrit} is fulfilled. This gives Algorithm \ref{alg:IPM}.
 
@@
 \caption{IPM implementation}
-\label{alg:seq}
+\label{alg:IPM}
 \end{algorithm}

### 2019-04-11 Paper1/one_shot_ipm.tex
Paper continued

@@
-\section{One-shot IPM}
+\subsection{One-shot IPM}
 \label{sec:OneShotIPM}
-For classical IPM, the iteration scheme for the moment vectors is
-\begin{align}\label{eq:momentIteration}
-\bm{u}_j^{n+1} = \bm{c}\left(\bm{\lambda}(\bm{u}_{j-1}^n),\bm{\lambda}(\bm{u}_{j}^n),\bm{\lambda}(\bm{u}_{j+1}^n)\right),
-\end{align}
-where $\bm{c}:\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$ is given by
-\begin{align*}
-\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle u(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}\rangle - \frac{\Delta t}{\Delta x}\left(\langle g(u(\bm{\lambda}_c^T\bm{\varphi}),u(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}\rangle-\langle g(u(\bm{\lambda}_{\ell}^T\bm{\varphi}),u(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}\rangle\right).
-\end{align*}
-The map from the moment vector to the dual variables is given by the exact fix point of $\bm{d}:\mathbb{R}^{N+1}\times\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$,
-\begin{align*}
-\bm{d}(\bm{\lambda},\bm{u}):= \bm{\lambda}-\bm{B}\cdot \left(\langle u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle)-\bm{u}\right),
-\end{align*}
-where $\bm{B}$ is a preconditioner, which is used in the numerical scheme to ensure convergence of the fix point iteration (which we call dual iteration in the following)
-\begin{align}\label{eq:dualIteration}
-\bm{\lambda}_j^{m+1} = \bm{d}(\bm{\lambda}_j^m,\bm{u}_j^{n}).
-\end{align}
-Note that here, $\bm{u}_j^{n}$ is a fixed parameter.
-By letting the dual iteration converge, i.e.
-$m\rightarrow\infty$, we obtain $\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^m,\bm{u}_j^{n}) =:\bm{\lambda}_j^{n} \equiv \bm\lambda(\bm{u}_j^{n})$, which are called the dual variables of $\bm{u}_j^{n}$.
-Hence, the numerical scheme for IPM converges the dual iteration \eqref{eq:dualIteration} to compute the mapping from the moment vectors to their dual variables and then performs one iteration of \eqref{eq:momentIteration} to obtain the time update of the moment vector.
-Note that this can become extremely expensive, since in each iteration of the moment vectors, we fully converge the dual iteration for all spatial cells.
 
-To simplify notation, we leave out the dependency on $\bm{\lambda}$ when writing the moment scheme in the following: Hence with 
-\begin{align*}
-\bm{\tilde c}\left(\bm{u}_{\ell}^n,\bm{u}_{c}^n,\bm{u}_{r}^n\right):=\bm{c}\left(\bm{\lambda}(\bm{u}_{\ell}^n),\bm{\lambda}(\bm{u}_{c}^n),\bm{\lambda}(\bm{u}_{r}^n)\right)
-\end{align*}
-we can rewrite the moment iteration as
-\begin{align}\label{eq:shortIPMIt}
-\bm{u}_j^{n+1} = \bm{\tilde c}\left(\bm{u}_{j-1}^n,\bm{u}_{j}^n,\bm{u}_{j+1}^n\right).
-\end{align}
-
-For steady problems, we assume that the IPM scheme converges to a fix point, i.e.
-we must have that $\rho (\bm{\tilde c}_{\bm{u}})<1$.
-The main idea of \textit{One-Shot IPM} is to not fully converge the dual iteration, since the moment vectors are not yet converged to the exact steady solution.
-So if we successively perform one update of the moment iteration and one update of the dual iteration, we obtain
+In this section we aim at breaking up the inner loop in the IPM algorithm \ref{alg:IPM}, i.e.
+to just perform one iteration of the dual problem in each time step.
+Consequently, the IPM reconstruction given by \eqref{eq:primalProblem} is not done exactly, meaning that the reconstructed solution does not minimize the entropy while not fulfilling the moment constraint.
+However, the fact that the moment vectors are not yet converged to the steady solution seems to permit such an inexact reconstruction.
+Hence, we aim at iterating the moments to steady state and the dual variables to the exact solution of the IPM optimization problem \eqref{eq:primalProblem} simultaneously.
+Hence, by successively performing one update of the moment iteration and one update of the dual iteration, we obtain 
 \begin{subequations}\label{eq:oneshotIPM}
@@
 \end{subequations}
-In the following, we study convergence of this scheme.
-For this we take an approach commonly chosen to prove the local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fix point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence.
+This gives algorithm
+\begin{algorithm}[H]
+\begin{algorithmic}[1]
+\For{$j=0$ to $N_x+1$}
+\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle dx$
+\EndFor
+\While{\eqref{eq:residualUnsteady} is violated}
+\For{$j=1$ to $N_x$}
+\State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
+\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$
+\EndFor
+\State $n \leftarrow n+1$
+\EndWhile
+\end{algorithmic}
+\caption{One-shot IPM implementation}
+\label{alg:osIPM}
+\end{algorithm}
+We call this method One-Shot IPM, since it is inspired by One-shot optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables.
+Note that the dual variables from the One-Shot iteration are written without a hat to indicate that they are not the exact solution of the dual problem.
+
+In the following, we will show that this iteration converges, if the chosen initial condition is sufficiently close to the steady state solution.
+For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fix point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence:
+
 \begin{theorem}\label{th:Contractive}
-Assume that the classical IPM iteration \eqref{eq:shortIPMIt} is contractive at its fix point $\bm{u}^*$.
-Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fix point $(\bm{\lambda}^*,\bm{u}^*)$ if the preconditioner $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$, i.e.
+Assume that the classical IPM iteration is contractive at its fix point $\bm{\hat u}^*$.
+Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fix point $(\bm{\lambda}^*,\bm{\hat u}^*)$ if the preconditioner $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$, i.e.
 the Hessian of the dual problem is used.
@@
 \begin{proof}
-We first point out that since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\bm{\tilde c}_{\bm{u}}(\bm{u}^*))<1$ with $\bm{\tilde c}_{\bm{u}}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by
+First, understand what contraction of the classical IPM iteration implies, we rewrite the moment iteration \eqref{eq:momentIteration} of the classical IPM scheme: When defining the update function
 \begin{align*}
-\bm{\tilde c}_{\bm{u}} = 
+\bm{\tilde c}\left(\bm{\hat{u}}_{\ell},\bm{\hat{u}}_{c},\bm{\hat{u}}_{r}\right):=\bm{c}\left(\bm{\hat{\lambda}}(\bm{\hat{u}}_{\ell}),\bm{\hat{\lambda}}(\bm{\hat{u}}_{c}),\bm{\hat{\lambda}}(\bm{\hat{u}}_{r})\right)
+\end{align*}
+we can rewrite the classical moment iteration as
+\begin{align}\label{eq:shortIPMIt}
+\bm{\hat u}_j^{n+1} = \bm{\tilde c}\left(\bm{\hat u}_{j-1}^n,\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^n\right).
+\end{align}
+Since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by
+\begin{align*}
+\nabla_{\bm{\hat u}}\bm{\tilde c} = 
 \begin{pmatrix} 
-    \partial_{\bm{u}_c}\bm{\tilde c}_{1} & \partial_{\bm{u}_r}\bm{\tilde c}_{1}& 0 & 0 & \dots \\
-    \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{2} & \partial_{\bm{u}_c}\bm{\tilde c}_{2} & \partial_{\bm{u}_r}\bm{\tilde c}_{2}& 0 & \dots \\
-    0 & \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{3} & \partial_{\bm{u}_c}\bm{\tilde c}_{3} & \partial_{\bm{u}_r}\bm{\tilde c}_{3}\\
+    \partial_{\bm{\hat u}_c}\bm{\tilde c}_{1} & \partial_{\bm{\hat u}_r}\bm{\tilde c}_{1}& 0 & 0 & \dots \\
+    \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_{2} & \partial_{\bm{\hat u}_c}\bm{\tilde c}_{2} & \partial_{\bm{\hat u}_r}\bm{\tilde c}_{2}& 0 & \dots \\
+    0 & \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_{3} & \partial_{\bm{\hat u}_c}\bm{\tilde c}_{3} & \partial_{\bm{\hat u}_r}\bm{\tilde c}_{3}\\
     \vdots & & & \ddots & \\
-    0 &\cdots &  0 & \partial_{\bm{u}_{\ell}}\bm{\tilde c}_{N_x} & \partial_{\bm{u}_c}\bm{\tilde c}_{N_x}
+    0 &\cdots &  0 & \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_{N_x} & \partial_{\bm{\hat u}_c}\bm{\tilde c}_{N_x}
     \end{pmatrix},
 \end{align*}
-where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{u}_{j-1}^*,\bm{u}_{j}^*,\bm{u}_{j+1}^*\right)$ for all $j$. Now we have for each term inside the matrix $\bm{\tilde c}_{\bm{u}}$
+where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{\hat u}_{j-1}^*,\bm{\hat u}_{j}^*,\bm{\hat u}_{j+1}^*\right)$ for all $j$.
+Now for each term inside the matrix $\nabla_{\bm{\hat u}}\bm{\tilde c}$ we have
 \begin{align}\label{eq:cTildeDer}
-\partial_{\bm{u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{\lambda}(\bm{u}_{j-1}^*)}{\partial \bm{u}},\enskip\partial_{\bm{u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_c}\frac{\partial \bm{\lambda}(\bm{u}_j^*)}{\partial \bm{u}},\enskip\partial_{\bm{u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_r}\frac{\partial \bm{\lambda}(\bm{u}_{j+1}^*)}{\partial \bm{u}}.
+\partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\hat \lambda}_{\ell}}\frac{\partial \bm{\hat \lambda}(\bm{\hat u}_{j-1}^*)}{\partial \bm{\hat u}},\enskip\partial_{\bm{\hat u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\hat \lambda}_c}\frac{\partial \bm{\hat \lambda}(\bm{\hat u}_j^*)}{\partial \bm{\hat u}},\enskip\partial_{\bm{\hat u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\hat \lambda}_r}\frac{\partial \bm{\hat \lambda}(\bm{\hat u}_{j+1}^*)}{\partial \bm{\hat u}}.
 \end{align}
-We first wish to understand the structure of the terms $\partial_{\bm{u}} \bm{\lambda}(\bm{u})$. For this, we note that the exact dual state fulfills
+We first wish to understand the structure of the terms $\partial_{\bm{\hat u}} \bm{\hat \lambda}(\bm{\hat u})$. For this, we note that the exact dual variables fulfill
 \begin{align}\label{eq:ulambda}
-\bm{u} = \langle u(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\rangle =: \bm{h}(\bm{\lambda}),
+\bm{\hat u} = \langle \bm{u}_s(\bm{\hat \lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle =: \bm{h}(\bm{\hat \lambda}),
 \end{align}
-which is why we have the mapping $\bm{u}:\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$, $\bm{u}(\bm{\lambda}) = \bm{h}(\bm{\lambda})$.
+which is why we have the mapping $\bm{\hat u}:\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$, $\bm{\hat u}(\bm{\hat \lambda}) = \bm{h}(\bm{\hat \lambda})$.
 Since the solution of the dual problem for a given moment vector is unique, this mapping is bijective and therefore we have an inverse function
 \begin{align}\label{eq:lambdau}
-\bm{\lambda} = \bm{h}^{-1}(\bm{u}(\bm{\lambda}))
+\bm{\hat \lambda} = \bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda})).
 \end{align}
-Now we differentiate both sides w.r.t. $\bm{\lambda}$ to get
+Now we differentiate both sides w.r.t. $\bm{\hat \lambda}$ to get
 \begin{align*}
-\bm{I}_{d} = \frac{\partial \bm{h}^{-1}(\bm{u}(\bm{\lambda}))}{\partial \bm{u}}\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}.
+\bm{I}_{d} = \frac{\partial \bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda}))}{\partial \bm{\hat u}}\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}.
 \end{align*}
-We multiply with the matrix inverse of $\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}$ to get
+We multiply with the matrix inverse of $\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}$ to get
 \begin{align*}
-\left(\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}\right)^{-1} = \frac{\partial \bm{h}^{-1}(\bm{u}(\bm{\lambda}))}{\partial \bm{u}}.
+\left(\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}\right)^{-1} = \frac{\partial \bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda}))}{\partial \bm{\hat u}}.
 \end{align*}
 Note that on the left-hand-side we have the inverse of a matrix and on the right-hand-side, we have the inverse of a multi-dimensional function.
-By rewriting $\bm{h}^{-1}(\bm{u}(\bm{\lambda}))$ as $\bm{\lambda}(\bm{u})$ and simply computing the term $\frac{\partial \bm{u}(\bm{\lambda})}{\partial \bm{\lambda}}$ by differentiating \eqref{eq:ulambda} w.r.t.
-$\bm{\lambda}$, one obtains
+By rewriting $\bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda}))$ as $\bm{\hat \lambda}(\bm{\hat u})$ and simply computing the term $\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}$ by differentiating \eqref{eq:ulambda} w.r.t.
+$\bm{\hat \lambda}$, one obtains
 \begin{align}\label{eq:dudlambdaex}
-\partial_{\bm{u}} \bm{\lambda}(\bm{u}) = \langle u'(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}.
+\partial_{\bm{\hat u}} \bm{\hat \lambda}(\bm{\hat u}) = \langle \nabla\bm{u}_s(\bm{\hat \lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-T}.
 \end{align}
@@
 \begin{align*}
-&\bm{\lambda}_j^{n+1} = \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all j} \\
-&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{u}_{j+1}^{n})\right)
+&\bm{\lambda}_j^{n+1} = \bm{d}(\bm{\lambda}_j^{n},\bm{\hat u}_j^{n}) \enskip \text{ for all j} \\
+&\bm{\hat u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{\hat u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{\hat u}_{j+1}^{n})\right)
 \end{align*}
@@
 \begin{pmatrix}
- \partial_{\bm{\lambda}} \bm{d} & \partial_{\bm{u}} \bm{d}  \\
-\partial_{\bm{\lambda}} \bm{c} & \partial_{\bm{u}} \bm{c}
+ \partial_{\bm{\lambda}} \bm{d} & \partial_{\bm{\hat u}} \bm{d}  \\
+\partial_{\bm{\lambda}} \bm{c} & \partial_{\bm{\hat u}} \bm{c}
 \end{pmatrix},
@@
 \begin{align*}
-\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{u}_j^n) &= \bm{I}_d - \bm{B} \cdot \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle - \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle u(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi} \rangle - \bm{u}\right) \\
-&=- \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle u(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi} \rangle - \bm{u}\right),
+\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{B} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right) \\
+&=- \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right),
 \end{align*}
-where we used $\bm{B} = \langle u'(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-1}$.
-Recall that at the fix point $(\bm{\lambda}^*,\bm{u}^*)$, we have $\langle u(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi} \rangle = \bm{u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
-For the block $\partial_{\bm{u}} \bm{d}$, we get
+where we used $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T}$.
+Recall that at the fix point $(\bm{\lambda}^*,\bm{\hat u}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
+For the block $\partial_{\bm{\hat u}} \bm{d}$, we get
 \begin{align*}
-\partial_{\bm{u}} \bm{d}(\bm{\lambda}_j^n,\bm{u}_j^n) = \bm{B},
+\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{B},
 \end{align*}
-hence $\partial_{\bm{u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$:
+hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$:
 \begin{align*}
-\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{u}_{j-1}^{n})}{\partial \bm{\lambda}} = \bm{0},
+\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\lambda}} = \bm{0},
 \end{align*}
@@
 \begin{align*}
-\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{u}_{j-1}^{n})}{\partial \bm{u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{B} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle u'(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-1} = \partial_{\bm{u}_{\ell}}\bm{\tilde c}_j
+\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{B} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
 \end{align*}
 by the choice of $\bm{B}$ as well as \eqref{eq:cTildeDer} and \eqref{eq:dudlambdaex}.
-Hence, we have that $\partial_{\bm{u}} \bm{c} = \bm{\tilde c}_{\bm{u}}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM iteration is contractive.
-Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fix point.
+We obtain an analogous result for the second and third input.
+Hence, we have that $\partial_{\bm{\hat u}} \bm{c} = \nabla_{\bm{\hat u}}\bm{\tilde c}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM iteration is contractive.
+Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{\hat u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fix point.
 \end{proof}
@@
 there exists a $\delta>0$ s.t.
-for all starting points $(\bm{\lambda}^0,\bm{u}^0)\in B_{\delta}(\bm{\lambda}^*,\bm{u}^*)$ we have
+for all starting points $(\bm{\lambda}^0,\bm{\hat u}^0)\in B_{\delta}(\bm{\lambda}^*,\bm{\hat u}^*)$ we have
 \begin{align*}
-\Vert (\bm{\lambda}^n,\bm{u}^n) - (\bm{\lambda}^*,\bm{u}^*)\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
+\Vert (\bm{\lambda}^n,\bm{\hat u}^n) - (\bm{\lambda}^*,\bm{\hat u}^*)\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
 \end{align*}
@@
 \end{align*}
-Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{u})$.
+Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{\hat u})$.
 Hence there exists an open ball with radius $\delta$ around the fix point in which the eigenvalues remain in the interval $(-1,1)$.
@@
 \begin{remark}
-Theorem \ref{th:localConvergence} gives some insights in how to set up the One-Shot IPM iteration: If we look at the Jacobian \eqref{eq:Jacobian} at an arbitrary point, we obtain
-
-Let us assume that the classical IPM scheme converges globally.
-
-\begin{itemize}
-\item Choose a moment vector $\bm{u}^0$ and compute the corresponding exact dual variables $\bm{\lambda}^0:=\bm{\lambda}(\bm{u}^0)$ as starting vector.
-In this case, the term $\partial_{\lambda}\bm{d}$ is equal to zero for the starting conditions.
-\end{itemize}
+Since the preconditioning step of the Collocation-accelerated IPM method generates initial conditions which are close to the steady state solution, using One-Shot IPM instead of classical IPM is well suited.
+However, our numerical calculations show that one-shot IPM converges even if the solution is far away from its steady state.
 \end{remark}

### 2019-04-11 code/src/settings.cpp
refinement levels in config file

@@ -196,10 +196,9 @@ Settings::Settings( std::string inputFile ) : _inputFile( inputFile ), _hasExact
                         "StochasticGalerkin, Euler, Euler2D,L2Filter,LassoFilter" );
             validConfig = false;
         }
-        auto nMoments = moment_system->get_as<unsigned>( "moments" );
+        auto nMoments = moment_system->get_array_of<int64_t>( "moments" );
         if( nMoments ) {
-            _nMoments = *nMoments;
-
+            _nMoments = unsigned( ( *nMoments )[nMoments->size() - 1] );
             // compute nTotal
             _useMaxDegree = false;
             _nTotal       = 0;
@@ -213,17 +212,9 @@ Settings::Settings( std::string inputFile ) : _inputFile( inputFile ), _hasExact
                 if( totalDegree < _nMoments || _useMaxDegree ) ++_nTotal;
             }
             // set vector containing nMoments for each refinement level
-            _nRefinementLevels                             = 8;
-            _nTotalRefinementLevel                         = VectorU( _nRefinementLevels );
-            _nTotalRefinementLevel[_nRefinementLevels - 1] = 10;
-            _nTotalRefinementLevel[6]                      = 9;
-            _nTotalRefinementLevel[5]                      = 8;
-            _nTotalRefinementLevel[4]                      = 7;
-            _nTotalRefinementLevel[3]                      = 6;
-            _nTotalRefinementLevel[2]                      = 5;
-            _nTotalRefinementLevel[1]                      = 4;
-            _nTotalRefinementLevel[0]                      = 3;
-            //_nTotalRefinementLevel[0]                      = 2;
+            _nRefinementLevels     = unsigned( nMoments->size() );
+            _nTotalRefinementLevel = VectorU( _nRefinementLevels );
+            for( unsigned l = 0; l < _nRefinementLevels; ++l ) _nTotalRefinementLevel[l] = unsigned( ( *nMoments )[l] );
         }
         else {
             log->error( "[inputfile] [moment_system] 'moments' not set!" );
@@ -446,10 +437,10 @@ Settings::Settings( const std::istringstream& inputStream ) {
                         "StochasticGalerkin, Euler, Euler2D,L2Filter,LassoFilter" );
             validConfig = false;
         }
-        auto nMoments = moment_system->get_as<unsigned>( "moments" );
-        if( nMoments ) {
-            _nMoments = *nMoments;
 
[…]

### 2019-06-04 Paper1/introduction.tex
2d stuff added
plot script new colorbar, finer output
Paper corrections

@@
 
-Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods.
+Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive techniques.
 A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
@@
 \begin{align*}
-\bm{\hat u_i} = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{N_q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi_k})f_{\Xi}(\bm{\xi}_k).
+\bm{\hat u_i} = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{N_q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
 \end{align*} 
-Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a big implementation effort.
+Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
 Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
@@
 Furthermore, collocation methods require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation}.
-Despite their easy implementation, collocation methods can become cumbersome for unsteady problems, since in this case, the solution needs to be written out, i.e.
+Despite their easy implementation, collocation methods face challenges when examining unsteady problems, since in this case, the solution needs to be written out, i.e.
 stored in an external file, at all quadrature points in every time step to compute the time evolution of the moments.
@@
 \end{align*}
-A commonly used closure is given by stochastic-Galerkin \cite{ghanem2003stochastic}, which represents the solution by a polynomial:
+A commonly used closure is given by stochastic-Galerkin (SG) \cite{ghanem2003stochastic}, which represents the solution by a polynomial:
 \begin{align*}
@@
 where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_0,\cdots,\bm{\hat{\lambda}}_N)^T\in \mathbb{R}^{(N+1) \times p}$ are called the dual variables.
-The solution to \eqref{eq:primalProblem} is then obtained by
+The solution to \eqref{eq:primalProblem} is then given by
 \begin{align}\label{eq:ansatz}
@@
 \end{subequations}
+with $|i|\leq M$.
 The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
@@
 The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
-Hence, the desirable properties of IPM come along with a significantly increased run time.
-However, IPM and minimal entropy methods in general are well suited for modern HPC architecture, which can be used to significantly reduce runtime \cite{garrett2015optimization}.
+Hence, the desirable properties of IPM come along with significantly increased numerical costs.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architecture, which can be used to reduce the run time \cite{garrett2015optimization}.
 
@@
 The main task here is to find an adequate refinement indicator.
-In this paper we compare methods without using local refinement strategies.\\
+\\
 
@@
 \end{itemize}
+In addition to these two strategies for steady problems, we make us of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
+We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case.
 

### 2019-06-04 code/src/euler2d.cpp
modified Euler 2D Pipe settings

@@ -152,18 +152,24 @@ Vector Euler2D::IC( const Vector& x, const Vector& xi ) {
 
         double rhoFarfield = p / ( R * T );
 
-        y[0]                  = rhoFarfield;
+        y[0] = rhoFarfield;
+        if( xi.size() > 1 ) {
+            y[0] += _sigma[1] * xi[1];
+        }
         y[1]                  = rhoFarfield * uF;
         y[2]                  = rhoFarfield * vF;
-        double kineticEnergyL = 0.5 * rhoFarfield * ( pow( uF, 2 ) + pow( vF, 2 ) );
-        double innerEnergyL   = ( p / ( rhoFarfield * ( gamma - 1 ) ) ) * rhoFarfield;
+        double kineticEnergyL = 0.5 * y[0] * ( pow( uF, 2 ) + pow( vF, 2 ) );
+        double innerEnergyL   = ( p / ( y[0] * ( gamma - 1 ) ) ) * y[0];
         y[3]                  = kineticEnergyL + innerEnergyL;
 
         if( x[1] < 1.1 + _sigma[0] * xi[0] ) {
             y[0] = 0.5 * rhoFarfield;
-            y[1] = rhoFarfield * uF;
-            y[2] = rhoFarfield * vF;
-            y[3] = 0.5 * ( kineticEnergyL + innerEnergyL );
+            if( xi.size() > 2 ) y[0] += _sigma[2] * xi[2];
+            y[1]           = rhoFarfield * uF;
+            y[2]           = rhoFarfield * vF;
+            kineticEnergyL = 0.5 * rhoFarfield * ( pow( uF, 2 ) + pow( vF, 2 ) );
+            innerEnergyL   = ( p / ( rhoFarfield * ( gamma - 1 ) ) ) * rhoFarfield;
+            y[3]           = 0.5 * ( kineticEnergyL + innerEnergyL );
         }
         /*
          if( x[1] < 1.1 + sigma && x[1] < 1.1 - sigma ) {

### 2019-07-06 Paper1/adaptivity.tex
adaptivity section in paper started + multi D notation intro corrected

\section{Adaptivity}
\label{sec:adaptivity}

The following section presents the adaptivity strategy used in this work. Since stochastic hyperbolic problems generally experience shocks in a small portion of the space-time domain, the idea is to perform arising computations on a high accuracy level in this small area, while keeping a low level of accuracy in the remainder. The hope is to achieve the finest level accuracy with this adaptive strategy, i.e. the same error is obtained while using a significantly reduced number of unknowns.

In the following, we discuss the building blocks of the IPM method for accuracy levels $\ell = 1,\cdots,N_{\text{ad}}$. At a given level $\ell$, the total degree of the basis function is given by $M_{\ell}$ with a corresponding number of moments $N_{\ell}$. The number of quadrature points at level $\ell$ is denoted by $Q_{\ell}$. To determine the accuracy level of a given moment vector $\bm{\hat u}$ we choose techniques used in discontinuous Galerkin (DG) methods. Adaptivity is a common strategy to accelerate this class of methods and several indicators to determine the smoothness of the solution exist. We make use of a strategy from \cite{persson2006sub}, which uses the highest degree moments as indicator. Translating the idea of the discontinuity sensor used in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at level $\ell$ as
\begin{align*}
\bm{\tilde u}_{\ell} := \sum_{|i|\leq M_{\ell}} \bm{\hat{u}}_i \varphi_i.
\end{align*}
Now the indicator for a moment vector at level $\ell$ is defined as
\begin{align}\label{eq:errorIndicator}
\bm S_{\ell} := \frac{\langle \left(\bm{\tilde u}_{\ell} - \bm{\tilde u}_{\ell-1}\right)^2\rangle}{\langle \bm{\tilde u}_{\ell}^2\rangle},
\end{align}
where divisions and multiplications are performed element-wise. In this work, we use the first entry in $\bm S_{\ell}$ to determine the refinement level, i.e. in the case of gas dynamics, the regularity of the density is chosen to indicate an adequate refinement level. If the moment vector in a given cell at a certain timestep is initially at refinement level $\ell$, this level is kept if if the error indicator \eqref{eq:errorIndicator} lies in the interval $I_{\delta}:=[\delta_{-},\delta_{+}]$. Here $\delta_{\pm}$ are user determined parameters. If the indicator is smaller than $\delta_-$, the refinement level is decreased, if it lies above $\delta_+$, it is increased.

Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders on different cells. Let first add dimensions to the definition of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIteration}.

### 2019-07-06 Paper1/framework.tex
Paper corrections
adaptivity section in paper started + multi D notation intro corrected

@@
 for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
-The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\hat{\bm{u}},\hat{\bm{u}})=\bm{F}(\hat{\bm{u}})$.
+The numerical flux is assumed to be consistent, i.e. $\bm{G}(\hat{\bm{u}},\hat{\bm{u}})=\bm{F}(\hat{\bm{u}})$.
 
@@
 which is usually done by an iterative method. Defining the dual iteration function $\bm{d}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$,
-\begin{align*}
+\begin{align}\label{eq:dualIterationFunction}
 \bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T-\bm{\hat{u}}\right),
-\end{align*}
+\end{align}
 with an adequate preconditioner $\bm{B}$, the iteration process for spatial cell $j$ is given by
@@
 
+TODO: Extension to 2D $\rightarrow$ triangular meshes
+
 \subsection{Code framework}
+
+TODO: Code structure, why is it easy to include different equations?

### 2019-07-06 Paper1/introduction.tex
adaptivity section in paper started + multi D notation intro corrected

@@
 More generally, one is interested in determining the moments of the solution for a given set of basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
+Note that this yields
+\begin{align*}
+N:=\begin{pmatrix}
+M+s \\ s
+\end{pmatrix}
+\end{align*}
+basis polynomials when defining $|i|:=\sum_{n = 1}^s |i_n|$.
 Commonly one chooses orthonormal polynomials as basis functions \cite{xiu2002wiener}, i.e.
-$\langle \varphi_n \varphi_m \rangle = \delta_{nm}$.
-The moments are then given by $\bm{\hat u_i} := \langle \bm{u}\varphi_i \rangle$.
+$\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^s\delta_{i_nj_n}$.
+The corresponding moments are then given by $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
 Besides facilitating the computation of the expectation value and variance, the moments can be used to span the solution in $\bm\xi$ \cite{wiener1938homogeneous}.
@@
 
-Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive techniques.
+Numerical methods for approximating the moment $\bm{\hat u}_i$ can be divided into intrusive and non-intrusive techniques.
 A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
-\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $N_q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
+\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
 \begin{align*}
-\bm{\hat u_i} = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{N_q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
+\bm{\hat u}_i = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{Q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
 \end{align*} 
@@
 \begin{align*}
-\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi).
+\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u},\cdots,\bm{\hat u}_N;\bm\xi).
 \end{align*}
+Here, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\bm{\hat{u}}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N \times p}$.
+The dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted in the following for sake of readability.
 A commonly used closure is given by stochastic-Galerkin (SG) \cite{ghanem2003stochastic}, which represents the solution by a polynomial:
 \begin{align*}
-\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
+\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
 \end{align*}
-Here, we collect the moments for which $\vert i \vert \leq M$ holds in the moment matrix $\bm{\hat{u}}:=(\bm{\hat{u}}_0,\cdots,\bm{\hat{u}}_N)^T\in\mathbb{R}^{(N+1) \times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_0,\cdots,\varphi_N)^T\in\mathbb{R}^{N+1}$.
+Corresponding to the definition of $\bm{\hat u}$, we collect the $N$ basis function for which $\vert i \vert \leq M$ holds in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N+1}$.
 When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty} and the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
@@
 \begin{align}\label{eq:primalProblem}
-\mathcal{U}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
+\mathcal{U}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } |i| \leq M.
 \end{align}
-Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstraint optimization problem, given by
+Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstraint optimization problem.
+Defining the variables $\bm{\lambda}_i\in\mathbb{R}^p$ where $i$ is again a multi index, gives the unconstrained dual problem
 \begin{align}\label{eq:dualProblem}
- \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{(N+1) \times p}}
-  \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
+ \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{N \times p}}
+  \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{|i|\leq M}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
 \end{align}
-where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_0,\cdots,\bm{\hat{\lambda}}_N)^T\in \mathbb{R}^{(N+1) \times p}$ are called the dual variables.
+where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_i)_{|i|\leq M}\in \mathbb{R}^{N \times p}$ are called the dual variables.
 The solution to \eqref{eq:primalProblem} is then given by

### 2019-07-07 Paper1/adaptivity.tex
paper continued
- N notation further corrections
- quadrature added
- adaptivity continued

@@
 Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders on different cells.
-Let first add dimensions to the definition of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIteration}.
+Let us first add dimensions to the notation of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIterationFunction}.
+Now, we have
+$\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times p}\times\mathbb{R}^{N_{\ell}\times p}\to\mathbb{R}^{N_{\ell}\times p}$, given by
+\begin{align}\label{eq:dualIterationFunction}
+\bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}_{\ell}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
+\end{align}
+where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $\ell$. The preconditioner $\bm{B}_{\ell}$ is given by

### 2019-07-07 Paper1/framework.tex
paper continued
- N notation further corrections
- quadrature added
- adaptivity continued

@@
 \end{align*}
-with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
+with the flux $\bm{F}:\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
@@
 \end{align*}
-If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$ as
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times p}\times\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$ as
 \begin{align}\label{eq:IPMDiscretization}
@@
 \begin{align*}
- \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T.
+ \bm{\tilde G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T.
+\end{align*}
+Commonly, this integral cannot be evaluated analytically and therefore needs to be approximated by a quadrature rule
+\begin{align*}
+\langle h \rangle \approx \langle h \rangle_{Q} := \sum_{k=1}^Q w_k h(\xi_k).
+\end{align*}
+The approximated numerical flux then becomes
+\begin{align*}
+ \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T_Q.
 \end{align*}
@@
 \begin{align*}
-L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i
+L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle_Q - \sum_{i\leq M}\bm{\lambda}_i^T \bm{\hat u}_i
 \end{align*}
@@
 \begin{align*}
-\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle^T - \bm{\hat u}_i,
+\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}_i,
 \end{align*}
-which is usually done by an iterative method. Defining the dual iteration function $\bm{d}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$,
+which is usually done by an iterative method. Defining the dual iteration function $\bm{d}:\mathbb{R}^{N\times p}\times\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$,
 \begin{align}\label{eq:dualIterationFunction}
-\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T-\bm{\hat{u}}\right),
+\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T-\bm{\hat{u}}\right),
 \end{align}
@@
 \begin{align*}
-\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-T}.
+\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle_Q^{-T}.
 \end{align*}
@@
 \begin{align*}
-\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle \bm u_{s}(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}^T\rangle^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}^T\rangle^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{\ell}^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}^T\rangle^T\right).
+\bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle \bm u_{s}(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}^T\rangle_Q^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{\ell}^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}^T\rangle_Q^T\right).
 \end{align*}
@@
 \For{$j=0$ to $N_x+1$}
-\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle dx$
+\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle_Q dx$
 \EndFor

### 2019-07-07 Paper1/adaptivity.tex
adaptivity + algorithm done

@@
 $\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times p}\times\mathbb{R}^{N_{\ell}\times p}\to\mathbb{R}^{N_{\ell}\times p}$, given by
-\begin{align}\label{eq:dualIterationFunction}
+\begin{align}\label{eq:dualIterationFunctionAd}
 \bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}_{\ell}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
@@
 where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $\ell$. The preconditioner $\bm{B}_{\ell}$ is given by 
+\begin{align*}
+\bm{B}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{-T}.
+\end{align*}
+An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
+\begin{align*}
+\bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= \langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
+\end{align*}
+Hence, the index vector $\bm\ell'\in\mathbb{N}^{3}$ denotes the refinement levels of the stencil cells, which are used to compute the time updated moment vector at level $\ell$.
+
+The strategy now is to perform the dual update for a set of moment vectors $\bm{\hat u}_j^n$ at refinement levels $\ell_j^n$ for $j = 1,\cdots,N_x$.
+Hence, the dual iteration makes use of the iteration function \eqref{eq:dualIterationFunctionAd} at refinement level $\ell_j^n$.
+After that, the refinement level at the next time step $\ell_j^{n+1}$ is determined by making use of the smoothness indicator \eqref{eq:errorIndicator}.
+The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$.
+The IPM algorithm with adaptivity then reads
+\begin{algorithm}[H]
+\begin{algorithmic}[1]
+\For{$j=0$ to $N_x+1$}
+\State $\ell_j^0 \leftarrow$ choose initial refinement level
+\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi}_{\ell_j^0} \rangle_{Q_{\ell_j^0}} dx$
+\EndFor
+\For{$n=0$ to $N_t$}
+\For{$j=0$ to $N_x+1$}
+\State $\bm{\lambda}_j^{(0)} \leftarrow \bm{\hat \lambda}_j^{n}$
+\While{\eqref{eq:tauCrit} is violated}
+\State $\bm{\lambda}_j^{(m+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n})$
+\State $m \leftarrow m+1$
+\EndWhile
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(m)}$
+\State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\hat \lambda}_j^{n+1}\right)$
+\EndFor
+\For{$j=1$ to $N_x$}
+\State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$
+\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}_{\ell_j^{n+1}}^{\bm\ell'}(\bm{\hat \lambda}_{j-1}^{n+1},\bm{\hat \lambda}_j^{n+1},\bm{\hat \lambda}_{j+1}^{n+1})$
+\EndFor
+\EndFor
+\end{algorithmic}
+\caption{Adaptive IPM implementation}
+\label{alg:ad-IPM}
+\end{algorithm}

### 2019-07-07 Paper1/adaptivity.tex
typos corrected

@@
 
-Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders on different cells.
+Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders in different cells.
 Let us first add dimensions to the notation of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIterationFunction}.
@@
 \end{align}
-where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $\ell$. The preconditioner $\bm{B}_{\ell}$ is given by 
+where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $M_\ell$. The preconditioner $\bm{B}_{\ell}$ is given by 
 \begin{align*}
@@
 \end{algorithm}
+%TODO Auch adaptiver Algorithmus fuer osIPM?

### 2019-07-07 Paper1/introduction.tex
typos corrected

@@
 \begin{align*}
-\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u},\cdots,\bm{\hat u}_N;\bm\xi).
+\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u};\bm\xi).
 \end{align*}
@@
 \begin{align*}
-\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
+\mathcal{U}_{\text{SG}}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
 \end{align*}
@@
 \partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
+\bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
 \end{align}
@@
 We thereby provide the ability to recycle existing implementations of deterministic solvers.
-Furthermore, we investigate intrusive methods for steady problems and compare them to collocation methods.
+Furthermore, we investigate intrusive methods for steady problems and compare them against collocation methods.
 The steady setting provides different opportunities to take advantage of features of intrusive methods:
@@
 \end{itemize}
-In addition to these two strategies for steady problems, we make us of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
+In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
 We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case.
@@
 Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration.
-A comparison of results computed with the presented methods is given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.
+Section~\ref{sec:adaptivity} extends the presented numerical framework to an algorithm making use of adaptivity.
+A comparison of results computed with the presented methods is then given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.

### 2019-07-07 Paper1/one_shot_ipm.tex
paper continued
- N notation further corrections
- quadrature added
- adaptivity continued
typos corrected

@@
 \begin{proof}
-First, understand what contraction of the classical IPM iteration implies, we rewrite the moment iteration \eqref{eq:momentIteration} of the classical IPM scheme: When defining the update function
+First, to understand what contraction of the classical IPM iteration implies, we rewrite the moment iteration \eqref{eq:momentIteration} of the classical IPM scheme: When defining the update function
 \begin{align*}

### 2019-07-07 Paper1/framework.tex
adaptivity + algorithm done
comment on kinetic flux

@@
 \begin{align*}
-\langle h \rangle \approx \langle h \rangle_{Q} := \sum_{k=1}^Q w_k h(\xi_k).
+\langle h \rangle \approx \langle h \rangle_{Q} := \sum_{k=1}^Q w_k h(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
 \end{align*}
 The approximated numerical flux then becomes
-\begin{align*}
+\begin{align}\label{eq:numericalFluxIPM}
  \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T_Q.
-\end{align*}
+\end{align}
 Note that the numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^s \to \mathbb{R}^s$,
@@
 We now write down the entire scheme: To obtain a more compact notation, one defines
-\begin{align*}
+\begin{align}\label{eq:momentIterationFunction}
 \bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle \bm u_{s}(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}^T\rangle_Q^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{\ell}^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}^T\rangle_Q^T\right).
-\end{align*}
+\end{align}
 The moment iteration is then given by
@@
 
+\subsection{Costs of the numerical flux}
+A straight forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}.
+By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
+However, the need to approximate the flux by an integral seems to be numerically expensive, especially when considering that the stochastic-Galerkin method often allows an analytic computation of all arising integrals.
+In the following, we discuss this issue and show the advantages of our numerical flux choice by considering the stochastic Burgers' equation
+\begin{align*}
+\partial_t u + \partial_x \frac{u^2}{2} &= 0\\
+u(t=0,x,\xi) &= u_{IC}(x,\xi).
+\end{align*}
+The scalar random variable $\xi$ is uniformly distributed in the $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
+Testing with these basis functions and choosing the SG closure \eqref{eq:SGClosure} yields the closed SG moment system
+\begin{align*}
+\partial_t \hat u_i + \partial_x \frac12\sum_{n,m = 0}^M \hat u_n \hat u_m \langle \varphi_n\varphi_m\varphi_i \rangle = 0.
+\end{align*}
+Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{M\times M}$ gives
+\begin{align*}
+\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = 0
+\end{align*}
+with $F_i(\bm{\hat u}) = \frac12\bm{\hat u}^T\bm C_i\bm{\hat u}$. Note that $\bm{C}_i$ can be computed analytically, hence choosing a Lax-Friedrichs flux
+\begin{align}\label{eq:numFluxAnalytic}
+G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i.
+\end{align}
+requires no integral evaluations. Recall, that the numerical flux choice made in this work gives
+\begin{align}\label{eq:numericalFluxIPMBurgers}
+ \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \langle g(\mathcal{U}_{SG}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}_{SG}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}^T\rangle^T_Q.
+\end{align}
+When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the integral brackets is of order $3M$.
+Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 M +\frac12$ quadrature points suffice for an exact computation of the numerical flux.
+Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent.
+Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When precomputing the values $\mathcal{U}_{SG}(\bm{\hat u};\xi_k)=\sum_{i=0}^N\hat u_i\varphi_i(\xi_k)$ which requires $M\cdot Q$ evaluations, the computation of \eqref{eq:numericalFluxIPMBurgers} again requires $O(M\cdot Q)$ operations, i.e.
+the costs are $O(M^2)$, whereas the evaluation of \eqref{eq:numFluxAnalytic} requires $O(M^3)$ operations.
+
 TODO: Extension to 2D $\rightarrow$ triangular meshes

### 2019-07-11 Paper1/framework.tex
paper

@@
 \begin{align*}
-\partial_t u + \partial_x \frac{u^2}{2} &= 0\\
+\partial_t u + \partial_x \frac{u^2}{2} &= 0,\\
 u(t=0,x,\xi) &= u_{IC}(x,\xi).
 \end{align*}
-The scalar random variable $\xi$ is uniformly distributed in the $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
+The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
 Testing with these basis functions and choosing the SG closure \eqref{eq:SGClosure} yields the closed SG moment system
@@
 \begin{align}\label{eq:numFluxAnalytic}
-G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i.
+G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i
 \end{align}

### 2019-07-11 Paper1/introduction.tex
comment on kinetic flux
paper small corrections

@@
 Also for steady problems, the fact that the moments are known in each iteration enables the computation of the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
-However intrusive methods are in general more difficult to implement and come along with higher numerical costs.
+However, intrusive methods are in general more difficult to implement and come along with higher numerical costs.
 The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
@@
 A commonly used closure is given by stochastic-Galerkin (SG) \cite{ghanem2003stochastic}, which represents the solution by a polynomial:
-\begin{align*}
+\begin{align}\label{eq:SGClosure}
 \mathcal{U}_{\text{SG}}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
-\end{align*}
+\end{align}
 Corresponding to the definition of $\bm{\hat u}$, we collect the $N$ basis function for which $\vert i \vert \leq M$ holds in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N+1}$.
@@
 In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
-We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case.
+We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.
 

### 2019-07-11 Paper1/main.tex
Paper continued
vtk plot and paper continued
adaptivity section in paper started + multi D notation intro corrected
paper small corrections

@@
 Additionally, we introduce methods to decrease numerical costs of intrusive methods for steady problems.
-We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as the shallow water equations.
+We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment.
 \end{abstract}
@@
 \input{framework}
+\section{Strategies for steady problems}
 \input{colloc_ipm}
 \input{one_shot_ipm}
+\input{adaptivity}
 \input{results}
@@
 
+\newpage
 \bibliographystyle{unsrt}  

### 2019-07-11 Paper1/framework.tex
added adaptive one shot to paper and polished up some parts

@@
 \begin{align}\label{eq:numericalFluxIPMBurgers}
- \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \langle g(\mathcal{U}_{SG}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}_{SG}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}^T\rangle^T_Q.
+ \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}_{SG}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}_{SG}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k).
 \end{align}
-When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the integral brackets is of order $3M$.
-Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 M +\frac12$ quadrature points suffice for an exact computation of the numerical flux.
+When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the sum is of order $3M=3(N-1)$.
+Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 N -1$ quadrature points suffice for an exact computation of the numerical flux.
 Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent.
-Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When precomputing the values $\mathcal{U}_{SG}(\bm{\hat u};\xi_k)=\sum_{i=0}^N\hat u_i\varphi_i(\xi_k)$ which requires $M\cdot Q$ evaluations, the computation of \eqref{eq:numericalFluxIPMBurgers} again requires $O(M\cdot Q)$ operations, i.e.
-the costs are $O(M^2)$, whereas the evaluation of \eqref{eq:numFluxAnalytic} requires $O(M^3)$ operations.
-
-TODO: Extension to 2D $\rightarrow$ triangular meshes
+Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When computing and storing the values in a matrix $\bm A\in\mathbb{R}^{Q\times N}$ with entries $a_{ki} = \varphi_i(\xi_k)$ before running the program, the numerical flux \eqref{eq:numericalFluxIPMBurgers} can be split into two parts.
+First, we determine the SG solution at all quadrature points, i.e.
+we compute $\bm{u}^{(\ell)} := \bm A \bm{\hat u}_{\ell}$ and $\bm{u}^{(r)} := \bm A \bm{\hat u}_{r}$ which requires $O(N\cdot Q)$ operations.
+These solution values are then used to compute the numerical flux
+\begin{align*}
+G_i(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) &= \sum_{k=1}^Q w_k g(u^{(\ell)}_k,u^{(r)}_k)a_{ki}f_{\Xi}(\xi_k),
+\end{align*}
+which again requires $O(N\cdot Q)$ operations, i.e. the costs are $O(N^2)$. The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
 

### 2019-07-11 Paper1/introduction.tex
added adaptive one shot to paper and polished up some parts

@@
 When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
-These non-physical structures tend to dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}.
-Here, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies which locally increase the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or add artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur.
+These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}.
+Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{durrwachter2018hyperbolicity}.
+These structures commonly arise on a small portion of the space-time domain.
+Therefore, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies.
+By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs.
 The main task here is to find an adequate refinement indicator.
@@
 \begin{itemize}
-\item Accelerate convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
+\item Accelerate the convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
 \item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e.

### 2019-07-11 Paper1/one_shot_ipm.tex
added adaptive one shot to paper and polished up some parts

@@
 \end{align}
-Since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by
+Since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
 \begin{align*}
@@
 \end{align*}
-The Jacobian $\bm{J}\in\mathbb{R}^{2(N+1)\cdot N_x \times 2(N+1)\cdot N_x}$ has the form
+The Jacobian $\bm{J}\in\mathbb{R}^{2N\cdot N_x \times 2N\cdot N_x}$ has the form
 \begin{align}\label{eq:Jacobian}
@@
 \begin{align*}
-\text{det}(\bm{\tilde J}) = \sum_{\sigma} \text{sgn}(\sigma)\prod_{i = 1}^{2 N_x (N+1)} \tilde J_{\sigma(i),i}.
+\text{det}(\bm{\tilde J}) = \sum_{\sigma} \text{sgn}(\sigma)\prod_{i = 1}^{2 N_x N} \tilde J_{\sigma(i),i}.
 \end{align*}

### 2019-07-11 code/src/main.cpp
bugfix and I/O cleaned

@@ -151,7 +151,7 @@ void initLogger( spdlog::level::level_enum terminalLogLvl, spdlog::level::level_
     fileSink->set_pattern( "%Y-%m-%d %H:%M:%S.%f | %v" );
 
     std::vector<spdlog::sink_ptr> sinks;
-    sinks.push_back( terminalSink );
+    if( mype == 0 ) sinks.push_back( terminalSink );
     if( fileLogLvl != spdlog::level::off ) sinks.push_back( fileSink );
 
     auto event_logger = std::make_shared<spdlog::logger>( "event", begin( sinks ), end( sinks ) );
@@ -163,12 +163,6 @@ void initLogger( spdlog::level::level_enum terminalLogLvl, spdlog::level::level_
     momentFileSink->set_pattern( "%v" );
     auto moment_logger = std::make_shared<spdlog::logger>( "moments", momentFileSink );
     spdlog::register_logger( moment_logger );
-
-    auto dualsFileSink = std::make_shared<spdlog::sinks::basic_file_sink_mt>( outputDir + "/logs/" + currentDateTime() + "_duals" );
-    dualsFileSink->set_level( spdlog::level::info );
-    dualsFileSink->set_pattern( "%v" );
-    auto duals_logger = std::make_shared<spdlog::logger>( "duals", dualsFileSink );
-    spdlog::register_logger( duals_logger );
 }
 
 void PrintInit( std::string configFile ) {
@@ -250,7 +244,6 @@ int main( int argc, char* argv[] ) {
     // perform reduction to obtain full moments on all PEs
     std::vector<int> PEforCell = settings->GetPEforCell();
     for( unsigned j = 0; j < nCells; ++j ) {
-        u[j].reset();
         MPI_Reduce( uMoments[j].GetPointer(), u[j].GetPointer(), int( nStates * nTotal ), MPI_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD );
     }
 
@@ -263,7 +256,10 @@ int main( int argc, char* argv[] ) {
     // compute errors
     std::vector<std::vector<double>> referenceSolution;
     unsigned nQTotal = settings->GetNQTotal();
-    if( settings->GetMyPE() != 0 ) return EXIT_SUCCESS;
+    if( settings->GetMyPE() != 0 ) {
+        MPI_Finalize();
+        return EXIT_SUCCESS;
+    }
     Matrix meanAndVar;
[…]

### 2019-07-12 Paper1/adaptivity.tex
paper
added adaptive one shot to paper and polished up some parts
paper continued

@@
 in the case of gas dynamics, the regularity of the density is chosen to indicate an adequate refinement level.
-If the moment vector in a given cell at a certain timestep is initially at refinement level $\ell$, this level is kept if if the error indicator \eqref{eq:errorIndicator} lies in the interval $I_{\delta}:=[\delta_{-},\delta_{+}]$.
+If the moment vector in a given cell at a certain timestep is initially at refinement level $\ell$, this level is kept if the error indicator \eqref{eq:errorIndicator} lies in the interval $I_{\delta}:=[\delta_{-},\delta_{+}]$.
 Here $\delta_{\pm}$ are user determined parameters.
@@
 \end{algorithm}
+The same strategy can be applied for one-shot IPM. In this case, the algorithm is given by
+\begin{algorithm}[H]
+\begin{algorithmic}[1]
+\For{$j=0$ to $N_x+1$}
+\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle_Q dx$
+\EndFor
+\While{\eqref{eq:residualUnsteady} is violated}
+\For{$j=1$ to $N_x$}
+\State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
+\State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right)$
+\State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$
+\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$
+\EndFor
+\State $n \leftarrow n+1$
+\EndWhile
+\end{algorithmic}
+\caption{Adaptive one-shot IPM implementation}
+\label{alg:osIPM}
+\end{algorithm}
 %TODO Auch adaptiver Algorithmus fuer osIPM?

### 2019-07-12 Paper1/results.tex
Paper continued
paper continued

@@
 We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 profile computed with different methods.
-To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $50$ quadrature points and compare against Collocation with $5$ collocation points as well as SG and IPM with $5$ moments and $10$ quadrature points.
+The stochastic Euler equations in two dimensions are given by
+\begin{align*}
+\partial_t
+\begin{pmatrix}
+\rho \\ \rho v_1 \\ \rho v_2 \\ \rho e
+\end{pmatrix}
++\partial_{x_1}
+\begin{pmatrix}
+\rho v_1 \\ \rho v_1^2 +p \\ \rho v_1 v_2 \\  v_1 (\rho e+p)
+\end{pmatrix}
++\partial_{x_2}
+\begin{pmatrix}
+\rho v_2 \\ \rho v_1 v_2 \\ \rho v_2^2+p \\ v_2 (\rho e+p)
+\end{pmatrix}
+=\bm{0}.
+\end{align*}
+These equations determine the time evolution of the conserved variables $(\rho,rho \bm v, rho e)$, i.e. density, momentum and energy. A closure for the pressure $p$ is given by
+\begin{align*}
+p = (\gamma-1)\rho\left(e-\frac12(v_1^2+v_2^2)\right).
+\end{align*}
+Since the the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$.
+The spatial mesh discretizes the flow domain around the airfoil.
+At the airfoil boundary $\Gamma_{0}$, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal.
+At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101,325$ Pa and a temperature of $273.15$ K.
+Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees.
+I.e.
+we choose $\phi(\xi) = 1.25 + 0.5\xi$ where $\xi\sim U(-1,1)$.
+The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ with different methods.
+The IPM methods makes use of the acceleration strategies proposed in this work.
+To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points.
+
+ and compare against Collocation with $5$ collocation points as well as SG and IPM with $5$ moments and $10$ quadrature points.
 Furthermore, we use convergence accelerated (caIPM) as well as convergence accelerated one-shot IPM (caosIPM), which we introduced in Sections \ref{sec:collIPM} and \ref{sec:OneShotIPM} with $10$ moments and $15$ quadrature points to converge the collocation solution with $5$ quadrature points to an entropy solution with increased accuracy.
@@
 		\centering
-		\includegraphics[scale=0.5]{figs/{convergence_runtime_residual}.pdf}
+		\includegraphics[scale=0.7]{figs/{L2_error_E[rho]osIPMIPMCol}.pdf}
+		\label{fig:sub1}
+	\caption{L2 error at airfoil with 5 MPI and 2 OMP threads.}
+	\label{fig:Obstacles2D}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+		\centering
+		\includegraphics[scale=0.7]{figs/{L2_error_E[rho]}.pdf}
+		\label{fig:sub1}
+	\caption{L2 error at airfoil with 5 MPI and 2 OMP threads.}
+	\label{fig:Obstacles2D}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+		\centering
+		\includegraphics[scale=0.7]{figs/{convergence_runtime_residual}.pdf}
 		\label{fig:sub1}

### 2019-07-12 Paper1/appendix.tex
results started + details IPM added

\appendix
\section{IPM for the 2D Euler equations}
\label{app:IPM2DEuler}

In the following, we provide details on the implementation of IPM for the 2D Euler equations. We for clarity, we denote the momentum by $m_1 := \rho v_1$ and $m_2:=\rho v_2$ and the energy by $E:=\rho e$. Then, the vector of conserved variables is $\bm u = (\rho,m_1,m_2,E)^T$. The entropy used is
\begin{align*}
s(\bm u) = -\rho \ln \left(\rho^{-\gamma} \left(E - \frac{m_1^2 + m_2^2}{2
\rho}\right)\right).
\end{align*}
Now the gradient of the entropy $\nabla_{\bm u} s$ has the components
\begin{align*}
\frac{\partial s}{\partial \rho} &= -\ln \left(\rho^{-\gamma}\left(E-\frac{m_1^2+m_2^2}{2 \rho }\right)\right)+\frac{m_1^2+m_2^2}{-2 \rho  E+m_1^2+m_2^2}+\gamma, \\
\frac{\partial s}{\partial m_i} &= -\frac{2\rho  m_i}{-2 \rho  E+m_1^2+m_2^2}, \\
\frac{\partial s}{\partial E} &=-\frac1\rho\left(E-\frac{m_1^2+m_2^2}{2 \rho }\right).
\end{align*}
To compute $\bm u_s(\bm\Lambda) = (\nabla_{\bm u}s)^{-1}(\bm \Lambda)$, we set $\bm \Lambda = \nabla_{\bm u}s(\bm u)$ and rearrange with respect to $\bm u$. Let us define
%\begin{align*}
%\alpha(\bm\Lambda) := \text{exp}\left(-\frac{0.5 \left(-10 \Lambda_1 \Lambda_4+5 \Lambda_2^2+5 \Lambda_3^2+14 \Lambda_4\right)}{\Lambda_4}\right) \cdot (-\Lambda_4)^{\frac{1}{1+\gamma}}.
%\end{align*}
\begin{align*}
\alpha(\bm\Lambda) := \text{exp}\left(\frac{ \Lambda_2^2 + \Lambda_3^2 - 2\Lambda_1  \Lambda_4 - 2 \Lambda_4  \gamma }{ 2 \Lambda_4(1-\gamma) } \right) \cdot (-\Lambda_4)^{\frac{1}{1-\gamma}}
\end{align*}
Then the solution ansatz $\bm u_s$ is given by
\begin{align*}
\rho(\bm\Lambda) &= \alpha(\bm\Lambda),\qquad m_1(\bm\Lambda) = -\frac{\Lambda_2 \alpha(\bm{\Lambda})}{\Lambda_4},\qquad m_2(\bm\Lambda) = -\frac{\Lambda_3 \alpha(\bm{\Lambda})}{\Lambda_4}, \\
E(\bm\Lambda) &= -\frac{  \alpha(\bm{\Lambda}) ( -\Lambda_2^2 - \Lambda_3^2 + 2\Lambda_4 ) }{ 2 \Lambda_4^2}.
\end{align*}

### 2019-07-12 Paper1/results.tex
results started + details IPM added

@@
 To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points.
+
+In the following, we compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques.
+Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
+For more information on the chosen entropy and the resulting solution ansatz for IPM, see \ref{app:IPM2DEuler}.
 

### 2019-07-12 code/src/main.cpp
MPI_FINALIZE removed

@@ -256,10 +256,6 @@ int main( int argc, char* argv[] ) {
     // compute errors
     std::vector<std::vector<double>> referenceSolution;
     unsigned nQTotal = settings->GetNQTotal();
-    if( settings->GetMyPE() != 0 ) {
-        MPI_Finalize();
-        return EXIT_SUCCESS;
-    }
     Matrix meanAndVar;
     Matrix meanAndVarErrors = Matrix( 2 * nStates, mesh->GetNumCells(), 0.0 );
     if( settings->HasExactSolution() ) {

### 2019-07-12 code/src/main.cpp
MPI_FINALIZE at end

@@ -256,134 +256,135 @@ int main( int argc, char* argv[] ) {
     // compute errors
     std::vector<std::vector<double>> referenceSolution;
     unsigned nQTotal = settings->GetNQTotal();
-    Matrix meanAndVar;
-    Matrix meanAndVarErrors = Matrix( 2 * nStates, mesh->GetNumCells(), 0.0 );
-    if( settings->HasExactSolution() ) {
-        meanAndVar = Matrix( 4 * nStates, mesh->GetNumCells(), 0.0 );
-    }
-    else {
-        meanAndVar = Matrix( 2 * nStates, mesh->GetNumCells(), 0.0 );
-    }
-    Matrix phiTildeWf = closure->GetPhiTildeWf();
-    Vector tmp( nStates, 0.0 );
-    for( unsigned j = 0; j < nCells; ++j ) {
-        // expected value
-        for( unsigned k = 0; k < settings->GetNQTotal(); ++k ) {
-            for( unsigned i = 0; i < nStates; ++i ) {
-                meanAndVar( i, j ) += uQFinal[j]( i, k ) * phiTildeWf( k, 0 );
-            }
-        }
-
-        // variance
-        for( unsigned k = 0; k < nQTotal; ++k ) {
-            for( unsigned i = 0; i < nStates; ++i ) {
-                meanAndVar( i + nStates, j ) += pow( uQFinal[j]( i, k ) - meanAndVar( i, j ), 2 ) * phiTildeWf( k, 0 );
-            }
-        }
-    }
-    // compute distance to exact solution
-    if( settings->HasExactSolution() ) {
-        // store xGrid on vector
-        Matrix xGrid( nCells, settings->GetMeshDimension() );
-        for( unsigned j = 0; j < nCells; ++j ) {
-            Vector midPj = mesh->GetCenterPos( j );
-            for( unsigned s = 0; s < settings->GetMeshDimension(); ++s ) {
-                xGrid( j, s ) = midPj[s];
-            }
-        }
-        Vector xiEta( settings->GetNDimXi() );
-        std::vector<Polynomial*> quad( 2 );
-        unsigned nQuadFine = 5 * settings->GetNQuadPoints();    // define fine quadrature for exact solution
-        quad[0]            = Polynomial::Create( settings, nQuadFine, DistributionType::D_LEGENDRE );
-        quad[1]            = Polynomial::Create( settings, nQuadFine, DistributionType::D_HERMITE );
-        unsigned n;
-
[…]

### 2019-07-13 Paper1/adaptivity.tex
refinement retardation added

@@
 \end{algorithm}
+Adaptivity can be used for intrusive methods in general as well as for steady and unsteady problems.
+In the case of steady problems, we can make use of a strategy, which we call \textit{refinement retardation}.
+Recall that the convergence to an admissible steady state solution is expensive and a high accuracy and desirable solution properties are only required at the end of this iteration process.
+Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualUnsteady} is fulfilled.
+For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of th required iterations on a lower but cheaper accuracy level.
 The same strategy can be applied for one-shot IPM. In this case, the algorithm is given by
@@
 \For{$j=1$ to $N_x$}
-\State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
-\State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right)$
+\State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
+\State $\ell_j^{n+1}\leftarrow \max\{\text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right),\ell_i^*\}$
 \State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$
-\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$
+\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}_{\ell_j^{n+1}}^{\bm\ell'}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$
 \EndFor
 \State $n \leftarrow n+1$
+\If{\eqref{eq:residualUnsteady} fulfills the stopping criterion $\varepsilon_m^*$}
+\State $m \leftarrow m+1$
+\EndIf
 \EndWhile
 \end{algorithmic}
-\caption{Adaptive one-shot IPM implementation}
-\label{alg:osIPM}
+\caption{Adaptive one-shot IPM implementation with refinement retardation}
+\label{alg:adosIPM}
 \end{algorithm}
-%TODO Auch adaptiver Algorithmus fuer osIPM?

### 2019-07-13 code/src/main.cpp
old version but with sparse grids

@@ -248,194 +248,24 @@ int main( int argc, char* argv[] ) {
         MPI_Reduce( uMoments[j].GetPointer(), u[j].GetPointer(), int( nStates * nTotal ), MPI_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD );
     }
 
-    std::cout << "Moments saved on u for PE " << settings->GetMyPE() << std::endl;
-
-    // send UQFinal to master thread //
-    unsigned nQTotal = settings->GetNQTotal();
-    std::vector<MatVec> uQFinalMat;
-    MatVec uQFinal( nCells, Matrix( nStates, nQTotal ) );
-    uQFinalMat.resize( nQTotal );
-    for( unsigned k = 0; k < nQTotal; ++k ) uQFinalMat[k] = MatVec( nCells, Matrix( nStates, 1 ) );
-
-    // store partial solution on full data field
-    for( unsigned k = settings->GetKStart(); k <= settings->GetKEnd(); ++k ) {
-        uQFinalMat[k] = uQ[k - settings->GetKStart()];
-    }
-
-    std::cout << "uQ saved on Full Matrix for PE " << settings->GetMyPE() << std::endl;
-
-    int PEhasK;
-    int usePE;
-    for( unsigned k = 0; k < nQTotal; ++k ) {
-        if( settings->GetMyPE() == 1 ) std::cout << "k = " << k << " of " << nQTotal << std::endl;
-        // which PE has this quadrature interval? store on usePE
-        if( k <= settings->GetKEnd() && k >= settings->GetKStart() ) {
-            PEhasK = settings->GetMyPE();
-        }
-        else {
-            PEhasK = 0;
-        }
-        usePE = 0;
-        MPI_Allreduce( &PEhasK, &usePE, 1, MPI_INT, MPI_SUM, MPI_COMM_WORLD );
-        // if( settings->GetMyPE() == 1 )
-        std::cout << "Reduction DONE." << std::endl;
-        // send from usePE to all other PEs
-        for( unsigned j = 0; j < nCells; ++j ) {
-            // std::cout << j << std::endl;
-            MPI_Bcast( uQFinalMat[k][j].GetPointer(), int( nStates ), MPI_DOUBLE, usePE, MPI_COMM_WORLD );
-            // if( settings->GetMyPE() == 1 )
-            std::cout << "BCast DONE." << std::endl;
-            // if( settings->GetMyPE() == 1 )
[…]

### 2019-07-13 Paper1/introduction.tex
small changes introduction

@@
 In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
-We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.
+In the case of steady problems, one can perform a large number of iteration to the steady state solution on a low number of moments and increase the maximal refinement level when the distance to the steady state has reached a specified barrier.
+Similar to the idea of using IPM as a post processing step, a large number of iterations will be performed by a cheap method, i.e.
+we can reduce numerical costs.
+The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.
 

### 2019-07-14 Paper1/colloc_ipm.tex
paper
started correcting Martins comments
smaller stuff fixed

@@
 \end{align}
-which is again needed for the construction of intrusive methdos.
+which is again needed for the construction of intrusive methods.
 By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:IPMmomentSystem}, i.e.
 Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
-Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the residual
+Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the condition
 \begin{align}\label{eq:residualUnsteady}
@@
 Commonly, a great amount of iterations in pseudo-time are needed to converge to a steady state solution.
-Consequently, the IPM method which requires solving the dual problem \eqref{eq:dualProblem} in every spacial cell in each iteration becomes prohibitively expensive.
+Consequently, the IPM method which requires solving the dual problem \eqref{eq:dualProblem} in every spatial cell in each iteration becomes prohibitively expensive.
 We tackle this problem by using IPM only as a postprocesssing step for the steady solution obtained by a cheap method.

### 2019-07-14 Paper1/framework.tex
started correcting Martins comments
smaller stuff fixed

@@
 \end{align}
-The exact dual state is then obtained by computing the fix point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
+The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
 $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
@@
 
-We now write down the entire scheme: To obtain a more compact notation, one defines
+We now write down the entire scheme: To obtain a more compact notation, we define
 \begin{align}\label{eq:momentIterationFunction}
@@
 \end{algorithmic}
-\caption{IPM implementation}
+\caption{IPM algorithm}
 \label{alg:IPM}
@@
 
-\subsection{Costs of the numerical flux}
-A straight forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}.
+\subsection{Costs of evaluating the numerical flux}
+A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}.
 By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
@@
 which again requires $O(N\cdot Q)$ operations, i.e. the costs are $O(N^2)$. The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
-
-\subsection{Code framework}
-
-TODO: Code structure, why is it easy to include different equations?

### 2019-07-14 Paper1/introduction.tex
started correcting Martins comments
smaller stuff fixed

@@
 \end{align*} 
+Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
+Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $Q = O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly.
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
@@
 A downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
-Furthermore, collocation methods require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation}.
+Therefore, collocation methods typically require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation} to reach a given accuracy.
 Despite their easy implementation, collocation methods face challenges when examining unsteady problems, since in this case, the solution needs to be written out, i.e.
@@
 When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty} and the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
-A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem.
+A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constrained optimization problem.
 For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is given by
@@
 \end{align}
-Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstraint optimization problem.
+Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstrained optimization problem.
 Defining the variables $\bm{\lambda}_i\in\mathbb{R}^p$ where $i$ is again a multi index, gives the unconstrained dual problem
@@
 This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution values, which can then cause a failure of the method.
-The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
+The main weakness of the IPM method is its computational effort, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
 Hence, the desirable properties of IPM come along with significantly increased numerical costs.

### 2019-07-14 Paper1/main.tex
paper continued
results started + details IPM added
started correcting Martins comments
smaller stuff fixed

@@
 % Declare title and authors, without \thanks
-\newcommand{\TheTitle}{A semi-intrusive Code Framework for Uncertainty Quantification}
+\newcommand{\TheTitle}{Acceleration strategies for intrusive methods in Uncertainty Quantification}
 
@@
 A main drawback of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver.
-Furthermore, intrusive methods come with increased computational costs, making it hard to compete with non-intrusive methods.
+Moreover, non-intrusive methods are embarrassingly parallel and therefore easy to parallelize.
 In this work, we present a code framework, which facilitates the implementation of intrusive methods by recycling parts of deterministic codes.
@@
 \input{summary_outlook}
+\input{appendix}
 

### 2019-07-14 Paper1/one_shot_ipm.tex
started correcting Martins comments
smaller stuff fixed

@@
 In the following, we will show that this iteration converges, if the chosen initial condition is sufficiently close to the steady state solution.
-For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fix point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence:
+For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence:
 
 \begin{theorem}\label{th:Contractive}
-Assume that the classical IPM iteration is contractive at its fix point $\bm{\hat u}^*$.
-Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fix point $(\bm{\lambda}^*,\bm{\hat u}^*)$ if the preconditioner $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$, i.e.
+Assume that the classical IPM iteration is contractive at its fixed point $\bm{\hat u}^*$.
+Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$ if the preconditioner $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$, i.e.
 the Hessian of the dual problem is used.
@@
 \end{align}
-Since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
+Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
 \begin{align*}
@@
 Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
-Note that in its current form this iteration is not really a fix point iteration, since it uses the time updated dual variables in \eqref{eq:oneshotIPMmoment}.
-To obtain a fix point iteration, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
+Note that in its current form this iteration is not really a fixed point iteration, since it uses the time updated dual variables in \eqref{eq:oneshotIPMmoment}.
+To obtain a fixed point iteration, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
 \begin{align*}
@@
 where we used $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T}$.
-Recall that at the fix point $(\bm{\lambda}^*,\bm{\hat u}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
+Recall that at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
 For the block $\partial_{\bm{\hat u}} \bm{d}$, we get
@@
 Hence, we have that $\partial_{\bm{\hat u}} \bm{c} = \nabla_{\bm{\hat u}}\bm{\tilde c}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM iteration is contractive.
-Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{\hat u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fix point.
+Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{\hat u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fixed point.
 \end{proof}
@@
 \begin{proof}
-By Theorem \ref{th:Contractive}, the One-Shot scheme is contractive at its fix point.
+By Theorem \ref{th:Contractive}, the One-Shot scheme is contractive at its fixed point.
 Since we assumed convergence of the classical IPM scheme, we can conclude that all entries in the Jacobian $\bm{J}$ are continuous functions.
@@
 Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{\hat u})$.
-Hence there exists an open ball with radius $\delta$ around the fix point in which the eigenvalues remain in the interval $(-1,1)$.
+Hence there exists an open ball with radius $\delta$ around the fixed point in which the eigenvalues remain in the interval $(-1,1)$.
 \end{proof}

### 2019-07-14 Paper1/figs/TestC.jl

@@ -1,0 +1,31 @@
+using PyPlot
+
+NMax = 50;
+
+entries = zeros(NMax);
+for N = 1:NMax
+    #C = zeros(N,N,N);
+    counterNonZero = 0;
+    for i = 1:N
+        for j = 1:N
+            for k = 1:N
+                s = 0.5*(i+j+k)
+                if s >= i && s >= j && s >= k
+                    counterNonZero += 1;
+                    #C[i,j,k] = factorial(i)*factorial(j)*factorial(k)/(factorial(s-i)*factorial(s-j)*factorial(s-k));
+                else
+                    #C[i,j,k] = 0;
+                end
+            end
+        end
+    end
+    entries[N] = counterNonZero;
+end
+
+fig, ax = subplots(figsize=(15, 8), dpi=100)#, facecolor='w', edgecolor='k') # dpi Aufloesung
+ax.plot((1:NMax).^3,entries, "r--o", linewidth=2, alpha=0.6)
+ax.set_xlabel(L"$N^3$", fontsize=20);
+ax.set_ylabel("entries in C", fontsize=20);
+ax.tick_params("both",labelsize=20)
+ax.set_xlim([0,NMax^3])
+ax.set_ylim([0,entries[end]+100])

### 2019-07-14 Paper1/introduction.tex
rewritten as standard UQ derivation

@@
 where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_s$.
-More generally, one is interested in determining the moments of the solution for a given set of basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
+To approximate quantities of interest (such as expectation value, variance or higher order moments), the solution is spanned with a set of polynomial basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
 Note that this yields
@@
 \end{align*}
-basis polynomials when defining $|i|:=\sum_{n = 1}^s |i_n|$.
-Commonly one chooses orthonormal polynomials as basis functions \cite{xiu2002wiener}, i.e.
+basis functions when defining $|i|:=\sum_{n = 1}^s |i_n|$.
+Commonly, these functions are chosen to be orthonormal polynomials \cite{wiener1938homogeneous} with respect to the probability function, i.e.
 $\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^s\delta_{i_nj_n}$.
-The corresponding moments are then given by $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
-Besides facilitating the computation of the expectation value and variance, the moments can be used to span the solution in $\bm\xi$ \cite{wiener1938homogeneous}.
-\\
+The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates by the solution by
+\begin{align}\label{eq:SGClosure}
+\mathcal{U}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
+\end{align}
+where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^p$ are called moments.
+To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i\leq M|}\in\mathbb{R}^{N\times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
+In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
+The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
+The solution ansatz \eqref{eq:SGClosure} can then be used to compute the quantities of interest.
+Indeed, we have that
+\begin{align*}
+\text{E}[\mathcal{U}(\bm{\hat u})] = \bm{\hat u}_0,\quad \text{Var}[\mathcal{U}(\bm{\hat u})] = \text{E}[\mathcal{U}(\bm{\hat u})^2] - \text{E}[\mathcal{U}(\bm{\hat u})]^2 = \left(\sum_{i = 1}^N \hat{u}_{\ell i}^2\right)_{\ell = 1,\cdots,p}.
+\end{align*}
 
@@
 Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
-Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $Q = O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly.
+Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly.
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
@@
 However, intrusive methods are in general more difficult to implement and come along with higher numerical costs.
-The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
-\begin{subequations}\label{eq:nonClosedMomentSystem}
+The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin (SG) \cite{ghanem2003stochastic} moment system
+\begin{subequations}\label{eq:SGMomentSystem}
 \begin{align}
-\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\bm{u}(t,\bm{x},\cdot)) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
+\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
+\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle
 \end{align}
 \end{subequations}
-To obtain a closed set of equations, one needs to derive a closure $\mathcal{U}$ such that 
-\begin{align*}
-\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u};\bm\xi).
-\end{align*}
-Here, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\bm{\hat{u}}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N \times p}$.
-The dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted in the following for sake of readability.
-A commonly used closure is given by stochastic-Galerkin (SG) \cite{ghanem2003stochastic}, which represents the solution by a polynomial:
-\begin{align}\label{eq:SGClosure}
-\mathcal{U}_{\text{SG}}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
-\end{align}
-Corresponding to the definition of $\bm{\hat u}$, we collect the $N$ basis function for which $\vert i \vert \leq M$ holds in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N+1}$.
-When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty} and the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
-A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure \cite{poette2009uncertainty}: The closure is given by a constrained optimization problem.
+with $|i|\leq M$.
+This system of $N$ equations can be solved by a deterministic method to determine the time evolution of the moments.
+A drawback of the stochastic-Galerkin method is that the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
+In order to remain hyperbolic, the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
+A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
+The main difference to SG is that the solution ansatz of IPM is given by an optimization problem instead of a polynomial expansion \eqref{eq:SGClosure}.
 For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is given by
@@
 \end{align}
-Plugging the derived closure into the moment system \eqref{eq:nonClosedMomentSystem}, one obtains the IPM moment system
+Plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the IPM system
 \begin{subequations}\label{eq:IPMmomentSystem}
@@
 \partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
+\bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
 \end{align}
@@
 \end{itemize}
-In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
+For both steady and unsteady problems, we use adaptivity: 
+\begin{itemize}
+\item The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
 In the case of steady problems, one can perform a large number of iteration to the steady state solution on a low number of moments and increase the maximal refinement level when the distance to the steady state has reached a specified barrier.
@@
 we can reduce numerical costs.
+\end{itemize}
 The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.

### 2019-07-15 Paper1/introduction.tex
continued paper

@@
 Note that this yields
-\begin{align*}
+\begin{align}\label{eq:numberBasisFcts}
 N:=\begin{pmatrix}
@@
 \end{pmatrix}
-\end{align*}
+\end{align}
 basis functions when defining $|i|:=\sum_{n = 1}^s |i_n|$.
@@
 $\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^s\delta_{i_nj_n}$.
-The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates by the solution by
+The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates the solution by
 \begin{align}\label{eq:SGClosure}
@@
 Therefore, collocation methods typically require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation} to reach a given accuracy.
-Despite their easy implementation, collocation methods face challenges when examining unsteady problems, since in this case, the solution needs to be written out, i.e.
+Despite their easy implementation, collocation methods face challenges when examining unsteady problems.
+In this case, the solution needs to be written out, i.e.
 stored in an external file, at all quadrature points in every time step to compute the time evolution of the moments.
@@
 \partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle
+\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
 \end{align}
@@
 By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs.
-The main task here is to find an adequate refinement indicator.
+In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments $N$ given by \eqref{eq:numberBasisFcts} is asymptotically smaller than the number of quadrature points, which increases with $O(M(\log_2(M)^{s-1}))$.
+Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex, high-dimensional settings.
 \\
 
-In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods.
-The framework only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used).
-We thereby provide the ability to recycle existing implementations of deterministic solvers.
-Furthermore, we investigate intrusive methods for steady problems and compare them against collocation methods.
+In this paper, we investigate intrusive methods for steady problems and compare them against collocation methods. For both steady and unsteady problems, we use adaptivity: 
+\begin{itemize}
+\item The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
+\end{itemize}
 The steady setting provides different opportunities to take advantage of features of intrusive methods: 
 \begin{itemize}
+\item When using adaptivity, one can perform a large number of iteration to the steady state solution on a low number of moments and increase the maximal refinement level when the distance to the steady state has reached a specified barrier.
+Consequently, a large number of iterations will be performed by a cheap method, i.e.
+we can reduce numerical costs.
 \item Accelerate the convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
@@
 \end{itemize}
-For both steady and unsteady problems, we use adaptivity: 
-\begin{itemize}
-\item The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
-In the case of steady problems, one can perform a large number of iteration to the steady state solution on a low number of moments and increase the maximal refinement level when the distance to the steady state has reached a specified barrier.
-Similar to the idea of using IPM as a post processing step, a large number of iterations will be performed by a cheap method, i.e.
-we can reduce numerical costs.
-\end{itemize}
+
 The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.
+Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods.
+The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used).
+We thereby provide the ability to recycle existing implementations of deterministic solvers.
 
-The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive framework in section \ref{sec:framework}.
+The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive method in section \ref{sec:framework}.
 Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration.

### 2019-07-15 Paper1/main.tex
continued paper

@@
 A main drawback of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver.
-Moreover, non-intrusive methods are embarrassingly parallel and therefore easy to parallelize.
-In this work, we present a code framework, which facilitates the implementation of intrusive methods by recycling parts of deterministic codes.
-Additionally, we introduce methods to decrease numerical costs of intrusive methods for steady problems.
+Unlike intrusive methods, non-intrusive methods yield a decoupled system of equations, meaning that they are embarrassingly parallel.
+Moreover, the coupled system of the intrusive stochastic-Galerkin (SG) method is not necessarily hyperbolic.
+The Intrusive Polynomial Moment (IPM) method fixes this issue at the cost of having to solve an optimization problem in every spatial cell and every time step.
+In this work, we discuss advantages of intrusive methods and propose acceleration techniques for SG and IPM.
+When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from shape optimization.
+By integrating the iteration to solve the optimization problem into the moment update, we can guarantee local convergence while reducing numerical costs.
+Furthermore, we propose to use non-intrusive methods as a preconditioner for intrusive methods.
+Consequently, a large amount of iterations to the steady solution are performed by a cheap method, while maintaining the increased accuracy of an intrusive method.
+Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
 We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment.

### 2019-07-15 Paper1/introduction.tex
a bit more

@@
 In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments $N$ given by \eqref{eq:numberBasisFcts} is asymptotically smaller than the number of quadrature points, which increases with $O(M(\log_2(M)^{s-1}))$.
-Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex, high-dimensional settings.
+Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex and high-dimensional settings.
 \\
@@
 we can reduce numerical costs.
-\item Accelerate the convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
+\item Accelerate the convergence to the steady state IPM solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
 \item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e.
 they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
+Consequently, the dual variables and the moments are converges simultaneously to their steady state, which is similar to the idea of one-shot optimization in shape optimization \cite{hazra2005aerodynamic}.
 \end{itemize}

### 2019-07-15 Paper1/framework.tex
some more corrections

@@
-\section{Discretization and code framework}
+\section{Discretization of the IPM system}
 \label{sec:framework}
-\subsection{Discretization}
+\subsection{Finite Volume Discretization}
 In the following, we discretize the moment system in space and time according to \cite{kusch2017maximum}.
@@
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
-First we perform the discretization of the spatial domain: We choose the discrete unknowns which represent the solution to be the spatial averages over each cell at time $t_n$, given by
+We choose the discrete unknowns which represent the solution to be the spatial averages over each cell at time $t_n$, given by
 \begin{align*}
@@
 \end{align}
-for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
-The numerical flux is assumed to be consistent, i.e. $\bm{G}(\hat{\bm{u}},\hat{\bm{u}})=\bm{F}(\hat{\bm{u}})$.
+for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$. Here, the number of spatial cells is denoted by $N_x$ and the number of time steps by $N_t$.
+The numerical flux is assumed to be consistent, i.e. $\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$.
 
@@
 \end{align}
-Note that the numerical flux requires evaluating the closure $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^s \to \mathbb{R}^s$,
+Note that the numerical flux requires evaluating the ansatz $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^s \to \mathbb{R}^s$,
 \begin{align*}
@@
 \end{align*}
-meaning that the closure \eqref{eq:ansatz} at cell $j$ in timestep $n$ can be written as
+meaning that the IPM ansatz \eqref{eq:ansatz} at cell $j$ in timestep $n$ can be written as
 \begin{align*}
@@
 The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
-Testing with these basis functions and choosing the SG closure \eqref{eq:SGClosure} yields the closed SG moment system
+Choosing the SG ansatz \eqref{eq:SGClosure} and testing with the gPC basis functions yields the SG moment system
 \begin{align*}
@@
 \begin{align}\label{eq:numericalFluxIPMBurgers}
- \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}_{SG}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}_{SG}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k).
+ \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k),
 \end{align}
-When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the sum is of order $3M=3(N-1)$.
+where $\mathcal{U}$ is the SG ansatz \eqref{eq:SGClosure}.
+When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the sum is $3M=3(N-1)$.
 Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 N -1$ quadrature points suffice for an exact computation of the numerical flux.

### 2019-07-15 Paper1/introduction.tex
some more corrections

@@
 Despite their easy implementation, collocation methods face challenges when examining unsteady problems.
-In this case, the solution needs to be written out, i.e.
-stored in an external file, at all quadrature points in every time step to compute the time evolution of the moments.
+In this case, the solution needs to be cached at all quadrature points in every time step to compute the time evolution of the moments.
 This contradicts modern HPC paradigms, which aim at reducing the amount of data produced by numerical methods.

### 2019-07-15 Paper1/one_shot_ipm.tex
some more corrections

@@
 Hence, we aim at iterating the moments to steady state and the dual variables to the exact solution of the IPM optimization problem \eqref{eq:primalProblem} simultaneously.
-Hence, by successively performing one update of the moment iteration and one update of the dual iteration, we obtain 
+By successively performing one update of the moment iteration and one update of the dual iteration, we obtain 
 \begin{subequations}\label{eq:oneshotIPM}
@@
 \end{align}
-which is why we have the mapping $\bm{\hat u}:\mathbb{R}^{N+1}\to\mathbb{R}^{N+1}$, $\bm{\hat u}(\bm{\hat \lambda}) = \bm{h}(\bm{\hat \lambda})$.
+which is why we have the mapping $\bm{\hat u}:\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$, $\bm{\hat u}(\bm{\hat \lambda}) = \bm{h}(\bm{\hat \lambda})$.
 Since the solution of the dual problem for a given moment vector is unique, this mapping is bijective and therefore we have an inverse function

### 2019-07-15 Paper1/main.tex
abstract rewritten

@@
 Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
-A main drawback of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver.
-Unlike intrusive methods, non-intrusive methods yield a decoupled system of equations, meaning that they are embarrassingly parallel.
-Moreover, the coupled system of the intrusive stochastic-Galerkin (SG) method is not necessarily hyperbolic.
-The Intrusive Polynomial Moment (IPM) method fixes this issue at the cost of having to solve an optimization problem in every spatial cell and every time step.
-In this work, we discuss advantages of intrusive methods and propose acceleration techniques for SG and IPM.
+Non-intrusive methods allow using a given deterministic solver as a black box while being embarrassingly parallel.
+Intrusive methods on the other hand do not suffer from aliasing effects and provide a more general framework for adaptive refinement techniques.
+Since the moment system solved by intrusive methods is not necessarily hyperbolic, the Intrusive Polynomial Moment (IPM) method has been introduced.
+IPM maintains hyperbolicity but comes at the cost of having to solve an optimization problem in every spatial cell and every time step.
+In this work, we propose acceleration techniques for intrusive methods.
 When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from shape optimization.
-By integrating the iteration to solve the optimization problem into the moment update, we can guarantee local convergence while reducing numerical costs.
+Integrating the iteration to solve the optimization problem into the moment update guarantees local convergence while reducing numerical costs.
 Furthermore, we propose to use non-intrusive methods as a preconditioner for intrusive methods.
@@
 Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
-We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment.
+We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
 \end{abstract}

### 2019-07-16 code/src/mathtools.cpp
added selectable refinement thresholds

@@ -105,7 +105,7 @@ std::pair<Vector, Matrix> MathTools::ComputeEigenValTriDiagMatrix( const Matrix&
 
 int MathTools::Factorial( int i ) {
     int fact = 1;
-    for( unsigned n = 1; n <= i; n++ ) {
+    for( int n = 1; n <= i; n++ ) {
         fact *= i;
     }
     return fact;

### 2019-07-16 code/src/momentsolver.cpp

@@ -96,9 +96,10 @@ void MomentSolver::Solve() {
             // if( _mesh->GetBoundaryType( _cellIndexPE[j] ) == BoundaryType::DIRICHLET && timeIndex > 0 ) continue;
             double indicator = ComputeRefIndicator( refinementLevel, u[_cellIndexPE[j]], refinementLevel[_cellIndexPE[j]] );
 
-            if( indicator > 0.005 && refinementLevel[_cellIndexPE[j]] < _settings->GetNRefinementLevels( retCounter ) - 1 )
+            if( indicator > _settings->GetRefinementThreshold() &&
+                refinementLevel[_cellIndexPE[j]] < _settings->GetNRefinementLevels( retCounter ) - 1 )
                 refinementLevel[_cellIndexPE[j]] += 1;
-            else if( indicator < 0.0005 && refinementLevel[_cellIndexPE[j]] > 0 )
+            else if( indicator < _settings->GetCoarsenThreshold() && refinementLevel[_cellIndexPE[j]] > 0 )
                 refinementLevel[_cellIndexPE[j]] -= 1;
         }
 

### 2019-07-16 code/src/settings.cpp

@@ -88,7 +88,7 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
 
         if( !restart ) {
             auto refFile    = general->get_as<std::string>( "referenceSolution" );
-            _writeFrequency = general->get_as<int64_t>( "writeFrequency" ).value_or( 1000 );
+            _writeFrequency = general->get_as<int>( "writeFrequency" ).value_or( 1000 );
             if( refFile ) {
                 _referenceFile = _inputDir.string() + "/" + *refFile;
             }
@@ -258,7 +258,7 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
                         _polyIndices.push_back( indexTest );
                     }
                 }
-                _nTotalRefinementLevel[level] = _polyIndices.size();
+                _nTotalRefinementLevel[level] = static_cast<unsigned>( _polyIndices.size() );
                 if( UsesMaxDegree() ) break;
                 previousDegree = int( GetPolyDegreeforRefLevel( level ) );
             }
@@ -300,7 +300,6 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
         }
 
         // read convergence retardation
-        std::cout << "Reading cRetardation" << std::endl;
         auto cRetardation = moment_system->get_array_of<cpptoml::array>( "cRetardation" );
         if( cRetardation ) {
             auto retardationSteps    = ( *cRetardation )[0]->get_array_of<int64_t>();
@@ -320,7 +319,13 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
             _residualRetardation = Vector( 1, _minResidual );
             _nRetardationLevels  = 1;
         }
-        std::cout << "Reading cRetardation DONE." << std::endl;
+        if( _nRefinementLevels > 1 ) {
+            auto refinementThresholds = moment_system->get_array_of<double>( "refinementThresholds" );
+            assert( refinementThresholds->size() == 2 );
+            _refinementThreshold = ( *refinementThresholds )[0];
+            _coarsenThreshold    = ( *refinementThresholds )[1];
+        }
+
         _maxIterations = moment_system->get_as<unsigned>( "maxIterations" ).value_or( 1000 );
         _epsilon       = moment_system->get_as<double>( "epsilon" ).value_or( 5e-5 );
         _hasSource     = false;
[…]

### 2019-07-16 code/src/settings.h

@@ -60,6 +60,8 @@ class Settings
     Vector _residualRetardation;     // stores residual at different refinement levels
     VectorU _retardationSteps;       // stores maximal truncation orders at retardation step
     unsigned _nRetardationLevels;    // number of retardation levels
+    double _refinementThreshold;     // threshold for refinementIndicator above which moment/quadrature will be refined
+    double _coarsenThreshold;        // threshold for refinementIndicator below which moment/quadrature will be coarsened
 
     int _mype;                             // PE number
     int _npes;                             // number of all PEs
@@ -162,6 +164,8 @@ class Settings
     unsigned GetNRefinementLevels() const;
     unsigned GetNRefinementLevels( unsigned retardation ) const;
     unsigned GetNRetardationLevels() const;
+    double GetRefinementThreshold() const;
+    double GetCoarsenThreshold() const;
     double GetResidualRetardation( unsigned retardation ) const;
     unsigned GetNTotalforRefLevel( unsigned level ) const;
     unsigned GetPolyDegreeforRefLevel( unsigned level ) const;

### 2019-07-16 Paper1/adaptivity.tex
started correcting Martins comments
smaller stuff fixed
a bit more
implementation section added

@@
 
-Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders in different cells.
+Now we need to specify how the different building blocks of IPM can be modified to work with varying truncation orders in different cells.
 Let us first add dimensions to the notation of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIterationFunction}.
@@
 An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
-\begin{align*}
+\begin{align}\label{eq:adaptiveFVUpdate}
 \bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= \langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
-\end{align*}
+\end{align}
 Hence, the index vector $\bm\ell'\in\mathbb{N}^{3}$ denotes the refinement levels of the stencil cells, which are used to compute the time updated moment vector at level $\ell$.
@@
 \State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
-\State $\ell_j^{n+1}\leftarrow \max\{\text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right),\ell_i^*\}$
+\State $\ell_j^{n+1}\leftarrow \max\{\text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right),\ell_m^*\}$
 \State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$

### 2019-07-16 Paper1/framework.tex
implementation section added

@@
 \subsection{Costs of evaluating the numerical flux}
+\label{sec:costNumFlux}
 A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}.

### 2019-07-16 Paper1/implementation.tex
implementation section added

\section{Parallelization and Implementation}
\label{sec:parallel}
It remains to discuss the parallelization of the presented algorithms. In order to minimize the parallelization overhead, our goal is to minimize the need to send data between processors. Note that the dual problem (line 8 in Algorithm~\ref{alg:ad-IPM} and line 5 in Algorithm~\ref{alg:adosIPM}) does not have a stencil, i.e. it suffices to distribute the spatial cells between processors. In contrast to that, the finite volume update (line 14 in Algorithm~\ref{alg:ad-IPM} and line 8 in Algorithm~\ref{alg:adosIPM}) has a stencil. Hence, distributing the spatial mesh between processors will yield communication overhead since data needs to be sent whenever a stencil cell lies on a different processor. Therefore, we choose to parallelize the quadrature points, which again minimizes communication effort. As mentioned in Section~\ref{sec:costNumFlux}, we first compute the solution at stencil cells for all quadrature points. I.e. we determine $\bm u^{(j)}_k\in\mathbb{R}^p$ and the corresponding stencil cells for $k = 1,\cdots,Q$ by
\begin{align*}
\bm u^{(j-1)}_k := \bm u_{s}(\bm{\lambda}_{j-1}^T\bm{\varphi}_{\ell_1'}(\bm \xi_k)), \enskip \bm u^{(j)}_k := \bm u_{s}(\bm{\lambda}_{j}^T\bm{\varphi}_{\ell_2'}(\bm \xi_k)), \enskip \bm u^{(j+1)}_k := \bm u_{s}(\bm{\lambda}_{j+1}^T\bm{\varphi}_{\ell_3'}(\bm \xi_k)).
\end{align*}
Then, the finite volume update function \eqref{eq:adaptiveFVUpdate} can be written as
\begin{align}\label{eq:momentUpQuadrature}
\bm{c}_{\ell}^{\bm{\ell}'}&\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right)=\sum_{k=1}^Q w_k \left[\bm u^{(j)}_k- \frac{\Delta t}{\Delta x}\left(\bm g( \bm u^{(j)}_k,\bm u^{(j+1)}_k )- \bm g( \bm u^{(j-1)}_k,\bm u^{(j)}_k )\right)\right]\bm{\varphi}_{\ell}(\bm \xi_k)^T.
\end{align}
Instead of distributing the mesh on the different processors, we now distribute the quadrature set, i.e. the sum in \eqref{eq:momentUpQuadrature} can be computed in parallel. Now after having performed the dual update, the dual variables are send to all processors. With these variables, each processor computes the solution on its portion of the quadrature set and then computed its part of the sum in \eqref{eq:momentUpQuadrature} on all spacial cells. All parts from the different processors are then added together and the full time-updated moments are distributed to all processors. From here, the dual update can again be performed. The standard IPM Algorithm~\ref{alg:IPM} and one-shot IPM \ref{alg:osIPM} use this parallelization strategy accordingly. Again, we point out that stochastic-Galerkin is a variant of IPM, i.e. all presented techniques for IPM can also be used for SG. The SC algorithm that we use to compare intrusive with non-intrusive methods uses a given deterministic solver as a black box. Here, we distribute the quadrature set between all processors. Note that both, SC and IPM are based on the same deterministic solver, i.e. we use the same deterministic numerical flux $\bm g$.

### 2019-07-16 code/src/pnequations.cpp
added radiation hydro 1D

@@ -1,6 +1,6 @@
 #include "pnequations.h"
 
-PNEquations::PNEquations( Settings* settings ) : Problem( settings ), _N( 13 ) {
+PNEquations::PNEquations( Settings* settings ) : Problem( settings ), _N( 2 ) {
     _nStates = unsigned( GlobalIndex( _N, _N ) + 1 );
     _settings->SetNStates( _nStates );
     _settings->SetSource( true );

### 2019-07-16 code/src/problem.cpp

@@ -6,6 +6,7 @@
 #include "pnequations.h"
 #include "problem.h"
 #include "radiationhydrodynamics.h"
+#include "radiationhydrodynamics1d.h"
 #include "shallowwater.h"
 #include "shallowwater2d.h"
 
@@ -38,6 +39,9 @@ Problem* Problem::Create( Settings* settings ) {
     }
     else if( settings->GetProblemType() == ProblemType::P_RADIATIONHYDRO_2D ) {
         return new RadiationHydrodynamics( settings );
+    }
+    else if( settings->GetProblemType() == ProblemType::P_RADIATIONHYDRO_1D ) {
+        return new RadiationHydrodynamics1D( settings );
     }
     else {
         log->error( "[Problem] Invalid problem type!" );

### 2019-07-16 code/src/radiationhydrodynamics1d.cpp

@@ -1,0 +1,423 @@
+#include "radiationhydrodynamics1d.h"
+
+RadiationHydrodynamics1D::RadiationHydrodynamics1D( Settings* settings ) : PNEquations( settings, true ) {
+    _N            = 1;    // set moment order
+    _gamma        = 5.0 / 3.0;
+    _R            = 287.87;
+    double cLight = 299792458.0 * 100.0;    // speed of light in [cm/s]
+    // initial chock states
+    double lRef = 1000;
+    double rhoL = 5.45887 * 1e-13;
+    double uL   = 2.3545 * 1e5;
+    double TL   = 100.0;
+    double rhoR = 1.2479 * 1e-12;
+    double uR   = 1.03 * 1e5;
+    double TR   = 207.757;
+    double pRef = TL * ( _R * rhoL );
+    double aRef = sqrt( _gamma * pRef / rhoL );
+    _c          = cLight / aRef;
+    _P          = 0.0;    // 0.5;
+
+    if( _N == 1 )    // GlobalIndex has different ordering here
+        _nMoments = 4;
+    else
+        _nMoments = unsigned( GlobalIndex( _N, _N ) + 1 );
+    _nStates = _nMoments + 3;    // total number of states in equations in P_N + Euler
+    _settings->SetNStates( _nStates );
+    _settings->SetSource( true );    // TODO:DEBUGGING
+    _sigmaA = 3.93 * 1e-5 * lRef;    // absorption coefficient
+    _sigmaT = 0.848902 * lRef;       // 7.5657 * 1e-15;
+    _sigmaS = _sigmaT - _sigmaA;     // scattering coefficient
+    try {
+        auto file    = cpptoml::parse_file( _settings->GetInputFile() );
+        auto problem = file->get_table( "problem" );
+        SetupSystemMatrices();
+    } catch( const cpptoml::parse_exception& e ) {
+        _log->error( "[pnequations] Failed to parse {0}: {1}", _settings->GetInputFile(), e.what() );
+        exit( EXIT_FAILURE );
+    }
+}
+
+Matrix RadiationHydrodynamics1D::Source( const Matrix& uQ ) const {
[…]

### 2019-07-16 code/src/radiationhydrodynamics1d.h

@@ -1,0 +1,38 @@
+#ifndef RADIATIONHYDRODYNAMICS1D_H
+#define RADIATIONHYDRODYNAMICS1D_H
+
+#include "mathtools.h"
+#include "pnequations.h"
+#include <cmath>
+
+class RadiationHydrodynamics1D : public PNEquations
+{
+    double _c;    // radiation velocity
+    double _P;
+    double _R;
+    double _gamma;
+    unsigned _nMoments;
+    double Delta( int l, int k ) const;
+
+    void SetupSystemMatrices();
+    double Er0( const Vector& u ) const;
+    Vector Fr0( const Vector& u ) const;
+    double SE( const Vector& u ) const;
+    Vector SF( const Vector& u ) const;
+    Matrix FRadiation( const Vector& u ) const;
+    Matrix FEuler( const Vector& u ) const;
+    Matrix F( const Vector& u ) const;
+    virtual int GlobalIndex( int l, int k ) const;
+
+  public:
+    RadiationHydrodynamics1D( Settings* settings );
+    virtual double ComputeDt( const Matrix& u, double dx, unsigned level ) const;
+    inline Vector G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n );
+    virtual Matrix G( const Matrix& u, const Matrix& v, const Vector& nUnit, const Vector& n, unsigned level );
+    virtual Matrix Source( const Matrix& uQ ) const;
+    virtual Vector IC( const Vector& x, const Vector& xi );
+    virtual Matrix BoundaryFlux( const Matrix& u, const Vector& nUnit, const Vector& n, unsigned level ) const;
+};
+
+#endif    // RADIATIONHYDRODYNAMICS1D_H
+

### 2019-07-16 code/src/settings.cpp

@@ -29,6 +29,25 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
         _cwd      = std::filesystem::current_path();
         _inputDir = std::experimental::filesystem::path( _inputFile.parent_path() );
 
+        // section mesh
+        auto mesh          = file->get_table( "mesh" );
+        auto meshDimension = mesh->get_as<unsigned>( "dimension" );
+        if( meshDimension ) {
+            _meshDimension = *meshDimension;
+        }
+        else {
+            log->error( "[inputfile] [mesh] 'dimension' not set!" );
+            validConfig = false;
+        }
+        auto outputFile = mesh->get_as<std::string>( "outputFile" );
+        if( outputFile ) {
+            _outputFile = _inputDir.string() + "/" + _outputDir.string() + "/" + *outputFile;
+        }
+        else {
+            log->error( "[inputfile] [mesh] 'outputFile' not set!" );
+            validConfig = false;
+        }
+
         // section general
         auto general           = file->get_table( "general" );
         auto problemTypeString = general->get_as<std::string>( "problem" );
@@ -52,12 +71,15 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
             else if( problemTypeString->compare( "PNEquations2D" ) == 0 ) {
                 _problemType = ProblemType::P_PNEQUATIONS_2D;
             }
-            else if( problemTypeString->compare( "RadiationHydrodynamics2D" ) == 0 ) {
-                _problemType = ProblemType::P_RADIATIONHYDRO_2D;
+            else if( problemTypeString->compare( "RadiationHydrodynamics" ) == 0 ) {
+                if( _meshDimension == 1 )
+                    _problemType = ProblemType::P_RADIATIONHYDRO_1D;
+                else if( _meshDimension == 2 )
+                    _problemType = ProblemType::P_RADIATIONHYDRO_2D;
             }
             else {
                 log->error( "[inputfile] [general] 'problem' is invalid!\nPlease set one of the following types: Burgers1D, Euler1D, "
-                            "Euler2D,ShallowWater1D,ShallowWater2D" );
+                            "Euler2D,ShallowWater1D,ShallowWater2D,PNEquations2D,RadiationHydrodynamics" );
                 validConfig = false;
             }
         }
@@ -92,25 +114,6 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
             if( refFile ) {
[…]

### 2019-07-16 code/src/settings.h

@@ -10,7 +10,16 @@
 
 #include "typedefs.h"
 
-enum ProblemType { P_BURGERS_1D, P_EULER_1D, P_EULER_2D, P_SHALLOWWATER_1D, P_SHALLOWWATER_2D, P_PNEQUATIONS_2D, P_RADIATIONHYDRO_2D };
+enum ProblemType {
+    P_BURGERS_1D,
+    P_EULER_1D,
+    P_EULER_2D,
+    P_SHALLOWWATER_1D,
+    P_SHALLOWWATER_2D,
+    P_PNEQUATIONS_2D,
+    P_RADIATIONHYDRO_1D,
+    P_RADIATIONHYDRO_2D
+};
 enum ClosureType {
     C_BOUNDEDBARRIER,
     C_LOGSIN,

### 2019-07-17 Paper1/colloc_ipm.tex
some more corrections
paper continued

@@
 which is again needed for the construction of intrusive methods.
-By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:IPMmomentSystem}, i.e.
+By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
 Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
@@
 \label{sec:collIPM}
-Commonly, a great amount of iterations in pseudo-time are needed to converge to a steady state solution.
+Commonly, a great amount of iterations in pseudo-time is needed to converge to a steady state solution.
 Consequently, the IPM method which requires solving the dual problem \eqref{eq:dualProblem} in every spatial cell in each iteration becomes prohibitively expensive.

### 2019-07-17 Paper1/framework.tex
paper continued

@@
 In the following, we discretize the moment system in space and time according to \cite{kusch2017maximum}.
-Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:IPMmomentSystem}.
+Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:SGMomentSystem}.
 Omitting initial conditions and assuming a one-dimensional spatial domain, we can write this system  as
@@
 Hence, to determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
+\begin{align}\label{eq:L}
+L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle_Q - \sum_{i\leq M}\bm{\lambda}_i^T \bm{\hat u}_i
+\end{align}
+needs to be minimized. Hence, one needs to find the root of
 \begin{align*}
-L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle_Q - \sum_{i\leq M}\bm{\lambda}_i^T \bm{\hat u}_i
+\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}_i = \langle \bm u_s(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}_i,
 \end{align*}
-needs to be minimized, i.e. one needs to find the root of
+where we used $\nabla s_* \equiv \bm u_s$.
+The root is usually determined by using Newton's method.
+For simplicity, let us define the full gradient of the Lagrangian to be $\nabla_{\bm{\lambda}}L(\bm{\lambda};\bm{\hat{u}})\in\mathbb{R}^{N\cdot p}$, i.e.
+we store all entries in a vector.
+Newton's method uses the iteration function $\bm{d}:\mathbb{R}^{N\times p}\times\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$,
+\begin{align}\label{eq:dualIterationFunction}
+\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{H}(\bm{\lambda})^{-1}\cdot\nabla_{\bm{\lambda}}L(\bm{\lambda};\bm{\hat{u}}),
+\end{align}
+where $\bm H\in\mathbb{R}^{p\cdot N\times p\cdot N}$ is the Hessian of \eqref{eq:L}, given by
 \begin{align*}
-\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}_i,
+\bm{H}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi})\otimes\bm{\varphi}\bm{\varphi}^T\rangle_Q^{T}.
 \end{align*}
-which is usually done by an iterative method. Defining the dual iteration function $\bm{d}:\mathbb{R}^{N\times p}\times\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$,
-\begin{align}\label{eq:dualIterationFunction}
-\bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T-\bm{\hat{u}}\right),
-\end{align}
-with an adequate preconditioner $\bm{B}$, the iteration process for spatial cell $j$ is given by
+The function $\bm d$ will in the following be called dual iteration function. Now, the Newton iteration for spatial cell $j$ is given by
 \begin{align}\label{eq:dualIteration1}
@@
 $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
-A common preconditioner is the inverse Hessian of the dual problem, i.e.
-\begin{align*}
-\bm{B}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle_Q^{-T}.
-\end{align*}
 To obtain a finite number of iterations for the iteration in cell $j$, a stopping criterion 
@@
 \label{sec:costNumFlux}
-A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}.
+A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}.
+This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux}.
 By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.

### 2019-07-17 Paper1/introduction.tex
paper continued

@@
 The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
-The solution ansatz \eqref{eq:SGClosure} can then be used to compute the quantities of interest.
+One can use the solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
 Indeed, we have that
@@
 
-Numerical methods for approximating the moment $\bm{\hat u}_i$ can be divided into intrusive and non-intrusive techniques.
+Numerical methods for approximating the moments $\bm{\hat u}$ can be divided into intrusive and non-intrusive techniques.
 A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
@@
 Therefore, collocation methods typically require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation} to reach a given accuracy.
-Despite their easy implementation, collocation methods face challenges when examining unsteady problems.
-In this case, the solution needs to be cached at all quadrature points in every time step to compute the time evolution of the moments.
-This contradicts modern HPC paradigms, which aim at reducing the amount of data produced by numerical methods.
 
-Intrusive methods do not suffer from this problem, since the computation is directly carried out on the moments, i.e.
-their time evolution can be recorded during the computation.
-Also for steady problems, the fact that the moments are known in each iteration enables the computation of the stochastic residual, which indicates when to stop the iteration towards the steady state solution.
-However, intrusive methods are in general more difficult to implement and come along with higher numerical costs.
+Intrusive methods are in general more difficult to implement and come along with higher numerical costs.
 The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin (SG) \cite{ghanem2003stochastic} moment system
@@
 This system of $N$ equations can be solved by a deterministic method to determine the time evolution of the moments.
-A drawback of the stochastic-Galerkin method is that the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
-In order to remain hyperbolic, the solution needs to be manipulated \cite{schlachter2018hyperbolicity} in order to prevent a failure of the method.
+One significant drawback of the stochastic-Galerkin method is that the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
@@
 \end{align}
-Plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the IPM system
-\begin{subequations}\label{eq:IPMmomentSystem}
-\begin{align}
-\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
-\end{align}
-\end{subequations}
-with $|i|\leq M$.
+When plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero again yields the moment system \eqref{eq:SGMomentSystem}, but with the ansatz \eqref{eq:ansatz} instead of \eqref{eq:SGClosure}.
 The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.

### 2019-07-17 Paper1/one_shot_ipm.tex
paper continued

@@
 Assume that the classical IPM iteration is contractive at its fixed point $\bm{\hat u}^*$.
-Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$ if the preconditioner $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^{n})\bm{\varphi}\bm{\varphi}^T\rangle^{-1}$, i.e.
-the Hessian of the dual problem is used.
+Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$.
 \end{theorem}
@@
 \begin{align*}
-\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{B} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right) \\
-&=- \partial_{\bm{\lambda}}\bm{B} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right),
+\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\bm\lambda)^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\bm\lambda)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right) \\
+&=- \partial_{\bm{\lambda}}\bm{H}(\bm\lambda)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right).
 \end{align*}
-where we used $\bm{B} = \langle \nabla \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T}$.
 Recall that at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
@@
 \begin{align*}
-\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{B},
+\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\bm\lambda)^{-1},
 \end{align*}
@@
 \end{align*}
-since we already showed that by the choice of $\bm{B}$ the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
+since we already showed that by the choice of $\bm{H}(\bm\lambda)^{-1}$ the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
 We can show the same result for all spatial cells and all inputs of $\bm{c}$ analogously, hence $\partial_{\bm{\lambda}} \bm{c} = \bm{0}$.
@@
 \begin{align*}
-\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{B} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
+\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
 \end{align*}
-by the choice of $\bm{B}$ as well as \eqref{eq:cTildeDer} and \eqref{eq:dudlambdaex}.
+by the choice of $\bm{H}(\bm\lambda)^{-1}$ as well as \eqref{eq:cTildeDer} and \eqref{eq:dudlambdaex}.
 We obtain an analogous result for the second and third input.

### 2019-07-19 Paper1/framework.tex
reference SG flux added

@@
 \label{sec:costNumFlux}
+
 A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}.
@@
 By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
-However, the need to approximate the flux by an integral seems to be numerically expensive, especially when considering that the stochastic-Galerkin method often allows an analytic computation of all arising integrals.
+However, the need to approximate the flux by an integral seems to be numerically expensive, especially when considering that the stochastic-Galerkin method often allows an analytic computation of all arising integrals, see for example \cite{hu2015stochastic}.
 In the following, we discuss this issue and show the advantages of our numerical flux choice by considering the stochastic Burgers' equation
@@
 \end{align*}
-Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{M\times M}$ gives
+Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{N\times N}$ gives
 \begin{align*}

### 2019-07-22 Paper1/framework.tex
kinetic flux continued

@@
 By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
-However, the need to approximate the flux by an integral seems to be numerically expensive, especially when considering that the stochastic-Galerkin method often allows an analytic computation of all arising integrals, see for example \cite{hu2015stochastic}.
-In the following, we discuss this issue and show the advantages of our numerical flux choice by considering the stochastic Burgers' equation
+Therefore, we call the proposed numerical method \textit{semi-intrusive}.
+
+Moreover, by computing the numerical flux with a quadrature rule, we minimize numerical costs: When physical fluxes or source terms in the deterministic equation are polynomials, numerical fluxes for the SG moment system \eqref{eq:SGMomentSystem} can be computed analytically before running the program.
+See for example \cite{hu2015stochastic,tryoen2010instrusive} for the numerical flux computation or \cite{hu2016stochastic} for the computation of source terms.
+Even when terms in the deterministic problem are not polynomials, an approximation which again allows a precomputation can be constructed \cite{debusschere2004numerical}.
+However, analytically precomputing terms in the SG moment system is often less efficient than using techniques as stratified or latin hypercube sampling \cite[Section 6]{ghanem1998stochastic}.
+In the following, we briefly discuss the number of operations needed when precomputing integrals versus the use of a kinetic flux, i.e.
+the sampling is performed by a quadrature rule.
+
+We consider the stochastic Burgers' equation
 \begin{align*}

### 2019-07-22 Paper1/framework.tex
example more compact

@@
 
-\subsection{Costs of evaluating the numerical flux}
+\subsection{Properties of the kinetic flux}
 \label{sec:costNumFlux}
@@
 This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux}.
-By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
+By simply taking moments of a given deterministic flux, the method can easily be applied to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
 Therefore, we call the proposed numerical method \textit{semi-intrusive}.
@@
 Moreover, by computing the numerical flux with a quadrature rule, we minimize numerical costs: When physical fluxes or source terms in the deterministic equation are polynomials, numerical fluxes for the SG moment system \eqref{eq:SGMomentSystem} can be computed analytically before running the program.
+Even when terms in the deterministic problem are not polynomials, an approximation which again allows a precomputation can be constructed \cite{debusschere2004numerical}.
 See for example \cite{hu2015stochastic,tryoen2010instrusive} for the numerical flux computation or \cite{hu2016stochastic} for the computation of source terms.
-Even when terms in the deterministic problem are not polynomials, an approximation which again allows a precomputation can be constructed \cite{debusschere2004numerical}.
 However, analytically precomputing terms in the SG moment system is often less efficient than using techniques as stratified or latin hypercube sampling \cite[Section 6]{ghanem1998stochastic}.
-In the following, we briefly discuss the number of operations needed when precomputing integrals versus the use of a kinetic flux, i.e.
-the sampling is performed by a quadrature rule.
+In the following, we briefly discuss the number of operations needed when precomputing integrals versus the use of a kinetic flux for Burgers' equation.
 
-We consider the stochastic Burgers' equation
-\begin{align*}
-\partial_t u + \partial_x \frac{u^2}{2} &= 0,\\
-u(t=0,x,\xi) &= u_{IC}(x,\xi).
-\end{align*}
-The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
-Choosing the SG ansatz \eqref{eq:SGClosure} and testing with the gPC basis functions yields the SG moment system
-\begin{align*}
-\partial_t \hat u_i + \partial_x \frac12\sum_{n,m = 0}^M \hat u_n \hat u_m \langle \varphi_n\varphi_m\varphi_i \rangle = 0.
-\end{align*}
-Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{N\times N}$ gives
+With the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{N\times N}$ the Burgers' SG moment system reads
 \begin{align*}
@@
 \end{align*}
-which again requires $O(N\cdot Q)$ operations, i.e. the costs are $O(N^2)$. The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
+which again requires $O(N\cdot Q)$ operations, i.e.
+the costs are $O(N^2)$.
+The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
+When not using a quadratic entropy in the IPM method, integrals cannot be evaluated analytically and quadrature rules are not exact.
+We will study the approximation properties of IPM with different quadrature orders in Section~TODO

### 2019-07-22 Paper1/appendix.tex
numerical flux now in appendix

@@
 \appendix
+\section{Costs of evaluating the numerical flux}
+\label{app:costNumFlux}
+In the following, we briefly discuss the number of operations needed when precomputing integrals versus the use of a kinetic flux for Burgers' equation. The stochastic Burgers' equation reads
+\begin{align*}
+\partial_t u + \partial_x \frac{u^2}{2} &= 0,\\
+u(t=0,x,\xi) &= u_{IC}(x,\xi).
+\end{align*}
+The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
+Choosing the SG ansatz \eqref{eq:SGClosure} and testing with the gPC basis functions yields the SG moment system
+\begin{align*}
+\partial_t \hat u_i + \partial_x \frac12\sum_{n,m = 0}^M \hat u_n \hat u_m \langle \varphi_n\varphi_m\varphi_i \rangle = 0.
+\end{align*}
+Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{N\times N}$ gives
+\begin{align*}
+\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = 0
+\end{align*}
+with $F_i(\bm{\hat u}) = \frac12\bm{\hat u}^T\bm C_i\bm{\hat u}$. Note that $\bm{C}_i$ can be computed analytically, hence choosing a Lax-Friedrichs flux
+\begin{align}\label{eq:numFluxAnalytic}
+G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i
+\end{align}
+requires no integral evaluations. Recall, that the numerical flux choice made in this work gives
+\begin{align}\label{eq:numericalFluxIPMBurgers}
+ \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k),
+\end{align}
+where $\mathcal{U}$ is the SG ansatz \eqref{eq:SGClosure}.
+When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the sum is $3M=3(N-1)$.
+Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 N -1$ quadrature points suffice for an exact computation of the numerical flux.
+Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent.
+Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When computing and storing the values in a matrix $\bm A\in\mathbb{R}^{Q\times N}$ with entries $a_{ki} = \varphi_i(\xi_k)$ before running the program, the numerical flux \eqref{eq:numericalFluxIPMBurgers} can be split into two parts.
+First, we determine the SG solution at all quadrature points, i.e.
+we compute $\bm{u}^{(\ell)} := \bm A \bm{\hat u}_{\ell}$ and $\bm{u}^{(r)} := \bm A \bm{\hat u}_{r}$ which requires $O(N\cdot Q)$ operations.
+These solution values are then used to compute the numerical flux
+\begin{align*}
+G_i(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) &= \sum_{k=1}^Q w_k g(u^{(\ell)}_k,u^{(r)}_k)a_{ki}f_{\Xi}(\xi_k),
+\end{align*}
+which again requires $O(N\cdot Q)$ operations, i.e. the costs are $O(N^2)$. The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
+
 \section{IPM for the 2D Euler equations}

### 2019-07-22 Paper1/framework.tex
numerical flux now in appendix

@@
 See for example \cite{hu2015stochastic,tryoen2010instrusive} for the numerical flux computation or \cite{hu2016stochastic} for the computation of source terms.
-However, analytically precomputing terms in the SG moment system is often less efficient than using techniques as stratified or latin hypercube sampling \cite[Section 6]{ghanem1998stochastic}.
-In the following, we briefly discuss the number of operations needed when precomputing integrals versus the use of a kinetic flux for Burgers' equation.
-
-With the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{N\times N}$ the Burgers' SG moment system reads
-\begin{align*}
-\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = 0
-\end{align*}
-with $F_i(\bm{\hat u}) = \frac12\bm{\hat u}^T\bm C_i\bm{\hat u}$. Note that $\bm{C}_i$ can be computed analytically, hence choosing a Lax-Friedrichs flux
-\begin{align}\label{eq:numFluxAnalytic}
-G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i
-\end{align}
-requires no integral evaluations. Recall, that the numerical flux choice made in this work gives
-\begin{align}\label{eq:numericalFluxIPMBurgers}
- \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k),
-\end{align}
-where $\mathcal{U}$ is the SG ansatz \eqref{eq:SGClosure}.
-When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the sum is $3M=3(N-1)$.
-Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 N -1$ quadrature points suffice for an exact computation of the numerical flux.
-Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent.
-Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When computing and storing the values in a matrix $\bm A\in\mathbb{R}^{Q\times N}$ with entries $a_{ki} = \varphi_i(\xi_k)$ before running the program, the numerical flux \eqref{eq:numericalFluxIPMBurgers} can be split into two parts.
-First, we determine the SG solution at all quadrature points, i.e.
-we compute $\bm{u}^{(\ell)} := \bm A \bm{\hat u}_{\ell}$ and $\bm{u}^{(r)} := \bm A \bm{\hat u}_{r}$ which requires $O(N\cdot Q)$ operations.
-These solution values are then used to compute the numerical flux
-\begin{align*}
-G_i(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) &= \sum_{k=1}^Q w_k g(u^{(\ell)}_k,u^{(r)}_k)a_{ki}f_{\Xi}(\xi_k),
-\end{align*}
-which again requires $O(N\cdot Q)$ operations, i.e.
-the costs are $O(N^2)$.
-The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
+However, analytically precomputing terms in the SG moment system is often less efficient than using techniques as stratified or latin hypercube sampling \cite[Section 6]{ghanem1998stochastic}, see Appendix~\ref{app:costNumFlux}.
 When not using a quadratic entropy in the IPM method, integrals cannot be evaluated analytically and quadrature rules are not exact.

### 2019-07-25 Paper1/introduction.tex
results continued, intro rearranged a bit

@@
 Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
-A downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
-Therefore, collocation methods typically require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation} to reach a given accuracy.
 
@@
 In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments $N$ given by \eqref{eq:numberBasisFcts} is asymptotically smaller than the number of quadrature points, which increases with $O(M(\log_2(M)^{s-1}))$.
+An additional downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
+Consequently, collocation methods typically require a higher number of unknowns than intrusive methods to reach a given accuracy \cite{xiu2009fast,alekseev2011estimation}.
 Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex and high-dimensional settings.
@@
 We thereby provide the ability to recycle existing implementations of deterministic solvers.
+Our numerical studies show the following main results:
+\begin{itemize}
+\item The standard IPM method cannot compete with SC and SG. However when using the discussed acceleration techniques, IPM reaches a given accuracy with the shortest run time.
+\item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM leads to better variance approximations.
+\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods, even for one spatial dimension to reach the same accuracy level.
+\end{itemize}
 

### 2019-07-25 Paper1/results.tex
results continued, intro rearranged a bit

@@
 
-\subsection{2D Euler equations}
+\subsection{2D Euler equations with a one dimensional uncertainty}
 We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 profile computed with different methods.
@@
 we choose $\phi(\xi) = 1.25 + 0.5\xi$ where $\xi\sim U(-1,1)$.
+As commonly done, the initial condition is equal to the far field boundary values.
+Consequently, the wall condition at the airfoil is violated and will correct the flow solution.
+The computational domain is a circle with a diameter of $40$ meters.
+In the center, the NACA0012 airfoil with a length of one meter is located.
+The discretization is composed of a coarsely refined far field and a finely resolved region around the airfoil, since we are interested in the flow solution at the airfoil.
+Altogether, the mesh consists of 22361 triangular elements.
+
 The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ with different methods.
@@
 To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points.
+In the following, we investigate the L$^2$ error of the variance and the expectation value.
+Hence, when denoting the reference solution by $\bm u$, we have
+\begin{align*}
+\Vert E[\bm u] - E[\mathcal{U}] \Vert := \sqrt{\sum_{j=0}^{N_x} \Delta x_j \left( E[\bm u] - E[\mathcal{U}] \right)^2}, \enskip \Vert Var[\bm u] - Var[\mathcal{U}] \Vert := \sqrt{\sum_{j=0}^{N_x} \Delta x_j \left( Var[\bm u] - Var[\mathcal{U}] \right)^2}.
+\end{align*}
+The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent that small fluctuations in the coarsely refined far field influence the error.
 
@@
 For more information on the chosen entropy and the resulting solution ansatz for IPM, see \ref{app:IPM2DEuler}.
+
+Recall that the numerical flux \eqref{eq:numericalFluxIPM} uses a quadrature rule to approximate integrals.
+We start by investigating the effects of this quadrature has on the solution accuracy.
+For this, we run the IPM method with a moment order ranging from three to seven \comment{[use IPM instead of osIPM, add accuracy of SC$^3$]} using a Clenshaw-Curtis quadrature rule with level three (i.e.
+9 quadrature points) and level four (i.e.
+17 quadrature points).
+The results are given in Figure TODO.
+It can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
+This behavior results from aliasing effects, which dominate the accuracy level.
+Note that SC can be interpreted as an intrusive method which uses as many moments as quadrature points to compute the numerical flux.
+Therefore, we can conclude that the error of SC is dominated by aliasing effects and will not reach the same accuracy level as an intrusive method with an sufficiently accurate quadrature rule.
+Especially for high dimensional problems
 

### 2019-07-25 Paper1/results.tex
results continued

@@
 At the airfoil boundary $\Gamma_{0}$, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal.
-At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101,325$ Pa and a temperature of $273.15$ K.
+At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of $273.15$ K.
 Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees.
@@
 In the following, we investigate the L$^2$ error of the variance and the expectation value.
-Hence, when denoting the reference solution by $\bm u$, we have
+The L$^2$ of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
 \begin{align*}
-\Vert E[\bm u] - E[\mathcal{U}] \Vert := \sqrt{\sum_{j=0}^{N_x} \Delta x_j \left( E[\bm u] - E[\mathcal{U}] \right)^2}, \enskip \Vert Var[\bm u] - Var[\mathcal{U}] \Vert := \sqrt{\sum_{j=0}^{N_x} \Delta x_j \left( Var[\bm u] - Var[\mathcal{U}] \right)^2}.
+\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}
 \end{align*}
-The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent that small fluctuations in the coarsely refined far field influence the error.
+Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\hat{\bm u}_{\Delta}$, we investigate the relative error
+\begin{align*}
+\frac{\Vert E[\bm u_{\Delta}] - E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert}{\Vert E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert} \qquad \text{ and }\qquad \frac{\Vert Var[\bm u_{\Delta}] - Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert}{\Vert Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert}.
+\end{align*}
+The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely refined far field from affecting the error.
 
-In the following, we compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques.
-Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
+The quantities of interests are now computed with the different methods.
+All methods in this section have been computed using five MPI threads.
 For more information on the chosen entropy and the resulting solution ansatz for IPM, see \ref{app:IPM2DEuler}.
@@
 Recall that the numerical flux \eqref{eq:numericalFluxIPM} uses a quadrature rule to approximate integrals.
-We start by investigating the effects of this quadrature has on the solution accuracy.
+We start by investigating the effects this quadrature has on the solution accuracy.
 For this, we run the IPM method with a moment order ranging from three to seven \comment{[use IPM instead of osIPM, add accuracy of SC$^3$]} using a Clenshaw-Curtis quadrature rule with level three (i.e.
@@
 Note that SC can be interpreted as an intrusive method which uses as many moments as quadrature points to compute the numerical flux.
-Therefore, we can conclude that the error of SC is dominated by aliasing effects and will not reach the same accuracy level as an intrusive method with an sufficiently accurate quadrature rule.
-Especially for high dimensional problems
+Therefore, we can conclude that the error of SC is dominated by aliasing effects and will not reach the same accuracy level as an intrusive method with a sufficiently accurate quadrature rule.
+Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
+However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ evaluations, the numerical costs will depend on the chosen quadrature rule.
+
+Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques.
+Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
+The methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 10^{-5}$.
+Note that the adaptive SG method fails, since it yields negative densities during the iteration.
+The standard SG method however remains in the admissible set.
+The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:Obstacles2D}.
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rho]}.pdf}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rho]}.pdf}
+		\label{fig:sub2}
+	\end{subfigure}
+	\caption{Relative L$^2$ error with 5 MPI threads.}
+	\label{fig:Obstacles2D}
+\end{figure}
+
+Note that in the case of SC, the error cannot be recorded without destroying the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
+To account for the fact that sparse grids need to be used in higher dimensions, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
+Stochasti-Collocation uses quadrature levels 2,3 and 4, which corresponds to 5,9 and 17 quadrature points.
+All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
+3 to 10 moments).
+Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
+The remaining methods have been computed with a level 4 quadrature rule.
+
+First, it sticks out that again, the intrusive methods yield a smaller error level for the same number of unknowns compared to SC.
+For the same number of unknowns, the run time of the intrusive methods is always bigger than the corresponding SC method.
+However, the error of SG and IPM lies in the area of SC$^{17}$ at which the adaptive method has a smaller run time that SC.
+\comment{SG kann nicht schneller sein als SC17} While SG shows a slightly better approximation of the expectation value, IPM shows a smaller error of the variance.
+When comparing the different methods without acceleration techniques,
 

### 2019-07-25 Paper1/summary_outlook.tex
results continued

@@
 \label{sec:summary_outlook}
+
+In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations.
+Similar to the one-shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to steady state.
+Hessian approximations that can be interesting are BFGS and sparse BFGS, which construct the Hessian from previously computed gradients.
+Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
+Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, it can be important to intrusively modify the code in order to fully exploit all acceleration potentials.
+Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods.
+In this case one can switch to a fine quadrature level in a certain spatial cell by for example computing moments with the given coarse set of collocation points.
+From these moments one can compute an IPM reconstruction, which one can then evaluate at a finer quadrature set.
+Another example of breaking up the non-intrusive nature of Monte Carlo methods can be found in \cite{poette2019gpc}, where the generation of random samples is combined with the sampling after collisions to increase efficiency.

### 2019-07-25 Paper1/adaptivity.tex
paper continued
results continued, intro rearranged a bit
small fix

@@
 where divisions and multiplications are performed element-wise.
+Note that a similar indicator has been used in \cite{kroker2012finite} for intrusive methods in uncertainty quantification.
 In this work, we use the first entry in $\bm S_{\ell}$ to determine the refinement level, i.e.
@@
 \begin{align}\label{eq:dualIterationFunctionAd}
-\bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}_{\ell}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
+\bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{H}_{\ell}^{-1}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
 \end{align}
-where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $M_\ell$. The preconditioner $\bm{B}_{\ell}$ is given by 
+where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $M_\ell$. The Hessian $\bm{H}_{\ell}$ is given by 
 \begin{align*}
-\bm{B}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{-T}.
+\bm{H}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi}_{\ell})\otimes\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{T}.
 \end{align*}
@@
 Recall that the convergence to an admissible steady state solution is expensive and a high accuracy and desirable solution properties are only required at the end of this iteration process.
-Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualUnsteady} is fulfilled.
+Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualSteady} is fulfilled.
 For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of th required iterations on a lower but cheaper accuracy level.
@@
 \EndFor
-\While{\eqref{eq:residualUnsteady} is violated}
+\While{\eqref{eq:residualSteady} is violated}
 \For{$j=1$ to $N_x$}
@@
 \State $n \leftarrow n+1$
-\If{\eqref{eq:residualUnsteady} fulfills the stopping criterion $\varepsilon_m^*$}
+\If{\eqref{eq:residualSteady} fulfills the stopping criterion $\varepsilon_m^*$}
 \State $m \leftarrow m+1$

### 2019-07-25 Paper1/introduction.tex
minor fixes introduction

@@
 
-In this paper, we investigate intrusive methods for steady problems and compare them against collocation methods. For both steady and unsteady problems, we use adaptivity: 
+In this paper, we propose acceleration techniques for intrusive methods and compare them against stochastic-Collocation. For steady and unsteady problems, we use adaptivity: 
 \begin{itemize}
@@
 \end{itemize}
-The steady setting provides different opportunities to take advantage of features of intrusive methods: 
+A steady problem provides different opportunities to take advantage of features of intrusive methods: 
 \begin{itemize}
@@
 The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.
-Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods.
-The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used).
-We thereby provide the ability to recycle existing implementations of deterministic solvers.
 Our numerical studies show the following main results:
@@
 \end{itemize}
+Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods.
+The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used).
+We thereby provide the ability to recycle existing implementations of deterministic solvers.
 

### 2019-07-26 Paper1/framework.tex
introduction and framework nicer

@@
 This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux}.
-By simply taking moments of a given deterministic flux, the method can easily be applied to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
+By simply taking moments of a given numerical flux for the deterministic problem, the method can easily be applied to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.
 Therefore, we call the proposed numerical method \textit{semi-intrusive}.
 
-Moreover, by computing the numerical flux with a quadrature rule, we minimize numerical costs: When physical fluxes or source terms in the deterministic equation are polynomials, numerical fluxes for the SG moment system \eqref{eq:SGMomentSystem} can be computed analytically before running the program.
-Even when terms in the deterministic problem are not polynomials, an approximation which again allows a precomputation can be constructed \cite{debusschere2004numerical}.
-See for example \cite{hu2015stochastic,tryoen2010instrusive} for the numerical flux computation or \cite{hu2016stochastic} for the computation of source terms.
-However, analytically precomputing terms in the SG moment system is often less efficient than using techniques as stratified or latin hypercube sampling \cite[Section 6]{ghanem1998stochastic}, see Appendix~\ref{app:costNumFlux}.
-When not using a quadratic entropy in the IPM method, integrals cannot be evaluated analytically and quadrature rules are not exact.
+Intrusive numerical methods which compute arising integrals analytically and therefore directly depend on the moments (i.e.
+they do not necessitate the evaluation of the gPC expansion on quadrature points) can be constructed by performing a gPC expansion on the system flux directly \cite{debusschere2004numerical}.
+Examples can be found in \cite{hu2015stochastic,hu2016stochastic,tryoen2010instrusive}.
+While the analytic computation of arising integrals is not always more efficient \cite[Section 6]{ghanem1998stochastic}, it can also complicate recycling a deterministic solver.
+See \ref{app:costNumFlux} for a comparison of numerical costs when using Burgers' equation.
+However, when not using a quadratic entropy in the IPM method or when the physical flux of the deterministic problem is not a polynomial, it is not clear how many quadrature points the numerical quadrature rule requires to guarantee a sufficiently small quadrature error.
 We will study the approximation properties of IPM with different quadrature orders in Section~TODO

### 2019-07-26 Paper1/introduction.tex
introduction and framework nicer

@@
 \begin{itemize}
-\item The standard IPM method cannot compete with SC and SG. However when using the discussed acceleration techniques, IPM reaches a given accuracy with the shortest run time.
-\item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM leads to better variance approximations.
-\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods, even for one spatial dimension to reach the same accuracy level.
+\item In our test cases, the need to solve the dual problem \eqref{eq:dualProblem} when using the IPM method leads to a significantly higher run time than SC and SG.
+However when using the discussed acceleration techniques, IPM requires the shortest time to reach a given accuracy.
+\item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM shows improved variance approximations.
+\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for one spatial dimension) to reach the same accuracy level.
 \end{itemize}
-Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods.
-The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used).
-We thereby provide the ability to recycle existing implementations of deterministic solvers.
+The IPM and SG calculations use a semi-intrusive numerical method, since the discretization allows recycling a given deterministic code to generate the IPM solver.
+While facilitating the task of implementing general intrusive methods, this framework allows a fair comparison of intrusive and non-intrusive methods, since the intrusive method can be based on the same deterministic code as the code used in a black-box fashion for stochastic-Collocation.
 

### 2019-07-27 Paper1/results.tex
minor changes, new pics

@@
 
+\begin{figure}[h!]
+\centering
+		\centering
+		\includegraphics[scale=0.7]{figs/{ConvergenceQuadrature}.pdf}
+		\label{fig:sub1}
+	\caption{Relative L$^2$ error with different quadrature levels.}
+	\label{fig:Obstacles2D}
+\end{figure}
+
 Recall that the numerical flux \eqref{eq:numericalFluxIPM} uses a quadrature rule to approximate integrals.
 We start by investigating the effects this quadrature has on the solution accuracy.
-For this, we run the IPM method with a moment order ranging from three to seven \comment{[use IPM instead of osIPM, add accuracy of SC$^3$]} using a Clenshaw-Curtis quadrature rule with level three (i.e.
+For this, we run the IPM method with a moment order ranging from three to seven \comment{[use IPM instead of osIPM]} using a Clenshaw-Curtis quadrature rule with level three (i.e.
 9 quadrature points) and level four (i.e.
@@
 Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
-The methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 10^{-5}$.
+The methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-5}$.
 Note that the adaptive SG method fails, since it yields negative densities during the iteration.
 The standard SG method however remains in the admissible set.
-The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:Obstacles2D}.
+The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{Relative L$^2$ error with 5 MPI threads.}
-	\label{fig:Obstacles2D}
+
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rhoE]}.pdf}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rhoE]}.pdf}
+		\label{fig:sub2}
+	\end{subfigure}
+	\caption{Relative L$^2$ error with 5 MPI threads for density and energy.}
+	\label{fig:L2ErrorSolution}
 \end{figure}
@@
 To account for the fact that sparse grids need to be used in higher dimensions, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
-Stochasti-Collocation uses quadrature levels 2,3 and 4, which corresponds to 5,9 and 17 quadrature points.
+Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
 All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
@@
 For the same number of unknowns, the run time of the intrusive methods is always bigger than the corresponding SC method.
-However, the error of SG and IPM lies in the area of SC$^{17}$ at which the adaptive method has a smaller run time that SC.
-\comment{SG kann nicht schneller sein als SC17} While SG shows a slightly better approximation of the expectation value, IPM shows a smaller error of the variance.
+However, the error of SG and IPM lies in the area of SC$^{17}$ at which the adaptive method has a smaller run time than SC.
+While SG shows a slightly better approximation of the expectation value, IPM shows a smaller error of the variance.
 When comparing the different methods without acceleration techniques,
@@
 Furthermore, we use convergence accelerated (caIPM) as well as convergence accelerated one-shot IPM (caosIPM), which we introduced in Sections \ref{sec:collIPM} and \ref{sec:OneShotIPM} with $10$ moments and $15$ quadrature points to converge the collocation solution with $5$ quadrature points to an entropy solution with increased accuracy.
-
-\begin{figure}[h!]
-\centering
-		\centering
-		\includegraphics[scale=0.7]{figs/{L2_error_E[rho]osIPMIPMCol}.pdf}
-		\label{fig:sub1}
-	\caption{L2 error at airfoil with 5 MPI and 2 OMP threads.}
-	\label{fig:Obstacles2D}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-		\centering
-		\includegraphics[scale=0.7]{figs/{L2_error_E[rho]}.pdf}
-		\label{fig:sub1}
-	\caption{L2 error at airfoil with 5 MPI and 2 OMP threads.}
-	\label{fig:Obstacles2D}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-		\centering
-		\includegraphics[scale=0.7]{figs/{convergence_runtime_residual}.pdf}
-		\label{fig:sub1}
-	\caption{Convergence to steady state with 5 MPI and 2 OMP threads.}
-	\label{fig:Obstacles2D}
-\end{figure}
 

### 2019-08-05 Paper1/introduction.tex
started rewriting introduction and background

@@
 A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
-\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
-\begin{align*}
-\bm{\hat u}_i = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{Q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
-\end{align*} 
+\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule.
 Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
@@
 Intrusive methods are in general more difficult to implement and come along with higher numerical costs.
-The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin (SG) \cite{ghanem2003stochastic} moment system
-\begin{subequations}\label{eq:SGMomentSystem}
-\begin{align}
-\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
-\end{align}
-\end{subequations}
-with $|i|\leq M$.
+The main idea of these methods is to derive a system of equations for the moments $\bm{\hat{u}}_i$ with $|i|\leq M$.
 This system of $N$ equations can be solved by a deterministic method to determine the time evolution of the moments.
-One significant drawback of the stochastic-Galerkin method is that the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
+
+The stochastic-Galerkin (SG) \cite{ghanem2003stochastic} method uses the solution ansatz \eqref{eq:SGClosure} to represent the solution.
+Plugging this ansatz into \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the SG moment system.
+One significant drawback of this strategy is that the moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
-The main difference to SG is that the solution ansatz of IPM is given by an optimization problem instead of a polynomial expansion \eqref{eq:SGClosure}.
-For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is given by
-\begin{align}\label{eq:primalProblem}
-\mathcal{U}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } |i| \leq M.
-\end{align}
-Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstrained optimization problem.
-Defining the variables $\bm{\lambda}_i\in\mathbb{R}^p$ where $i$ is again a multi index, gives the unconstrained dual problem
-\begin{align}\label{eq:dualProblem}
- \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{N \times p}}
-  \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{|i|\leq M}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
-\end{align}
-where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_i)_{|i|\leq M}\in \mathbb{R}^{N \times p}$ are called the dual variables.
-The solution to \eqref{eq:primalProblem} is then given by
-\begin{align}\label{eq:ansatz}
- \mathcal{U}(\bm{\hat u}) = \left( \nabla_{\bm{u}} s \right)^{-1}(\bm{\hat{\lambda}}(\bm{\hat u})^T \bm{\varphi}).
-\end{align}
-When plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero again yields the moment system \eqref{eq:SGMomentSystem}, but with the ansatz \eqref{eq:ansatz} instead of \eqref{eq:SGClosure}.
-The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
-IPM generalizes different intrusive methods.
+The main difference to SG is that the solution ansatz of IPM is given by a constraint optimization problem instead of a polynomial expansion \eqref{eq:SGClosure}.
+This more general solution ansatz allows the generation of a hyperbolic moment system.
+Besides hyperbolicity, the IPM method has several advantages: First, it generalizes the SG method, i.e.
+the optimization problem can be choosen such that the resulting IPM system is again the SG moment system.
 Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
@@
 A comparison of results computed with the presented methods is then given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.
+
+\section{Background}
+Numerical methods for approximating the moments $\bm{\hat u}$ can be divided into intrusive and non-intrusive techniques.
+A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
+\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
+\begin{align*}
+\bm{\hat u}_i = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{Q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
+\end{align*} 
+Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
+Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly.
+Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
+Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
+
+Intrusive methods are in general more difficult to implement and come along with higher numerical costs.
+The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin (SG) \cite{ghanem2003stochastic} moment system
+\begin{subequations}\label{eq:SGMomentSystem}
+\begin{align}
+\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
+\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
+\end{align}
+\end{subequations}
+with $|i|\leq M$.
+This system of $N$ equations can be solved by a deterministic method to determine the time evolution of the moments.
+One significant drawback of the stochastic-Galerkin method is that the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
+A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
+The main difference to SG is that the solution ansatz of IPM is given by an optimization problem instead of a polynomial expansion \eqref{eq:SGClosure}.
+For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is given by
+\begin{align}\label{eq:primalProblem}
+\mathcal{U}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } |i| \leq M.
+\end{align}
+Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstrained optimization problem.
+Defining the variables $\bm{\lambda}_i\in\mathbb{R}^p$ where $i$ is again a multi index, gives the unconstrained dual problem
+\begin{align}\label{eq:dualProblem}
+ \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{N \times p}}
+  \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{|i|\leq M}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
+\end{align}
+where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_i)_{|i|\leq M}\in \mathbb{R}^{N \times p}$ are called the dual variables.
+The solution to \eqref{eq:primalProblem} is then given by
+\begin{align}\label{eq:ansatz}
+ \mathcal{U}(\bm{\hat u}) = \left( \nabla_{\bm{u}} s \right)^{-1}(\bm{\hat{\lambda}}(\bm{\hat u})^T \bm{\varphi}).
+\end{align}
+When plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero again yields the moment system \eqref{eq:SGMomentSystem}, but with the ansatz \eqref{eq:ansatz} instead of \eqref{eq:SGClosure}.
+The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
+IPM generalizes different intrusive methods.
+Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
+Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the expectation value of the entropy, which is
+\begin{align*}
+S(\bm{\hat u}) := \langle s( \mathcal{U}(\bm{\hat u}))\rangle,
+\end{align*}
+i.e.
+the IPM method yields a physically correct entropy solution.
+This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution values, which can then cause a failure of the method.
+The main weakness of the IPM method is its computational effort, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
+Hence, the desirable properties of IPM come along with significantly increased numerical costs.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architecture, which can be used to reduce the run time \cite{garrett2015optimization}.

### 2019-08-05 Paper1/introduction.tex
continued writing intro

@@
 
-A general hyperbolic set of equations with random initial data can be written as
-\begin{subequations}\label{eq:hyperbolicProblem}
-\begin{align}
-\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D, \\ \label{eq:ic}
-\bm{u}(t=0,\bm{x},&\bm{\xi}) = \bm{u}_{\text{IC}}(\bm{x},\bm{\xi}),
-\end{align}
-\end{subequations}
-where the solution $\bm u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
-Hence, the probability density function of $\bm{\xi}$ is then given by $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
-The physical flux is given by $\bm{f}:\mathbb{R}^p\to\mathbb{R}^{d\times p}$.
-To simplify notation, we assume that only the initial condition is random, i.e.
-$\bm{\xi}$ enters through the definition of $\bm{u}_{IC}$.
-Equations \eqref{eq:hyperbolicProblem} are usually supplemented with boundary conditions, which we will specify later for the individual problems.
-
-Due to the randomness of the solution, one is interested in determining the expectation value or the variance, i.e.
-\begin{align*}
-\text{E}[\bm{u}] = \langle \bm{u} \rangle,\qquad \text{Var}[\bm{u}] = \langle \left( \bm{u}-\text{E}[\bm{u}]\right)^2\rangle,
-\end{align*}
-where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_s$.
-To approximate quantities of interest (such as expectation value, variance or higher order moments), the solution is spanned with a set of polynomial basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
-Note that this yields
-\begin{align}\label{eq:numberBasisFcts}
-N:=\begin{pmatrix}
-M+s \\ s
-\end{pmatrix}
-\end{align}
-basis functions when defining $|i|:=\sum_{n = 1}^s |i_n|$.
-Commonly, these functions are chosen to be orthonormal polynomials \cite{wiener1938homogeneous} with respect to the probability function, i.e.
-$\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^s\delta_{i_nj_n}$.
-The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates the solution by
-\begin{align}\label{eq:SGClosure}
-\mathcal{U}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
-\end{align}
-where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^p$ are called moments.
-To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i\leq M|}\in\mathbb{R}^{N\times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
-In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
-The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
-One can use the solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
-Indeed, we have that
-\begin{align*}
-\text{E}[\mathcal{U}(\bm{\hat u})] = \bm{\hat u}_0,\quad \text{Var}[\mathcal{U}(\bm{\hat u})] = \text{E}[\mathcal{U}(\bm{\hat u})^2] - \text{E}[\mathcal{U}(\bm{\hat u})]^2 = \left(\sum_{i = 1}^N \hat{u}_{\ell i}^2\right)_{\ell = 1,\cdots,p}.
-\end{align*}
+The core idea of methods to quantify the effect of uncertainties is to represent the uncertainty by a polynomial. To determine
 
@@
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
-The main difference to SG is that the solution ansatz of IPM is given by a constraint optimization problem instead of a polynomial expansion \eqref{eq:SGClosure}.
+The main difference to SG is that not the solution but the entropy variables $\bm v = \nabla_{\bm u} s(\bm u)$ are represented by a polynomial.
+Here, the function $s:\mathbb{R}^{p}\rightarrow\mathbb{R}$ is called the entropy.
 This more general solution ansatz allows the generation of a hyperbolic moment system.
-Besides hyperbolicity, the IPM method has several advantages: First, it generalizes the SG method, i.e.
-the optimization problem can be choosen such that the resulting IPM system is again the SG moment system.
+Besides hyperbolicity the IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
+IPM generalizes different intrusive methods.
 Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
@@
 \section{Background}
-Numerical methods for approximating the moments $\bm{\hat u}$ can be divided into intrusive and non-intrusive techniques.
-A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
-\cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
+In the following, we briefly introduce the methods used in this work.
+
+A general hyperbolic set of equations with random initial data can be written as
+\begin{subequations}\label{eq:hyperbolicProblem}
+\begin{align}
+\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D, \\ \label{eq:ic}
+\bm{u}(t=0,\bm{x},&\bm{\xi}) = \bm{u}_{\text{IC}}(\bm{x},\bm{\xi}),
+\end{align}
+\end{subequations}
+where the solution $\bm u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
+Hence, the probability density function of $\bm{\xi}$ is then given by $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
+The physical flux is given by $\bm{f}:\mathbb{R}^p\to\mathbb{R}^{d\times p}$.
+To simplify notation, we assume that only the initial condition is random, i.e.
+$\bm{\xi}$ enters through the definition of $\bm{u}_{IC}$.
+Equations \eqref{eq:hyperbolicProblem} are usually supplemented with boundary conditions, which we will specify later for the individual problems.
+
+Due to the randomness of the solution, one is interested in determining the expectation value or the variance, i.e.
+\begin{align*}
+\text{E}[\bm{u}] = \langle \bm{u} \rangle,\qquad \text{Var}[\bm{u}] = \langle \left( \bm{u}-\text{E}[\bm{u}]\right)^2\rangle,
+\end{align*}
+where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_s$.
+To approximate quantities of interest (such as expectation value, variance or higher order moments), the solution is spanned with a set of polynomial basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
+Note that this yields
+\begin{align}\label{eq:numberBasisFcts}
+N:=\begin{pmatrix}
+M+s \\ s
+\end{pmatrix}
+\end{align}
+basis functions when defining $|i|:=\sum_{n = 1}^s |i_n|$.
+Commonly, these functions are chosen to be orthonormal polynomials \cite{wiener1938homogeneous} with respect to the probability function, i.e.
+$\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^s\delta_{i_nj_n}$.
+The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates the solution by
+\begin{align}\label{eq:SGClosure}
+\mathcal{U}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
+\end{align}
+where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^p$ are called moments.
+To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i\leq M|}\in\mathbb{R}^{N\times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
+In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
+The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
+One can use the solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
+Indeed, we have that
+\begin{align*}
+\text{E}[\mathcal{U}(\bm{\hat u})] = \bm{\hat u}_0,\quad \text{Var}[\mathcal{U}(\bm{\hat u})] = \text{E}[\mathcal{U}(\bm{\hat u})^2] - \text{E}[\mathcal{U}(\bm{\hat u})]^2 = \left(\sum_{i = 1}^N \hat{u}_{\ell i}^2\right)_{\ell = 1,\cdots,p}.
+\end{align*}
+
+The core idea of the stochastic-Collocation method is to compute the moments in the gPC expansion with a quadrature rule.
+Given a set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
 \begin{align*}
@@
 \end{align*}
+
 Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.

### 2019-08-05 Paper1/introduction.tex
introduction restructured, but needs some polishing

@@
 
-The core idea of methods to quantify the effect of uncertainties is to represent the uncertainty by a polynomial. To determine
-
-Numerical methods for approximating the moments $\bm{\hat u}$ can be divided into intrusive and non-intrusive techniques.
+Commonly, methods to quantify uncertainties use a polynomial chaos (PC) expansion \cite{wiener1938homogeneous,xiu2002wiener} to represent the solution, i.e.
+the uncertain space is spanned by a polynomial.
+The task is then to determine adequate expansion coefficients, which are often called moments.
+Numerical methods for approximating the moments can be divided into intrusive and non-intrusive techniques.
 A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
@@
 Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
-Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly.
+When the problem depends on $s$ uncertainties, tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of total degree $M$ exactly.
+Sparse grids on the other hand only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
@@
 Intrusive methods are in general more difficult to implement and come along with higher numerical costs.
-The main idea of these methods is to derive a system of equations for the moments $\bm{\hat{u}}_i$ with $|i|\leq M$.
-This system of $N$ equations can be solved by a deterministic method to determine the time evolution of the moments.
-
-The stochastic-Galerkin (SG) \cite{ghanem2003stochastic} method uses the solution ansatz \eqref{eq:SGClosure} to represent the solution.
-Plugging this ansatz into \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the SG moment system.
+The main idea of these methods is to derive a system of equations for the desired moments.
+This system can then be solved by a deterministic method to determine the time evolution of the moments.
+The stochastic-Galerkin (SG) method \cite{ghanem2003stochastic} uses a Galerkin projection to derive a moment system.
 One significant drawback of this strategy is that the moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
-The main difference to SG is that not the solution but the entropy variables $\bm v = \nabla_{\bm u} s(\bm u)$ are represented by a polynomial.
-Here, the function $s:\mathbb{R}^{p}\rightarrow\mathbb{R}$ is called the entropy.
+The main difference to SG is that not the solution but the entropy variables are represented by a polynomial.
 This more general solution ansatz allows the generation of a hyperbolic moment system.
-Besides hyperbolicity the IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
+Besides hyperbolicity, the IPM method has several advantages: Choosing a quadratic entropy yields the stochastic-Galerkin closure, i.e.
 IPM generalizes different intrusive methods.
 Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
-Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the expectation value of the entropy, which is
-\begin{align*}
-S(\bm{\hat u}) := \langle s( \mathcal{U}(\bm{\hat u}))\rangle,
-\end{align*}
-i.e.
+Also, as discussed in \cite{poette2009uncertainty}, when choosing a physically correct entropy of the deterministic problem, the IPM solution dissipates the expectation value of the entropy, i.e.
 the IPM method yields a physically correct entropy solution.
-This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution values, which can then cause a failure of the method.
-The main weakness of the IPM method is its computational effort, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
+The main weakness of the IPM method is its computational effort, since it requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
 Hence, the desirable properties of IPM come along with significantly increased numerical costs.
@@
 By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs.
-In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments $N$ given by \eqref{eq:numberBasisFcts} is asymptotically smaller than the number of quadrature points, which increases with $O(M(\log_2(M)^{s-1}))$.
+In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments needed to obtain a certain order is asymptotically smaller than the number of quadrature points.
 An additional downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
@@
 
-The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive method in section \ref{sec:framework}.
+The paper is structured as follows: After the introduction, we introduce the discussed methods in more detail in section~\ref{sec:background}.
+The numerical discretization as well as the implementation and structure of the semi-intrusive method is introduced in section \ref{sec:framework}.
 Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration.
@@
 \section{Background}
-In the following, we briefly introduce the methods used in this work.
-
-A general hyperbolic set of equations with random initial data can be written as
+\label{sec:background}
+In the following, we briefly introduce the notation and methods used in this work. A general hyperbolic set of equations with random initial data can be written as
 \begin{subequations}\label{eq:hyperbolicProblem}
@@
 \end{align*} 
-
-Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
-Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly.
-Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
-Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
-
-Intrusive methods are in general more difficult to implement and come along with higher numerical costs.
-The main idea of these methods is to derive a system of equations for the moments and then implementing a numerical solver for this system: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin (SG) \cite{ghanem2003stochastic} moment system
+Hence, the moments can be computed by running a given deterministic solver for the original problem at the quadrature points.
+Intrusive methods derive a system which directly describes the time evolution of the moments: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin moment system
 \begin{subequations}\label{eq:SGMomentSystem}
@@
 with $|i|\leq M$.
-This system of $N$ equations can be solved by a deterministic method to determine the time evolution of the moments.
-One significant drawback of the stochastic-Galerkin method is that the resulting moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
-A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
-The main difference to SG is that the solution ansatz of IPM is given by an optimization problem instead of a polynomial expansion \eqref{eq:SGClosure}.
-For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is given by
+To ensure hyperbolicity, the IPM method uses a solution ansatz which minimizes a given entropy under a moment constraint instead of a polynomial expansion \eqref{eq:SGClosure}.
+For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, the IPM solution ansatz is given by
 \begin{align}\label{eq:primalProblem}
@@
 When plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero again yields the moment system \eqref{eq:SGMomentSystem}, but with the ansatz \eqref{eq:ansatz} instead of \eqref{eq:SGClosure}.
-The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin closure, i.e.
-IPM generalizes different intrusive methods.
-Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}.
-Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the expectation value of the entropy, which is
-\begin{align*}
-S(\bm{\hat u}) := \langle s( \mathcal{U}(\bm{\hat u}))\rangle,
-\end{align*}
-i.e.
-the IPM method yields a physically correct entropy solution.
-This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution values, which can then cause a failure of the method.
-The main weakness of the IPM method is its computational effort, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves solving the optimization problem \eqref{eq:dualProblem}.
-Hence, the desirable properties of IPM come along with significantly increased numerical costs.
-However, IPM and minimal entropy methods in general are well suited for modern HPC architecture, which can be used to reduce the run time \cite{garrett2015optimization}.

### 2019-08-07 Paper1/introduction.tex
intro continued

@@
 Commonly, methods to quantify uncertainties use a polynomial chaos (PC) expansion \cite{wiener1938homogeneous,xiu2002wiener} to represent the solution, i.e.
-the uncertain space is spanned by a polynomial.
-The task is then to determine adequate expansion coefficients, which are often called moments.
+the uncertain space is spanned by polynomial basis functions.
+The remaining task is then to determine adequate expansion coefficients, often called moments or PC coefficients.
 Numerical methods for approximating the moments can be divided into intrusive and non-intrusive techniques.
@@
 \cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule.
-Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems.
-When the problem depends on $s$ uncertainties, tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of total degree $M$ exactly.
-Sparse grids on the other hand only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
+Commonly, SC uses sparse grids, since they posses a reduced number of collocation points for multi-dimensional problems.
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
@@
 
-Intrusive methods are in general more difficult to implement and come along with higher numerical costs.
-The main idea of these methods is to derive a system of equations for the desired moments.
-This system can then be solved by a deterministic method to determine the time evolution of the moments.
-The stochastic-Galerkin (SG) method \cite{ghanem2003stochastic} uses a Galerkin projection to derive a moment system.
-One significant drawback of this strategy is that the moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
+The main idea of intrusive methods is to derive a system of equations describing the time evolution of the moments which can then be solved with a deterministic numerical scheme.
+A popular approach to describe the moment system is the stochastic-Galerkin (SG) method \cite{ghanem2003stochastic}, which chooses a polynomial solution ansatz and performs a Galerkin projection to derive a closed system of equations.
+One significant drawback of SG is its not necessarily hyperbolic moment  \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
-The main difference to SG is that not the solution but the entropy variables are represented by a polynomial.
-This more general solution ansatz allows the generation of a hyperbolic moment system.
-Besides hyperbolicity, the IPM method has several advantages: Choosing a quadratic entropy yields the stochastic-Galerkin closure, i.e.
+Instead of performing the PC expansion on the solution, the IPM method represents the entropy variables with polynomials.
+Besides yielding a hyperbolic moment system, the IPM method has several advantages: Choosing a quadratic entropy yields the stochastic-Galerkin moment system, i.e.
 IPM generalizes different intrusive methods.
@@
 the IPM method yields a physically correct entropy solution.
-The main weakness of the IPM method is its computational effort, since it requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
+The main weakness of the IPM method is its computational cost, since it requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
 Hence, the desirable properties of IPM come along with significantly increased numerical costs.
@@
 By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs.
-In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments needed to obtain a certain order is asymptotically smaller than the number of quadrature points.
+In addition to that, the number of moments needed to obtain a certain order with intrusive methods is asymptotically smaller than the number of quadrature points for SC.
 An additional downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
@@
 \item The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
+To guarantee an efficient implementation, we propose an adaptive discretization strategy for IPM.
 \end{itemize}
@@
 \begin{itemize}
-\item When using adaptivity, one can perform a large number of iteration to the steady state solution on a low number of moments and increase the maximal refinement level when the distance to the steady state has reached a specified barrier.
-Consequently, a large number of iterations will be performed by a cheap method, i.e.
+\item When using adaptivity, one can perform a large number of iterations to the steady state solution on a low number of moments and increase the maximal truncation order when the distance to the steady state has reached a specified barrier.
+Consequently, a large number of iterations will be performed by a cheap, low order method, i.e.
 we can reduce numerical costs.
 \item Accelerate the convergence to the steady state IPM solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
-\item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e.
-they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}.
-Consequently, the dual variables and the moments are converges simultaneously to their steady state, which is similar to the idea of one-shot optimization in shape optimization \cite{hazra2005aerodynamic}.
+\item Perform an inexact map from the moments to the entropic expansion coefficients for IPM: Since the moments during the iteration process are inaccurate, i.e.
+they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes the IPM optimization problem.
+Consequently, the entropic expansion coefficients and the moments are converged simultaneously to their steady state, which is similar to the idea of one-shot optimization in shape optimization \cite{hazra2005aerodynamic}.
 \end{itemize}
@@
 
-The paper is structured as follows: After the introduction, we introduce the discussed methods in more detail in section~\ref{sec:background}.
+The paper is structured as follows: After the introduction, we present the discussed methods in more detail in section~\ref{sec:background}.
 The numerical discretization as well as the implementation and structure of the semi-intrusive method is introduced in section \ref{sec:framework}.
@@
 Section~\ref{sec:adaptivity} extends the presented numerical framework to an algorithm making use of adaptivity.
-A comparison of results computed with the presented methods is then given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.
+A comparison of results computed with the presented methods is then given in section \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.
 
@@
 where the solution $\bm u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
-Hence, the probability density function of $\bm{\xi}$ is then given by $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
+Hence, the probability density function of $\bm{\xi}$ is $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
 The physical flux is given by $\bm{f}:\mathbb{R}^p\to\mathbb{R}^{d\times p}$.
@@
 where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^p$ are called moments.
-To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i\leq M|}\in\mathbb{R}^{N\times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
+To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N\times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
 In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
-The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i := \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
+The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
 One can use the solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
@@
 \end{align*} 
-Hence, the moments can be computed by running a given deterministic solver for the original problem at the quadrature points.
+Hence, the moments can be computed by running a given deterministic solver for the original problem at each quadrature point.
+To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of total degree $M$ exactly, sparse grids only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
 Intrusive methods derive a system which directly describes the time evolution of the moments: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin moment system

### 2019-08-11 Paper1/introduction.tex
paper cont

@@
 The remaining task is then to determine adequate expansion coefficients, often called moments or PC coefficients.
-Numerical methods for approximating the moments can be divided into intrusive and non-intrusive techniques.
+Numerical methods for approximating these coefficients can be divided into intrusive and non-intrusive techniques.
 A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
@@
 A popular approach to describe the moment system is the stochastic-Galerkin (SG) method \cite{ghanem2003stochastic}, which chooses a polynomial solution ansatz and performs a Galerkin projection to derive a closed system of equations.
-One significant drawback of SG is its not necessarily hyperbolic moment  \cite{poette2009uncertainty}.
+One significant drawback of SG is its not necessarily hyperbolic moment system \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
@@
 the IPM method yields a physically correct entropy solution.
-The main weakness of the IPM method is its computational cost, since it requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
-Hence, the desirable properties of IPM come along with significantly increased numerical costs.
-However, IPM and minimal entropy methods in general are well suited for modern HPC architecture, which can be used to reduce the run time \cite{garrett2015optimization}.
+Unfortunately, the desirable properties of IPM come along with significantly increased numerical costs, since IPM requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architectures, which can be used to reduce the run time \cite{garrett2015optimization}.
 
@@
 \begin{itemize}
-\item In our test cases, the need to solve the dual problem \eqref{eq:dualProblem} when using the IPM method leads to a significantly higher run time than SC and SG.
+\item In our test cases, the need to solve an optimization problem when using the IPM method leads to a significantly higher run time than SC and SG.
 However when using the discussed acceleration techniques, IPM requires the shortest time to reach a given accuracy.
@@
 \item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for one spatial dimension) to reach the same accuracy level.
+\item Using sparse grids for the IPM discretization when the space of uncertainty is multi-dimensional, the number of quadrature points needed to guarantee sufficient regularity of the Hessian matrix is significantly increased.
+Therefore, techniques to regularize the IPM optimization problem are required.
 \end{itemize}

### 2019-08-11 Paper1/results.tex
paper cont

@@
 \centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.25]{figs/Euler1DPlots5/{sc100_ERho}.png}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.25]{figs/Euler1DPlots5/{sc100_VarRho}.png}
+		\label{fig:sub2}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\label{fig:L2ErrorSolution}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
 	\begin{subfigure}{0.33\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_sc_n5_s05_aoa}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_ERho}.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_ERho}.png}
 		\label{fig:sub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_ERho}.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_VarRho}.png}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_VarRho}.png}
 		\label{fig:sub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_VarRho}.png}
 		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/{euler2D_nacaCoarse_sc_n50_s05_aoa_legend}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\caption{E$[\rho]$ (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM, reference solution.}
+	\end{subfigure}
+	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
 	\label{fig:Obstacles2D}

### 2019-08-12 Paper1/results.tex
pics 10 added for paper

@@
 	\caption{Relative L$^2$ error with different quadrature levels.}
-	\label{fig:Obstacles2D}
+	\label{fig:ErrorDifferentQuad}
 \end{figure}
@@
 We start by investigating the effects this quadrature has on the solution accuracy.
-For this, we run the IPM method with a moment order ranging from three to seven \comment{[use IPM instead of osIPM]} using a Clenshaw-Curtis quadrature rule with level three (i.e.
+For this, we run the IPM method with a moment order ranging from three to seven using a Clenshaw-Curtis quadrature rule with level three (i.e.
 9 quadrature points) and level four (i.e.
 17 quadrature points).
-The results are given in Figure TODO.
+The results are given in Figure~\ref{fig:ErrorDifferentQuad}.
 It can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
@@
 	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
-	\label{fig:Obstacles2D}
+	\label{fig:ERhoVarRho5}
+\end{figure}
+
+\newgeometry{top=1.5cm, left=1.0cm}
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+				\includegraphics[scale=0.25]{figs/Euler1DPlots10/{EulerSC4-17Res1e-6_errors_ERho}.png}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+				\includegraphics[scale=0.25]{figs/Euler1DPlots10/{euler2D_nacaCoarse_sg_n9_sg4_s05_aoa_errors_ERho}.png}
+		\label{fig:sub2}
+	\end{subfigure}
+		\begin{subfigure}{0.33\linewidth}
+		\centering
+				\includegraphics[scale=0.25]{figs/Euler1DPlots10/{euler2D_nacaCoarse_ipm_n9_sg4_s05_aoa_errors_ERho}.png}
+		\label{fig:sub2}
+	\end{subfigure}
+	\caption{Errors E$[\rho]$ for SC$^{17}$, SG$_{10}$, IPM$_{10}$ (from left to right).}
+	\label{fig:ErrorsERho}
 \end{figure}

### 2019-08-12 code/src/momentsolver.cpp
mpi bug fixed

@@ -140,14 +140,8 @@ void MomentSolver::Solve() {
 
         // compute partial moment vectors on each PE (for inexact dual variables)
         for( unsigned j = 0; j < _nCells; ++j ) {
-            if( _settings->HasRegularization() ) {
-                // skip recomputation step if regularization is applied
-                u[j] = uNew[j];
-            }
-            else {
-                // recompute moments with inexact dual variables
-                u[j] = uQ[j] * _closure->GetPhiTildeWfAtRef( refinementLevel[j] );
-            }
+            // recompute moments with inexact dual variables
+            u[j] = uQ[j] * _closure->GetPhiTildeWfAtRef( refinementLevel[j] );
         }
 
         // determine time step size
@@ -170,6 +164,12 @@ void MomentSolver::Solve() {
         // perform reduction onto u to obtain full moments on PE PEforCell[j], which is the PE that solves the dual problem for cell j
         for( unsigned j = 0; j < _nCells; ++j ) {
             MPI_Reduce( uNew[j].GetPointer(), u[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, MPI_SUM, PEforCell[j], MPI_COMM_WORLD );
+        }
+
+        if( _settings->HasRegularization() ) {
+            for( unsigned j = 0; j < static_cast<unsigned>( _cellIndexPE.size() ); ++j ) {
+                u[_cellIndexPE[j]] = u[_cellIndexPE[j]] + _settings->GetRegularizationStrength() * _lambda[_cellIndexPE[j]];
+            }
         }
 
         // compute residual

### 2019-08-12 code/src/matrix.cpp
refinement with regularization started

@@ -20,6 +20,7 @@ template <class T> class Matrix    // column major
     ~Matrix();
     void operator=( const Matrix<T>& other );
     void resize( unsigned rows, unsigned columns );
+    Matrix<T> Add( const Matrix<T>& other, unsigned rows, unsigned columns ) const;
 
     T& operator()( unsigned i, unsigned j );
     const T& operator()( unsigned i, unsigned j ) const;
@@ -102,6 +103,17 @@ template <class T> void Matrix<T>::operator=( const Matrix<T>& other ) {
             ( *this )( i, j ) = other( i, j );
         }
     }
+}
+
+template <class T> Matrix<T> Matrix<T>::Add( const Matrix<T>& other, unsigned rows, unsigned columns ) const {
+    Matrix<T> res( _rows, _columns, true );
+
+    for( unsigned j = 0; j < columns; ++j ) {
+        for( unsigned i = 0; i < rows; ++i ) {
+            res( i, j ) = ( *this )( i, j ) + other( i, j );
+        }
+    }
+    return res;
 }
 
 template <class T> T& Matrix<T>::operator()( unsigned i, unsigned j ) { return _data[j * _rows + i]; }

### 2019-08-12 code/src/momentsolver.cpp

@@ -102,8 +102,11 @@ void MomentSolver::Solve() {
         // determine refinement level of cells on current PE
         for( unsigned j = 0; j < static_cast<unsigned>( _cellIndexPE.size() ); ++j ) {
             // if( _mesh->GetBoundaryType( _cellIndexPE[j] ) == BoundaryType::DIRICHLET && timeIndex > 0 ) continue;
-            // double indicator = ComputeRefIndicator( refinementLevel, u[_cellIndexPE[j]], refinementLevel[_cellIndexPE[j]] );
-            double indicator = 0.0;
+            double indicator;
+            if( _settings->GetProblemType() == P_RADIATIONHYDRO_1D )
+                indicator = 0.0;
+            else
+                indicator = ComputeRefIndicator( refinementLevel, u[_cellIndexPE[j]], refinementLevel[_cellIndexPE[j]] );
             if( indicator > _settings->GetRefinementThreshold() &&
                 refinementLevel[_cellIndexPE[j]] < _settings->GetNRefinementLevels( retCounter ) - 1 )
                 refinementLevel[_cellIndexPE[j]] += 1;
@@ -167,8 +170,10 @@ void MomentSolver::Solve() {
         }
 
         if( _settings->HasRegularization() ) {
+            // add eta*lambda to obtain old moments
             for( unsigned j = 0; j < static_cast<unsigned>( _cellIndexPE.size() ); ++j ) {
-                u[_cellIndexPE[j]] = u[_cellIndexPE[j]] + _settings->GetRegularizationStrength() * _lambda[_cellIndexPE[j]];
+                u[_cellIndexPE[j]] = u[_cellIndexPE[j]].Add(
+                    _settings->GetRegularizationStrength() * _lambda[_cellIndexPE[j]], _nStates, _nTotalForRef[refinementLevelOld[_cellIndexPE[j]]] );
             }
         }
 

### 2019-08-15 Paper1/results.tex
minor fix formula
smaller changes intro and results continued

@@
 \end{align*}
-These equations determine the time evolution of the conserved variables $(\rho,rho \bm v, rho e)$, i.e. density, momentum and energy. A closure for the pressure $p$ is given by
+These equations determine the time evolution of the conserved variables $(\rho,\rho \bm v, \rho e)$, i.e. density, momentum and energy. A closure for the pressure $p$ is given by
 \begin{align*}
@@
 
-Note that in the case of SC, the error cannot be recorded without destroying the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
-To account for the fact that sparse grids need to be used in higher dimensions, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
+Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
+To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
 Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
@@
 When comparing the different methods without acceleration techniques,
+
+\begin{center}
+\begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
+\hline
+residual \eqref{eq:residualSteady} & 5e-06 & 4e-06 & 3e-06 & 2e-06 & 1e-06 & 9e-07 & 8e-07 & 7e-07 & 6e-07 & 5e-07  \\ 
+\hline
+error E[$\rho$] in e-03 & 1.506 & 1.464 & 1.438 & 1.401 & 1.368 & 1.360 & 1.359 & 1.358 & 1.358 & 1.357  \\ 
+\hline
+error Var[$\rho$] in e-03 & 1.339 & 1.332 & 1.292 & 1.248 & 1.217 & 1.215 & 1.214 & 1.213 & 1.213 & 1.213  \\ 
+\hline
+\end{tabular}
+\end{center}
 
@@
 
-\newgeometry{top=1.5cm, left=1.0cm}
 \begin{figure}[h!]
@@
 		\centering
-				\includegraphics[scale=0.25]{figs/Euler1DPlots10/{EulerSC4-17Res1e-6_errors_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho}.png}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.25]{figs/Euler1DPlots10/{euler2D_nacaCoarse_sg_n9_sg4_s05_aoa_errors_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho}.png}
 		\label{fig:sub2}
-	\end{subfigure}
+	\end{subfigure}%
 	\begin{subfigure}{0.33\linewidth}
 		\centering
-				\includegraphics[scale=0.25]{figs/Euler1DPlots10/{euler2D_nacaCoarse_ipm_n9_sg4_s05_aoa_errors_ERho}.png}
-		\label{fig:sub2}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel}.png}
+		\label{fig:sub1}
 	\end{subfigure}
-	\caption{Errors E$[\rho]$ for SC$^{17}$, SG$_{10}$, IPM$_{10}$ (from left to right).}
-	\label{fig:ErrorsERho}
+	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{2-9}$}
+	\label{fig:ERhoVarRho5}
 \end{figure}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_sc_n5_s05_aoa_var}.png}
+				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{EulerSC4-17Res1e-6_errors_ERho}.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa_var}.png}
+				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_sg_n9_sg4_s05_aoa_errors_ERho}.png}
 		\label{fig:sub2}
-	\end{subfigure}%
+	\end{subfigure}
 		\begin{subfigure}{0.33\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa_var}.png}
-		\label{fig:sub1}
+				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_ipm_n9_sg4_s05_aoa_errors_ERho}.png}
+		\label{fig:sub2}
 	\end{subfigure}
-
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa_var}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa_var}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/{euler2D_nacaCoarse_sc_n50_s05_aoa_var_legend}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\caption{Var$[\rho]$ (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM, reference solution.}
-	\label{fig:Obstacles2D}
+	\caption{Errors E$[\rho]$ for SC$^{17}$, SG$_{10}$, IPM$_{10}$ (from left to right).}
+	\label{fig:ErrorsERho}
 \end{figure}
-
-%%% error plots %%%
-\begin{figure}[h!]
-\centering
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_sc_n5_s05_aoa_error_legend}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa_error_legend}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa_error_legend}.png}
-		\label{fig:sub1}
-	\end{subfigure}
-
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa_error_legend}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/expectedRho/errors/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa_error_legend}.png}
-		\label{fig:sub1}
-	\end{subfigure}
-	\caption{E$[\rho]$ distance to reference solution (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM.}
-	\label{fig:Obstacles2D}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_sc_n5_s05_aoa_var_error_legend}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_sg_n5_nq10_s05_aoa_var_error_legend}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_ipm_n5_nq10_s05_aoa_var_error_legend}.png}
-		\label{fig:sub1}
-	\end{subfigure}
-
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_caipm_n10_nq15_s05_aoa_var_error_legend}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/varRho/errors/{euler2D_nacaCoarse_caosipm_n10_nq15_s05_aoa_var_error_legend}.png}
-		\label{fig:sub1}
-	\end{subfigure}
-	\caption{Var$[\rho]$ distance to reference solution (from top left to bottom right) SC, SG, IPM, caIPM, caosIPM.}
-	\label{fig:Obstacles2D}
-\end{figure}
-\newgeometry{top=2.5cm, bottom=2.5cm}
+\newgeometry{top=2.5cm, left=2.5cm}

### 2019-08-15 code/src/euler2d.cpp
fixed warnings and removed and removed unsued variables

@@ -178,19 +178,22 @@ Vector Euler2D::IC( const Vector& x, const Vector& xi ) {
 
         double rhoFarfield = p / ( R * T );
 
-        y[0] = 1.0;
-        y[1] = 0.0;
-        y[2] = 0.0;
+        y[0] = rhoFarfield;
+        y[1] = rhoFarfield * uF;
+        y[2] = rhoFarfield * vF;
+        // double kineticEnergyL = 0.5 * rhoFarfield * ( pow( uF, 2 ) + pow( vF, 2 ) );
+        // double innerEnergyL   = ( p / ( rhoFarfield * ( gamma - 1 ) ) ) * rhoFarfield;
         y[3] = 1.0;
 
-        if( x[0] > 0.5 + _sigma[0] * xi[0] ) {
-            y[0] = 0.2;
+        if( x[1] < 1.1 + _sigma[0] * xi[0] ) {
+            y[0] = 0.5 * rhoFarfield;
             if( _settings->GetNDimXi() > 1 ) y[0] += _sigma[1] * xi[1];
-            y[1] = 0.0;
-            y[2] = 0.0;
-            y[3] = 0.2;
+            y[1] = rhoFarfield * uF;
+            y[2] = rhoFarfield * vF;
+            y[3] = 0.3;
             if( _settings->GetNDimXi() > 2 ) y[3] += _sigma[2] * xi[2];
         }
+
         /*
          if( x[1] < 1.1 + sigma && x[1] < 1.1 - sigma ) {
              y[0] = 0;

### 2019-08-15 code/src/momentsolver.cpp

@@ -63,8 +63,8 @@ void MomentSolver::Solve() {
     if( _settings->GetMyPE() == 0 ) log->info( "{:10}   {:10}", "t", "residual" );
 
     // init time and residual
-    double t      = _tStart;
-    int timeIndex = 0;
+    double t           = _tStart;
+    unsigned timeIndex = 0;
     double dt;
     double minResidual  = _settings->GetMinResidual();
     double residualFull = minResidual + 1.0;
@@ -457,6 +457,7 @@ Settings* MomentSolver::DeterminePreviousSettings() const {
         prevSettings->SetNStates( _settings->GetNStates() );
         prevSettings->SetGamma( _settings->GetGamma() );
         prevSettings->SetClosureType( _settings->GetClosureType() );
+        prevSettings->SetNumCells( _settings->GetNumCells() );
     }
     else {
         prevSettings = _settings;
@@ -468,6 +469,7 @@ Closure* MomentSolver::DeterminePreviousClosure( Settings* prevSettings ) const 
     Closure* prevClosure;
     if( prevSettings->GetNMoments() != _settings->GetNMoments() || prevSettings->GetNQTotal() != _settings->GetNQTotal() ) {
         prevClosure = Closure::Create( prevSettings );
+        std::cout << "foo" << std::endl;
     }
     else {
         prevClosure = _closure;
@@ -532,9 +534,9 @@ void MomentSolver::SetDuals( Settings* prevSettings, Closure* prevClosure, MatVe
         }
         _closure->SetMaxIterations( maxIterations );
         // delete reload closures and settings
-        delete prevSettings;
+        // delete prevSettings;
     }
-    if( prevSettings->GetNMoments() != _settings->GetNMoments() || prevSettings->GetNQTotal() != _settings->GetNQTotal() ) delete prevClosure;
+    // if( prevSettings->GetNMoments() != _settings->GetNMoments() || prevSettings->GetNQTotal() != _settings->GetNQTotal() ) delete prevClosure;
 }
 
 MatVec MomentSolver::SetupIC() const {
@@ -602,6 +604,9 @@ Settings* MomentSolver::ImportPrevSettings() const {
         if( line.find( "Config file" ) != std::string::npos ) {
             configSection = true;
             for( unsigned i = 0; i < 4; ++i ) std::getline( file, line );
+            // while( line.find( "==================================" ) != std::string::npos ) {
+            //    std::getline( file, line );
+            //}
         }
         else if( configSection && line.find( "==================================" ) != std::string::npos ) {
             break;

### 2019-08-15 code/src/radiationhydrodynamics1d.cpp

@@ -13,9 +13,9 @@ RadiationHydrodynamics1D::RadiationHydrodynamics1D( Settings* settings ) : PNEqu
     double rhoL = 5.45887 * 1e-13;
     double uL   = 2.3545 * 1e5;
     double TL   = 100.0;
-    double rhoR = 1.2479 * 1e-12;
-    double uR   = 1.03 * 1e5;
-    double TR   = 207.757;
+    // double rhoR = 1.2479 * 1e-12;
+    // double uR   = 1.03 * 1e5;
+    // double TR   = 207.757;
 
     // reference values
     double lRef = 1000;

### 2019-08-15 code/src/settings.cpp

@@ -101,7 +101,11 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
             _restartFile = _inputDir.string() + "/" + *restartFile;
             if( restart ) _loadLambda = general->get_as<bool>( "importDualState" ).value_or( false );
         }
-        if( !restart ) _loadLambda = general->get_as<bool>( "importDualState" ).value_or( false );
+        if( !restart )
+            _loadLambda = general->get_as<bool>( "importDualState" ).value_or( false );
+        else {
+            _loadLambda = false;
+        }
 
         auto icFile = general->get_as<std::string>( "icFile" );
         if( icFile ) {
@@ -110,7 +114,7 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
 
         if( !restart ) {
             auto refFile    = general->get_as<std::string>( "referenceSolution" );
-            _writeFrequency = general->get_as<int>( "writeFrequency" ).value_or( 1000 );
+            _writeFrequency = general->get_as<unsigned>( "writeFrequency" ).value_or( 1000 );
             if( refFile ) {
                 _referenceFile = _inputDir.string() + "/" + *refFile;
             }
@@ -286,7 +290,6 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
             if( quadratureType->at( 0 ).compare( "sparseGrid" ) == 0 ) {
                 _gridType = G_SPARSEGRID;
                 _nQTotal  = unsigned( std::pow( 2, _nQuadPoints ) ) + 1u;
-                _nQTotal  = 13;
             }
             else if( quadratureType->at( 0 ).compare( "tensorizedGrid" ) == 0 ) {    // tensorizedGrid
                 _gridType = G_TENSORIZEDGRID;
@@ -330,6 +333,10 @@ void Settings::Init( std::shared_ptr<cpptoml::table> file, bool restart ) {
             _refinementThreshold = ( *refinementThresholds )[0];
             _coarsenThreshold    = ( *refinementThresholds )[1];
         }
+        else {
+            _refinementThreshold = std::numeric_limits<double>::infinity();
+            _coarsenThreshold    = 0.0;
+        }
 
         _maxIterations = moment_system->get_as<unsigned>( "maxIterations" ).value_or( 1000 );
[…]

### 2019-08-15 code/src/settings.h

@@ -31,7 +31,6 @@ enum ClosureType {
     C_L2FILTER,
     C_LASSOFILTER
 };
-enum LimiterType { L_MINMOD, L_NONE };
 enum TimesteppingType { T_EXPLICITEULER };
 enum DistributionType { D_LEGENDRE, D_HERMITE };
 enum GridType { G_SPARSEGRID, G_TENSORIZEDGRID };
@@ -50,7 +49,7 @@ class Settings
     std::filesystem::path _referenceFile;
     bool _loadLambda;
 
-    int _writeFrequency;    // number of time steps until error to reference solution is computed
+    unsigned _writeFrequency;    // number of time steps until error to reference solution is computed
 
     // requied settings
     unsigned _meshDimension;
@@ -61,9 +60,7 @@ class Settings
     unsigned _nQTotal;               // number of quadrature points in all dimensions
     unsigned _nQPE;                  // number of total quadrature points on PE
     VectorU _nQPEAtRef;              // number of total quadrature points on PE for all refinement levels
-    unsigned _kStart;                // start point in quadrature point array for PE
     MatrixU _kIndicesAtRef;          // quadrature indices for PE at different refinement levels
-    unsigned _kEnd;                  // end point in quadrature point array for PE
     VectorU _quadLevel;              // quadrature level array
     VectorU _nQTotalForRef;          // number of quadrature points on different refinement levels
     Vector _residualRetardation;     // stores residual at different refinement levels
@@ -96,9 +93,7 @@ class Settings
     double _tEnd;
     double _minResidual;    // residual at which iteration is stopped for steady problems
     double _dt;             // timestepsize only required if no function ComputeDt provided in problem
-    unsigned _plotStepInterval;
-    double _plotTimeInterval;
-    LimiterType _limiterType;
+
     ClosureType _closureType;
     ProblemType _problemType;
     TimesteppingType _timesteppingType;
@@ -126,7 +121,7 @@ class Settings
     std::string GetInputFile() const;
     std::string GetInputDir() const;
     std::string GetOutputDir() const;
-    int GetWriteFrequency() const;
+    unsigned GetWriteFrequency() const;
 
     // mesh
     unsigned GetMeshDimension() const;
@@ -184,16 +179,10 @@ class Settings
     void SetNQuadPoints( unsigned nqNew );
     unsigned GetNQTotal() const;
     bool UsesMaxDegree() const;
[…]

### 2019-08-15 code/src/shallowwater2d.cpp

@@ -120,13 +120,13 @@ Matrix ShallowWater2D::BoundaryFlux( const Matrix& u, const Vector& nUnit, const
         Vector v( 2, 0.0 );
         v.reset();
 
-        v[0]           = u( 1, k ) / u( 0, k );
-        v[1]           = u( 2, k ) / u( 0, k );
-        double vn      = dot( nUnit, v );
-        Vector Vn      = vn * nUnit;
-        Vector Vb      = -Vn + v;
-        double velMagB = Vb[0] * Vb[0] + Vb[1] * Vb[1];
-        double velMag  = v[0] * v[0] + v[1] * v[1];
+        v[0]      = u( 1, k ) / u( 0, k );
+        v[1]      = u( 2, k ) / u( 0, k );
+        double vn = dot( nUnit, v );
+        Vector Vn = vn * nUnit;
+        Vector Vb = -Vn + v;
+        // double velMagB = Vb[0] * Vb[0] + Vb[1] * Vb[1];
+        // double velMag  = v[0] * v[0] + v[1] * v[1];
         double rho     = u( 0, k );
         uB[0]          = rho;
         uB[1]          = rho * ( Vb[0] );

### 2019-08-15 Paper1/framework.tex
some more smaller fixes

@@
 In the following, we discretize the moment system in space and time according to \cite{kusch2017maximum}.
-Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:SGMomentSystem}.
-Omitting initial conditions and assuming a one-dimensional spatial domain, we can write this system  as
+Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system.
+Hence, we discretize the system \eqref{eq:SGMomentSystem} with the more general IPM solution ansatz \eqref{eq:ansatz}.
+Omitting initial conditions and assuming a one-dimensional spatial domain, we can write the IPM system  as
 \begin{align*}
@@
 
-When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the original problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take
+When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the original problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take the numerical flux
 \begin{align*}
- \bm{\tilde G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T.
+ \bm{\tilde G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T
 \end{align*}
-Commonly, this integral cannot be evaluated analytically and therefore needs to be approximated by a quadrature rule
+in \eqref{eq:IPMDiscretization}. Commonly, this integral cannot be evaluated analytically and therefore needs to be approximated by a quadrature rule
 \begin{align*}
@@
 \end{align*}
-The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
+The computation of the dual variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
 Hence, to determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function

### 2019-08-15 Paper1/introduction.tex
smaller changes intro and results continued
some more smaller fixes

@@
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
-Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.
+Furthermore, SC is embarrassingly parallel, since the required computations can be carried out simultaneously on different cores.
 
@@
 \item Perform an inexact map from the moments to the entropic expansion coefficients for IPM: Since the moments during the iteration process are inaccurate, i.e.
-they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes the IPM optimization problem.
+they are not the correct steady state solution, we propose to not fully converge the dual iteration, which solves the IPM optimization problem.
 Consequently, the entropic expansion coefficients and the moments are converged simultaneously to their steady state, which is similar to the idea of one-shot optimization in shape optimization \cite{hazra2005aerodynamic}.
@@
 \end{itemize}
-The IPM and SG calculations use a semi-intrusive numerical method, since the discretization allows recycling a given deterministic code to generate the IPM solver.
+The IPM and SG calculations use a semi-intrusive numerical method, meaning that the discretization allows recycling a given deterministic code to generate the IPM solver.
 While facilitating the task of implementing general intrusive methods, this framework allows a fair comparison of intrusive and non-intrusive methods, since the intrusive method can be based on the same deterministic code as the code used in a black-box fashion for stochastic-Collocation.
@@
 To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of total degree $M$ exactly, sparse grids only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
+
 Intrusive methods derive a system which directly describes the time evolution of the moments: Plugging the solution ansatz \eqref{eq:SGClosure} into the set of equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero yields the stochastic-Galerkin moment system
@@
 with $|i|\leq M$.
+As already mentioned, the SG moment system is not necessarily hyperbolic.
 To ensure hyperbolicity, the IPM method uses a solution ansatz which minimizes a given entropy under a moment constraint instead of a polynomial expansion \eqref{eq:SGClosure}.

### 2019-08-16 Paper1/introduction.tex
reference mult Element added

@@
 These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}.
-Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{durrwachter2018hyperbolicity}.
+Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
 These structures commonly arise on a small portion of the space-time domain.

### 2019-08-17 Paper1/results.tex
new pics for paper
paper continued

@@
 \centering
-		\includegraphics[scale=0.7]{figs/{ConvergenceQuadrature}.pdf}
-		\label{fig:sub1}
-	\caption{Relative L$^2$ error with different quadrature levels.}
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.544]{figs/StudyQuadrature/{L2_error_E[rho]}.pdf}
+		\caption{}
+		\label{fig:ErrorDifferentQuadA}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.544]{figs/StudyQuadrature/{L2_error_Var[rho]}.pdf}
+		\caption{}
+		\label{fig:ErrorDifferentQuadB}
+	\end{subfigure}
+	\caption{Relative L$^2$ error with different quadrature levels. The subscript denotes the moment order, the superscript denotes the Clenshaw-Curtis quadrature level.}
 	\label{fig:ErrorDifferentQuad}
@@
 17 quadrature points).
-The results are given in Figure~\ref{fig:ErrorDifferentQuad}.
-It can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
-This behavior results from aliasing effects, which dominate the accuracy level.
-Note that SC can be interpreted as an intrusive method which uses as many moments as quadrature points to compute the numerical flux.
-Therefore, we can conclude that the error of SC is dominated by aliasing effects and will not reach the same accuracy level as an intrusive method with a sufficiently accurate quadrature rule.
+A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}.
+When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
+This behavior results from aliasing effects in the numerical flux, which dominate the accuracy level.
+Furthermore, Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, whereas the error of the variance in Figure~\ref{fig:ErrorDifferentQuadB} shows that IPM with 9 quadrature points yields an improved variance results over SC$^9$ even with only four moments.
+Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC.
+This result will show throughout the results of this work.
 Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
@@
 
-Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques.
+Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM at a fixed moment order 9 as well as its proposed acceleration techniques.
 Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
-The methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-5}$.
-Note that the adaptive SG method fails, since it yields negative densities during the iteration.
-The standard SG method however remains in the admissible set.
-The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
+All intrusive methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
+For every quadrature point, SC iterates the solution until the density reaches a certain residual level $\varepsilon$.
+We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following table for the SC$^{17}$ solution.
+\begin{center}
+\begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
+\hline
+residual $\rho$ & 5e-06 & 4e-06 & 3e-06 & 2e-06 & 1e-06 & 9e-07 & 8e-07 & 7e-07 & 6e-07 & 5e-07  \\ 
+\hline
+error E[$\rho$] in e-03 & 1.506 & 1.464 & 1.438 & 1.401 & 1.368 & 1.360 & 1.359 & 1.358 & 1.358 & 1.357  \\ 
+\hline
+error Var[$\rho$] in e-03 & 1.339 & 1.332 & 1.292 & 1.248 & 1.217 & 1.215 & 1.214 & 1.213 & 1.213 & 1.213  \\ 
+\hline
+\end{tabular}
+\end{center}
+Hence, we can deduce that SC requires a smaller
 \begin{figure}[h!]
@@
 
+First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration.
+The standard SG method however preserves positivity of mass, energy and pressure.
+The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
 Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
@@
 
-\begin{center}
-\begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
-\hline
-residual \eqref{eq:residualSteady} & 5e-06 & 4e-06 & 3e-06 & 2e-06 & 1e-06 & 9e-07 & 8e-07 & 7e-07 & 6e-07 & 5e-07  \\ 
-\hline
-error E[$\rho$] in e-03 & 1.506 & 1.464 & 1.438 & 1.401 & 1.368 & 1.360 & 1.359 & 1.358 & 1.358 & 1.357  \\ 
-\hline
-error Var[$\rho$] in e-03 & 1.339 & 1.332 & 1.292 & 1.248 & 1.217 & 1.215 & 1.214 & 1.213 & 1.213 & 1.213  \\ 
-\hline
-\end{tabular}
-\end{center}
-
  and compare against Collocation with $5$ collocation points as well as SG and IPM with $5$ moments and $10$ quadrature points.
@@
 		\centering
-				\includegraphics[scale=0.25]{figs/Euler1DPlots5/{sc100_ERho}.png}
+				\includegraphics[scale=0.15]{figs/Euler1DPlots5/{sc100_ERho}.png}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.25]{figs/Euler1DPlots5/{sc100_VarRho}.png}
+				\includegraphics[scale=0.15]{figs/Euler1DPlots5/{sc100_VarRho}.png}
 		\label{fig:sub2}
@@
 \end{figure}
-
 \begin{figure}[h!]
@@
 		\label{fig:sub2}
-	\end{subfigure}
+	\end{subfigure}%
 		\begin{subfigure}{0.33\linewidth}

### 2019-09-13 Paper1/results.tex
results continued

@@
 Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
-However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ evaluations, the numerical costs will depend on the chosen quadrature rule.
+However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
 
@@
 Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
+To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
+Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
+All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
+3 to 10 moments).
+Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
+The remaining methods have been computed with a level 4 quadrature rule.
 All intrusive methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
@@
 \end{center}
-Hence, we can deduce that SC requires a smaller
+Hence, we can deduce that SC requires a smaller residual to converge to a steady state solution.
+Therefore, the results for SC$^{17}$ will be presented for a residual of $\varepsilon = 5\cdot 10^{-6}$ as well as a reduced residual of $\varepsilon = 10^{-6}$.
+Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
+%One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
 \begin{figure}[h!]
@@
 The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
-Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
-To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
-Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
-All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
-3 to 10 moments).
-Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
-The remaining methods have been computed with a level 4 quadrature rule.
+When comparing intrusive methods without acceleration techniques as well as SC, the following properties stick out:
+\begin{itemize}
+\item SG comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
+\item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
+\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.
+Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
+\end{itemize}
+The proposed acceleration techniques show the following behavior:
+\begin{itemize}
+\item The one shot IPM (osIPM) method proposed in in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
+\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the one-shot idea, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM).
+This method reaches the steady state IPM solution with a faster runtime than SG.
+\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime.
+Here, we use two different strategies: ...
+\end{itemize}
 
-First, it sticks out that again, the intrusive methods yield a smaller error level for the same number of unknowns compared to SC.
-For the same number of unknowns, the run time of the intrusive methods is always bigger than the corresponding SC method.
-However, the error of SG and IPM lies in the area of SC$^{17}$ at which the adaptive method has a smaller run time than SC.
-While SG shows a slightly better approximation of the expectation value, IPM shows a smaller error of the variance.
-When comparing the different methods without acceleration techniques,
+Let us now discuss the different acceleration techniques used in this work.
+The IPM method combined with the one-shot idea discussed in Section~\ref{sec:OneShotIPM} is denoted by one-shot IPM (osIPM).
+When looking at the error in Figure~\ref{fig:L2ErrorSolution}, one can see that osIPM$_9$ reaches the same steady state error with a reduced runtime.
+When using the one-shot idea in combination with adaptivity, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM).
+\comment{TODO: eins der beiden adosIPM Ergebnisse rausnehmen}
 
@@
 \newgeometry{top=2.5cm, left=2.5cm}
+
+\subsection{2D Euler equations with a two dimensional uncertainty}
+
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_E[rho]}.pdf}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_Var[rho]}.pdf}
+		\label{fig:sub2}
+	\end{subfigure}
+
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_E[rhoE]}.pdf}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_Var[rhoE]}.pdf}
+		\label{fig:sub2}
+	\end{subfigure}
+	\caption{Relative L$^2$ error with 17 MPI threads for density and energy.}
+	\label{fig:L2ErrorSolution}
+\end{figure}
+
+It sticks out that again, the intrusive methods (SG$_9$,IPM$_9$) yield a smaller error level for the same number of unknowns compared to SC.
+For the same number of unknowns, the run time of the intrusive methods is always bigger than the corresponding SC method.
+However, the error of SG and IPM lies in the area of SC$^{17}$ at which the adaptive method has a smaller run time than SC.
+While SG shows a slightly better approximation of the expectation value, IPM shows a smaller error of the variance.

### 2019-09-19 Paper1/results.tex
changes black forest

@@
 \subsection{2D Euler equations with a one dimensional uncertainty}
-We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 profile computed with different methods.
+\label{sec:resultsNACA1D}
+We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 airfoil computed with different methods.
 The stochastic Euler equations in two dimensions are given by
@@
 \end{align*}
-Since the the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$.
+Since the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$.
 The spatial mesh discretizes the flow domain around the airfoil.
@@
 Consequently, the wall condition at the airfoil is violated and will correct the flow solution.
+
 The computational domain is a circle with a diameter of $40$ meters.
@@
 
-The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ with different methods.
-The IPM methods makes use of the acceleration strategies proposed in this work.
+The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods.
 To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points.
 In the following, we investigate the L$^2$ error of the variance and the expectation value.
-The L$^2$ of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
+The L$^2$ error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
 \begin{align*}
-\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}
+\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}.
 \end{align*}
@@
 \begin{align*}
-\frac{\Vert E[\bm u_{\Delta}] - E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert}{\Vert E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert} \qquad \text{ and }\qquad \frac{\Vert Var[\bm u_{\Delta}] - Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert}{\Vert Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert}.
+\frac{\Vert E[\bm u_{\Delta}] - E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert Var[\bm u_{\Delta}] - Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}.
 \end{align*}
@@
 A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}.
-When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
+\begin{itemize}
+\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
 This behavior results from aliasing effects in the numerical flux, which dominate the accuracy level.
-Furthermore, Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, whereas the error of the variance in Figure~\ref{fig:ErrorDifferentQuadB} shows that IPM with 9 quadrature points yields an improved variance results over SC$^9$ even with only four moments.
+\item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy.
+Note that this holds for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$.
+Hence, the variance is more sensitive to aliasing errors.
+\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, whereas the error of the variance in Figure~\ref{fig:ErrorDifferentQuadB} shows that IPM with 9 quadrature points yields an improved variance results over SC$^9$ even with only four moments.
 Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC.
-This result will show throughout the results of this work.
+This result can be observed throughout the results of this work.
 Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
 However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
-
-Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM at a fixed moment order 9 as well as its proposed acceleration techniques.
-Note that since IPM generalizes SG, all proposed techniques can be used for this method as well.
+\end{itemize}
+Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order 9.
+Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well.
 To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
@@
 The remaining methods have been computed with a level 4 quadrature rule.
+
 All intrusive methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.

### 2019-09-20 Paper1/results.tex
results updated, section naca continued

@@
 
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sc100_ERho}.png}
+		\label{fig:referenceSolutionsub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sc100_VarRho}.png}
+		\label{fig:referenceSolutionsub2}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\label{fig:referenceSolution}
+\end{figure}
+
 The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods.
-To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points.
+To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
 In the following, we investigate the L$^2$ error of the variance and the expectation value.
@@
 \item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime.
-Here, we use two different strategies: ...
+Here, we use two different strategies: First, we steadily increase the maximal truncation order when the residual approaches zero.
+To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order.
+The residual values are then determined to be $0.00006,0.00003,0.000022$ and $0.00002$.
+The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$.
+Strategy 1 is depicted in red, Strategy 2 is depicted in green.
+It can be seen that both approaches reach the IPM$_9$ error for the same run time.
+Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
 \end{itemize}
-
-Let us now discuss the different acceleration techniques used in this work.
-The IPM method combined with the one-shot idea discussed in Section~\ref{sec:OneShotIPM} is denoted by one-shot IPM (osIPM).
-When looking at the error in Figure~\ref{fig:L2ErrorSolution}, one can see that osIPM$_9$ reaches the same steady state error with a reduced runtime.
-When using the one-shot idea in combination with adaptivity, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM).
-\comment{TODO: eins der beiden adosIPM Ergebnisse rausnehmen}
-
- and compare against Collocation with $5$ collocation points as well as SG and IPM with $5$ moments and $10$ quadrature points.
-Furthermore, we use convergence accelerated (caIPM) as well as convergence accelerated one-shot IPM (caosIPM), which we introduced in Sections \ref{sec:collIPM} and \ref{sec:OneShotIPM} with $10$ moments and $15$ quadrature points to converge the collocation solution with $5$ quadrature points to an entropy solution with increased accuracy.
-
-\newgeometry{top=1.5cm, left=1.0cm}
-\begin{figure}[h!]
-\centering
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-				\includegraphics[scale=0.15]{figs/Euler1DPlots5/{sc100_ERho}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-				\includegraphics[scale=0.15]{figs/Euler1DPlots5/{sc100_VarRho}.png}
-		\label{fig:sub2}
-	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
-	\label{fig:L2ErrorSolution}
-\end{figure}
 \begin{figure}[h!]
@@
 \end{figure}
-
+Let us finally take a look at the expectation value and variance computed with different methods.
+All results are depicted for a zoomed view around the airfoil.
+Figure~\ref{fig:ERhoVarRho5} shows the expectation value (first row) and variance (second row) computed with $5$ quadrature points for SC and $5$ moments for SG and IPM.
+One can observe the following
+\begin{itemize}
+\item All methods yield non-physical step-like profiles of the expectation value and variance along the airfoil. 
+\item The jump position of the intrusive solution profiles capture the exact behavior more accurately.
+\end{itemize}
+The readosIPM results as well as the corresponding refinement levels are depicted in Figure~\ref{fig:readosIPMEVar}.
 \begin{figure}[h!]
@@
 	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{2-9}$}
-	\label{fig:ERhoVarRho5}
+	\label{fig:readosIPMEVar}
 \end{figure}
@@
 \end{figure}
-\newgeometry{top=2.5cm, left=2.5cm}
+%\newgeometry{top=2.5cm, left=2.5cm}
 

### 2019-09-20 Paper1/results.tex
smaller changes results

@@
 A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}.
+To denote the quadrature level, we use a superscript and the moment order is denoted by a subscript.
 \begin{itemize}
@@
 Hence, the variance is more sensitive to aliasing errors.
-\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, whereas the error of the variance in Figure~\ref{fig:ErrorDifferentQuadB} shows that IPM with 9 quadrature points yields an improved variance results over SC$^9$ even with only four moments.
+\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, i.e.
+the aliasing error heavily affects the accuracy.
+This behavior becomes more dominant when looking at the variance error in Figure~\ref{fig:ErrorDifferentQuadB}.
+Here, IPM$_8^4$ yields a significantly improved result compared to IPM$_8^3$.
+Again, the IPM$_8^3$ result stagnates at the SC$^9$ accuracy level, which can be reached with IPM when only using four moments (combined with a sufficiently accurate quadrature level).
 Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC.
@@
 Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
+When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
 The remaining methods have been computed with a level 4 quadrature rule.
@@
 \begin{itemize}
-\item SG comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
+\item Compared to IPM, stochastic-Galerkin comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
 \item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
@@
 The readosIPM results as well as the corresponding refinement levels are depicted in Figure~\ref{fig:readosIPMEVar}.
+One observes that the solution no longer shows the previously observed discontinuous profile and yield a satisfactory agreement with the reference solution depicted in Figure~\ref{fig:referenceSolution}.
+The refinement level shows that away from the airfoil, a refinement level of 0 (i.e.
+a truncation order of 2) suffices to yield the IPM$_9$ solution.
+The region with a high variance requires a refinement level of 7 (truncation order 9).
 \begin{figure}[h!]

### 2019-09-25 Paper1/colloc_ipm.tex
results continued
no more caIPM, results 2d started

@@
-In the following, we look at steady state problems, i.e. we have
-\begin{align}\label{eq:hyperbolicProblemSteady}
-\nabla\cdot\bm{f}(\bm{u}(\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D
-\end{align}
-with adequate boundary conditions.
-A general strategy for computing the steady state solution to \eqref{eq:hyperbolicProblemSteady} is to introduce a pseudo-time and numerically treat \eqref{eq:hyperbolicProblemSteady} as an unsteady problem.
-A steady state solution is then obtained by iterating in pseudo-time until the solution remains constant.
-It is important to point out that the time it takes to converge to a steady state solution is crucially affected by the chosen initial condition and its distance to the steady state solution.
-Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system for \eqref{eq:hyperbolicProblemSteady} given by
-\begin{align}\label{eq:MomentSystemSteady}
-\nabla\cdot\langle\bm{f}(\bm{u}(\bm{x},\bm{\xi}))\bm{\varphi}^T\rangle^T = \bm{0} \enskip \text{ in } D
-\end{align}
-which is again needed for the construction of intrusive methods.
-By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
-Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
-Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the condition
-\begin{align}\label{eq:residualUnsteady}
-\sum_{j = 1}^{N_x} \Delta x_j \Vert \bm{\hat{u}}_j^n - \bm{\hat{u}}_j^{n-1} \Vert \leq \varepsilon
-\end{align}
-is fulfilled. Since one is generally interested in low order moments such as the expectation value, this residual can be modified by only accounting for the zero order moments.
-
-\subsection{Collocation accelerated IPM}
-\label{sec:collIPM}
-Commonly, a great amount of iterations in pseudo-time is needed to converge to a steady state solution.
+%\subsection{Collocation accelerated IPM}
+%\label{sec:collIPM}
+%Commonly, a great amount of iterations in pseudo-time is needed to converge to a steady state solution.
 Consequently, the IPM method which requires solving the dual problem \eqref{eq:dualProblem} in every spatial cell in each iteration becomes prohibitively expensive.
@@
 
-Different variants of this method are possible:
-\begin{itemize}
-\item Since the IPM iterations will again modify the steady state Collocation solution, it is not necessary to converge Collocation to the exact steady state solution before starting IPM.
+%Different variants of this method are possible:
+%\begin{itemize}
+%\item Since the IPM iterations will again modify the steady state Collocation solution, it is not necessary to converge Collocation to the exact steady state solution before starting IPM.
 Here, one needs to determine an indicator to choose at which residual the collocation iteration is sufficiently accurate and can therefore be switched to IPM.
-\item The main idea is to use a cheap but inexact method as a preconditioner for an expensive but accurate method.
+%\item The main idea is to use a cheap but inexact method as a preconditioner for an expensive but accurate method.
 Here, one is not limited in choosing SC and IPM, but one can for example choose Monte Carlo methods as an accelerator or SC with an increased number of quadrature points as the expensive method.
 Note that in the latter case, a map from the moments to the solution at the increased quadrature set is required, for which the IPM reconstruction \eqref{eq:ansatz} would be a suitable choice.
-\end{itemize}
+%\end{itemize}

### 2019-09-25 Paper1/introduction.tex
changes black forest
no more caIPM, results 2d started

@@
 Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
-These structures commonly arise on a small portion of the space-time domain.
+Discontinuous structures commonly arise on a small portion of the space-time domain.
 Therefore, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies.
@@
 we can reduce numerical costs.
-\item Accelerate the convergence to the steady state IPM solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state.
 \item Perform an inexact map from the moments to the entropic expansion coefficients for IPM: Since the moments during the iteration process are inaccurate, i.e.
@@
 \item Using sparse grids for the IPM discretization when the space of uncertainty is multi-dimensional, the number of quadrature points needed to guarantee sufficient regularity of the Hessian matrix is significantly increased.
-Therefore, techniques to regularize the IPM optimization problem are required.
 \end{itemize}
@@
 \partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
-\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
+\bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
 \end{align}

### 2019-09-25 Paper1/main.tex
implementation section added
results continued, intro rearranged a bit
introduction restructured, but needs some polishing
smaller changes intro and results continued
no more caIPM, results 2d started

@@
 \usepackage{epstopdf}
+\usepackage{array}
 
@@
 % Declare title and authors, without \thanks
-\newcommand{\TheTitle}{Acceleration strategies for intrusive methods in Uncertainty Quantification}
+\newcommand{\TheTitle}{Intrusive acceleration strategies for Uncertainty Quantification}
 
@@
 
-\newcommand{\commentRyan}[1]{{\Large{\textcolor{red}{\newline[RYAN: #1]\newline}}}}
-\newcommand{\commentJonas}[1]{{\Large{\textcolor{blue}{\newline[JONAS: #1]\newline}}}}
+\newcommand{\commentJannick}[1]{{\textcolor{red}{#1}}}
+\newcommand{\comment}[1]{{\textcolor{blue}{#1}}}
 
@@
 Integrating the iteration to solve the optimization problem into the moment update guarantees local convergence while reducing numerical costs.
-Furthermore, we propose to use non-intrusive methods as a preconditioner for intrusive methods.
-Consequently, a large amount of iterations to the steady solution are performed by a cheap method, while maintaining the increased accuracy of an intrusive method.
 Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
@@
 \input{framework}
-\section{Strategies for steady problems}
-\input{colloc_ipm}
+%\section{Strategies for steady problems}
+%\input{colloc_ipm}
 \input{one_shot_ipm}
 \input{adaptivity}
+\input{implementation}
 \input{results}
 \input{summary_outlook}
-\input{appendix}
 
@@
 \bibliography{references}
+\input{appendix}
 \end{document}

### 2019-09-25 Paper1/one_shot_ipm.tex
some more smaller fixes
no more caIPM, results 2d started

@@
-\subsection{One-shot IPM}
+\section{One-shot IPM}
 \label{sec:OneShotIPM}
+
+In the following, we look at steady state problems, i.e. we have
+\begin{align}\label{eq:hyperbolicProblemSteady}
+\nabla\cdot\bm{f}(\bm{u}(\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D
+\end{align}
+with adequate boundary conditions.
+A general strategy for computing the steady state solution to \eqref{eq:hyperbolicProblemSteady} is to introduce a pseudo-time and numerically treat \eqref{eq:hyperbolicProblemSteady} as an unsteady problem.
+A steady state solution is then obtained by iterating in pseudo-time until the solution remains constant.
+It is important to point out that the time it takes to converge to a steady state solution is crucially affected by the chosen initial condition and its distance to the steady state solution.
+Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system for \eqref{eq:hyperbolicProblemSteady} given by
+\begin{align}\label{eq:MomentSystemSteady}
+\nabla\cdot\langle\bm{f}(\bm{u}(\bm{x},\bm{\xi}))\bm{\varphi}^T\rangle^T = \bm{0} \enskip \text{ in } D
+\end{align}
+which is again needed for the construction of intrusive methods.
+By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
+Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
+Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the condition
+\begin{align}\label{eq:residualSteady}
+\sum_{j = 1}^{N_x} \Delta x_j \Vert \bm{\hat{u}}_j^n - \bm{\hat{u}}_j^{n-1} \Vert \leq \varepsilon
+\end{align}
+is fulfilled. Since one is generally interested in low order moments such as the expectation value, this residual can be modified by only accounting for the zero order moments.
 

### 2019-09-25 Paper1/results.tex
no more caIPM, results 2d started

@@
 
-\begin{figure}[h!]
-\centering
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{EulerSC4-17Res1e-6_errors_ERho}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_sg_n9_sg4_s05_aoa_errors_ERho}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-		\begin{subfigure}{0.33\linewidth}
-		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_ipm_n9_sg4_s05_aoa_errors_ERho}.png}
-		\label{fig:sub2}
-	\end{subfigure}
-	\caption{Errors E$[\rho]$ for SC$^{17}$, SG$_{10}$, IPM$_{10}$ (from left to right).}
-	\label{fig:ErrorsERho}
-\end{figure}
+%\begin{figure}[h!]
+%\centering
+%	\begin{subfigure}{0.33\linewidth}
+%		\centering
+%				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{EulerSC4-17Res1e-6_errors_ERho}.png}
+%		\label{fig:sub1}
+%	\end{subfigure}%
+%	\begin{subfigure}{0.33\linewidth}
+%		\centering
+%				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_sg_n9_sg4_s05_aoa_errors_ERho}.png}
+%		\label{fig:sub2}
+%	\end{subfigure}%
+%		\begin{subfigure}{0.33\linewidth}
+%		\centering
+%				\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_ipm_n9_sg4_s05_aoa_errors_ERho}.png}
+%		\label{fig:sub2}
+%	\end{subfigure}
+%	\caption{Errors E$[\rho]$ for SC$^{17}$, SG$_{10}$, IPM$_{10}$ (from left to right).}
+%	\label{fig:ErrorsERho}
+%\end{figure}
 %\newgeometry{top=2.5cm, left=2.5cm}
@@
 
+In the following, we assume a fixed angle of attack with $\phi = 1.25$ degrees and study the effect of two sources of uncertainties, namely the farfield pressure and Mach number.
+The farfield pressure is $p \sim U(100\;325,102\;325)$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
+Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which proved to be more efficient than a sparse grid quadrature.
+For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the one-shot strategy).
+The IPM method uses moment orders ranging from $1$ to $9$ adaptively with refinement barriers of $\delta_{-} = 1\cdot 10^{-4}$ and $\delta_{+} = 1\cdot 10^{-5}$.
+The refinement retardation allows the truncation order to have a maximal total degree of $1$ until a residual of $\varepsilon = 1.5\cdot 10^{-5}$ and then increases the truncation order by one when the residual is reduced by an amount of $5\cdot 10^{-6}$.
+Hence, the maximal truncation order of $9$ is reached when the residual is below $\varepsilon = 7\cdot 10^{-6}$.
+Again, the refinement strategy let to negative densities for SG resulting in a failure of the method.
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{Relative L$^2$ error with 17 MPI threads for density and energy.}
-	\label{fig:L2ErrorSolution}
+	\caption{Relative L$^2$ error with 16 MPI threads for density and energy.}
+	\label{fig:L2ErrorSolution2D}
 \end{figure}
-
-It sticks out that again, the intrusive methods (SG$_9$,IPM$_9$) yield a smaller error level for the same number of unknowns compared to SC.
-For the same number of unknowns, the run time of the intrusive methods is always bigger than the corresponding SC method.
-However, the error of SG and IPM lies in the area of SC$^{17}$ at which the adaptive method has a smaller run time than SC.
-While SG shows a slightly better approximation of the expectation value, IPM shows a smaller error of the variance.
+The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}.
+As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method.
+Furthermore, the error obtained with the intrusive methods requires a total degree of $M = 9$, i.e.
+$N = 55$ moments to reach the error level of SC with 17 quadrature points per dimension i.e.
+$17^2 = 289$ collocation points.
+Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns needed for intrusive methods to obtain a certain error weighs in more heavily.
+This shows one promising characteristic of intrusive techniques and points to their applicability for higher dimensional problems.
+In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance.
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.2]{figs/Euler2D/{sc50UncertainMaSigma0-025PSigma1000-0Dim2_ERho}.png}
+		\label{fig:referenceSolutionsub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.2]{figs/Euler2D/{sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho}.png}
+		\label{fig:referenceSolutionsub2}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\label{fig:referenceSolution2D}
+\end{figure}
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_ERho}.png}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_VarRho}.png}
+		\label{fig:sub2}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_refinementLevel}.png}
+		\label{fig:sub1}
+	\end{subfigure}
+	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{1-9}$}
+	\label{fig:readosIPMEVar2D}
+\end{figure}

### 2019-09-28 Paper1/introduction.tex
Paper done without pipe testcase

@@
 The numerical discretization as well as the implementation and structure of the semi-intrusive method is introduced in section \ref{sec:framework}.
-Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration.
+In section~\ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration.
 Section~\ref{sec:adaptivity} extends the presented numerical framework to an algorithm making use of adaptivity.
+Implementation and parallelization details are given in section~\ref{sec:parallel}.
 A comparison of results computed with the presented methods is then given in section \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.

### 2019-09-28 Paper1/results.tex
Paper done without pipe testcase

@@
 \end{figure}
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rho]}.pdf}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rho]}.pdf}
+		\label{fig:sub2}
+	\end{subfigure}
+
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rhoE]}.pdf}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rhoE]}.pdf}
+		\label{fig:sub2}
+	\end{subfigure}
+	\caption{Relative L$^2$ error with 5 MPI threads for density and energy.}
+	\label{fig:L2ErrorSolution}
+\end{figure}
+\begin{figure}[h!]
+\centering
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_ERho}.png}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_ERho}.png}
+		\label{fig:sub2}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_ERho}.png}
+		\label{fig:sub1}
+	\end{subfigure}
+
+		\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_VarRho}.png}
+		\label{fig:sub1}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_VarRho}.png}
+		\label{fig:sub2}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_VarRho}.png}
+		\label{fig:sub1}
+	\end{subfigure}
+	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
+	\label{fig:ERhoVarRho5}
+\end{figure}
 
@@
 %One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
-\begin{figure}[h!]
-\centering
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rho]}.pdf}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rho]}.pdf}
-		\label{fig:sub2}
-	\end{subfigure}
-
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rhoE]}.pdf}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rhoE]}.pdf}
-		\label{fig:sub2}
-	\end{subfigure}
-	\caption{Relative L$^2$ error with 5 MPI threads for density and energy.}
-	\label{fig:L2ErrorSolution}
-\end{figure}
 
@@
 \end{itemize}
-\begin{figure}[h!]
-\centering
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_ERho}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_ERho}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_ERho}.png}
-		\label{fig:sub1}
-	\end{subfigure}
 
-		\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_VarRho}.png}
-		\label{fig:sub1}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_VarRho}.png}
-		\label{fig:sub2}
-	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
-		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_VarRho}.png}
-		\label{fig:sub1}
-	\end{subfigure}
-	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
-	\label{fig:ERhoVarRho5}
-\end{figure}
 Let us finally take a look at the expectation value and variance computed with different methods.
@@
 The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}.
+To determine the error, we used Stochastic-Collocation with a $50$ by $50$ Gauss quadrature rule.
 As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method.
@@
 In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance.
+Furthermore, when comparing IPM with and without one-shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case.
+This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased.
+I.e.
+using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the one-shot strategy an effective approach, especially for problems with high uncertain dimension.
 \begin{figure}[h!]
@@
 		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_ERho}.png}
+		\caption{}
 		\label{fig:sub1}
@@
 		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_VarRho}.png}
+		\caption{}
 		\label{fig:sub2}
@@
 		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_refinementLevel}.png}
-		\label{fig:sub1}
+		\caption{}
+		\label{fig:readosIPMEVar2DRefinement}
 	\end{subfigure}
-	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{1-9}$}
+	\caption{(a) E$[\rho]$, (b) Var$[\rho]$ and (c) refinement level for readosIPM$_{1-9}$}
 	\label{fig:readosIPMEVar2D}
 \end{figure}
+When looking at the computed IPM expectation value and variance in Figure~\ref{fig:readosIPMEVar2D} and compare the results against the reference solution in Figure~\ref{fig:referenceSolution2D}, one can observe that since the uncertainty effects the result more heavily than in the one-dimensional case, the IPM solution still shows a step-like profile.
+However, we are able to capture the main features of the reference solution.
+The increased effect of the uncertainty can again be seen in Figure~\ref{fig:readosIPMEVar2DRefinement}, where a refinement level of $8$, i.e.
+a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case.
+
+\subsection{Pipe Testcase}
+\comment{TODO: Pipe testcase}

### 2019-09-28 Paper1/summary_outlook.tex
Paper done without pipe testcase

@@
 \label{sec:summary_outlook}
+In this work, we proposed acceleration techniques, which can be applied to the intrusive framework of IPM and SG: We use a one-shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously.
+By not fully converging the dual iteration, we can reduce computational costs.
+Additionally, we make use of adaptivity and propose to keep a low maximal truncation order for most of the iteration to the steady state solution.
+Since complicated structures in the uncertain dimension only appear on a small portion of the spacial mesh, we are able to heavily reduce computational costs.
+The effects of the proposed techniques have been demonstrated by comparing results obtained with IPM as well as SG against SC.
+In our test cases, the intrusive methods yield the same error as SC for a reduced runtime, especially since intrusive methods require less unknowns to achieve a certain accuracy due to aliasing errors.
+In higher-dimensional problems, this effect is amplified since the number of unknowns to achieve a certain total degree is asymptotically smaller than the number of quadrature points in a tensorized or sparse grid.
+Furthermore, we could observe that the required residual at which the solution of intrusive methods reaches a steady state is smaller than for SC.
+Additionally, the ability to adaptively change the truncation order helps intrusive methods to compete with SC.
 
 In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations.
-Similar to the one-shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to steady state.
+Similar to the one-shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state.
 Hessian approximations that can be interesting are BFGS and sparse BFGS, which construct the Hessian from previously computed gradients.
@@
 Another example of breaking up the non-intrusive nature of Monte Carlo methods can be found in \cite{poette2019gpc}, where the generation of random samples is combined with the sampling after collisions to increase efficiency.
+Furthermore, we aim at applying the proposed acceleration techniques to SG methods with hyperbolicity limiters \cite{wu2017stochastic,schlachter2018hyperbolicity}.

### 2019-09-28 code/src/navierstokes.cpp
add ns flux solver

@@ -18,21 +18,42 @@ NavierStokes::NavierStokes( Settings* settings ) : Problem( settings ) {
 
 NavierStokes::~NavierStokes() {}
 
-Vector NavierStokes::G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n ) {
-    double rhoInv = 1.0 / u[0];
-    double vU     = u[1] * rhoInv;
-    double p      = ( _gamma - 1.0 ) * ( u[2] - 0.5 * u[0] * pow( vU, 2 ) );
-    double aU     = sqrt( _gamma * p * rhoInv );
-    rhoInv        = 1.0 / v[0];
-    double vV     = v[1] * rhoInv;
-    p             = ( _gamma - 1.0 ) * ( v[2] - 0.5 * v[0] * pow( vV, 2 ) );
-    double aV     = sqrt( _gamma * p * rhoInv );
+Vector NavierStokes::G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n, const double& length ) {
 
-    double uUProjected = nUnit[0] * vU;
-    double uVProjected = nUnit[0] * vV;
+    // gas property
+    double inK = (3.0 - _gamma) / (_gamma - 1.0);
+    double vis = 0.001;
+    
+    // left interface
+    double wL[3], primL[3];
+    
+    wL[0] = u[0];
+    wL[1] = u[1];
+    wL[2] = u[2];
+    
+    get_primitive(primL, wL, _gamma);
+    
+    double pL = ( _gamma - 1.0 ) * ( wL[2] - 0.5 * wL[0] * pow( wL[1], 2 ) );
+    double ssL = sqrt( _gamma * pL / primL[0] );
+    
+    // right interface
+    double wR[3], primR[3];
+    
+    wR[0] = v[0];
+    wR[1] = v[1];
+    wR[2] = v[2];
+    
+    get_primitive(primR, wR, _gamma);
+    
+    double pR = ( _gamma - 1.0 ) * ( wR[2] - 0.5 * wR[0] * pow( wR[1], 2 ) );
+    double ssR = sqrt( _gamma * pR / primR[0] );
 
-    double lambdaMin = uUProjected - aU;
-    double lambdaMax = uVProjected + aV;
+    // projection
+    double uUProjected = nUnit[0] * primL[1];
+    double uVProjected = nUnit[0] * primR[1];
+
+    double lambdaMin = uUProjected - ssL;
[…]

### 2019-10-31 Paper1/adaptivity.tex
paper remarks jannick

@@
 Since stochastic hyperbolic problems generally experience shocks in a small portion of the space-time domain, the idea is to perform arising computations on a high accuracy level in this small area, while keeping a low level of accuracy in the remainder.
-The hope is to achieve the finest level accuracy with this adaptive strategy, i.e.
-the same error is obtained while using a significantly reduced number of unknowns.
+The hope is to \comja{automatically select the lowest order moment capable of approximating the solution with given accuracy, i.e....} achieve the finest level accuracy with this adaptive strategy, i.e.
+the same error is obtained while using a significantly reduced number of unknowns \comja{in most parts of the computational domain}.
 
@@
 An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
+\comja{
+\begin{align}\label{eq:adaptiveFVUpdate}
+\bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= &\langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T \\&- \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
+\nonumber
+\end{align}
+}
 \begin{align}\label{eq:adaptiveFVUpdate}
@@
 The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$.
-The IPM algorithm with adaptivity then reads
+The IPM algorithm with adaptivity then reads \comja{results in Algorithm \ref{alg:ad-IPM}.}
 \begin{algorithm}[H]
@@
 For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of th required iterations on a lower but cheaper accuracy level.
-The same strategy can be applied for one-shot IPM. In this case, the algorithm is given by
+The same strategy can be applied for one-shot IPM. In this case, the algorithm is given by \comja{Algorithm \ref{alg:adosIPM}.}
 \begin{algorithm}[H]

### 2019-10-31 Paper1/introduction.tex
paper remarks jannick

@@
 Hyperbolic equations play an important role in various research areas such as fluid dynamics or plasma physics.
-Efficient numerical methods combined with robust implementations are available for these problems, however they do not account for uncertainties which can arise in measurement data or modeling assumptions.
+Efficient numerical methods combined with robust implementations are \comja{widely} available for these problems, however they do not account for uncertainties which can arise in measurement data or modeling assumptions.
 Including the effects of uncertainties in differential equations has become an important topic in the last decades.
@@
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
-Furthermore, SC is embarrassingly parallel, since the required computations can be carried out simultaneously on different cores.
+Furthermore, SC is embarrassingly parallel, since the required computations \comja{decouple and the workload can easily be distributed across several machines.} can be carried out simultaneously on different cores.
 
 The main idea of intrusive methods is to derive a system of equations describing the time evolution of the moments which can then be solved with a deterministic numerical scheme.
-A popular approach to describe the moment system is the stochastic-Galerkin (SG) method \cite{ghanem2003stochastic}, which chooses a polynomial solution ansatz and performs a Galerkin projection to derive a closed system of equations.
-One significant drawback of SG is its not necessarily hyperbolic moment system \cite{poette2009uncertainty}.
+A popular approach to describe the moment system is the stochastic-Galerkin (SG) method \cite{ghanem2003stochastic}, which chooses a polynomial \comja{basis ansatz of the solution?} solution ansatz and performs a Galerkin projection to derive a closed system of equations.
+One significant drawback of SG is \comja{, that its moment system is not necessarily hyperbolic.} its not necessarily hyperbolic moment system \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
@@
 Unfortunately, the desirable properties of IPM come along with significantly increased numerical costs, since IPM requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
-However, IPM and minimal entropy methods in general are well suited for modern HPC architectures, which can be used to reduce the run time \cite{garrett2015optimization}.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architectures \comja{würde den nächsten Halbsatz weglassen}, which can be used to reduce the run time \cite{garrett2015optimization}.
 
 When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
-These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}.
+These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through \comja{either} the spatial numerical method \cite{offner2017stability} or \comja{by} filters \cite{kusch2018filtered}.
 Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
@@
 In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
-The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
-One can use the solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
+The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal \comja{$L^2$-optimal,} if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
+One can \comja{also} use the \comja{nur ansatz} solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
 Indeed, we have that

### 2019-10-31 Paper1/main.tex
paper remarks jannick

@@
 
-\newcommand{\commentJannick}[1]{{\textcolor{red}{#1}}}
+\def\comja#1{{\color{red}{\bfseries \fbox{Remark {\color{red}Jannick}}}\enspace #1}}
 \newcommand{\comment}[1]{{\textcolor{blue}{#1}}}
@@
 Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
-Non-intrusive methods allow using a given deterministic solver as a black box while being embarrassingly parallel.
+Non-intrusive methods allow \comja{the usage of} using a given deterministic solver \comja{in a black-box manner} as a black box\comja{,} while being embarrassingly parallel.
 Intrusive methods on the other hand do not suffer from aliasing effects and provide a more general framework for adaptive refinement techniques.
 Since the moment system solved by intrusive methods is not necessarily hyperbolic, the Intrusive Polynomial Moment (IPM) method has been introduced.
-IPM maintains hyperbolicity but comes at the cost of having to solve an optimization problem in every spatial cell and every time step.
+IPM maintains hyperbolicity \comja{,} but comes at the cost of \comja{solving an additional} having to solve an optimization problem in every spatial cell and every time step.
 In this work, we propose acceleration techniques for intrusive methods.
 When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from shape optimization.
-Integrating the iteration to solve the optimization problem into the moment update guarantees local convergence while reducing numerical costs.
+Integrating the iteration \comja{from the numerical treatment of the opt...} to solve the optimization problem into the moment update \comja{reduces numerical costs, while still guaranteeing local convergence.} guarantees local convergence while reducing numerical costs.
 Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
-We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
+We demonstrate the effectiveness of the proposed strategies \comja{for multi-dimensional uncertainty spaces} by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
 \end{abstract}

### 2019-10-31 Paper1/one_shot_ipm.tex
paper remarks jannick

@@
 
-In the following, we look at steady state problems, i.e. we have
+In the following \comja{section we only consider steady...}, we look at steady state problems, i.e. we have
 \begin{align}\label{eq:hyperbolicProblemSteady}
@@
 which is again needed for the construction of intrusive methods.
-By adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
+By \comja{introducing} adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
 Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
@@
 
-In this section we aim at breaking up the inner loop in the IPM algorithm \ref{alg:IPM}, i.e.
+In this section we aim at breaking up the inner loop in the IPM \comja{Algorithm} algorithm \ref{alg:IPM}, i.e.
 to just perform one iteration of the dual problem in each time step.
@@
 \end{subequations}
-This gives algorithm
+This gives algorithm \comja{yields Algotihm \ref{alg:osIPM}.}
 \begin{algorithm}[H]
@@
 \end{algorithm}
-We call this method One-Shot IPM, since it is inspired by One-shot optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables.
+We call this method One-Shot IPM, since it is inspired by One-shot \comja{Shot} optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables.
 Note that the dual variables from the One-Shot iteration are written without a hat to indicate that they are not the exact solution of the dual problem.
@@
 &\bm{\lambda}_j^{n+1} = \bm{d}(\bm{\lambda}_j^{n},\bm{\hat u}_j^{n}) \enskip \text{ for all j} \\
-&\bm{\hat u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{\hat u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{\hat u}_{j+1}^{n})\right)
+&\bm{\hat u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{\hat u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{\hat u}_{j+1}^{n})\right).
 \end{align*}
@@
 \end{proof}
+\comja{ca-IPM ist raus - remark streichen?}
 \begin{remark}

### 2019-10-31 Paper1/results.tex
paper remarks jannick

@@
 At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of $273.15$ K.
-Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees.
+Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees \comja{, i.e.}.
 I.e.
@@
 To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
-In the following, we investigate the L$^2$ error of the variance and the expectation value.
-The L$^2$ error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
+In the following, we investigate the \comja{L$^2$-error}L$^2$ error of the variance and the expectation value.
+The \comja{s.o.} L$^2$ error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
 \begin{align*}
@@
 The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely refined far field from affecting the error.
+\comja{hier sollten wir vll in Fig 1 die Box gestrichelt markieren}
 
@@
 	\end{subfigure}
-	\caption{Relative L$^2$ error with 5 MPI threads for density and energy.}
+	\caption{Relative L$^2$ error with 5 MPI threads for density and energy. \comja{ados reados am besten auch in der caption erklären - kommt erst spät im text}}
 	\label{fig:L2ErrorSolution}
@@
 \centering
-	\begin{subfigure}{0.33\linewidth}
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 	\end{subfigure}%
-	\begin{subfigure}{0.33\linewidth}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 		\label{fig:sub1}
-	\end{subfigure}
-
-		\begin{subfigure}{0.33\linewidth}
+	\end{subfigure}\\
+	\vspace{-0.35cm}
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 	\end{subfigure}
-	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
+	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).\comja{Bilder auf Abstand gesetzt}}
 	\label{fig:ERhoVarRho5}
@@
 Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
-When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
+\comja{When the smoothness indicator of the refinement retardation augmented IPM ...}When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
 The remaining methods have been computed with a level 4 quadrature rule.
@@
 For every quadrature point, SC iterates the solution until the density reaches a certain residual level $\varepsilon$.
-We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following table for the SC$^{17}$ solution.
+We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following table \comja{Table \ref{tab:resSC17}} for the SC$^{17}$ solution.
 \begin{center}
+\begin{table}
+\resizebox{\columnwidth}{!}{%
 \begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
@@
 \end{tabular}
+}
+\caption{\comja{Residual level SC$^{17}$ + Größe skaliert}}
+\label{tab:resSC17}
+\end{table}
+
 \end{center}
@@
 The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
-When comparing intrusive methods without acceleration techniques as well as SC, the following properties stick out:
+When comparing intrusive methods without acceleration techniques as well as SC, the following properties stick out \comja{emerge}:
 \begin{itemize}
-\item Compared to IPM, stochastic-Galerkin comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
+\item Compared to IPM, stochastic-Galerkin comes at a heavily \comja{significantly} reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
 \item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
@@
 \begin{itemize}
-\item The one shot IPM (osIPM) method proposed in in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
+\item The \comja{One-Shot} one shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
 \item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the one-shot idea, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM).
@@
 To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order.
-The residual values are then determined to be $0.00006,0.00003,0.000022$ and $0.00002$.
+The residual values are then determined to be $0.00006,0.00003,0.000022$ and $0.00002$ \comja{scientific notation?}.
 The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$.
@@
 In the following, we assume a fixed angle of attack with $\phi = 1.25$ degrees and study the effect of two sources of uncertainties, namely the farfield pressure and Mach number.
-The farfield pressure is $p \sim U(100\;325,102\;325)$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
+The farfield pressure is $p \sim U(100\;325,102\;325)$ \comja{no newline} Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
 Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which proved to be more efficient than a sparse grid quadrature.
@@
 This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased.
-I.e.
+\comja{This means, }I.e.
 using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the one-shot strategy an effective approach, especially for problems with high uncertain dimension.

### 2019-10-31 code/src/problem.cpp
nonlinear thermal radiative added

@@ -13,6 +13,7 @@
 #include "shallowwater.h"
 #include "shallowwater2d.h"
 #include "thermalradiative.h"
+#include "thermalradiativegeneral.h"
 
 Problem::Problem( Settings* settings ) : _settings( settings ) {
     _log = spdlog::get( "event" );
@@ -54,7 +55,7 @@ Problem* Problem::Create( Settings* settings ) {
         return new RadiationHydrodynamics1D( settings );
     }
     else if( settings->GetProblemType() == ProblemType::P_THERMALRAD_1D ) {
-        return new ThermalRadiative( settings );
+        return new ThermalRadiativeGeneral( settings );
     }
     else if( settings->GetProblemType() == ProblemType::P_NAVIERSTOKES_1D ) {
         return new NavierStokes( settings );

### 2019-10-31 code/src/thermalradiativegeneral.cpp

@@ -1,7 +1,7 @@
-#include "thermalradiative.h"
+#include "thermalradiativegeneral.h"
 #include "quadraturegrid.h"
 
-ThermalRadiative::ThermalRadiative( Settings* settings ) : Problem( settings ) {
+ThermalRadiativeGeneral::ThermalRadiativeGeneral( Settings* settings ) : Problem( settings ) {
     _nStates = 3;
     settings->SetNStates( _nStates );
     _settings->SetExactSolution( false );
@@ -31,20 +31,20 @@ ThermalRadiative::ThermalRadiative( Settings* settings ) : Problem( settings ) {
         auto file    = cpptoml::parse_file( _settings->GetInputFile() );
         auto problem = file->get_table( "problem" );
     } catch( const cpptoml::parse_exception& e ) {
-        _log->error( "[ThermalRadiative] Failed to parse {0}: {1}", _settings->GetInputFile(), e.what() );
+        _log->error( "[ThermalRadiativeGeneral] Failed to parse {0}: {1}", _settings->GetInputFile(), e.what() );
         exit( EXIT_FAILURE );
     }
 }
 
-ThermalRadiative::~ThermalRadiative() {}
+ThermalRadiativeGeneral::~ThermalRadiativeGeneral() {}
 
-Vector ThermalRadiative::G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n ) {
+Vector ThermalRadiativeGeneral::G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n ) {
     Vector g = 0.5 * ( F( u ) + F( v ) ) * nUnit - 0.5 * ( v - u ) * norm( n ) / _settings->GetDT();
     g[2]     = 0.0;    // set temperature flux to zero
     return g;
 }
 
-Matrix ThermalRadiative::G( const Matrix& u, const Matrix& v, const Vector& nUnit, const Vector& n, unsigned level ) {
+Matrix ThermalRadiativeGeneral::G( const Matrix& u, const Matrix& v, const Vector& nUnit, const Vector& n, unsigned level ) {
     unsigned nStates = u.rows();
     unsigned Nq      = _settings->GetNqPEAtRef( level );
     Matrix y( nStates, Nq );
[…]

### 2019-10-31 code/src/thermalradiativegeneral.h

@@ -1,0 +1,33 @@
+#ifndef THERMALRADIATIVEGENERAL_H
+#define THERMALRADIATIVEGENERAL_H
+
+#include "problem.h"
+
+class ThermalRadiativeGeneral : public Problem
+{
+  private:
+    double _epsilon;
+    double _c;
+    double _a;
+    double _TRef;
+    double _sigma;
+    double _alpha;
+    std::vector<Vector> _xiQuad;
+    std::vector<double> _variances;
+    double ScaledInternalEnergy(double TTilde )const;
+    double ScaledTemperature(double eTilde ) const;
+
+  public:
+    ThermalRadiativeGeneral( Settings* settings );
+    virtual ~ThermalRadiativeGeneral();
+    inline Vector G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n );
+    virtual Matrix G( const Matrix& u, const Matrix& v, const Vector& nUnit, const Vector& n, unsigned level );
+    Matrix F( const Vector& u );
+    Matrix F( const Matrix& u );
+    virtual Matrix Source( const Matrix& uQ, const Vector& x, double t, unsigned level ) const;
+    virtual double ComputeDt( const Matrix& u, double dx, unsigned level ) const;
+    virtual Vector IC( const Vector& x, const Vector& xi );
+    virtual Vector LoadIC( const Vector& x, const Vector& xi );
+};
+
+#endif // THERMALRADIATIVEGENERAL_H

### 2019-11-02 Paper1/adaptivity.tex
integrated remarks

@@
 Since stochastic hyperbolic problems generally experience shocks in a small portion of the space-time domain, the idea is to perform arising computations on a high accuracy level in this small area, while keeping a low level of accuracy in the remainder.
-The hope is to \comja{automatically select the lowest order moment capable of approximating the solution with given accuracy, i.e....} achieve the finest level accuracy with this adaptive strategy, i.e.
-the same error is obtained while using a significantly reduced number of unknowns \comja{in most parts of the computational domain}.
+The hope is to automatically select the lowest order moment capable of approximating the solution with given accuracy, i.e.
+the same error is obtained while using a significantly reduced number of unknowns in most parts of the computational domain.
 
@@
 An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
-\comja{
 \begin{align}\label{eq:adaptiveFVUpdate}
@@
 \nonumber
-\end{align}
-}
-\begin{align}\label{eq:adaptiveFVUpdate}
-\bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= \langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
 \end{align}
@@
 The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$.
-The IPM algorithm with adaptivity then reads \comja{results in Algorithm \ref{alg:ad-IPM}.}
+The IPM algorithm with adaptivity results in Algorithm \ref{alg:ad-IPM}.
 \begin{algorithm}[H]
@@
 For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of th required iterations on a lower but cheaper accuracy level.
-The same strategy can be applied for one-shot IPM. In this case, the algorithm is given by \comja{Algorithm \ref{alg:adosIPM}.}
+The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm \ref{alg:adosIPM}.
 \begin{algorithm}[H]
@@
 \end{algorithmic}
-\caption{Adaptive one-shot IPM implementation with refinement retardation}
+\caption{Adaptive One-Shot IPM implementation with refinement retardation}
 \label{alg:adosIPM}

### 2019-11-02 Paper1/implementation.tex
results continued
integrated remarks

@@
 From here, the dual update can again be performed.
-The standard IPM Algorithm~\ref{alg:IPM} and one-shot IPM \ref{alg:osIPM} use this parallelization strategy accordingly.
+The standard IPM Algorithm~\ref{alg:IPM} and One-Shot IPM \ref{alg:osIPM} use this parallelization strategy accordingly.
 Again, we point out that stochastic-Galerkin is a variant of IPM, i.e.
@@
 we use the same deterministic numerical flux $\bm g$.
+This allows a fair comparison of the different intrusive and non-intrusive techniques.

### 2019-11-02 Paper1/introduction.tex
integrated remarks

--- 2019-09-28 Paper1/introduction.tex
@@
 Hyperbolic equations play an important role in various research areas such as fluid dynamics or plasma physics.
-Efficient numerical methods combined with robust implementations are available for these problems, however they do not account for uncertainties which can arise in measurement data or modeling assumptions.
+Efficient numerical methods combined with robust implementations are widely available for these problems, however they do not account for uncertainties which can arise in measurement data or modeling assumptions.
 Including the effects of uncertainties in differential equations has become an important topic in the last decades.
@@
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
-Furthermore, SC is embarrassingly parallel, since the required computations can be carried out simultaneously on different cores.
+Furthermore, SC is embarrassingly parallel, since the required computations decouple and the workload can easily be distributed across several processors.
 
 The main idea of intrusive methods is to derive a system of equations describing the time evolution of the moments which can then be solved with a deterministic numerical scheme.
-A popular approach to describe the moment system is the stochastic-Galerkin (SG) method \cite{ghanem2003stochastic}, which chooses a polynomial solution ansatz and performs a Galerkin projection to derive a closed system of equations.
-One significant drawback of SG is its not necessarily hyperbolic moment system \cite{poette2009uncertainty}.
+A popular approach to describe the moment system is the stochastic-Galerkin (SG) method \cite{ghanem2003stochastic}, which chooses a polynomial basis ansatz of the solution and performs a Galerkin projection to derive a closed system of equations.
+One significant drawback of SG is, that its moment system is not necessarily hyperbolic \cite{poette2009uncertainty}.
 A generalization of stochastic-Galerkin, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) method \cite{poette2009uncertainty}.
@@
 Unfortunately, the desirable properties of IPM come along with significantly increased numerical costs, since IPM requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
-However, IPM and minimal entropy methods in general are well suited for modern HPC architectures, which can be used to reduce the run time \cite{garrett2015optimization}.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architectures.
 
 When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
-These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}.
+These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through either the spatial numerical method \cite{offner2017stability} or by filters \cite{kusch2018filtered}.
 Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
@@
 In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
-The solution ansatz \eqref{eq:SGClosure} is $L^2$ optimal if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
-One can use the solution ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
+The solution ansatz \eqref{eq:SGClosure} is $L^2$-optimal, if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
+One can also use the ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
 Indeed, we have that

### 2019-11-02 Paper1/main.tex
integrated remarks

--- 2019-09-25 Paper1/main.tex
@@
 
-\newcommand{\commentJannick}[1]{{\textcolor{red}{#1}}}
+\def\remja#1{{\color{red}{\bfseries \fbox{Remark {\color{red}Jannick}}}\enspace #1}}
+\def\frama#1{{\color{red}{\bfseries \fbox{{\color{blue}Martin}}}\enspace #1}}
 \newcommand{\comment}[1]{{\textcolor{blue}{#1}}}
@@
 Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques.
-Non-intrusive methods allow using a given deterministic solver as a black box while being embarrassingly parallel.
+Non-intrusive methods allow the usage of a given deterministic solver in a black-box manner, while being embarrassingly parallel.
 Intrusive methods on the other hand do not suffer from aliasing effects and provide a more general framework for adaptive refinement techniques.
 Since the moment system solved by intrusive methods is not necessarily hyperbolic, the Intrusive Polynomial Moment (IPM) method has been introduced.
-IPM maintains hyperbolicity but comes at the cost of having to solve an optimization problem in every spatial cell and every time step.
+IPM maintains hyperbolicity, but comes at the cost of solving an additional problem in every spatial cell and every time step.
 In this work, we propose acceleration techniques for intrusive methods.
 When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from shape optimization.
-Integrating the iteration to solve the optimization problem into the moment update guarantees local convergence while reducing numerical costs.
+Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while still guaranteeing local convergence.
 Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
-We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
+We demonstrate the effectiveness of the proposed strategies for multi-dimensional uncertainty spaces by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
 \end{abstract}

### 2019-11-02 Paper1/one_shot_ipm.tex
integrated remarks

@@
-\section{One-shot IPM}
+\section{One-Shot IPM}
 \label{sec:OneShotIPM}
 
-In the following \comja{section we only consider steady...}, we look at steady state problems, i.e. we have
+In the following section we only consider steady state problems, i.e. we have
 \begin{align}\label{eq:hyperbolicProblemSteady}
@@
 which is again needed for the construction of intrusive methods.
-By \comja{introducing} adding a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
+By introducing a pseudo-time and using the IPM closure, we obtain the same system as in \eqref{eq:SGMomentSystem}, i.e.
 Algorithm \ref{alg:IPM} can be used to iterate to a steady state solution.
@@
 
-In this section we aim at breaking up the inner loop in the IPM \comja{Algorithm} algorithm \ref{alg:IPM}, i.e.
+In this section we aim at breaking up the inner loop in the IPM Algorithm \ref{alg:IPM}, i.e.
 to just perform one iteration of the dual problem in each time step.
@@
 \end{subequations}
-This gives algorithm \comja{yields Algotihm \ref{alg:osIPM}.}
+This yields Algotihm \ref{alg:osIPM}.
 \begin{algorithm}[H]
@@
 \end{algorithmic}
-\caption{One-shot IPM implementation}
+\caption{One-Shot IPM implementation}
 \label{alg:osIPM}
 \end{algorithm}
-We call this method One-Shot IPM, since it is inspired by One-shot \comja{Shot} optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables.
+We call this method One-Shot IPM, since it is inspired by One-Shot optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables.
 Note that the dual variables from the One-Shot iteration are written without a hat to indicate that they are not the exact solution of the dual problem.
@@
 \end{proof}
-\comja{ca-IPM ist raus - remark streichen?}
-\begin{remark}
-Since the preconditioning step of the Collocation-accelerated IPM method generates initial conditions which are close to the steady state solution, using One-Shot IPM instead of classical IPM is well suited.
-However, our numerical calculations show that one-shot IPM converges even if the solution is far away from its steady state.
-\end{remark}
+%\begin{remark}
+%Since the preconditioning step of the Collocation-accelerated IPM method generates initial conditions which are close to the steady state solution, using One-Shot IPM instead of classical IPM is well suited.
+However, our numerical calculations show that One-Shot IPM converges even if the solution is far away from its steady state.
+%\end{remark}

### 2019-11-02 Paper1/results.tex
integrated remarks

--- 2019-09-28 Paper1/results.tex
@@
 At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of $273.15$ K.
-Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees.
-I.e.
+Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees, i.e.
 we choose $\phi(\xi) = 1.25 + 0.5\xi$ where $\xi\sim U(-1,1)$.
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$. \comment{TODO: Error-Box (gestrichelt) hinzufügen}}
 	\label{fig:referenceSolution}
@@
 To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
-In the following, we investigate the L$^2$ error of the variance and the expectation value.
-The L$^2$ error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
+In the following, we investigate the L$^2$-error of the variance and the expectation value.
+The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
 \begin{align*}
@@
 	\end{subfigure}
-	\caption{Relative L$^2$ error with different quadrature levels. The subscript denotes the moment order, the superscript denotes the Clenshaw-Curtis quadrature level.}
+	\caption{Relative L$^2$-error with different quadrature levels. The subscript denotes the moment order, the superscript denotes the Clenshaw-Curtis quadrature level.}
 	\label{fig:ErrorDifferentQuad}
@@
 	\end{subfigure}
-	\caption{Relative L$^2$ error with 5 MPI threads for density and energy.}
+	\caption{Relative L$^2$-error with 5 MPI threads for density and energy.}
 	\label{fig:L2ErrorSolution}
@@
 Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
-When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
+When the smoothness indicator of the refinement retardation augmented IPM \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
 The remaining methods have been computed with a level 4 quadrature rule.
@@
 For every quadrature point, SC iterates the solution until the density reaches a certain residual level $\varepsilon$.
-We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following table for the SC$^{17}$ solution.
+We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following Table \ref{tab:resSC17} for the SC$^{17}$ solution.
 \begin{center}
+\begin{table}
+\resizebox{\columnwidth}{!}{%
 \begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
@@
 \end{tabular}
+}
+\label{tab:resSC17}
+\end{table}
+
 \end{center}
@@
 The standard SG method however preserves positivity of mass, energy and pressure.
-The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
-When comparing intrusive methods without acceleration techniques as well as SC, the following properties stick out:
+The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
+When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
 \begin{itemize}
-\item Compared to IPM, stochastic-Galerkin comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
+\item Compared to IPM, stochastic-Galerkin comes at a significantly reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
 \item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
@@
 \begin{itemize}
-\item The one shot IPM (osIPM) method proposed in in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
-\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the one-shot idea, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM).
+\item The One-Shot one shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
+\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the One-Shot idea, the method is denoted by \textit{adaptive One-Shot IPM} (adosIPM).
 This method reaches the steady state IPM solution with a faster runtime than SG.
@@
 To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order.
-The residual values are then determined to be $0.00006,0.00003,0.000022$ and $0.00002$.
+The residual values are then determined to be $6e-05,3e-05,2.2e-5$ and $2e-05$.
 The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$.
@@
 In the following, we assume a fixed angle of attack with $\phi = 1.25$ degrees and study the effect of two sources of uncertainties, namely the farfield pressure and Mach number.
-The farfield pressure is $p \sim U(100\;325,102\;325)$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
+The farfield pressure is ${p \sim U(100\;325,102\;325)}$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
 Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which proved to be more efficient than a sparse grid quadrature.
-For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the one-shot strategy).
+For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the One-Shot strategy).
 The IPM method uses moment orders ranging from $1$ to $9$ adaptively with refinement barriers of $\delta_{-} = 1\cdot 10^{-4}$ and $\delta_{+} = 1\cdot 10^{-5}$.
@@
 	\end{subfigure}
-	\caption{Relative L$^2$ error with 16 MPI threads for density and energy.}
+	\caption{Relative L$^2$-error with 16 MPI threads for density and energy.}
 	\label{fig:L2ErrorSolution2D}
@@
 In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance.
-Furthermore, when comparing IPM with and without one-shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case.
+Furthermore, when comparing IPM with and without One-Shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case.
 This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased.
-I.e.
-using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the one-shot strategy an effective approach, especially for problems with high uncertain dimension.
+This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy an effective approach, especially for problems with high uncertain dimension.
 \begin{figure}[h!]

### 2019-11-02 code/src/closure.cpp
thermalradiation closure added

@@ -15,6 +15,7 @@
 #include "shallowwaterclosure2d.h"
 #include "stochasticgalerkin.h"
 #include "tensorizedquadrature.h"
+#include "thermalradiationclosure.h"
 #include "uniformsparsegrid.h"
 
 Closure::Closure( Settings* settings )
@@ -150,6 +151,9 @@ Closure* Closure::Create( Settings* settings ) {
     }
     else if( closureType == ClosureType::C_RADHYDRO ) {
         return new RadiHydroClosure1D( settings );
+    }
+    else if( closureType == ClosureType::C_THERMALRAD_1D ) {
+        return new ThermalRadiationClosure( settings );
     }
     else if( closureType == ClosureType::C_M1_1D ) {
         return new M1IPMClosure( settings );

### 2019-11-02 code/src/settings.h

@@ -35,6 +35,7 @@ enum ClosureType {
     C_L2FILTER,
     C_LASSOFILTER,
     C_RADHYDRO,
+    C_THERMALRAD_1D,
     C_M1_1D
 };
 enum LimiterType { L_MINMOD, L_NONE };

### 2019-11-02 code/src/thermalradiationclosure.cpp

@@ -1,0 +1,198 @@
+#include "thermalradiationclosure.h"
+
+ThermalRadiationClosure::ThermalRadiationClosure( Settings* settings ) : Closure( settings ), _nHydroStates( 1 ), _nMoments( 2 ) { _alpha = 1.0; }
+
+ThermalRadiationClosure::~ThermalRadiationClosure() {}
+
+void ThermalRadiationClosure::U( Vector& out, const Vector& Lambda ) {
+    out[0] = Lambda[0];
+    out[1] = Lambda[1];
+    out[2] = exp( Lambda[2] );
+}
+
+void ThermalRadiationClosure::U( Vector& out, const Vector& Lambda, bool dummy ) { out[0] = exp( Lambda[0] ); }
+
+void ThermalRadiationClosure::U( Matrix& out, const Matrix& Lambda ) {
+    for( unsigned k = 0; k < Lambda.columns(); ++k ) {
+        out( 0, k ) = Lambda( 0, k );
+        out( 1, k ) = Lambda( 1, k );
+        out( 2, k ) = exp( Lambda( 2, k ) );
+    }
+}
+
+Matrix ThermalRadiationClosure::U( const Matrix& Lambda ) {
+    Matrix y( _nStates, Lambda.columns(), 0.0 );
+    for( unsigned k = 0; k < Lambda.columns(); ++k ) {
+        y( 0, k ) = Lambda( 0, k );
+        y( 1, k ) = Lambda( 1, k );
+        y( 2, k ) = exp( Lambda( 2, k ) );
+    }
+
+    return y;
+}
+
+void ThermalRadiationClosure::DU( Matrix& y, const Vector& Lambda ) { y( 0, 0 ) = exp( Lambda[0] ); }
+
+void ThermalRadiationClosure::Gradient( Vector& g, const Matrix& lambda, const Matrix& u, unsigned refLevel ) {
+    Vector uKinetic( _nHydroStates, 0.0 );
+    unsigned nTotal = _nTotalForRef[refLevel];
+    g.reset();
+
+    for( unsigned k = 0; k < _nQTotalForRef[refLevel]; ++k ) {
+        U( uKinetic, EvaluateLambda( lambda, k, nTotal ), true );
+        for( unsigned i = 0; i < nTotal; ++i ) {
+            for( unsigned l = 0; l < _nHydroStates; ++l ) {
[…]

### 2019-11-02 code/src/thermalradiationclosure.h

@@ -1,0 +1,32 @@
+#ifndef THERMALRADIATIONCLOSURE_H
+#define THERMALRADIATIONCLOSURE_H
+
+#include "closure.h"
+
+class ThermalRadiationClosure : public Closure
+{
+  private:
+    ThermalRadiationClosure() = delete;
+    unsigned _nHydroStates;
+    unsigned _nMoments;
+
+  public:
+    ThermalRadiationClosure( Settings* settings );
+    virtual ~ThermalRadiationClosure();
+
+    virtual void U( Vector& out, const Vector& Lambda );
+    virtual void U( Vector& out, const Vector& Lambda, bool dummy );
+    virtual void U( Matrix& out, const Matrix& Lambda );
+    virtual Matrix U( const Matrix& Lambda );
+    virtual void DU( Matrix& y, const Vector& Lambda );
+    virtual void SolveClosure( Matrix& lambdaFull, const Matrix& uFull, unsigned refLevel );
+    virtual void SolveClosureSafe( Matrix& lambdaFull, const Matrix& uFull, unsigned refLevel );
+    virtual void Gradient( Vector& g, const Matrix& lambda, const Matrix& u, unsigned refLevel );
+    virtual void Hessian( Matrix& H, const Matrix& lambda, unsigned refLevel );
+    virtual void AddMatrixVectorToMatrix( const Matrix& A, const Vector& b, Matrix& y, unsigned nTotal ) const;
+    virtual void SubstractVectorMatrixOnVector( Vector& b, const Matrix& A, unsigned nTotal ) const;
+    virtual double CalcNorm( Vector& test, unsigned nTotal ) const;
+};
+
+
+#endif // THERMALRADIATIONCLOSURE_H

### 2019-11-02 code/src/thermalradiativegeneral.cpp

@@ -136,7 +136,7 @@ Vector ThermalRadiativeGeneral::IC( const Vector& x, const Vector& xi ) {
                        //_a * pow( 50.0, 2 ) / ( 8.0 * M_PI * pow( sigmaXi + 2.0, 2 ) ) *
                        //  exp( -0.5 * pow( 50.0 * ( x[0] - x0 ), 2 ) / pow( sigmaXi + 2.0, 2 ) ) );
     double F              = 0;
-    double internalEnergy = 1e-3 * _a * pow( _TRef, 4 );    // fix to ensure positive values of the inner energy
+    double internalEnergy = 1e-7 * _a * pow( _TRef, 4 );    // fix to ensure positive values of the inner energy - use 1e-3 without IPM
 
     y[0] = E / _a / pow( _TRef, 4 );
     y[1] = F / _a / pow( _TRef, 4 );

### 2019-11-02 Paper1/adaptivity.tex
changes Paper Jonas + questions to Martin

@@
 Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualSteady} is fulfilled.
-For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of th required iterations on a lower but cheaper accuracy level.
+For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper accuracy level.
 The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm \ref{alg:adosIPM}.

### 2019-11-02 Paper1/implementation.tex
changes Paper Jonas + questions to Martin

@@
 Therefore, we choose to parallelize the quadrature points, which again minimizes communication effort.
-As mentioned in Section~\ref{sec:costNumFlux}, we first compute the solution at stencil cells for all quadrature points.
+As mentioned in \ref{app:costNumFlux}, we first compute the solution at stencil cells for all quadrature points.
 I.e.
@@
 \end{align}
-Instead of distributing the mesh on the different processors, we now distribute the quadrature set, i.e.
+Instead of distributing the spatial mesh on the different processors, we now distribute the quadrature set, i.e.
 the sum in \eqref{eq:momentUpQuadrature} can be computed in parallel.

### 2019-11-02 Paper1/results.tex
changes Paper Jonas + questions to Martin

@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$. \comment{TODO: Error-Box (gestrichelt) hinzufügen}}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
 	\label{fig:referenceSolution}
@@
 	\end{subfigure}
-	\caption{Relative L$^2$-error with different quadrature levels. The subscript denotes the moment order, the superscript denotes the Clenshaw-Curtis quadrature level.}
+	\caption{Relative L$^2$-error with different quadrature levels.
+The subscript denotes the moment order, the superscript denotes the Clenshaw-Curtis quadrature level.\frama{Sollen wir einige Linien hier raus nehmen, damit es \"ubersichtlicher bleibt?}}
 	\label{fig:ErrorDifferentQuad}
@@
 To denote the quadrature level, we use a superscript and the moment order is denoted by a subscript.
+We observe the following:
 \begin{itemize}
@@
 This behavior results from aliasing effects in the numerical flux, which dominate the accuracy level.
+\frama{Sollen wir hier Strang's Lemma erw\"ahnen?
+Also der Fehler der Quadratur sollte so gross sein wie der Fehler durch die moment truncation?}
 \item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy.
@@
 Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
-When the smoothness indicator of the refinement retardation augmented IPM \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
+When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the adaptive methods decrease the truncation order, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
 The remaining methods have been computed with a level 4 quadrature rule.
@@
 We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following Table \ref{tab:resSC17} for the SC$^{17}$ solution.
+\frama{Sollen wir hier die Ergebnisse zu Frage 1 im zweiten Dokument einbauen?}
 \begin{center}
@@
 }
+\caption{Error values at various residual levels when using SC.}
 \label{tab:resSC17}
@@
 \begin{itemize}
-\item The One-Shot one shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
+\item The One-Shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
 \item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the One-Shot idea, the method is denoted by \textit{adaptive One-Shot IPM} (adosIPM).

### 2019-11-08 Paper1/framework.tex
changes black forest
document to Martin

@@
 with the flux $\bm{F}:\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
+Note that the inner transpose represent a dyadic product, i.e.
+the outer transpose is applied to a matrix.
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
@@
 However, when not using a quadratic entropy in the IPM method or when the physical flux of the deterministic problem is not a polynomial, it is not clear how many quadrature points the numerical quadrature rule requires to guarantee a sufficiently small quadrature error.
-We will study the approximation properties of IPM with different quadrature orders in Section~TODO
+We will study the approximation properties of IPM with different quadrature orders in Section~\ref{sec:resultsNACA1D}.

### 2019-11-08 Paper1/one_shot_ipm.tex
document to Martin

@@
 In the following, we will show that this iteration converges, if the chosen initial condition is sufficiently close to the steady state solution.
-For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence:
-
+For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence: \frama{Nico meinte wir sollten hier in die Voraussetzung packen, dass wir nah genug an der richtigen Loesung sind (Einzugsbereich des Newton) um das Wort 'local' zu vermeiden.
+Manchmal stören sich die Reviewer daran.}
 \begin{theorem}\label{th:Contractive}

### 2019-11-08 Paper1/results.tex
document to Martin

@@
 The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods.
-To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
+To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
 In the following, we investigate the L$^2$-error of the variance and the expectation value.
@@
 	\end{subfigure}
-	\caption{Relative L$^2$-error with 5 MPI threads for density and energy.}
+	\caption{Relative L$^2$-error with 5 MPI threads for density and energy.\frama{Eventuell diese Plots aufteilen? Oder sind die Ergebnisse verst\"andlich?}}
 	\label{fig:L2ErrorSolution}
@@
 \item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
-This behavior results from aliasing effects in the numerical flux, which dominate the accuracy level.
+Hence, the aliasing effects that result from an inaccurate quadrature rule in the numerical flux and not the truncation error of the moment system dominate the accuracy level.
 \frama{Sollen wir hier Strang's Lemma erw\"ahnen?
@@
 Hence, the variance is more sensitive to aliasing errors.
-\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, i.e.
+\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points (level 3) stagnates at the error level of SC$^9$, i.e.
 the aliasing error heavily affects the accuracy.
@@
 Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well.
-To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions.
 Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
@@
 \item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime.
-Here, we use two different strategies: First, we steadily increase the maximal truncation order when the residual approaches zero.
+Here, we use two different strategies to increase the accuracy: First, we steadily increase the maximal truncation order when the residual approaches zero.
 To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order.
@@
 The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}.
-To determine the error, we used Stochastic-Collocation with a $50$ by $50$ Gauss quadrature rule.
+To determine the error, we used Stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule.
 As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method.

### 2019-11-20 Paper1/appendix.tex
first corrections from Martin added

@@
 Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent.
-Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When computing and storing the values in a matrix $\bm A\in\mathbb{R}^{Q\times N}$ with entries $a_{ki} = \varphi_i(\xi_k)$ before running the program, the numerical flux \eqref{eq:numericalFluxIPMBurgers} can be split into two parts.
+Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} uses $O(N^2)$ operations whereas \eqref{eq:numFluxAnalytic} requires $O(N^3)$ operations: When computing and storing the values in a matrix $\bm A\in\mathbb{R}^{Q\times N}$ with entries $a_{ki} = \varphi_i(\xi_k)$ before running the program, the numerical flux \eqref{eq:numericalFluxIPMBurgers} can be split into two parts.
 First, we determine the SG solution at all quadrature points, i.e.

### 2019-11-20 Paper1/implementation.tex
first corrections from Martin added

@@
 In order to minimize the parallelization overhead, our goal is to minimize the need to send data between processors.
-Note that the dual problem (line 8 in Algorithm~\ref{alg:ad-IPM} and line 5 in Algorithm~\ref{alg:adosIPM}) does not have a stencil, i.e.
+Note that the dual problem (line 8 in Algorithm~\ref{alg:ad-IPM} and line 5 in Algorithm~\ref{alg:adosIPM}) does not require communication, i.e.
 it suffices to distribute the spatial cells between processors.
-In contrast to that, the finite volume update (line 14 in Algorithm~\ref{alg:ad-IPM} and line 8 in Algorithm~\ref{alg:adosIPM}) has a stencil.
+In contrast to that, the finite volume update (line 14 in Algorithm~\ref{alg:ad-IPM} and line 8 in Algorithm~\ref{alg:adosIPM}) requires communication, since values at neighboring cells need to be evaluated.
 Hence, distributing the spatial mesh between processors will yield communication overhead since data needs to be sent whenever a stencil cell lies on a different processor.
@@
 From here, the dual update can again be performed.
-The standard IPM Algorithm~\ref{alg:IPM} and One-Shot IPM \ref{alg:osIPM} use this parallelization strategy accordingly.
+The standard IPM Algorithm~\ref{alg:IPM} and One-Shot IPM Algorithm~\ref{alg:osIPM} use this parallelization strategy accordingly.
 Again, we point out that stochastic-Galerkin is a variant of IPM, i.e.

### 2019-11-20 Paper1/introduction.tex
changes Paper Jonas + questions to Martin
first corrections from Martin added

@@
 
-Commonly, methods to quantify uncertainties use a polynomial chaos (PC) expansion \cite{wiener1938homogeneous,xiu2002wiener} to represent the solution, i.e.
+One strategy to represent the solution's dependence on uncertainties is to use a polynomial chaos (PC) expansion \cite{wiener1938homogeneous,xiu2002wiener}, i.e.
 the uncertain space is spanned by polynomial basis functions.
@@
 Unfortunately, the desirable properties of IPM come along with significantly increased numerical costs, since IPM requires the repeated computation of the entropic expansion coefficients from the moment vector, which involves solving a convex optimization problem.
-However, IPM and minimal entropy methods in general are well suited for modern HPC architectures.
+However, IPM and minimal entropy methods in general are well suited for modern HPC architectures \cite{garrett2015optimization}.
 
@@
 By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs.
-In addition to that, the number of moments needed to obtain a certain order with intrusive methods is asymptotically smaller than the number of quadrature points for SC.
+In addition to that, the number of moments needed to obtain a certain order with intrusive methods is smaller than the number of quadrature points for SC.
 An additional downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
@@
 Hence, the moments can be computed by running a given deterministic solver for the original problem at each quadrature point.
-To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of total degree $M$ exactly, sparse grids only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
+To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of maximal degree $M$ exactly, sparse grids are designed to integrate polynomials of total degree $M$, for which they only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
 

### 2019-11-20 Paper1/main.tex
changes Paper Jonas + questions to Martin
first corrections from Martin added

@@
 % Declare title and authors, without \thanks
-\newcommand{\TheTitle}{Intrusive acceleration strategies for Uncertainty Quantification}
+\newcommand{\TheTitle}{Intrusive acceleration strategies for Uncertainty Quantification for hyperbolic systems of conservation laws}
 
@@
 In this work, we propose acceleration techniques for intrusive methods.
-When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from shape optimization.
+When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from PDE-constrained optimization.
 Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while still guaranteeing local convergence.
@@
 We demonstrate the effectiveness of the proposed strategies for multi-dimensional uncertainty spaces by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
+\frama{Sollen die Erkenntnisse vom Vergleich mit SC schon hier erw\"ahnt werden?}
 \end{abstract}

### 2019-11-20 Paper1/results.tex
first corrections from Martin added

@@
 As commonly done, the initial condition is equal to the far field boundary values.
-Consequently, the wall condition at the airfoil is violated and will correct the flow solution.
+Consequently, the wall condition at the airfoil is violated initially and will correct the flow solution.
 
@@
 In the center, the NACA0012 airfoil with a length of one meter is located.
-The discretization is composed of a coarsely refined far field and a finely resolved region around the airfoil, since we are interested in the flow solution at the airfoil.
+The spatial mesh is composed of a coarsely discretized far field and a finely resolved region around the airfoil, since we are interested in the flow solution at the airfoil.
 Altogether, the mesh consists of 22361 triangular elements.
@@
 \end{align*}
-The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely refined far field from affecting the error.
+The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.
 
@@
 	\end{subfigure}
-	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
+	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the refernce solution shown in Figure \ref{fig:referenceSolution}.}
 	\label{fig:ERhoVarRho5}
@@
 We start by investigating the effects this quadrature has on the solution accuracy.
-For this, we run the IPM method with a moment order ranging from three to seven using a Clenshaw-Curtis quadrature rule with level three (i.e.
+For this, we run the IPM method with a moment order ranging from 3 to 7 using a Clenshaw-Curtis quadrature rule with level three (i.e.
 9 quadrature points) and level four (i.e.
@@
 \begin{itemize}
-\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
+\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ (the subscript indicates the polynomial order, the superscript indicates \comment{the number of quadrature points}) it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
 Hence, the aliasing effects that result from an inaccurate quadrature rule in the numerical flux and not the truncation error of the moment system dominate the accuracy level.
-\frama{Sollen wir hier Strang's Lemma erw\"ahnen?
-Also der Fehler der Quadratur sollte so gross sein wie der Fehler durch die moment truncation?}
 \item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy.
@@
 The farfield pressure is ${p \sim U(100\;325,102\;325)}$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
-Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which proved to be more efficient than a sparse grid quadrature.
+Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which in our experiments proved to be more efficient than a sparse grid quadrature.
 For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the One-Shot strategy).

### 2019-11-20 Paper1/summary_outlook.tex
paper remarks jannick
integrated remarks
first corrections from Martin added

@@
 Furthermore, we could observe that the required residual at which the solution of intrusive methods reaches a steady state is smaller than for SC.
-Additionally, the ability to adaptively change the truncation order helps intrusive methods to compete with SC.
+Additionally, the ability to adaptively change the truncation order helps intrusive methods to compete with SC in terms of computational runtime.
 
@@
 Similar to the one-shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state.
-Hessian approximations that can be interesting are BFGS and sparse BFGS, which construct the Hessian from previously computed gradients.
+Hessian approximations that can be interesting are BFGS and sparse BFGS \cite[Chapter~6.1]{nocedal2006numerical}, which construct the Hessian from previously computed gradients.
 Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.

### 2019-11-20 Paper1/framework.tex
continued

@@
 \end{align*}
-with the flux $\bm{F}:\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
+with the flux $\bm{F}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
 Note that the inner transpose represent a dyadic product, i.e.
@@
 \end{align*}
-If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times p}\times\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$ as
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
 \begin{align}\label{eq:IPMDiscretization}
@@
 
-When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the original problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take the numerical flux
+When a consistent numerical flux $\bm g:\mathbb{R}^m\times\mathbb{R}^m\to\mathbb{R}^m$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the original problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take the numerical flux
 \begin{align*}
@@
 we store all entries in a vector.
-Newton's method uses the iteration function $\bm{d}:\mathbb{R}^{N\times p}\times\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$,
+Newton's method uses the iteration function $\bm{d}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$,
 \begin{align}\label{eq:dualIterationFunction}
@@
 \end{align}
-where $\bm H\in\mathbb{R}^{p\cdot N\times p\cdot N}$ is the Hessian of \eqref{eq:L}, given by
+where $\bm H\in\mathbb{R}^{p\cdot N\times m\cdot N}$ is the Hessian of \eqref{eq:L}, given by
 \begin{align*}
-\bm{H}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi})\otimes\bm{\varphi}\bm{\varphi}^T\rangle_Q^{T}.
+\bm{H}(\bm{\lambda}) := \langle \nabla \bm{u}_{s} (\bm{\lambda}^T\bm{\varphi})\otimes\bm{\varphi}\bm{\varphi}^T\rangle_Q^{T}.
 \end{align*}
@@
 \begin{align}\label{eq:tauCrit}
-\sum_{i=0}^p\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(m)};\bm{\hat{u}}_j^n) \right\Vert < \tau
+\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(m)};\bm{\hat{u}}_j^n) \right\Vert < \tau
 \end{align}

### 2019-11-20 Paper1/introduction.tex
continued

@@
 \end{subequations}
-where the solution $\bm u\in\mathbb{R}^p$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
+where the solution $\bm u\in\mathbb{R}^m$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
 Hence, the probability density function of $\bm{\xi}$ is $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
-The physical flux is given by $\bm{f}:\mathbb{R}^p\to\mathbb{R}^{d\times p}$.
+The physical flux is given by $\bm{f}:\mathbb{R}^m\to\mathbb{R}^{d\times m}$.
 To simplify notation, we assume that only the initial condition is random, i.e.
@@
 \end{align}
-where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^p$ are called moments.
-To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N\times p}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
+where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^m$ are called moments.
+To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N\times m}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
 In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.
-The solution ansatz \eqref{eq:SGClosure} is $L^2$-optimal, if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^p$.
+The solution ansatz \eqref{eq:SGClosure} is $L^2$-optimal, if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^m$.
 One can also use the ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
@@
 \begin{align*}
-\text{E}[\mathcal{U}(\bm{\hat u})] = \bm{\hat u}_0,\quad \text{Var}[\mathcal{U}(\bm{\hat u})] = \text{E}[\mathcal{U}(\bm{\hat u})^2] - \text{E}[\mathcal{U}(\bm{\hat u})]^2 = \left(\sum_{i = 1}^N \hat{u}_{\ell i}^2\right)_{\ell = 1,\cdots,p}.
+\text{E}[\mathcal{U}(\bm{\hat u})] = \bm{\hat u}_0,\quad \text{Var}[\mathcal{U}(\bm{\hat u})] = \text{E}[\mathcal{U}(\bm{\hat u})^2] - \text{E}[\mathcal{U}(\bm{\hat u})]^2 = \left(\sum_{i = 1}^N \hat{u}_{\ell i}^2\right)_{\ell = 1,\cdots,m}.
 \end{align*}
@@
 To ensure hyperbolicity, the IPM method uses a solution ansatz which minimizes a given entropy under a moment constraint instead of a polynomial expansion \eqref{eq:SGClosure}.
-For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, the IPM solution ansatz is given by
+For a given convex entropy $s:\mathbb{R}^m\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, the IPM solution ansatz is given by
 \begin{align}\label{eq:primalProblem}
@@
 Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstrained optimization problem.
-Defining the variables $\bm{\lambda}_i\in\mathbb{R}^p$ where $i$ is again a multi index, gives the unconstrained dual problem
+Defining the variables $\bm{\lambda}_i\in\mathbb{R}^m$ where $i$ is again a multi index, gives the unconstrained dual problem
 \begin{align}\label{eq:dualProblem}
- \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{N \times p}}
+ \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{N \times m}}
   \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{|i|\leq M}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
 \end{align}
-where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_i)_{|i|\leq M}\in \mathbb{R}^{N \times p}$ are called the dual variables.
+where $s_*:\mathbb{R}^m\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_i)_{|i|\leq M}\in \mathbb{R}^{N \times m}$ are called the dual variables.
 The solution to \eqref{eq:primalProblem} is then given by

### 2019-11-20 Paper1/one_shot_ipm.tex
continued

@@
 \end{align}
-which is why we have the mapping $\bm{\hat u}:\mathbb{R}^{N\times p}\to\mathbb{R}^{N\times p}$, $\bm{\hat u}(\bm{\hat \lambda}) = \bm{h}(\bm{\hat \lambda})$.
+which is why we have the mapping $\bm{\hat u}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{\hat u}(\bm{\hat \lambda}) = \bm{h}(\bm{\hat \lambda})$.
 Since the solution of the dual problem for a given moment vector is unique, this mapping is bijective and therefore we have an inverse function

### 2019-11-21 Paper1/adaptivity.tex
adaptivity m corrected

@@
 Now, we have
-$\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times p}\times\mathbb{R}^{N_{\ell}\times p}\to\mathbb{R}^{N_{\ell}\times p}$, given by
+$\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times m}\times\mathbb{R}^{N_{\ell}\times m}\to\mathbb{R}^{N_{\ell}\times m}$, given by
 \begin{align}\label{eq:dualIterationFunctionAd}
@@
 \end{align*}
-An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
+An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times m}\times \mathbb{R}^{N_{\ell_2'}\times m}\times \mathbb{R}^{N_{\ell_3'}\times m}\rightarrow \mathbb{R}^{N_{\ell}\times m}$ and given by
 \begin{align}\label{eq:adaptiveFVUpdate}

### 2019-11-21 Paper1/adaptivity.tex
changing variables

@@
 \While{\eqref{eq:tauCrit} is violated}
-\State $\bm{\lambda}_j^{(m+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n})$
-\State $m \leftarrow m+1$
+\State $\bm{\lambda}_j^{(l+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n})$
+\State $l \leftarrow l+1$
 \EndWhile
-\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(m)}$
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(l)}$
 \State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\hat \lambda}_j^{n+1}\right)$
@@
 Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualSteady} is fulfilled.
-For a given set of maximal refinement levels $\ell_m^*$ and a set of residuals $\varepsilon_m^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper accuracy level.
+For a given set of maximal refinement levels $\ell_l^*$ and a set of residuals $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper accuracy level.
 The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm \ref{alg:adosIPM}.
@@
 \State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
-\State $\ell_j^{n+1}\leftarrow \max\{\text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right),\ell_m^*\}$
+\State $\ell_j^{n+1}\leftarrow \max\{\text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right),\ell_l^*\}$
 \State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$
@@
 \State $n \leftarrow n+1$
-\If{\eqref{eq:residualSteady} fulfills the stopping criterion $\varepsilon_m^*$}
-\State $m \leftarrow m+1$
+\If{\eqref{eq:residualSteady} fulfills the stopping criterion $\varepsilon_l^*$}
+\State $l \leftarrow l+1$
 \EndIf

### 2019-11-21 Paper1/framework.tex
changing variables

@@
 \end{align*}
-If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N\times m}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
 \begin{align}\label{eq:IPMDiscretization}
@@
 \end{align}
-Note that the numerical flux requires evaluating the ansatz $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^s \to \mathbb{R}^s$,
+Note that the numerical flux requires evaluating the ansatz $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^p \to \mathbb{R}^p$,
 \begin{align*}
@@
 The root is usually determined by using Newton's method.
-For simplicity, let us define the full gradient of the Lagrangian to be $\nabla_{\bm{\lambda}}L(\bm{\lambda};\bm{\hat{u}})\in\mathbb{R}^{N\cdot p}$, i.e.
+For simplicity, let us define the full gradient of the Lagrangian to be $\nabla_{\bm{\lambda}}L(\bm{\lambda};\bm{\hat{u}})\in\mathbb{R}^{N\cdot m}$, i.e.
 we store all entries in a vector.
@@
 \end{align}
-where $\bm H\in\mathbb{R}^{p\cdot N\times m\cdot N}$ is the Hessian of \eqref{eq:L}, given by
+where $\bm H\in\mathbb{R}^{m \cdot  N\times m\cdot N}$ is the Hessian of \eqref{eq:L}, given by
 \begin{align*}
@@
 \begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{(m+1)}_j = \bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j}).
+\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j}).
 \end{align}
 The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
-$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(m)},\bm{\hat{u}_j^n})$.
+$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j^n})$.
 To obtain a finite number of iterations for the iteration in cell $j$, a stopping criterion 
 \begin{align}\label{eq:tauCrit}
-\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(m)};\bm{\hat{u}}_j^n) \right\Vert < \tau
+\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(l)};\bm{\hat{u}}_j^n) \right\Vert < \tau
 \end{align}
@@
 \begin{align}\label{eq:dualIteration}
-\bm{\lambda}_j^{(m+1)} = \bm{d}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n}).
+\bm{\lambda}_j^{(l+1)} = \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n}).
 \end{align}
@@
 \While{\eqref{eq:tauCrit} is violated}
-\State $\bm{\lambda}_j^{(m+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n})$
-\State $m \leftarrow m+1$
+\State $\bm{\lambda}_j^{(l+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n})$
+\State $l \leftarrow l+1$
 \EndWhile
-\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(m)}$
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(l)}$
 \EndFor

### 2019-11-21 Paper1/introduction.tex
changing variables

@@
 \end{subequations}
-where the solution $\bm u\in\mathbb{R}^m$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^s$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,s$.
-Hence, the probability density function of $\bm{\xi}$ is $f_{\Xi}(\bm\xi):=\prod_{i=1}^s f_{\Xi,i}(\xi_i)$.
+where the solution $\bm u\in\mathbb{R}^m$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^p$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,p$.
+Hence, the probability density function of $\bm{\xi}$ is $f_{\Xi}(\bm\xi):=\prod_{i=1}^p f_{\Xi,i}(\xi_i)$.
 The physical flux is given by $\bm{f}:\mathbb{R}^m\to\mathbb{R}^{d\times m}$.
@@
 \end{align*}
-where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_s$.
-To approximate quantities of interest (such as expectation value, variance or higher order moments), the solution is spanned with a set of polynomial basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_s)$ we have $|i| \leq M$.
+where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_p$.
+To approximate quantities of interest (such as expectation value, variance or higher order moments), the solution is spanned with a set of polynomial basis functions $\varphi_{i}:\Theta\to\mathbb{R}$ such that for the multi-index $i = (i_1,\cdots,i_p)$ we have $|i| \leq M$.
 Note that this yields
@@
 N:=\begin{pmatrix}
-M+s \\ s
+M+p \\ p
 \end{pmatrix}
 \end{align}
-basis functions when defining $|i|:=\sum_{n = 1}^s |i_n|$.
+basis functions when defining $|i|:=\sum_{n = 1}^p |i_n|$.
 Commonly, these functions are chosen to be orthonormal polynomials \cite{wiener1938homogeneous} with respect to the probability function, i.e.
-$\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^s\delta_{i_nj_n}$.
+$\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^p\delta_{i_nj_n}$.
 The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates the solution by
@@
 Hence, the moments can be computed by running a given deterministic solver for the original problem at each quadrature point.
-To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^s)$ quadrature points to integrate polynomials of maximal degree $M$ exactly, sparse grids are designed to integrate polynomials of total degree $M$, for which they only require $O(M(\log_2(M)^{s-1}))$ quadrature points.
+To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^p)$ quadrature points to integrate polynomials of maximal degree $M$ exactly, sparse grids are designed to integrate polynomials of total degree $M$, for which they only require $O(M(\log_2(M)^{p-1}))$ quadrature points.
 

### 2019-11-21 Paper1/results.tex
updated plots for ipm aliaising study and added reference solution for the pipe testcase

@@
 		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sc100_ERho}.png}
+				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_ERho.png}
 		\label{fig:referenceSolutionsub1}
@@
 		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sc100_VarRho}.png}
+				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_VarRho.png}
 		\label{fig:referenceSolutionsub2}
@@
 		\centering
-				\includegraphics[scale=0.544]{figs/StudyQuadrature/{L2_error_E[rho]}.pdf}
+				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_E[rho]_ipm.pdf}
 		\caption{}
-		\label{fig:ErrorDifferentQuadA}
+		\label{fig:ErrorDifferentQuadMeanIPM}
 	\end{subfigure}%
@@
 		\centering
-				\includegraphics[scale=0.544]{figs/StudyQuadrature/{L2_error_Var[rho]}.pdf}
+				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_Var[rho]_ipm.pdf}
 		\caption{}
-		\label{fig:ErrorDifferentQuadB}
+		\label{fig:ErrorDifferentQuadVarIPM}
 	\end{subfigure}
-	\caption{Relative L$^2$-error with different quadrature levels.
-The subscript denotes the moment order, the superscript denotes the Clenshaw-Curtis quadrature level.\frama{Sollen wir einige Linien hier raus nehmen, damit es \"ubersichtlicher bleibt?}}
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+		\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_E[rho]_mixed.pdf}
+		\caption{}
+		\label{fig:ErrorDifferentQuadMeanMixed}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+		\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_Var[rho]_mixed.pdf}
+		\caption{}
+		\label{fig:ErrorDifferentQuadVarMixed}
+	\end{subfigure}
+	\caption{Relative L$^2$-error with different quadrature levels for solely IPM (Figure \ref{fig:ErrorDifferentQuadMeanIPM} and \ref{fig:ErrorDifferentQuadVarIPM}) and in comparison with SC (Figure \ref{fig:ErrorDifferentQuadMeanMixed} and \ref{fig:ErrorDifferentQuadVarMixed}).
+The subscript denotes the moment order, the superscript denotes the number of Clenshaw-Curtis quadrature points.}
 	\label{fig:ErrorDifferentQuad}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rho]}.pdf}
+				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho].pdf}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rho]}.pdf}
+				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho].pdf}
 		\label{fig:sub2}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_E[rhoE]}.pdf}
+				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rhoE].pdf}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/{L2_error_Var[rhoE]}.pdf}
+				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rhoE].pdf}
 		\label{fig:sub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_ERho.png}
 		\label{fig:sub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_VarRho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_VarRho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
 		\label{fig:sub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_VarRho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho.png}
 		\label{fig:sub1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho.png}
 		\label{fig:sub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel}.png}
+		\includegraphics[scale=0.2]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel.png}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_E[rho]}.pdf}
+				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_E[rho].pdf}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_Var[rho]}.pdf}
+				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_Var[rho].pdf}
 		\label{fig:sub2}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_E[rhoE]}.pdf}
+				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_E[rhoE].pdf}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/{L2_error_Var[rhoE]}.pdf}
+				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_Var[rhoE].pdf}
 		\label{fig:sub2}
@@
 		\centering
-				\includegraphics[scale=0.2]{figs/Euler2D/{sc50UncertainMaSigma0-025PSigma1000-0Dim2_ERho}.png}
+				\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_ERho.png}
 		\label{fig:referenceSolutionsub1}
@@
 		\centering
-				\includegraphics[scale=0.2]{figs/Euler2D/{sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho}.png}
+				\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho.png}
 		\label{fig:referenceSolutionsub2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_ERho}.png}
+		\includegraphics[scale=0.2]{figs/Euler2D/retadosIPM9-4MaP_4_ERho.png}
 		\caption{}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_VarRho}.png}
+		\includegraphics[scale=0.2]{figs/Euler2D/retadosIPM9-4MaP_4_VarRho.png}
 		\caption{}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/{retadosIPM9-4MaP_4_refinementLevel}.png}
+		\includegraphics[scale=0.2]{figs/Euler2D/retadosIPM9-4MaP_4_refinementLevel.png}
 		\caption{}
@@
 \subsection{Pipe Testcase}
+\begin{figure}[H]
+	\centering
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
+		\label{fig:referenceSolutionsPipe1}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
+		\label{fig:referenceSolutionsPipe2}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\label{fig:referenceSolution3D}
+\end{figure}
 \comment{TODO: Pipe testcase}

### 2019-11-21 code/src/burgers.cpp
hard coded for 2d uniform

@@ -5,7 +5,7 @@ Burgers::Burgers( Settings* settings ) : Problem( settings ) {
     _settings->SetNStates( _nStates );
 
     // for the one dimensional case an exact solution is specified
-    if( _settings->GetNDimXi() == 1 ) {
+    if( _settings->GetNDimXi() == 1 || true ) {
         _settings->SetExactSolution( true );
     }
 
@@ -48,7 +48,7 @@ Matrix Burgers::F( const Matrix& u ) { return 0.5 * pow( u, 2 ); }
 Vector Burgers::IC( const Vector& x, const Vector& xi ) {
     Vector y( _nStates );
     _sigma = _settings->GetSigma();
-    if( xi.size() == 1 ) {
+    if( xi.size() == 1 || true ) {
         _x0 = 0.5;
         _x1 = 1.5;
         _uL = 12.0;

### 2019-11-21 code/src/momentsolver.cpp

@@ -371,7 +371,7 @@ Matrix MomentSolver::WriteMeanAndVar( const VectorU& refinementLevel, double t, 
             for( unsigned i = 0; i < _nStates; ++i ) {
                 double wf = phiTildeWf( k, k );    // dirty fix for pSC, works only for uniform distributions
                 auto w    = _closure->GetWGridAtRef( refinementLevel[j] );
-                meanAndVar( i, j ) += tmp[i] * w[k] * 0.5;
+                meanAndVar( i, j ) += tmp[i] * w[k] * 0.25;    // hard code pdf: 2d uniform is 0.25
                 // if( j == 4 ) {
                 //  std::cout << "uQ = " << tmp[0] << std::endl;
                 // std::cout << "wf = " << wf << std::endl;
@@ -385,7 +385,7 @@ Matrix MomentSolver::WriteMeanAndVar( const VectorU& refinementLevel, double t, 
             for( unsigned i = 0; i < _nStates; ++i ) {
                 double wf = phiTildeWf( k, k );    // dirty fix for pSC, works only for uniform distributions + pSC
                 auto w    = _closure->GetWGridAtRef( refinementLevel[j] );
-                meanAndVar( i + _nStates, j ) += pow( tmp[i] - meanAndVar( i, j ), 2 ) * w[k] * 0.5;
+                meanAndVar( i + _nStates, j ) += pow( tmp[i] - meanAndVar( i, j ), 2 ) * w[k] * 0.25;    // hard code pdf: 2d uniform is 0.25
             }
         }
     }

### 2019-11-22 Paper1/introduction.tex
new abstract

@@
 \item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM shows improved variance approximations.
-\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for one spatial dimension) to reach the same accuracy level.
+\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for a one-dimension uncertainty) to reach the same accuracy level.
 \item Using sparse grids for the IPM discretization when the space of uncertainty is multi-dimensional, the number of quadrature points needed to guarantee sufficient regularity of the Hessian matrix is significantly increased.

### 2019-11-22 Paper1/main.tex
new abstract

@@
 Non-intrusive methods allow the usage of a given deterministic solver in a black-box manner, while being embarrassingly parallel.
-Intrusive methods on the other hand do not suffer from aliasing effects and provide a more general framework for adaptive refinement techniques.
-Since the moment system solved by intrusive methods is not necessarily hyperbolic, the Intrusive Polynomial Moment (IPM) method has been introduced.
-IPM maintains hyperbolicity, but comes at the cost of solving an additional problem in every spatial cell and every time step.
-In this work, we propose acceleration techniques for intrusive methods.
-When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from PDE-constrained optimization.
-Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while still guaranteeing local convergence.
-Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
-We demonstrate the effectiveness of the proposed strategies for multi-dimensional uncertainty spaces by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
-\frama{Sollen die Erkenntnisse vom Vergleich mit SC schon hier erw\"ahnt werden?}
+However, abandoning the idea of intrusively modifying a given solver takes away the ability to use several numerical acceleration tools, which have an intrusive nature.
+One example is the use of adaptivity for which intrusive methods provide a convenient framework.
+Moreover, intrusive methods are expected to reach a given accuracy with a smaller number of unknowns than non-intrusive techniques.
+This effect is amplified in settings with high dimensional uncertainty where the number of polynomials is smaller than the number of points in a sparse grid to achieve a certain total degree.
+A downside of intrusive methods is however the need to guarantee hyperbolicity of the moment system as well as numerical costs: Compared to stochastic-Galerkin (SG), the Intrusive Polynomial Moment (IPM) method maintains hyperbolicity, but requires solving an optimization problem in every spatial cell and every time step.
+\\
+In this work, we propose acceleration techniques for intrusive methods and study their advantages and shortcomings compared to the non-intrusive stochastic-Collocation method.
+When solving steady problems with IPM, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from PDE-constrained optimization.
+Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while preserving local convergence.
+Additionally, we propose an adaptive implementation and an efficient parallelization of the IPM method.
+The effectiveness of the proposed strategies is demonstrated for multi-dimensional uncertainties in fluid dynamics applications.
+In our numerical experiments, we observe the expected behavior of requiring a smaller number of unknowns to achieve a given accuracy when using intrusive methods.
+Using the proposed acceleration techniques, our implementation reaches a given accuracy faster than stochastic-Collocation.
 \end{abstract}

### 2019-11-22 Paper1/results.tex
updated errorEuler and StudyQuadrature plots; removed errorEuler energy plots and updated the text accordingly

@@
 		\centering
-				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_E[rho]_ipm.pdf}
+				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_E[rho]_new.pdf}
 		\caption{}
-		\label{fig:ErrorDifferentQuadMeanIPM}
+		\label{fig:ErrorDifferentQuadA}
 	\end{subfigure}%
@@
 		\centering
-				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_Var[rho]_ipm.pdf}
+				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_Var[rho]_new.pdf}
 		\caption{}
-		\label{fig:ErrorDifferentQuadVarIPM}
+		\label{fig:ErrorDifferentQuadB}
 	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-		\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_E[rho]_mixed.pdf}
-		\caption{}
-		\label{fig:ErrorDifferentQuadMeanMixed}
-	\end{subfigure}%
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-		\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_Var[rho]_mixed.pdf}
-		\caption{}
-		\label{fig:ErrorDifferentQuadVarMixed}
-	\end{subfigure}
-	\caption{Relative L$^2$-error with different quadrature levels for solely IPM (Figure \ref{fig:ErrorDifferentQuadMeanIPM} and \ref{fig:ErrorDifferentQuadVarIPM}) and in comparison with SC (Figure \ref{fig:ErrorDifferentQuadMeanMixed} and \ref{fig:ErrorDifferentQuadVarMixed}).
+	\caption{Relative L$^2$-error with different quadrature levels for IPM in comparison with SC.
 The subscript denotes the moment order, the superscript denotes the number of Clenshaw-Curtis quadrature points.}
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho].pdf}
-		\label{fig:sub1}
+		\label{fig:sub31}
+		\caption{}
 	\end{subfigure}%
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho].pdf}
-		\label{fig:sub2}
+		\label{fig:sub32}
+		\caption{}
 	\end{subfigure}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rhoE].pdf}
-		\label{fig:sub1}
+				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
+		\label{fig:sub33}
+		\caption{}
 	\end{subfigure}%
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rhoE].pdf}
-		\label{fig:sub2}
+				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
+		\label{fig:sub34}
+		\caption{}
 	\end{subfigure}
-	\caption{Relative L$^2$-error with 5 MPI threads for density and energy.\frama{Eventuell diese Plots aufteilen? Oder sind die Ergebnisse verst\"andlich?}}
+	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
+Computed with 5 MPI threads.}
 	\label{fig:L2ErrorSolution}

### 2019-11-22 Paper1/introduction.tex
introduction adaptivity discussed

@@
 
-In this paper, we propose acceleration techniques for intrusive methods and compare them against stochastic-Collocation. For steady and unsteady problems, we use adaptivity:
+In this paper, we propose acceleration techniques for intrusive methods and compare them against stochastic-Collocation.
+For steady and unsteady problems, we use adaptivity, for which intrusive methods provide a convenient framework: %Since complex structures in the uncertain domain tend to arise in small portions of the spatial mesh, our aim is to locally increase the accuracy of the stochastic discretization in region that show a complex structure in the random domain, while choosing a low order method in the remainder:
 \begin{itemize}
-\item The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not).
+\item Since complex structures in the uncertain domain tend to arise in small portions of the spatial mesh, our aim is to locally increase the accuracy of the stochastic discretization in region that show a complex structure in the random domain, while choosing a low order method in the remainder.
+Such an adaptive treatment cannot be realized with non--intrusive methods, since one needs to break up the black-box approach.
 To guarantee an efficient implementation, we propose an adaptive discretization strategy for IPM.

### 2019-11-22 Paper1/introduction.tex
intro fair comparison

@@
 The IPM and SG calculations use a semi-intrusive numerical method, meaning that the discretization allows recycling a given deterministic code to generate the IPM solver.
-While facilitating the task of implementing general intrusive methods, this framework allows a fair comparison of intrusive and non-intrusive methods, since the intrusive method can be based on the same deterministic code as the code used in a black-box fashion for stochastic-Collocation.
+While facilitating the task of implementing general intrusive methods, this framework reduces the number of operations required to compute numerical fluxes.
+Also, it provides the ability to base the intrusive method on the same deterministic solver as the implementation used in a black-box fashion for stochastic-Collocation, which we believe allows a fair comparison between intrusive and non-intrusive methods.
+The code is publicly available to allow reproducibility \comment{[RefCode]}.
 

### 2019-11-22 Paper1/one_shot_ipm.tex
local convergence comment

@@
 In the following, we will show that this iteration converges, if the chosen initial condition is sufficiently close to the steady state solution.
-For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence: \frama{Nico meinte wir sollten hier in die Voraussetzung packen, dass wir nah genug an der richtigen Loesung sind (Einzugsbereich des Newton) um das Wort 'local' zu vermeiden.
-Manchmal stören sich die Reviewer daran.}
+For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence.
+Hence, we preserve the convergence results of the original IPM method, which uses Newton's method, i.e.
+converges locally as well.
 \begin{theorem}\label{th:Contractive}

### 2019-11-22 Paper1/adaptivity.tex

@@
 
-In the following, we discuss the building blocks of the IPM method for accuracy levels $\ell = 1,\cdots,N_{\text{ad}}$.
-At a given level $\ell$, the total degree of the basis function is given by $M_{\ell}$ with a corresponding number of moments $N_{\ell}$.
+In the following, we discuss the building blocks of the IPM method for refinement levels $\ell = 1,\cdots,N_{\text{ad}}$, where level $1$ uses the coarsest discretization and level $N_{\text{ad}}$ uses the finest discretization of the uncertain domain.
+At a given refinement level $\ell$, the total degree of the basis function is given by $M_{\ell}$ with a corresponding number of moments $N_{\ell}$.
 The number of quadrature points at level $\ell$ is denoted by $Q_{\ell}$.
-To determine the accuracy level of a given moment vector $\bm{\hat u}$ we choose techniques used in discontinuous Galerkin (DG) methods.
+To determine the refinement level of a given moment vector $\bm{\hat u}$ we choose techniques used in discontinuous Galerkin (DG) methods.
 Adaptivity is a common strategy to accelerate this class of methods and several indicators to determine the smoothness of the solution exist.
 We make use of a strategy from \cite{persson2006sub}, which uses the highest degree moments as indicator.
-Translating the idea of the discontinuity sensor used in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at level $\ell$ as
+Translating the idea of the discontinuity sensor used in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at refinement level $\ell$ as
 \begin{align*}
@@
 Here $\delta_{\pm}$ are user determined parameters.
-If the indicator is smaller than $\delta_-$, the refinement level is decreased, if it lies above $\delta_+$, it is increased.
+If the indicator is smaller than $\delta_-$, the refinement level is decreased by a value of one, if it lies above $\delta_+$, it is increased by one.
 
@@
 \begin{align*}
-\bm{H}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi}_{\ell})\otimes\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{T}.
+\bm{H}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{s} (\bm{\lambda}^T\bm{\varphi}_{\ell})\otimes\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{T}.
 \end{align*}
@@
 Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualSteady} is fulfilled.
-For a given set of maximal refinement levels $\ell_l^*$ and a set of residuals $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper accuracy level.
-The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm \ref{alg:adosIPM}.
+For a given set of maximal refinement levels $\ell_l^*$ and a set of residuals $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper refinement level.
+The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm~\ref{alg:adosIPM}.
 \begin{algorithm}[H]

### 2019-11-22 Paper1/adaptivity.tex
fluid air

@@
 Recall that the convergence to an admissible steady state solution is expensive and a high accuracy and desirable solution properties are only required at the end of this iteration process.
-Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualSteady} is fulfilled.
-For a given set of maximal refinement levels $\ell_l^*$ and a set of residuals $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper refinement level.
+Hence, we propose to iteratively increase the maximal refinement level whenever the residual \eqref{eq:residualSteady} lies below a certain tolerance level.
+For a given set of maximal refinement levels $\ell_l^*$ and a set of tolerance levels $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper refinement level.
 The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm~\ref{alg:adosIPM}.

### 2019-11-22 Paper1/main.tex

--- 2019-11-20 Paper1/main.tex
@@
 Non-intrusive methods allow the usage of a given deterministic solver in a black-box manner, while being embarrassingly parallel.
-Intrusive methods on the other hand do not suffer from aliasing effects and provide a more general framework for adaptive refinement techniques.
-Since the moment system solved by intrusive methods is not necessarily hyperbolic, the Intrusive Polynomial Moment (IPM) method has been introduced.
-IPM maintains hyperbolicity, but comes at the cost of solving an additional problem in every spatial cell and every time step.
-In this work, we propose acceleration techniques for intrusive methods.
-When solving steady problems, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from PDE-constrained optimization.
-Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while still guaranteeing local convergence.
-Additionally, we propose an adaptive implementation of the IPM method, which further reduces numerical costs.
-We demonstrate the effectiveness of the proposed strategies for multi-dimensional uncertainty spaces by comparing results of the uncertain NACA0012 testcase as well as a bent shock tube experiment with non-intrusive methods.
-\frama{Sollen die Erkenntnisse vom Vergleich mit SC schon hier erw\"ahnt werden?}
+However, not intrusively modifying a given solver takes away the ability to use several inherently intrusive numerical acceleration tools.
+Moreover, intrusive methods are expected to reach a given accuracy with a smaller number of unknowns compared to non-intrusive techniques.
+This effect is amplified in settings with high dimensional uncertainty.
+A downside of intrusive methods is however the need to guarantee hyperbolicity of the resulting moment system.
+In contrast to stochastic-Galerkin (SG), the Intrusive Polynomial Moment (IPM) method is able to maintain hyperbolicity at the cost of solving an optimization problem in every spatial cell and every time step.\\
+In this work, we propose several acceleration techniques for intrusive methods and study their advantages and shortcomings compared to the non-intrusive stochastic-Collocation method.
+When solving steady problems with IPM, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by using concepts from PDE-constrained optimization.
+Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while preserving local convergence.
+Additionally, we propose an adaptive implementation and efficient parallelization strategy of the IPM method.
+The effectiveness of the proposed adaptations is demonstrated for multi-dimensional uncertainties in fluid dynamics applications, resulting in the observation of requiring a smaller number of unknowns to achieve a given accuracy when using intrusive methods.
+Furthermore, using the proposed acceleration techniques, our implementation reaches a given accuracy faster than stochastic-Collocation in a fair comparison as both methods are using the same codebase.
 \end{abstract}

### 2019-11-22 Paper1/adaptivity.tex
tolerance level

--- 2019-11-22 Paper1/adaptivity.tex
@@
 Recall that the convergence to an admissible steady state solution is expensive and a high accuracy and desirable solution properties are only required at the end of this iteration process.
-Hence, we propose to iteratively increase the maximal refinement level whenever a certain residual \eqref{eq:residualSteady} is fulfilled.
-For a given set of maximal refinement levels $\ell_l^*$ and a set of residuals $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper refinement level.
+Hence, we propose to iteratively increase the maximal refinement level whenever the residual \eqref{eq:residualSteady} lies below a certain tolerance $\varepsilon$.
+For a given set of maximal refinement levels $\ell_l^*$ and a set of tolerances $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper refinement level.
 The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm~\ref{alg:adosIPM}.
@@
 \State $n \leftarrow n+1$
-\If{\eqref{eq:residualSteady} fulfills the stopping criterion $\varepsilon_l^*$}
+\If{the residual \eqref{eq:residualSteady} lies below $\varepsilon_l^*$}
 \State $l \leftarrow l+1$

### 2019-11-22 Paper1/main.tex
small modification of abstract

@@
 Non-intrusive methods allow the usage of a given deterministic solver in a black-box manner, while being embarrassingly parallel.
-However, not intrusively modifying a given solver takes away the ability to use several inherently intrusive numerical acceleration tools.
+However, avoiding intrusive modifications of a given solver takes away the ability to use several inherently intrusive numerical acceleration tools.
 Moreover, intrusive methods are expected to reach a given accuracy with a smaller number of unknowns compared to non-intrusive techniques.
@@
 The effectiveness of the proposed adaptations is demonstrated for multi-dimensional uncertainties in fluid dynamics applications, resulting in the observation of requiring a smaller number of unknowns to achieve a given accuracy when using intrusive methods.
-Furthermore, using the proposed acceleration techniques, our implementation reaches a given accuracy faster than stochastic-Collocation in a fair comparison as both methods are using the same codebase.
+Furthermore, using the proposed acceleration techniques, our implementation reaches a given accuracy faster than stochastic-Collocation.
 \end{abstract}

### 2019-11-22 Paper1/results.tex
introduction adaptivity discussed
text expectation values
fluid air
results SC

@@
 \end{align*}
-Since the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$.
+Here, the heat capacity ratio $\gamma$ is chosen to be $1.4$.
 The spatial mesh discretizes the flow domain around the airfoil.
@@
 In the following, we investigate the L$^2$-error of the variance and the expectation value.
-The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
+The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
 \begin{align*}
-\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}.
+\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
 \end{align*}
@@
 \begin{align*}
-\frac{\Vert E[\bm u_{\Delta}] - E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert E[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert Var[\bm u_{\Delta}] - Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert Var[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}.
+\frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
 \end{align*}
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho].pdf}
+		\caption{}
 		\label{fig:sub31}
-		\caption{}
 	\end{subfigure}%
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho].pdf}
+		\caption{}
 		\label{fig:sub32}
-		\caption{}
 	\end{subfigure}
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
+		\caption{}
 		\label{fig:sub33}
-		\caption{}
 	\end{subfigure}%
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
+		\caption{}
 		\label{fig:sub34}
-		\caption{}
 	\end{subfigure}
 	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-Computed with 5 MPI threads.}
+All intrusive methods converge to a residual $\varepsilon=5e-6$ whereas all non-intrusive methods converge to a residual $\varepsilon=1e-7$.
+All computations are performed with with 5 MPI threads.}
 	\label{fig:L2ErrorSolution}
@@
 	\end{subfigure}
-	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the refernce solution shown in Figure \ref{fig:referenceSolution}.}
+	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the reference solution shown in Figure \ref{fig:referenceSolution}.}
 	\label{fig:ERhoVarRho5}
@@
 A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}.
-To denote the quadrature level, we use a superscript and the moment order is denoted by a subscript.
+To denote the number of quadrature points, we use a superscript and the moment order is denoted by a subscript.
 We observe the following:
 \begin{itemize}
-\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ (the subscript indicates the polynomial order, the superscript indicates \comment{the number of quadrature points}) it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
+\item When for example comparing the error obtained with IPM$_8^9$ and IPM$_8^{17}$ (i.e.
+the polynomial order is $8$, meaning that $9$ moments are used and the computation is done using $9$ and $17$ quadrature points) it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
 Hence, the aliasing effects that result from an inaccurate quadrature rule in the numerical flux and not the truncation error of the moment system dominate the accuracy level.
 \item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy.
-Note that this holds for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$.
+We observe this behavior for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$.
 Hence, the variance is more sensitive to aliasing errors.
-\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points (level 3) stagnates at the error level of SC$^9$, i.e.
+\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, i.e.
 the aliasing error heavily affects the accuracy.
 This behavior becomes more dominant when looking at the variance error in Figure~\ref{fig:ErrorDifferentQuadB}.
-Here, IPM$_8^4$ yields a significantly improved result compared to IPM$_8^3$.
-Again, the IPM$_8^3$ result stagnates at the SC$^9$ accuracy level, which can be reached with IPM when only using four moments (combined with a sufficiently accurate quadrature level).
+Here, IPM$_8^{17}$ yields a significantly improved result compared to IPM$_8^9$.
+Again, the IPM$_8^9$ result stagnates at the SC$^9$ accuracy level, which can be reached with IPM when only using four moments (combined with a sufficiently accurate quadrature level).
 Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC.
-This result can be observed throughout the results of this work.
+This result can be observed throughout the numerical experiments of this work.
 Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
-However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
+However note, that since the numerical flux evaluation used by IPM requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
 \end{itemize}
-Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order 9.
+
+Let us now compare results obtained with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order $9$.
 Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well.
-Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
 All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
 3 to 10 moments).
-Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
+Order $2$ uses a $5$ quadrature points, orders $3$ to $6$ use $9$ quadrature points and orders $8$ and $9$ use $17$ quadrature points.
 When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the adaptive methods decrease the truncation order, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
-The remaining methods have been computed with a level 4 quadrature rule.
-
-All intrusive methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
-For every quadrature point, SC iterates the solution until the density reaches a certain residual level $\varepsilon$.
-We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following Table \ref{tab:resSC17} for the SC$^{17}$ solution.
-\frama{Sollen wir hier die Ergebnisse zu Frage 1 im zweiten Dokument einbauen?}
-\begin{center}
-\begin{table}
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
-	\hline
-	residual $\rho$ & 5e-06 & 4e-06 & 3e-06 & 2e-06 & 1e-06 & 9e-07 & 8e-07 & 7e-07 & 6e-07 & 5e-07  \\ 
-	\hline
-	error E[$\rho$] in e-03 & 1.506 & 1.464 & 1.438 & 1.401 & 1.368 & 1.360 & 1.359 & 1.358 & 1.358 & 1.357  \\ 
-	\hline
-	error Var[$\rho$] in e-03 & 1.339 & 1.332 & 1.292 & 1.248 & 1.217 & 1.215 & 1.214 & 1.213 & 1.213 & 1.213  \\ 
-	\hline
-\end{tabular}
-}
-\caption{Error values at various residual levels when using SC.}
-\label{tab:resSC17}
-\end{table}
-
-\end{center}
-Hence, we can deduce that SC requires a smaller residual to converge to a steady state solution.
-Therefore, the results for SC$^{17}$ will be presented for a residual of $\varepsilon = 5\cdot 10^{-6}$ as well as a reduced residual of $\varepsilon = 10^{-6}$.
-Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
-%One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
+The remaining methods have been computed with $17$ quadrature points.
+The iteration in pseudo-time is performed until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
 
@@
 The standard SG method however preserves positivity of mass, energy and pressure.
-The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
+The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation values and in Figure~\ref{fig:sub32}.
 When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
@@
 \item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
-\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.
-Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
 \end{itemize}
@@
 \end{itemize}
+
+Let us now compare the results from intrusive methods with those of SC.
+For every quadrature point, SC iterates the solution until the density density reaches lies below a threshold of $\varepsilon = 1e^{-7}$.
+Recording the error of intrusive methods during the iteration is straight forward.
+To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs.
+The collocation runtimes that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D} are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error.
+Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e.
+$5,9$ and $17$ quadrature points.
+The comparison of intrusive methods with non-intrusive methods shows the following:
+\begin{itemize}
+\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1e^{-7}$ compared to $\varepsilon = 5e^{-6}$ for intrusive methods).
+\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.
+Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
+\end{itemize}
+
+Hence, we can deduce that SC requires a smaller residual to converge to a steady state solution.
+Therefore, the results for SC$^{17}$ will be presented for a residual of $\varepsilon = 5\cdot 10^{-6}$ as well as a reduced residual of $\varepsilon = 10^{-6}$.
+Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
+%One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
 

### 2019-11-22 Paper1/results.tex
minor change

@@
 The standard SG method however preserves positivity of mass, energy and pressure.
-The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation values and in Figure~\ref{fig:sub32}.
+The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance.
 When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:

### 2019-11-25 Paper1/introduction.tex
paper further comments from Martin

@@
 
-When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
+When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered,poette2019contribution} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
 These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through either the spatial numerical method \cite{offner2017stability} or by filters \cite{kusch2018filtered}.
@@
 they are not the correct steady state solution, we propose to not fully converge the dual iteration, which solves the IPM optimization problem.
-Consequently, the entropic expansion coefficients and the moments are converged simultaneously to their steady state, which is similar to the idea of one-shot optimization in shape optimization \cite{hazra2005aerodynamic}.
+Consequently, the entropic expansion coefficients and the moments are converged simultaneously to their steady state, which is similar to the idea of One-Shot optimization in shape optimization \cite{hazra2005aerodynamic}.
 \end{itemize}

### 2019-11-25 Paper1/results.tex
paper further comments from Martin

@@
 The spatial mesh discretizes the flow domain around the airfoil.
-At the airfoil boundary $\Gamma_{0}$, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal.
+At the airfoil boundary, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal.
 At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of $273.15$ K.
@@
 the polynomial order is $8$, meaning that $9$ moments are used and the computation is done using $9$ and $17$ quadrature points) it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
-Hence, the aliasing effects that result from an inaccurate quadrature rule in the numerical flux and not the truncation error of the moment system dominate the accuracy level.
+Hence, the accuracy level is dominated by aliasing effects that result from an inaccurate quadrature rule in the numerical flux and not the truncation error of the moment system.
 \item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy.
@@
 Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
-However note, that since the numerical flux evaluation used by IPM requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
+%However note, that since the numerical flux evaluation used by IPM requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
 \end{itemize}
@@
 3 to 10 moments).
-Order $2$ uses a $5$ quadrature points, orders $3$ to $6$ use $9$ quadrature points and orders $8$ and $9$ use $17$ quadrature points.
+Order $2$ uses $5$ quadrature points, orders $3$ to $6$ use $9$ quadrature points and orders $8$ and $9$ use $17$ quadrature points.
 When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the adaptive methods decrease the truncation order, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
@@
 To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order.
-The residual values are then determined to be $6e-05,3e-05,2.2e-5$ and $2e-05$.
+The residual values are then determined to be $6\cdot 10^{-5},3\cdot 10^{-5},2.2\cdot 10^{-5}$ and $2\cdot 10^{-5}$.
 The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$.
-Strategy 1 is depicted in red, Strategy 2 is depicted in green.
+Strategy 1 is depicted in purple, Strategy 2 is depicted in yellow.
 It can be seen that both approaches reach the IPM$_9$ error for the same run time.
@@
 Let us now compare the results from intrusive methods with those of SC.
-For every quadrature point, SC iterates the solution until the density density reaches lies below a threshold of $\varepsilon = 1e^{-7}$.
+For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1e^{-7}$.
 Recording the error of intrusive methods during the iteration is straight forward.
@@
 
-Hence, we can deduce that SC requires a smaller residual to converge to a steady state solution.
-Therefore, the results for SC$^{17}$ will be presented for a residual of $\varepsilon = 5\cdot 10^{-6}$ as well as a reduced residual of $\varepsilon = 10^{-6}$.
-Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
-%One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
-
 Let us finally take a look at the expectation value and variance computed with different methods.
@@
 \item All methods yield non-physical step-like profiles of the expectation value and variance along the airfoil.
+This effect can be observed in various settings \cite{le2004uncertainty,kusch2018filtered,poette2019contribution,barth2013non,dwight2013adaptive} and stems from Gibb's phenomena in the polynomial description of the uncertainty, either in the gPC polynomials of intrusive methods or the Lagrange polynomials used in collocation techniques.
 \item The jump position of the intrusive solution profiles capture the exact behavior more accurately.
@@
 This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased.
-This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy an effective approach, especially for problems with high uncertain dimension.
+This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy to be an effective approach, especially for problems with high uncertain dimension.
 \begin{figure}[h!]

### 2019-11-25 Paper1/summary_outlook.tex
paper further comments from Martin

@@
 The effects of the proposed techniques have been demonstrated by comparing results obtained with IPM as well as SG against SC.
-In our test cases, the intrusive methods yield the same error as SC for a reduced runtime, especially since intrusive methods require less unknowns to achieve a certain accuracy due to aliasing errors.
+In our test cases, the intrusive methods yield the same error level as SC for a reduced runtime, especially since intrusive methods require less unknowns to achieve a certain accuracy due to aliasing errors.
 In higher-dimensional problems, this effect is amplified since the number of unknowns to achieve a certain total degree is asymptotically smaller than the number of quadrature points in a tensorized or sparse grid.

### 2019-11-25 Paper1/results.tex
added mesh plots and log var plot for euler 2d

--- 2019-11-22 Paper1/results.tex
@@
 \end{align*}
-Here, the heat capacity ratio $\gamma$ is chosen to be $1.4$.
+Since the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$.
 The spatial mesh discretizes the flow domain around the airfoil.
@@
 \centering
-	\begin{subfigure}{0.5\linewidth}
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 	\end{subfigure}%	
-	\begin{subfigure}{0.5\linewidth}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/Euler1DPlots5/naca_mesh.png}
+		\label{fig:referenceSolutionsMesh}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and mesh used in the computation of all presented methods.}
 	\label{fig:referenceSolution}
@@
 In the following, we investigate the L$^2$-error of the variance and the expectation value.
-The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
+The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ denotes the cell average in spatial cell $j$ is denoted by
 \begin{align*}
-\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
+\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}.
 \end{align*}
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho].pdf}
+		\label{fig:sub31}
 		\caption{}
-		\label{fig:sub31}
 	\end{subfigure}%
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho].pdf}
+		\label{fig:sub32}
 		\caption{}
-		\label{fig:sub32}
 	\end{subfigure}
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
+		\label{fig:sub33}
 		\caption{}
-		\label{fig:sub33}
 	\end{subfigure}%
@@
 				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
+		\label{fig:sub34}
 		\caption{}
-		\label{fig:sub34}
 	\end{subfigure}
 	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-All intrusive methods converge to a residual $\varepsilon=5e-6$ whereas all non-intrusive methods converge to a residual $\varepsilon=1e-7$.
-All computations are performed with with 5 MPI threads.}
+Computed with 5 MPI threads.}
 	\label{fig:L2ErrorSolution}
@@
 	\end{subfigure}
-	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the reference solution shown in Figure \ref{fig:referenceSolution}.}
+	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the refernce solution shown in Figure \ref{fig:referenceSolution}.}
 	\label{fig:ERhoVarRho5}
@@
 A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}.
-To denote the number of quadrature points, we use a superscript and the moment order is denoted by a subscript.
+To denote the quadrature level, we use a superscript and the moment order is denoted by a subscript.
 We observe the following:
 \begin{itemize}
-\item When for example comparing the error obtained with IPM$_8^9$ and IPM$_8^{17}$ (i.e.
-the polynomial order is $8$, meaning that $9$ moments are used and the computation is done using $9$ and $17$ quadrature points) it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
+\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ (the subscript indicates the polynomial order, the superscript indicates \comment{the number of quadrature points}) it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate.
 Hence, the aliasing effects that result from an inaccurate quadrature rule in the numerical flux and not the truncation error of the moment system dominate the accuracy level.
 \item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy.
-We observe this behavior for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$.
+Note that this holds for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$.
 Hence, the variance is more sensitive to aliasing errors.
-\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, i.e.
+\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points (level 3) stagnates at the error level of SC$^9$, i.e.
 the aliasing error heavily affects the accuracy.
 This behavior becomes more dominant when looking at the variance error in Figure~\ref{fig:ErrorDifferentQuadB}.
-Here, IPM$_8^{17}$ yields a significantly improved result compared to IPM$_8^9$.
-Again, the IPM$_8^9$ result stagnates at the SC$^9$ accuracy level, which can be reached with IPM when only using four moments (combined with a sufficiently accurate quadrature level).
+Here, IPM$_8^4$ yields a significantly improved result compared to IPM$_8^3$.
+Again, the IPM$_8^3$ result stagnates at the SC$^9$ accuracy level, which can be reached with IPM when only using four moments (combined with a sufficiently accurate quadrature level).
 Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC.
-This result can be observed throughout the numerical experiments of this work.
+This result can be observed throughout the results of this work.
 Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly.
-However note, that since the numerical flux evaluation used by IPM requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
+However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
 \end{itemize}
-
-Let us now compare results obtained with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order $9$.
+Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order 9.
 Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well.
+Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points.
 All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
 3 to 10 moments).
-Order $2$ uses a $5$ quadrature points, orders $3$ to $6$ use $9$ quadrature points and orders $8$ and $9$ use $17$ quadrature points.
+Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule.
 When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the adaptive methods decrease the truncation order, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased.
-The remaining methods have been computed with $17$ quadrature points.
-The iteration in pseudo-time is performed until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
+The remaining methods have been computed with a level 4 quadrature rule.
+
+All intrusive methods are iterated until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
+For every quadrature point, SC iterates the solution until the density reaches a certain residual level $\varepsilon$.
+We observed that the SC error requires a smaller residual level to reach a constant state and we have recorded this behavior in the following Table \ref{tab:resSC17} for the SC$^{17}$ solution.
+\frama{Sollen wir hier die Ergebnisse zu Frage 1 im zweiten Dokument einbauen?}
+\begin{center}
+\begin{table}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{ | m{3cm} || m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | m{0.9cm} | } 
+	\hline
+	residual $\rho$ & 5e-06 & 4e-06 & 3e-06 & 2e-06 & 1e-06 & 9e-07 & 8e-07 & 7e-07 & 6e-07 & 5e-07  \\ 
+	\hline
+	error E[$\rho$] in e-03 & 1.506 & 1.464 & 1.438 & 1.401 & 1.368 & 1.360 & 1.359 & 1.358 & 1.358 & 1.357  \\ 
+	\hline
+	error Var[$\rho$] in e-03 & 1.339 & 1.332 & 1.292 & 1.248 & 1.217 & 1.215 & 1.214 & 1.213 & 1.213 & 1.213  \\ 
+	\hline
+\end{tabular}
+}
+\caption{Error values at various residual levels when using SC.}
+\label{tab:resSC17}
+\end{table}
+
+\end{center}
+Hence, we can deduce that SC requires a smaller residual to converge to a steady state solution.
+Therefore, the results for SC$^{17}$ will be presented for a residual of $\varepsilon = 5\cdot 10^{-6}$ as well as a reduced residual of $\varepsilon = 10^{-6}$.
+Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
+%One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
 
@@
 The standard SG method however preserves positivity of mass, energy and pressure.
-The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation values and in Figure~\ref{fig:sub32}.
+The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}.
 When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
@@
 \item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
+\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.
+Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
 \end{itemize}
@@
 \end{itemize}
-
-Let us now compare the results from intrusive methods with those of SC.
-For every quadrature point, SC iterates the solution until the density density reaches lies below a threshold of $\varepsilon = 1e^{-7}$.
-Recording the error of intrusive methods during the iteration is straight forward.
-To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs.
-The collocation runtimes that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D} are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error.
-Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e.
-$5,9$ and $17$ quadrature points.
-The comparison of intrusive methods with non-intrusive methods shows the following:
-\begin{itemize}
-\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1e^{-7}$ compared to $\varepsilon = 5e^{-6}$ for intrusive methods).
-\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.
-Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
-\end{itemize}
-
-Hence, we can deduce that SC requires a smaller residual to converge to a steady state solution.
-Therefore, the results for SC$^{17}$ will be presented for a residual of $\varepsilon = 5\cdot 10^{-6}$ as well as a reduced residual of $\varepsilon = 10^{-6}$.
-Note that in the case of SC, the error cannot be recorded without violating the non-intrusive framework and adding extra costs, which is why we only plot the final error and the corresponding run time.
-%One possible reason for this behavior could be that SC is less synchronized, meaning that the solution certain quadrature points
 
@@
 \centering
-	\begin{subfigure}{0.5\linewidth}
+	\begin{subfigure}{0.3\linewidth}
 		\centering
 		\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_ERho.png}
-		\label{fig:referenceSolutionsub1}
+		\caption{}
+		\label{fig:referenceSolutionsub2DE}
 	\end{subfigure}%
-	\begin{subfigure}{0.5\linewidth}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
 		\centering
 		\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho.png}
-		\label{fig:referenceSolutionsub2}
+		\caption{}
+		\label{fig:referenceSolutionsub2DVar}
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
-	\label{fig:referenceSolution2D}
-\end{figure}
-\begin{figure}[h!]
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
 		\centering
+		\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho_log.png}
+		\caption{}
+		\label{fig:referenceSolutionsub2DVarLog}
+	\end{subfigure}\\
 	\begin{subfigure}{0.3\linewidth}
@@
 		\caption{}
-		\label{fig:sub1}
+		\label{fig:reados2DE}
 	\end{subfigure}%
@@
 		\caption{}
-		\label{fig:sub2}
+		\label{fig:reados2DVar}
 	\end{subfigure}%
@@
 	\end{subfigure}
-	\caption{(a) E$[\rho]$, (b) Var$[\rho]$ and (c) refinement level for readosIPM$_{1-9}$}
-	\label{fig:readosIPMEVar2D}
+	\caption{Reference solution (\subref{fig:referenceSolutionsub2DE}) E$[\rho]$ and (\subref{fig:referenceSolutionsub2DVar}) Var$[\rho]$ and with logarithmic scaled (\subref{fig:referenceSolutionsub2DVarLog}) Var$[\rho]$.
+readosIPM$_{1-9}$ solution for (\subref{fig:reados2DE}) E$[\rho]$ and (\subref{fig:reados2DVar}) Var$[\rho]$, with the resulting (\subref{fig:readosIPMEVar2DRefinement}) refinement levels.}
+	\label{fig:refAndreadosIPMEVar2D}
 \end{figure}
-When looking at the computed IPM expectation value and variance in Figure~\ref{fig:readosIPMEVar2D} and compare the results against the reference solution in Figure~\ref{fig:referenceSolution2D}, one can observe that since the uncertainty effects the result more heavily than in the one-dimensional case, the IPM solution still shows a step-like profile.
+When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainty effects the result more heavily than in the one-dimensional case, the IPM solution still shows a step-like profile.
 However, we are able to capture the main features of the reference solution.
@@
 	\centering
-	\begin{subfigure}{0.5\linewidth}
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 	\end{subfigure}%
-	\begin{subfigure}{0.5\linewidth}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
 		\centering
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$.}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.2]{figs/pipe/pipe_mesh.png}
+		\label{fig:referenceSolutionsPipe2}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh.}
 	\label{fig:referenceSolution3D}

### 2019-11-25 Paper1/adaptivity.tex

@@
 The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$.
+
+Note that we use nested quadrature rules, which facilitate the task of evaluating the quadrature in the moment update \eqref{eq:adaptiveFVUpdate}.
+Assume that we want to compute the moment update in cell $j$ with refinement level $\ell_j$ where a neighboring cell $j-1$ has refinement level $\ell_{j-1}$.
+Now if $\ell_{j-1}\geq\ell_j$, the solution of cell $j-1$ is known at all $Q_{\ell}$ quadrature points, hence the integral inside the moment update can be computed.
+Vice versa, if $\ell_{j-1}\leq\ell_j$, we need to evaluate the neighboring cell at the finer quadrature level $\ell_j$.
+Except from this, increasing or decreasing the refinement level does not lead to additional costs.
+
 The IPM algorithm with adaptivity results in Algorithm \ref{alg:ad-IPM}.

### 2019-11-25 code/src/euler2d.cpp
last changes

@@ -180,7 +180,7 @@ Vector Euler2D::IC( const Vector& x, const Vector& xi ) {
         y[2] = rhoL * vF;
         y[3] = pL / ( gamma - 1 );
 
-        if( x[0] > -0.5 + _sigma[0] * xi[0] ) {
+        if( x[0] > -0.5 + _sigma[0] * xi[0] && false ) {
             double rhoR = 0.8;
             double pR   = 0.125;    // 0.3;
             y[0]        = rhoR;
@@ -255,47 +255,47 @@ Vector Euler2D::IC( const Vector& x, const Vector& xi ) {
         return y;
     }
     else if( pipeTestCase ) {    // pipe testcase
-		double gamma = 1.4;
-		double R     = 287.87;
-		double T     = 273.15;
-		double p     = 101325.0;
-		double Ma    = 0.0;
-		double a     = sqrt( gamma * R * T );
+        double gamma = 1.4;
+        double R     = 287.87;
+        double T     = 273.15;
+        double p     = 101325.0;
+        double Ma    = 0.0;
+        double a     = sqrt( gamma * R * T );
 
-		double uMax  = Ma * a;
-		double angle = ( 1.25 + _sigma[0] * 0.0 ) * ( 2.0 * M_PI ) / 360.0;
-		double uF    = uMax * cos( angle );
-		double vF    = uMax * sin( angle );
+        double uMax  = Ma * a;
+        double angle = ( 1.25 + _sigma[0] * 0.0 ) * ( 2.0 * M_PI ) / 360.0;
+        double uF    = uMax * cos( angle );
+        double vF    = uMax * sin( angle );
 
-		double rhoFarfield = p / ( R * T );
+        double rhoFarfield = p / ( R * T );
 
-		y[0] = rhoFarfield;
-		y[1] = rhoFarfield * uF;
-		y[2] = rhoFarfield * vF;
-		// double kineticEnergyL = 0.5 * rhoFarfield * ( pow( uF, 2 ) + pow( vF, 2 ) );
-		// double innerEnergyL   = ( p / ( rhoFarfield * ( gamma - 1 ) ) ) * rhoFarfield;
-		y[3] = 1.0;
+        y[0] = rhoFarfield;
+        y[1] = rhoFarfield * uF;
[…]

### 2019-11-25 code/src/expliciteuler.cpp

@@ -21,13 +21,16 @@ void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, unsigne
         std::pair cells = _mesh->CellsAtEdge( j );
         unsigned I      = cells.first;
         unsigned J      = cells.second;
-        // std::cout << "edge " << j << ": " << I << " " << J << std::endl;
+        std::cout << "edge " << j << ": " << I << " " << J << std::endl;
         unsigned level = refLevel[I];    // take max ref level of cells I and J
         if( level < refLevel[J] ) level = refLevel[J];
         double area = norm( _mesh->GetNormalsAtEdge( j ) );
-        if( _mesh->BoundaryAtEdge( j ) != NOSLIP ) {
+        if( _mesh->BoundaryAtEdge( j ) != NOSLIP && _mesh->BoundaryAtEdge( j ) != DIRICHLET ) {
             fluxFunc( _flux[j], _problem->G( uQ[I], uQ[J], _mesh->GetNormalsAtEdge( j ) / area, _mesh->GetNormalsAtEdge( j ), level ), level );
-            // std::cout << _flux[j] << std::endl;
+            // std::cout << "flux = " << _flux[j] << std::endl;
+            // std::cout << "uQI " << uQ[I] << std::endl;
+            // std::cout << "uQJ " << uQ[J] << std::endl;
+            // std::cout << "numCells " << numCells << std::endl;
         }
         else if( _mesh->BoundaryAtEdge( j ) == NOSLIP ) {
             fluxFunc( _flux[j], _problem->BoundaryFlux( uQ[I], _mesh->GetNormalsAtEdge( j ) / area, _mesh->GetNormalsAtEdge( j ), level ), level );
@@ -57,11 +60,11 @@ void ExplicitEuler::Advance( std::function<void( Matrix&, const Matrix&, unsigne
             // std::cout << _mesh->CellsAtEdge( edges[l] ).first << " " << j << std::endl;
             if( _mesh->CellsAtEdge( edges[l] ).first == j ) {
                 // std::cout << "+" << std::endl;
-                rhs = rhs + _flux[edges[l]];
+                rhs = rhs - _flux[edges[l]];
             }
             else {
                 // std::cout << "-" << std::endl;1
-                rhs = rhs - _flux[edges[l]];
+                rhs = rhs + _flux[edges[l]];
             }
         }
         // std::cout << std::endl;

### 2019-11-25 code/src/mesh2d.cpp

@@ -333,6 +333,7 @@ void Mesh2D::DetermineNeighbors() {
                             this->AddNeighbor( ci, cj, indexI0, indexI1 );
                             this->AddNeighbor( cj, ci, indexJ0, indexJ1 );
                             _edges.push_back( std::make_pair( ci->GetID(), cj->GetID() ) );
+                            _boundaryTypeEdge.push_back( _cells[i]->GetBoundaryType() );
                             goto cnt;
                         }
                     }
@@ -342,6 +343,7 @@ void Mesh2D::DetermineNeighbors() {
         }
         if( _cells[i]->IsBoundaryCell() ) {
             _edges.push_back( std::make_pair( _cells[i]->GetID(), _numCells ) );
+            _boundaryTypeEdge.push_back( _cells[i]->GetBoundaryType() );
             _cells[i]->UpdateBoundaryNormal();
         }
     }
@@ -354,9 +356,9 @@ void Mesh2D::DetermineNeighbors() {
         _normals[j] = Vector( _dimension, 0.0 );
     }
 
-    std::cout << "Setting Normals" << std::endl;
+    // std::cout << "Setting Normals" << std::endl;
     for( unsigned j = 0; j < _edges.size(); ++j ) {
-        std::cout << "First " << _edges[j].first << std::endl;
+        // std::cout << "First " << _edges[j].first << std::endl;
         _normals[j] = _cells[_edges[j].first]->GetUnitNormalForNgh( _edges[j].second );
         if( _edges[j].first == _numCells ) continue;
         for( unsigned l = 0; l < 3; ++l ) {
@@ -365,7 +367,7 @@ void Mesh2D::DetermineNeighbors() {
                 break;
             }
         }
-        std::cout << "Second " << _edges[j].second << std::endl;
+        // std::cout << "Second " << _edges[j].second << std::endl;
         if( _edges[j].second == _numCells ) continue;
         for( unsigned l = 0; l < 3; ++l ) {
             if( _edgesAtCell[_edges[j].second][l] != _numCells + 1 ) {

### 2019-11-25 Paper1/framework.tex
local convergence comment
Paper Jakob added
last comments from Martin

@@
 \end{align*}
+%Inside the Newton update \eqref{eq:dualIterationFunction}, we abuse notation for better readability by making use of
+%\begin{align*}
+%\left(\bm{H}^{-1}\cdot\nabla_{\bm{\lambda}}L\right)_{ij} := \sum_{i' = 1}^N\sum_{j' = 1}^m \left(\bm{H}^{-1}\right)_{m(j-1) + i,m(j'-1) + i'}\cdot\nabla_{\bm{\lambda}}L_{i'j'}.
+%\end{align*}
 The function $\bm d$ will in the following be called dual iteration function. Now, the Newton iteration for spatial cell $j$ is given by
@@
 The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
-$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{m\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j^n})$.
+$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{l\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j^n})$.
 To obtain a finite number of iterations for the iteration in cell $j$, a stopping criterion 
@@
 they do not necessitate the evaluation of the gPC expansion on quadrature points) can be constructed by performing a gPC expansion on the system flux directly \cite{debusschere2004numerical}.
-Examples can be found in \cite{hu2015stochastic,hu2016stochastic,tryoen2010instrusive}.
+Examples can be found in \cite{hu2015stochastic,hu2016stochastic,tryoen2010instrusive,durrwachterahigh} for the computation of numerical fluxes and sources.
 While the analytic computation of arising integrals is not always more efficient \cite[Section 6]{ghanem1998stochastic}, it can also complicate recycling a deterministic solver.

### 2019-11-25 Paper1/implementation.tex
continued
last comments from Martin

@@
 I.e.
-we determine $\bm u^{(j)}_k\in\mathbb{R}^p$ and the corresponding stencil cells for $k = 1,\cdots,Q$ by
+we determine $\bm u^{(j)}_k\in\mathbb{R}^m$ and the corresponding stencil cells for $k = 1,\cdots,Q$ by
 \begin{align*}
@@
 Now after having performed the dual update, the dual variables are send to all processors.
-With these variables, each processor computes the solution on its portion of the quadrature set and then computed its part of the sum in \eqref{eq:momentUpQuadrature} on all spacial cells.
+With these variables, each processor computes the solution on its portion of the quadrature set and then computes its part of the sum in \eqref{eq:momentUpQuadrature} on all spacial cells.
 All parts from the different processors are then added together and the full time-updated moments are distributed to all processors.
@@
 we use the same deterministic numerical flux $\bm g$.
-This allows a fair comparison of the different intrusive and non-intrusive techniques.
+To our best knowledge, this allows a fair comparison of the different intrusive and non-intrusive techniques.

### 2019-11-25 Paper1/introduction.tex
last comments from Martin

@@
 Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
+Non-intrusive Monte-Carlo methods \cite{mishra2012multi,mishra2012sparse,mishra2016numerical}, which randomly sample input unceratinties to compute quantities of interest are robust, but suffer from a slow rate of convergence while again lacking the ability to use adaptivity to their full extend.
 Discontinuous structures commonly arise on a small portion of the space-time domain.
@@
 Also, it provides the ability to base the intrusive method on the same deterministic solver as the implementation used in a black-box fashion for stochastic-Collocation, which we believe allows a fair comparison between intrusive and non-intrusive methods.
-The code is publicly available to allow reproducibility \comment{[RefCode]}.
+The code is publicly available to allow reproducibility \cite{uqcreator}.
 

### 2019-11-25 Paper1/summary_outlook.tex
last comments from Martin

@@
 \label{sec:summary_outlook}
-In this work, we proposed acceleration techniques, which can be applied to the intrusive framework of IPM and SG: We use a one-shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously.
+In this work, we proposed acceleration techniques, which can be applied to the intrusive framework of IPM and SG: We use a One-Shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously.
 By not fully converging the dual iteration, we can reduce computational costs.
@@
 In our test cases, the intrusive methods yield the same error level as SC for a reduced runtime, especially since intrusive methods require less unknowns to achieve a certain accuracy due to aliasing errors.
-In higher-dimensional problems, this effect is amplified since the number of unknowns to achieve a certain total degree is asymptotically smaller than the number of quadrature points in a tensorized or sparse grid.
+In higher-dimensional problems, this effect is amplified since the number of unknowns to achieve a certain total degree is asymptotically smaller than the number of quadrature points in a tensorized or sparse grid (\eqref{eq:numberBasisFcts} instead of $O(M(\log_2(M)^{p-1}))$).
 Furthermore, we could observe that the required residual at which the solution of intrusive methods reaches a steady state is smaller than for SC.
@@
 In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations.
-Similar to the one-shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state.
+Similar to the One-Shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state.
 Hessian approximations that can be interesting are BFGS and sparse BFGS \cite[Chapter~6.1]{nocedal2006numerical}, which construct the Hessian from previously computed gradients.

### 2019-11-29 Paper1/adaptivity.tex
corrections and remarks Jannick

@@
 Since stochastic hyperbolic problems generally experience shocks in a small portion of the space-time domain, the idea is to perform arising computations on a high accuracy level in this small area, while keeping a low level of accuracy in the remainder.
-The hope is to automatically select the lowest order moment capable of approximating the solution with given accuracy, i.e.
-the same error is obtained while using a significantly reduced number of unknowns in most parts of the computational domain.
+The idea is to automatically select the lowest order moment capable of approximating the solution with given accuracy, i.e.
+the same error is obtained while using a significantly reduced number of unknowns in most parts of the computational domain and thus boost the performance of intrusive methods.
 
@@
 Adaptivity is a common strategy to accelerate this class of methods and several indicators to determine the smoothness of the solution exist.
-We make use of a strategy from \cite{persson2006sub}, which uses the highest degree moments as indicator.
+We make use of a strategy from \cite{persson2006sub}, which uses the two highest degree moments as an indicator.
 Translating the idea of the discontinuity sensor used in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at refinement level $\ell$ as
@@
 in the case of gas dynamics, the regularity of the density is chosen to indicate an adequate refinement level.
-If the moment vector in a given cell at a certain timestep is initially at refinement level $\ell$, this level is kept if the error indicator \eqref{eq:errorIndicator} lies in the interval $I_{\delta}:=[\delta_{-},\delta_{+}]$.
+If the moment vector in a given cell and at a certain timestep is initially at refinement level $\ell$, this level is kept if the error indicator \eqref{eq:errorIndicator} lies in the interval $I_{\delta}:=[\delta_{-},\delta_{+}]$.
 Here $\delta_{\pm}$ are user determined parameters.
-If the indicator is smaller than $\delta_-$, the refinement level is decreased by a value of one, if it lies above $\delta_+$, it is increased by one.
+If the indicator is smaller than $\delta_-$, the refinement level is decreased to the next lower level, if it lies above $\delta_+$, it is increased to the next higher level.
 
@@
 The strategy now is to perform the dual update for a set of moment vectors $\bm{\hat u}_j^n$ at refinement levels $\ell_j^n$ for $j = 1,\cdots,N_x$.
-Hence, the dual iteration makes use of the iteration function \eqref{eq:dualIterationFunctionAd} at refinement level $\ell_j^n$.
+Thus, the dual iteration makes use of the iteration function \eqref{eq:dualIterationFunctionAd} at refinement level $\ell_j^n$.
 After that, the refinement level at the next time step $\ell_j^{n+1}$ is determined by making use of the smoothness indicator \eqref{eq:errorIndicator}.
-The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$.
+The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$, utilizing the the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$.
 
@@
 \While{\eqref{eq:tauCrit} is violated}
-\State $\bm{\lambda}_j^{(l+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n})$
-\State $l \leftarrow l+1$
+\State $\bm{\lambda}_j^{(\zeta+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(\zeta)};\bm{\hat u}_j^{n})$
+\State $\zeta \leftarrow \zeta+1$
 \EndWhile
-\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(l)}$
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(\zeta)}$
 \State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\hat \lambda}_j^{n+1}\right)$
@@
 Hence, we propose to iteratively increase the maximal refinement level whenever the residual \eqref{eq:residualSteady} lies below a certain tolerance $\varepsilon$.
-For a given set of maximal refinement levels $\ell_l^*$ and a set of tolerances $\varepsilon_l^*$ at which the refinement level must be increased we can now perform a large amount of the required iterations on a lower but cheaper refinement level.
+For a given set of maximal refinement levels $\ell_l^*$ and a set of tolerances $\varepsilon_l^*$ at which the refinement level must be increased, we can now perform a large amount of the required iterations on a lower, but cheaper refinement level.
 The same strategy can be applied for One-Shot IPM. In this case, the algorithm is given by Algorithm~\ref{alg:adosIPM}.
@@
 \If{the residual \eqref{eq:residualSteady} lies below $\varepsilon_l^*$}
-\State $l \leftarrow l+1$
+\State $\zeta \leftarrow \zeta+1$
 \EndIf

### 2019-11-29 Paper1/framework.tex
corrections and remarks Jannick

@@
 with the flux $\bm{F}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
-Note that the inner transpose represent a dyadic product, i.e.
-the outer transpose is applied to a matrix.
+Note that the inner transpose represents a dyadic product and therefore the outer transpose is applied to a matrix.
 Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments.
@@
 \end{align*}
-If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N\times m}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N\times m}$ \remja{0...N -> N+1}, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
 \begin{align}\label{eq:IPMDiscretization}
@@
 \end{align}
-where $\bm H\in\mathbb{R}^{m \cdot  N\times m\cdot N}$ is the Hessian of \eqref{eq:L}, given by
+where $\bm H\in\mathbb{R}^{N \cdot n\times N\cdot m}$ is the Hessian of \eqref{eq:L}, given by
 \begin{align*}
@@
 %\end{align*}
-The function $\bm d$ will in the following be called dual iteration function. Now, the Newton iteration for spatial cell $j$ is given by
+The function $\bm d$ will in the following be called dual iteration function.
+Now, the Newton iteration $\zeta$ \remja{l ist später auch level für adIPM und left flux - habs erstmal überall auf zeta geändert} for spatial cell $j$ is given by
 \begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j}).
+\bm{\lambda}^{(\zeta+1)}_j = \bm{d}(\bm{\lambda}_j^{(\zeta)},\bm{\hat{u}_j}).
 \end{align}
 The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
-$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{l\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j^n})$.
-To obtain a finite number of iterations for the iteration in cell $j$, a stopping criterion 
+$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{\zeta\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(\zeta)},\bm{\hat{u}_j^n})$.
+To obtain a finite number of iterations for the iteration in cell $j$, the stopping criterion 
 \begin{align}\label{eq:tauCrit}
-\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(l)};\bm{\hat{u}}_j^n) \right\Vert < \tau
+\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(\zeta)};\bm{\hat{u}}_j^n) \right\Vert < \tau
 \end{align}
@@
 \begin{align}\label{eq:dualIteration}
-\bm{\lambda}_j^{(l+1)} = \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n}).
+\bm{\lambda}_j^{(\zeta+1)} = \bm{d}(\bm{\lambda}_{j}^{(\zeta)};\bm{\hat u}_j^{n}).
 \end{align}
@@
 \While{\eqref{eq:tauCrit} is violated}
-\State $\bm{\lambda}_j^{(l+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n})$
-\State $l \leftarrow l+1$
+\State $\bm{\lambda}_j^{(\zeta+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(\zeta)};\bm{\hat u}_j^{n})$
+\State $\zeta \leftarrow \zeta+1$
 \EndWhile
-\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(l)}$
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(\zeta)}$
 \EndFor

### 2019-11-29 Paper1/implementation.tex
corrections and remarks Jannick

@@
 It remains to discuss the parallelization of the presented algorithms.
-In order to minimize the parallelization overhead, our goal is to minimize the need to send data between processors.
+In order to minimize the parallelization overhead, our goal is to minimize the communication between processors.
 Note that the dual problem (line 8 in Algorithm~\ref{alg:ad-IPM} and line 5 in Algorithm~\ref{alg:adosIPM}) does not require communication, i.e.
@@
 Hence, distributing the spatial mesh between processors will yield communication overhead since data needs to be sent whenever a stencil cell lies on a different processor.
-Therefore, we choose to parallelize the quadrature points, which again minimizes communication effort.
+Therefore, we choose to parallelize the quadrature points, which minimizes the computational time spend on communication.
 As mentioned in \ref{app:costNumFlux}, we first compute the solution at stencil cells for all quadrature points.
 I.e.
-we determine $\bm u^{(j)}_k\in\mathbb{R}^m$ and the corresponding stencil cells for $k = 1,\cdots,Q$ by
+we determine $\bm u^{(j)}_k\in\mathbb{R}^m$ and the corresponding stencil cells for $k = 1,\dots,Q$ by
 \begin{align*}
@@
 \end{align*}
-Then, the finite volume update function \eqref{eq:adaptiveFVUpdate} can be written as
+Thus, the finite volume update function \eqref{eq:adaptiveFVUpdate} can be written as
 \begin{align}\label{eq:momentUpQuadrature}
@@
 the sum in \eqref{eq:momentUpQuadrature} can be computed in parallel.
-Now after having performed the dual update, the dual variables are send to all processors.
+Now, after having performed the dual update, the dual variables are send to all processors.
 With these variables, each processor computes the solution on its portion of the quadrature set and then computes its part of the sum in \eqref{eq:momentUpQuadrature} on all spacial cells.

### 2019-11-29 Paper1/introduction.tex
corrections and remarks Jannick

@@
 One strategy to represent the solution's dependence on uncertainties is to use a polynomial chaos (PC) expansion \cite{wiener1938homogeneous,xiu2002wiener}, i.e.
-the uncertain space is spanned by polynomial basis functions.
+the uncertainty space is spanned by polynomial basis functions.
 The remaining task is then to determine adequate expansion coefficients, often called moments or PC coefficients.
@@
 \item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM shows improved variance approximations.
-\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for a one-dimension uncertainty) to reach the same accuracy level.
+\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for a one-dimensional uncertainty) to reach the same accuracy level.
 \item Using sparse grids for the IPM discretization when the space of uncertainty is multi-dimensional, the number of quadrature points needed to guarantee sufficient regularity of the Hessian matrix is significantly increased.
@@
 While facilitating the task of implementing general intrusive methods, this framework reduces the number of operations required to compute numerical fluxes.
-Also, it provides the ability to base the intrusive method on the same deterministic solver as the implementation used in a black-box fashion for stochastic-Collocation, which we believe allows a fair comparison between intrusive and non-intrusive methods.
+Also, it provides the ability to base the intrusive method on the same deterministic solver as used in the implementation of a black-box fashion stochastic-Collocation code, which allows for, what we believe, a fair comparison between intrusive and non-intrusive methods.
 The code is publicly available to allow reproducibility \cite{uqcreator}.
@@
 \begin{align}
-\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D, \\ \label{eq:ic}
+\label{eq:fulleq}\partial_t \bm{u}(t,\bm{x},\bm{\xi}) + \nabla&\cdot\bm{f}(\bm{u}(t,\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D, \\ \label{eq:ic}
 \bm{u}(t=0,\bm{x},&\bm{\xi}) = \bm{u}_{\text{IC}}(\bm{x},\bm{\xi}),
@@
 The solution ansatz \eqref{eq:SGClosure} is $L^2$-optimal, if the moments are chosen to be the Fourier coefficients $\bm{\hat u}_i \equiv \langle \bm{u}\varphi_i \rangle\in\mathbb{R}^m$.
-One can also use the ansatz \eqref{eq:SGClosure} to compute the quantities of interest.
-Indeed, we have that
+One can also use the ansatz \eqref{eq:SGClosure} to compute the quantities of interest as
 \begin{align*}
@@
 To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^p)$ quadrature points to integrate polynomials of maximal degree $M$ exactly, sparse grids are designed to integrate polynomials of total degree $M$, for which they only require $O(M(\log_2(M)^{p-1}))$ quadrature points.
+\remja{Brauchen wir hier noch ne Quelle?}
 
@@
 Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstrained optimization problem.
-Defining the variables $\bm{\lambda}_i\in\mathbb{R}^m$ where $i$ is again a multi index, gives the unconstrained dual problem
+Defining the variables $\bm{\lambda}_i\in\mathbb{R}^m$, where $i$ is again a multi index, gives the unconstrained dual problem
 \begin{align}\label{eq:dualProblem}

### 2019-11-29 Paper1/main.tex
corrections and remarks Jannick

@@
 \usepackage{array}
+\usepackage{hyperref}
 
@@
 \begin{keyword}
-uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation, Intrusive Polynomial Moment Method
+uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation, Intrusive Polynomial Moment Method, \remja{sparse grids?}
 \end{keyword}

### 2019-11-29 Paper1/results.tex
last comments from Martin
corrections and remarks Jannick

@@
 Altogether, the mesh consists of 22361 triangular elements.
+\remja{Note, that the cell density of the mesh was adjusted in order to capture areas of high variance and therefore the mesh is significantly different to a mesh classically applied in computational fluid dynamics.}
 
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and mesh used in the computation of all presented methods.}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.}
 	\label{fig:referenceSolution}
@@
 	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-All intrusive methods converge to a residual $\varepsilon=5e-6$ whereas all non-intrusive methods converge to a residual $\varepsilon=1e-7$.
+All intrusive methods converge to a residual $\varepsilon=5\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=\cdot 10^{-7}$.
 All computations are performed with with 5 MPI threads.}
@@
 Let us now compare results obtained with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order $9$.
-Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well.
+Note, that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well.
 All adaptive methods use gPC polynomials of order 2 to 9 (i.e.
@@
 \item Compared to IPM, stochastic-Galerkin comes at a significantly reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
-\item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
+\item For the expectation value SG$_9$ shows a smaller error compared to IPM$_9$, while for the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
 \end{itemize}
@@
 This method reaches the steady state IPM solution with a faster runtime than SG.
-\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime.
+\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by \textit{retardation adosIPM} (readosIPM), which further decreases runtime.
 Here, we use two different strategies to increase the accuracy: First, we steadily increase the maximal truncation order when the residual approaches zero.
@@
 Let us now compare the results from intrusive methods with those of SC.
-For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1e^{-7}$.
+For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1\cdot 10^{-7}$.
 Recording the error of intrusive methods during the iteration is straight forward.
 To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs.
-The collocation runtimes that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D} are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error.
+The collocation runtimes, that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D}, are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error.
 Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e.
@@
 \begin{itemize}
-\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1e^{-7}$ compared to $\varepsilon = 5e^{-6}$ for intrusive methods).
+\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1\cdot 10^{-7}$ compared to $\varepsilon = 5\cdot 10^{-6}$ for intrusive methods).
 \item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.
@@
 The farfield pressure is ${p \sim U(100\;325,102\;325)}$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$.
-Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which in our experiments proved to be more efficient than a sparse grid quadrature.
+Since this problem only has a two-dimensional uncertainty, we use a tensorized quadrature set, which in our experiments proved to be more efficient than a sparse grid quadrature.
 For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the One-Shot strategy).
-The IPM method uses moment orders ranging from $1$ to $9$ adaptively with refinement barriers of $\delta_{-} = 1\cdot 10^{-4}$ and $\delta_{+} = 1\cdot 10^{-5}$.
+The IPM method uses moment orders ranging from $1$ to $9$ adaptively with refinement barriers $\delta_{-} = 1\cdot 10^{-4}$ and $\delta_{+} = 1\cdot 10^{-5}$.
 The refinement retardation allows the truncation order to have a maximal total degree of $1$ until a residual of $\varepsilon = 1.5\cdot 10^{-5}$ and then increases the truncation order by one when the residual is reduced by an amount of $5\cdot 10^{-6}$.
@@
 The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}.
-To determine the error, we used Stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule.
+To determine the error, we used stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule.
 As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method.
@@
 $17^2 = 289$ collocation points.
-Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns needed for intrusive methods to obtain a certain error weighs in more heavily.
+Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns required for intrusive methods to obtain a certain error becomes more apparent.
 This shows one promising characteristic of intrusive techniques and points to their applicability for higher dimensional problems.
@@
 		\label{fig:referenceSolutionsub2DE}
-	\end{subfigure}%
+	\end{subfigure}
 	\hfill
@@
 \end{figure}
-When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainty effects the result more heavily than in the one-dimensional case, the IPM solution still shows a step-like profile.
+When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainties have a bigger effect on the result than in the one-dimensional case, the IPM solution still shows a step-like profile.
 However, we are able to capture the main features of the reference solution.
@@
 a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case.
+\remja{Visualizing the variance on a logarithmic scale in Figure~\ref{fig:referenceSolutionsub2DVarLog} further shows that the applied refinement indicators work well and use higher order moments in areas of high variance.}
 

### 2019-11-29 Paper1/summary_outlook.tex
corrections and remarks Jannick

@@
 \label{sec:summary_outlook}
-In this work, we proposed acceleration techniques, which can be applied to the intrusive framework of IPM and SG: We use a One-Shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously.
+In this work, we proposed acceleration techniques, which are applicable to the intrusive nature of IPM and SG: We use a One-Shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously.
 By not fully converging the dual iteration, we can reduce computational costs.
@@
 Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
-Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, it can be important to intrusively modify the code in order to fully exploit all acceleration potentials.
+Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, as shown in this work, it can be helpful to intrusively modify the code in order to fully exploit all acceleration potentials.
 Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods.

### 2019-11-29 code/src/momentsolver.cpp
first performance fix -> no longer works with adaptivity

@@ -89,9 +89,7 @@ void MomentSolver::Solve() {
 #pragma omp parallel for schedule( dynamic, 10 )
         for( unsigned j = 0; j < static_cast<unsigned>( _cellIndexPE.size() ); ++j ) {
             if( _mesh->GetBoundaryType( _cellIndexPE[j] ) == BoundaryType::DIRICHLET && timeIndex > 0 ) continue;
-            // std::cout << "Cell " << _cellIndexPE[j] << ", lambda = " << _lambda[_cellIndexPE[j]] << ", u = " << u[_cellIndexPE[j]] << std::endl;
-            _closure->SolveClosureSafe( _lambda[_cellIndexPE[j]], u[_cellIndexPE[j]], refinementLevel[_cellIndexPE[j]] );
-            // std::cout << "result = " << _lambda[_cellIndexPE[j]] << std::endl;
+            _lambda[_cellIndexPE[j]] = u[_cellIndexPE[j]];
         }
 
         // MPI Broadcast lambdas to all PEs
@@ -100,51 +98,9 @@ void MomentSolver::Solve() {
             MPI_Bcast( _lambda[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, PEforCell[j], MPI_COMM_WORLD );
         }
 
-        // save old refinement levels
-        refinementLevelOld = refinementLevel;
-
-        // determine refinement level of cells on current PE
-        for( unsigned j = 0; j < static_cast<unsigned>( _cellIndexPE.size() ); ++j ) {
-            // if( _mesh->GetBoundaryType( _cellIndexPE[j] ) == BoundaryType::DIRICHLET && timeIndex > 0 ) continue;
-            double indicator;
-            if( _settings->GetProblemType() == P_RADIATIONHYDRO_1D )
-                indicator = 0.0;
-            else
-                indicator = ComputeRefIndicator( refinementLevel, u[_cellIndexPE[j]], refinementLevel[_cellIndexPE[j]] );
-            if( indicator > _settings->GetRefinementThreshold() &&
-                refinementLevel[_cellIndexPE[j]] < _settings->GetNRefinementLevels( retCounter ) - 1 &&
-                _mesh->GetBoundaryType( _cellIndexPE[j] ) != BoundaryType::DIRICHLET )
-                refinementLevel[_cellIndexPE[j]] += 1;
-            else if( indicator < _settings->GetCoarsenThreshold() && refinementLevel[_cellIndexPE[j]] > 0 &&
[…]

### 2019-11-29 code/src/momentsolver.cpp
no more sending around of lambda

@@ -85,19 +85,6 @@ void MomentSolver::Solve() {
         Vector residualVec( _settings->GetNQTotal(), 0.0 );
         double residual;
 
-        // Solve dual problem
-#pragma omp parallel for schedule( dynamic, 10 )
-        for( unsigned j = 0; j < static_cast<unsigned>( _cellIndexPE.size() ); ++j ) {
-            if( _mesh->GetBoundaryType( _cellIndexPE[j] ) == BoundaryType::DIRICHLET && timeIndex > 0 ) continue;
-            _lambda[_cellIndexPE[j]] = u[_cellIndexPE[j]];
-        }
-
-        // MPI Broadcast lambdas to all PEs
-        for( unsigned j = 0; j < _nCells; ++j ) {
-            uOld[j] = u[j];    // save old Moments for residual computation
-            MPI_Bcast( _lambda[j].GetPointer(), int( _nStates * _nTotal ), MPI_DOUBLE, PEforCell[j], MPI_COMM_WORLD );
-        }
-
         // compute solution at quad points
         for( unsigned j = 0; j < _nCells; ++j ) {
             uQ[j] = _closure->U( _closure->EvaluateLambdaOnPE( u[j], refinementLevelOld[j], refinementLevelTransition[j] ) );
@@ -168,7 +155,8 @@ void MomentSolver::Solve() {
     if( _settings->GetMyPE() != 0 ) return;
 
     // save final moments on uNew
-    uNew = u;
+    uNew    = u;
+    _lambda = uNew;
 
     std::chrono::steady_clock::time_point toc = std::chrono::steady_clock::now();
     log->info( "" );

### 2019-12-11 Paper1/introduction.tex
fixed small error
data for plot

@@
 
-When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered,poette2019contribution} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
+When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty,kusch2018filtered}, IPM \cite{kusch2018filtered,poette2019contribution} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
 These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through either the spatial numerical method \cite{offner2017stability} or by filters \cite{kusch2018filtered}.

### 2019-12-11 Paper1/results.tex
updated figures euler2D

@@
 	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-All intrusive methods converge to a residual $\varepsilon=5\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=\cdot 10^{-7}$.
+All intrusive methods converge to a residual $\varepsilon=5\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=1\cdot 10^{-7}$.
 All computations are performed with with 5 MPI threads.}

### 2019-12-12 Paper1/results.tex
fine tuned residual limit for euler1D

--- 2019-11-29 Paper1/results.tex
@@
 	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-All intrusive methods converge to a residual $\varepsilon=5\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=\cdot 10^{-7}$.
+All intrusive methods converge to a residual $\varepsilon=6\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=1\cdot 10^{-7}$.
 All computations are performed with with 5 MPI threads.}

### 2019-12-12 Paper1/results.tex
set new residual limit in the text

@@
 The remaining methods have been computed with $17$ quadrature points.
-The iteration in pseudo-time is performed until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
+The iteration in pseudo-time is performed until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 6\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.
 
@@
 \begin{itemize}
-\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1\cdot 10^{-7}$ compared to $\varepsilon = 5\cdot 10^{-6}$ for intrusive methods).
+\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1\cdot 10^{-7}$ compared to $\varepsilon = 6\cdot 10^{-6}$ for intrusive methods).
 \item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns.

### 2019-12-12 Paper1/introduction.tex
spellchecked

@@
 Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
-Non-intrusive Monte-Carlo methods \cite{mishra2012multi,mishra2012sparse,mishra2016numerical}, which randomly sample input unceratinties to compute quantities of interest are robust, but suffer from a slow rate of convergence while again lacking the ability to use adaptivity to their full extend.
+Non-intrusive Monte-Carlo methods \cite{mishra2012multi,mishra2012sparse,mishra2016numerical}, which randomly sample input uncertainties to compute quantities of interest are robust, but suffer from a slow rate of convergence while again lacking the ability to use adaptivity to their full extend.
 Discontinuous structures commonly arise on a small portion of the space-time domain.

### 2019-12-12 Paper1/one_shot_ipm.tex
corrections and remarks Jannick
spellchecked

@@
 
-In the following section we only consider steady state problems, i.e. we have
+In the following section we only consider steady state problems, i.e. equation \ref{eq:fulleq} reduces to
 \begin{align}\label{eq:hyperbolicProblemSteady}
@@
 \end{subequations}
-This yields Algotihm \ref{alg:osIPM}.
+This yields Algorithm \ref{alg:osIPM}.
 \begin{algorithm}[H]
@@
 For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence.
-Hence, we preserve the convergence results of the original IPM method, which uses Newton's method, i.e.
-converges locally as well.
+Hence, we preserve the convergence property of the original IPM method, which uses Newton's method and therefore only converges locally as well \remja{Quelle?}.
 \begin{theorem}\label{th:Contractive}
@@
 Hence, we have that $\partial_{\bm{\hat u}} \bm{c} = \nabla_{\bm{\hat u}}\bm{\tilde c}$, which only has eigenvalues between $-1$ and $1$ by the assumption that the classical IPM iteration is contractive.
-Since $\bm{J}$ is an upper triangluar block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{\hat u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fixed point.
+Since $\bm{J}$ is an upper triangular block matrix, the eigenvalues are given by $\lambda\left(\partial_{\bm{\lambda}} \bm{d}\right) = 0$ and $\lambda\left(\partial_{\bm{\hat u}} \bm{c}\right)\in(-1,1)$, hence the One-Shot IPM is contractive around its fixed point.
 \end{proof}

### 2019-12-12 code/src/euler.cpp
Euler testcase numhyp paper

@@ -141,7 +141,7 @@ Vector Euler::LoadIC( const Vector& x, const Vector& xi ) {
 }
 
 Matrix Euler::ExactSolution( double t, const Matrix& x, const Vector& xi ) const {
-    bool sodShockTube = false;
+    bool sodShockTube = true;
     double x0         = 0.5 + _sigma[0] * xi[0];    // initial shock position
     double rho_l      = 1.0;
     double P_l        = 1.0;

### 2019-12-12 Paper1/framework.tex
zeta back to l

@@
 The function $\bm d$ will in the following be called dual iteration function.
-Now, the Newton iteration $\zeta$ \remja{l ist später auch level für adIPM und left flux - habs erstmal überall auf zeta geändert} for spatial cell $j$ is given by
+Now, the Newton iteration $l$ \remja{l ist später auch level für adIPM und left flux - habs erstmal überall auf zeta geändert} for spatial cell $j$ is given by
 \begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{(\zeta+1)}_j = \bm{d}(\bm{\lambda}_j^{(\zeta)},\bm{\hat{u}_j}).
+\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j}).
 \end{align}
 The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
-$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{\zeta\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(\zeta)},\bm{\hat{u}_j^n})$.
+$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{l\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j^n})$.
 To obtain a finite number of iterations for the iteration in cell $j$, the stopping criterion 
 \begin{align}\label{eq:tauCrit}
-\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(\zeta)};\bm{\hat{u}}_j^n) \right\Vert < \tau
+\sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(l)};\bm{\hat{u}}_j^n) \right\Vert < \tau
 \end{align}
@@
 \begin{align}\label{eq:dualIteration}
-\bm{\lambda}_j^{(\zeta+1)} = \bm{d}(\bm{\lambda}_{j}^{(\zeta)};\bm{\hat u}_j^{n}).
+\bm{\lambda}_j^{(l+1)} = \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n}).
 \end{align}
@@
 \While{\eqref{eq:tauCrit} is violated}
-\State $\bm{\lambda}_j^{(\zeta+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(\zeta)};\bm{\hat u}_j^{n})$
-\State $\zeta \leftarrow \zeta+1$
+\State $\bm{\lambda}_j^{(l+1)} \leftarrow \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n})$
+\State $l \leftarrow l+1$
 \EndWhile
-\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(\zeta)}$
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(l)}$
 \EndFor

### 2019-12-12 Paper1/framework.tex
changes

@@
 \end{align*}
-If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N\times m}$ \remja{0...N -> N+1}, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
+If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{ij}^n)_{\vert i\vert\leq M}\in\mathbb{R}^{N\times m}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
 \begin{align}\label{eq:IPMDiscretization}
@@
 %\end{align*}
-The function $\bm d$ will in the following be called dual iteration function.
-Now, the Newton iteration $l$ \remja{l ist später auch level für adIPM und left flux - habs erstmal überall auf zeta geändert} for spatial cell $j$ is given by
+The function $\bm d$ will in the following be called dual iteration function. Now, the Newton iteration $l$ for spatial cell $j$ is given by
 \begin{align}\label{eq:dualIteration1}

### 2019-12-12 Paper1/introduction.tex
changes

@@
 \begin{align}\label{eq:SGClosure}
-\mathcal{U}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),
+\mathcal{U}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \bm{\hat{u}}^T\bm{\varphi}(\bm\xi),
 \end{align}
 where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^m$ are called moments.
-To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\hat{\bm u}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N\times m}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
+To allow a more compact notation, we collect the $N$ moments for which $\vert i \vert \leq M$ holds in the moment matrix $\bm{\hat{u}}:=(\bm{\hat{u}}_i)_{|i|\leq M}\in\mathbb{R}^{N\times m}$ and the corresponding basis functions in $\bm{\varphi}:=(\varphi_i)_{|i|\leq M}\in\mathbb{R}^{N}$.
 In the following, the dependency of $\mathcal{U}$ on $\bm \xi$ will occasionally be omitted for sake of readability.

### 2019-12-12 Paper1/adaptivity.tex
zeta back to l
more changes

@@
 Adaptivity is a common strategy to accelerate this class of methods and several indicators to determine the smoothness of the solution exist.
-We make use of a strategy from \cite{persson2006sub}, which uses the two highest degree moments as an indicator.
-Translating the idea of the discontinuity sensor used in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at refinement level $\ell$ as
+Translating the idea of the so-called discontinuity sensor which has been defined in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at refinement level $\ell$ as
 \begin{align*}
@@
 \While{\eqref{eq:tauCrit} is violated}
-\State $\bm{\lambda}_j^{(\zeta+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(\zeta)};\bm{\hat u}_j^{n})$
-\State $\zeta \leftarrow \zeta+1$
+\State $\bm{\lambda}_j^{(l+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n})$
+\State $l \leftarrow l+1$
 \EndWhile
-\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(\zeta)}$
+\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(l)}$
 \State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\hat \lambda}_j^{n+1}\right)$
@@
 \If{the residual \eqref{eq:residualSteady} lies below $\varepsilon_l^*$}
-\State $\zeta \leftarrow \zeta+1$
+\State $l \leftarrow l+1$
 \EndIf

### 2019-12-12 Paper1/framework.tex
more changes

@@
 The computation of the dual variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
-Hence, to determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
+Therefore, to determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
 \begin{align}\label{eq:L}
@@
 The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
-$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u_j^n})=\lim_{l\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j^n})$.
+$\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)=\lim_{l\rightarrow\infty}\bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j^n)$.
 To obtain a finite number of iterations for the iteration in cell $j$, the stopping criterion 
@@
 A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}.
-This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux}.
+This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux} or \textit{kinetic scheme}, see e.g.
+\cite{deshpande1986kinetic,harten1983upstream,perthame1990boltzmann,perthame1992second}.
 By simply taking moments of a given numerical flux for the deterministic problem, the method can easily be applied to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available.

### 2019-12-12 Paper1/introduction.tex
more changes

@@
 Hence, the moments can be computed by running a given deterministic solver for the original problem at each quadrature point.
-To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^p)$ quadrature points to integrate polynomials of maximal degree $M$ exactly, sparse grids are designed to integrate polynomials of total degree $M$, for which they only require $O(M(\log_2(M)^{p-1}))$ quadrature points.
-\remja{Brauchen wir hier noch ne Quelle?}
+To reduce numerical costs in multi-dimensional settings, SC commonly uses sparse grids as quadrature rule: While tensorized quadrature sets require $O(M^p)$ quadrature points to integrate polynomials of maximal degree $M$ exactly, sparse grids are designed to integrate polynomials of total degree $M$, for which they only require $O(M(\log_2(M)^{p-1}))$ quadrature points, see e.g.
+\cite{trefethen2017cubature}.
 

### 2019-12-12 Paper1/one_shot_ipm.tex
more changes

@@
 
-In the following section we only consider steady state problems, i.e. equation \ref{eq:fulleq} reduces to
+In the following section we only consider steady state problems, i.e. equation \eqref{eq:fulleq} reduces to
 \begin{align}\label{eq:hyperbolicProblemSteady}
@@
 For this we take an approach commonly chosen to prove local convergence properties of Newton's method: In Theorem \ref{th:Contractive}, we show that the iteration function is contractive at its fixed point and conclude in Theorem \ref{th:localConvergence} that this yields local convergence.
-Hence, we preserve the convergence property of the original IPM method, which uses Newton's method and therefore only converges locally as well \remja{Quelle?}.
+Hence, we preserve the convergence property of the original IPM method, which uses Newton's method and therefore only converges locally as well.
 \begin{theorem}\label{th:Contractive}
@@
 \begin{align}\label{eq:ulambda}
-\bm{\hat u} = \langle \bm{u}_s(\bm{\hat \lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle =: \bm{h}(\bm{\hat \lambda}),
+\bm{\hat u} = \langle \bm{u}_s(\bm{\hat \lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T =: \bm{h}(\bm{\hat \lambda}),
 \end{align}

### 2019-12-12 Paper1/results.tex
more changes

@@
 Altogether, the mesh consists of 22361 triangular elements.
-\remja{Note, that the cell density of the mesh was adjusted in order to capture areas of high variance and therefore the mesh is significantly different to a mesh classically applied in computational fluid dynamics.}
 
@@
 \end{align*}
-Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\hat{\bm u}_{\Delta}$, we investigate the relative error
+Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
 \begin{align*}

### 2019-12-13 Paper1/results.tex
pipe testcase first draft; figs not final

@@
 a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case.
-\remja{Visualizing the variance on a logarithmic scale in Figure~\ref{fig:referenceSolutionsub2DVarLog} further shows that the applied refinement indicators work well and use higher order moments in areas of high variance.}
+Visualizing the variance on a logarithmic scale in Figure~\ref{fig:referenceSolutionsub2DVarLog} further shows that the applied refinement indicators work well and enforce the usage of higher order moments in areas of high variance.
 
-\subsection{Pipe Testcase}
+\subsection{2D Euler equations with a three dimensional uncertainty}
+For the last numerical study we again use the Euler equations, but this time not applied to the NACA airfoil, but rather to a bend Sod shock tube experiment.
+The prescribed initial condition describes a gas at rest, but with a discontinuity, i.e.
+as sudden jump, in density as well as energy in the upper part of the tube.
+The density $\rho_u = p/(R \cdot T)$ in the upper part results in $\rho_u = 1.289$ with $p = 101\;325$ Pa, a temperature of $273.15$ K and the specific gas constant for air $R=287.87$.
+The energy $\rho e_u$ is set to exactly $1.0$.
+For the initial conditions in the lower part we set $\rho_l = 0.5 \cdot \rho_u$ and $\rho e_l = 0.3$.
+The heat capacity ratio $\gamma$ is set to $1.4$ as in the previous studies.
+We again augment the deterministic case by inflicting three uncertainties this time.
+The applied boundary conditions are Euler slip conditions $\bm{v}^T\bm{n}=0$ for the walls of and Dirichlet conditions set to the initial conditions at each end of the tube.
+The first two uncertainties are the density and energy of the lower part of the shock tube.
+Here we set $\rho_l \sim U(1.189,1389)$ and $\rho e_l \sim U(0.2,0.4)$.
+The third uncertainty enters through the position of the shock itself and thus $y_\text{shock} \sim U(1.0, 1.2)$.
+Note, that in contrast to the upper testcases, here we are not interested in the steady state of the system but rather the evolution of the expectancy and variance in time.
+The computational mesh (see Figure \ref{fig:referenceSolutionsPipeMesh}) is an unstructured mesh of $25458$ triangular cells, where all the cells are similar in size.
+The simulations were run until a time of $0.2$s was exceeded and the corresponding results can be seen in Figure \ref{fig:solution3D}.
+The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was again computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
+For this testcase we want to compare the results for sparse grids w.r.t.
+tensorized grids in higher stochastical dimensions.
+As we made used of adaptivity, the corresponding moment orders and used quadrature points can best be seen in Table \ref{tab:orders3D}.
+\begin{table}
+	\centering
+	\begin{tabular}{l|cccc}
+		Sparse grids&&&&\\\hline
+		\quad Moment order&1&2&-&-\\
+		\quad Number of quadrature points&25&441&-&-\\
+		Tensorized grids & & & & \\\hline
+		\quad Moment order&1&2&3&4\\
+		\quad Number of quadrature points&8&27&64&125\\
+	\end{tabular}
+	\caption{Order of moments and corresponding number of quadrature nodes used.}
+	\label{tab:orders3D}
+\end{table}
+The following properties emerge:
+\begin{itemize}
+	\item It was not possible to obtain a result for sparse grids with moments of order 3 or higher.
+Any level of the used sparse grids based on Clenshaw Curtis quadrature tested up to level 11 ($72\;705$ quadrature nodes) for order 3, resulted in a failure of the method while computing the Hessian during the Newton iteration.
+	\item Sparse grids required a significant higher number of quadrature points for the sample order of moment in comparison to tensorized grids.
+	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, proving that the nestedness property can potentially increase the methods performance.
+\end{itemize}
+Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids we have used, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and only give results for low orders of moments.
+However, as three stochastical dimensions are difficult to deal with in hyperbolical problems, but not for an UQ analysis in general, this might change for high dimensions.
 \begin{figure}[H]
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
+		\includegraphics[scale=0.18]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
+		\caption{}
 		\label{fig:referenceSolutionsPipe1}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
+		\includegraphics[scale=0.18]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
+		\caption{}
 		\label{fig:referenceSolutionsPipe2}
@@
 		\centering
-		\includegraphics[scale=0.2]{figs/pipe/pipe_mesh.png}
-		\label{fig:referenceSolutionsPipe2}
+		\includegraphics[scale=0.18]{figs/pipe/pipe_mesh.png}
+		\caption{}
+		\label{fig:referenceSolutionsPipeMesh}
+	\end{subfigure}\\
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_ERho.png}
+		\caption{}
+		\label{fig:adIPMSolutionsPipeSGE}
+	\end{subfigure}%
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_VarRho.png}
+		\caption{}
+		\label{fig:adIPMSolutionsPipeSGVar}
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh.}
-	\label{fig:referenceSolution3D}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_refinementLevel.png}
+		\caption{}
+		\label{fig:adIPMSolutionsPipeSGRI}
+	\end{subfigure}\\
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_ERho.png}
+		\caption{}
+		\label{fig:adIPMSolutionsPipeTGE}
+	\end{subfigure}%
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_VarRho.png}
+		\caption{}
+		\label{fig:adIPMSolutionsPipeTGVar}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_refinementLevel.png}
+		\caption{}
+		\label{fig:adIPMSolutionsPipeTGRI}
+	\end{subfigure}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step}
+	\label{fig:solution3D}
 \end{figure}
-\comment{TODO: Pipe testcase}

### 2019-12-13 Paper1/results.tex
updated pipe

@@
 \subsection{2D Euler equations with a three dimensional uncertainty}
-For the last numerical study we again use the Euler equations, but this time not applied to the NACA airfoil, but rather to a bend Sod shock tube experiment.
-The prescribed initial condition describes a gas at rest, but with a discontinuity, i.e.
-as sudden jump, in density as well as energy in the upper part of the tube.
-The density $\rho_u = p/(R \cdot T)$ in the upper part results in $\rho_u = 1.289$ with $p = 101\;325$ Pa, a temperature of $273.15$ K and the specific gas constant for air $R=287.87$.
+For the last numerical study we again use the Euler equations, but this time not in the context of the NACA airfoil, but rather with a bend Sod shock tube experiment.
+The prescribed initial condition for this testcase describes a gas at rest, but with a discontinuity, in density as well as energy in the upper part of the tube.
+The density $\rho_u = p/(R \cdot T)$ in the upper part is set to $\rho_u = 1.289$ with $p = 101\;325$ Pa, a temperature of $273.15$ K and the specific gas constant for dry air $R=287.87$.
 The energy $\rho e_u$ is set to exactly $1.0$.
@@
 The heat capacity ratio $\gamma$ is set to $1.4$ as in the previous studies.
-We again augment the deterministic case by inflicting three uncertainties this time.
-The applied boundary conditions are Euler slip conditions $\bm{v}^T\bm{n}=0$ for the walls of and Dirichlet conditions set to the initial conditions at each end of the tube.
+We again augment the deterministic case by inflicting uncertainties and again increase the number uncertainties to three.
+The applied boundary conditions are Euler slip conditions $\bm{v}^T\bm{n}=0$ for the walls and Dirichlet type boundary conditions at the ends of the tube which set to the deterministic initial conditions.
 The first two uncertainties are the density and energy of the lower part of the shock tube.
@@
 The third uncertainty enters through the position of the shock itself and thus $y_\text{shock} \sim U(1.0, 1.2)$.
-Note, that in contrast to the upper testcases, here we are not interested in the steady state of the system but rather the evolution of the expectancy and variance in time.
-The computational mesh (see Figure \ref{fig:referenceSolutionsPipeMesh}) is an unstructured mesh of $25458$ triangular cells, where all the cells are similar in size.
-The simulations were run until a time of $0.2$s was exceeded and the corresponding results can be seen in Figure \ref{fig:solution3D}.
-The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was again computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
+Note, that in contrast to the upper testcases, here we are not interested in the steady state of the system, but rather the of time evolution of the expectancy and variance.
+The computational mesh (see Figure \ref{fig:referenceSolutionsPipeMesh}) is an unstructured mesh of $25\;458$ triangular cells, where the cells are all similar in size, i.e.
+there are not refined regions as in the NACA testcase.
+The simulations were run until a time of $0.2$s.
+The corresponding results can be seen in Figure \ref{fig:solution3D}.
+The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
 For this testcase we want to compare the results for sparse grids w.r.t.
 tensorized grids in higher stochastical dimensions.
-As we made used of adaptivity, the corresponding moment orders and used quadrature points can best be seen in Table \ref{tab:orders3D}.
+As we made used of adaptivity, the corresponding moment orders and used quadrature points can be seen in Table \ref{tab:orders3D}.
 \begin{table}
@@
 	\item It was not possible to obtain a result for sparse grids with moments of order 3 or higher.
-Any level of the used sparse grids based on Clenshaw Curtis quadrature tested up to level 11 ($72\;705$ quadrature nodes) for order 3, resulted in a failure of the method while computing the Hessian during the Newton iteration.
-	\item Sparse grids required a significant higher number of quadrature points for the sample order of moment in comparison to tensorized grids.
-	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, proving that the nestedness property can potentially increase the methods performance.
+Any level up to level 11 ($72\;705$ quadrature nodes) of the used Clenshaw-Curtis sparse grids, resulted in a failure of the method during the computation of the mapping from the dual variables to the moment vector.
+	\item Sparse grids required a significantly higher number of quadrature points for the same order of moments in comparison to tensorized grids.
+	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, showing that the nestedness property can potentially increase the methods performance.
 \end{itemize}
-Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids we have used, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and only give results for low orders of moments.
-However, as three stochastical dimensions are difficult to deal with in hyperbolical problems, but not for an UQ analysis in general, this might change for high dimensions.
+Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and are only able to generate results for low orders of moments.
+Tensorized quadrature on the other hand performs similarly well as in the previous testcases and should thus be preferred for lower dimensions.
 \begin{figure}[H]

### 2019-12-16 Paper1/results.tex
updated figs adipm solution for pipe

@@
 there are not refined regions as in the NACA testcase.
-The simulations were run until a time of $0.2$s.
+The simulations were run until a time of $2.0$s.
 The corresponding results can be seen in Figure \ref{fig:solution3D}.
 The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
-For this testcase we want to compare the results for sparse grids w.r.t.
-tensorized grids in higher stochastical dimensions.
+For this testcase we want to compare the results for sparse grids with respect to tensorized grids for higher stochastical dimensions, where both quadrature sets are based on Clenshaw-Curtis nodes.
 As we made used of adaptivity, the corresponding moment orders and used quadrature points can be seen in Table \ref{tab:orders3D}.
@@
 		\quad Moment order&1&2&3&4\\
-		\quad Number of quadrature points&8&27&64&125\\
+		\quad Number of quadrature points&27&125&125&729\\
 	\end{tabular}
@@
 Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and are only able to generate results for low orders of moments.
-Tensorized quadrature on the other hand performs similarly well as in the previous testcases and should thus be preferred for lower dimensions.
+Tensorized quadrature on the other hand performs similarly well as in the previous testcases and manages to yield a closer solution for the expectation as well as the variance due to the usage of higher moments.
+Tensorized quadrature rules should thus be preferred for lower stochastical dimensions.
 \begin{figure}[H]
@@
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_ERho.png}
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_tg1-3_ERho.png}
 		\caption{}
@@
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_VarRho.png}
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_tg1-3_VarRho.png}
 		\caption{}
@@
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_refinementLevel.png}
+		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_tg1-3_refinementLevel.png}
 		\caption{}

### 2019-12-16 code/src/pnequations.cpp
stuff deleated

@@ -176,8 +176,6 @@ void PNEquations::SetupSystemMatrices() {
 Vector PNEquations::G( const Vector& u, const Vector& v, const Vector& nUnit, const Vector& n ) {
     // return F( 0.5 * ( u + v ) ) * n - 0.5 * ( v - u ) * norm( n );
     return 0.5 * ( F( u ) + F( v ) ) * n - 0.5 * _AbsAx * ( v - u ) * fabs( n[0] ) - 0.5 * _AbsAz * ( v - u ) * fabs( n[1] );
-    // F( 0.5 * ( u + v ) ) * n - 0.5 * ( ( v - u ) * norm( n ) * nUnit[0] + ( v - u ) * norm( n ) * nUnit[1] );
-    // return F( 0.5 * ( u + v ) ) * n - 0.5 * ( v - u ) * norm( n ) / _settings->GetDT(); // LF does not work since norm(n) != Area(j)
 }
 
 Matrix PNEquations::G( const Matrix& u, const Matrix& v, const Vector& nUnit, const Vector& n, unsigned level ) {

### 2019-12-16 Paper1/results.tex
added refinement barries for pipe
results minors

@@
 The simulations were run until a time of $2.0$s.
+For the refinement barries values of $\delta_{-} = 2\cdot 10^{-2}$ and $\delta_{+} = 4\cdot 10^{-3}$ were set for all adaptive simulations.
 The corresponding results can be seen in Figure \ref{fig:solution3D}.
@@
 	\item It was not possible to obtain a result for sparse grids with moments of order 3 or higher.
-Any level up to level 11 ($72\;705$ quadrature nodes) of the used Clenshaw-Curtis sparse grids, resulted in a failure of the method during the computation of the mapping from the dual variables to the moment vector.
+Any level up to level 11 ($72\;705$ quadrature nodes) of the used Clenshaw-Curtis sparse grids, resulted in an ill conditioned Hessian matrix of the dual problem and thereby to a failure of the method.
 	\item Sparse grids required a significantly higher number of quadrature points for the same order of moments in comparison to tensorized grids.
-	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, showing that the nestedness property can potentially increase the methods performance.
 \end{itemize}

### 2019-12-16 Paper1/results.tex
updated figure alignment

--- 2019-12-16 Paper1/results.tex
@@
 \centering
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_ERho.png}
+				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sc100_ERho.png}
 		\label{fig:referenceSolutionsub1}
-	\end{subfigure}%	
+	\end{subfigure}
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_VarRho.png}
+				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sc100_VarRho.png}
 		\label{fig:referenceSolutionsub2}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/naca_mesh.png}
+				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/naca_mesh.png}
 		\label{fig:referenceSolutionsMesh}
@@
 		\centering
-				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_E[rho]_new.pdf}
+				\includegraphics[width=\linewidth]{figs/StudyQuadrature/L2_error_E[rho]_new.pdf}
 		\caption{}
@@
 		\centering
-				\includegraphics[scale=0.544]{figs/StudyQuadrature/L2_error_Var[rho]_new.pdf}
+				\includegraphics[width=\linewidth]{figs/StudyQuadrature/L2_error_Var[rho]_new.pdf}
 		\caption{}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho].pdf}
+				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_E[rho].pdf}
 		\caption{}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho].pdf}
+				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_Var[rho].pdf}
 		\caption{}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
+				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
 		\caption{}
@@
 		\centering
-				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
+				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
 		\caption{}
@@
 \centering
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
 		\label{fig:sub1}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_ERho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sg_4_9_ERho.png}
 		\label{fig:sub2}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
 		\label{fig:sub1}
@@
 	\vspace{-0.35cm}
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
 		\label{fig:sub1}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
 		\label{fig:sub2}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
 		\label{fig:sub1}
@@
 \centering
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho.png}
 		\label{fig:sub1}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho.png}
 		\label{fig:sub2}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel.png}
+		\includegraphics[width=\linewidth]{figs/Euler1DPlots10/euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel.png}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_E[rho].pdf}
+				\includegraphics[width=\linewidth]{figs/Euler2D/L2_error_E[rho].pdf}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_Var[rho].pdf}
+				\includegraphics[width=\linewidth]{figs/Euler2D/L2_error_Var[rho].pdf}
 		\label{fig:sub2}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_E[rhoE].pdf}
+				\includegraphics[width=\linewidth]{figs/Euler2D/L2_error_E[rhoE].pdf}
 		\label{fig:sub1}
@@
 		\centering
-				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_Var[rhoE].pdf}
+				\includegraphics[width=\linewidth]{figs/Euler2D/L2_error_Var[rhoE].pdf}
 		\label{fig:sub2}
@@
 \centering
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_ERho.png}
+		\includegraphics[width=\linewidth]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_ERho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho_log.png}
+		\includegraphics[width=\linewidth]{figs/Euler2D/sc50UncertainMaSigma0-025PSigma1000-0Dim2_VarRho_log.png}
 		\caption{}
@@
 	\end{subfigure}\\
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/retadosIPM9-4MaP_4_ERho.png}
+		\includegraphics[width=\linewidth]{figs/Euler2D/retadosIPM9-4MaP_4_ERho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/retadosIPM9-4MaP_4_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/Euler2D/retadosIPM9-4MaP_4_VarRho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.329\linewidth}
 		\centering
-		\includegraphics[scale=0.2]{figs/Euler2D/retadosIPM9-4MaP_4_refinementLevel.png}
+		\includegraphics[width=\linewidth]{figs/Euler2D/retadosIPM9-4MaP_4_refinementLevel.png}
 		\caption{}
@@
 	\centering
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
+		\includegraphics[width=\linewidth]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
+		\includegraphics[width=\linewidth]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/pipe_mesh.png}
+		\includegraphics[width=\linewidth]{figs/pipe/pipe_mesh.png}
 		\caption{}
@@
 	\end{subfigure}\\
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_ERho.png}
+		\includegraphics[width=\linewidth]{figs/pipe/euler2D_pipe_adipm_n1-2_ERho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/pipe/euler2D_pipe_adipm_n1-2_VarRho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_refinementLevel.png}
+		\includegraphics[width=\linewidth]{figs/pipe/euler2D_pipe_adipm_n1-2_refinementLevel.png}
 		\caption{}
@@
 	\end{subfigure}\\
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_tg1-3_ERho.png}
+		\includegraphics[width=\linewidth]{figs/pipe/euler2D_pipe_n1-4_tg1-3_ERho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_tg1-3_VarRho.png}
+		\includegraphics[width=\linewidth]{figs/pipe/euler2D_pipe_n1-4_tg1-3_VarRho.png}
 		\caption{}
@@
 	\hfill
-	\begin{subfigure}{0.3\linewidth}
+	\begin{subfigure}{0.31\linewidth}
 		\centering
-		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_tg1-3_refinementLevel.png}
+		\includegraphics[width=\linewidth]{figs/pipe/euler2D_pipe_n1-4_tg1-3_refinementLevel.png}
 		\caption{}

### 2019-12-17 Paper1/introduction.tex

@@
 
-When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty,kusch2018filtered}, IPM \cite{kusch2018filtered,poette2019contribution} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
+When studying hyperbolic equations, the moment approximations of various methods such as stochastic-Galerkin \cite{le2004uncertainty,kusch2018filtered}, IPM \cite{kusch2018filtered,poette2019contribution} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
 These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through either the spatial numerical method \cite{offner2017stability} or by filters \cite{kusch2018filtered}.

### 2019-12-19 Paper1/implementation.tex
final changes before arxive?

@@
 To our best knowledge, this allows a fair comparison of the different intrusive and non-intrusive techniques.
+In the following section, we will study the convergence of the expectation and variance error in pseudo-time.
+Recording this error is straight forward with intrusive methods, however non-intrusive methods only yield expectation value and variance at the final, steady state solution.
+Therefore, to record the error for SC, we have implemented a collocation code, which couples all quadrature points in each time step, allowing the computation of the error in pseudo-time.
+Since this adds additional communication costs, we do not use the run time of this method, but instead make use of the run times from the black-box SC code.
+Thereby, we are able to record the convergence of expectation values and variances in pseudo-time for the non-intrusive SC method without including additional communication costs.

### 2019-12-19 Paper1/one_shot_ipm.tex
final changes before arxive?

@@
 \end{align}
-is fulfilled. Since one is generally interested in low order moments such as the expectation value, this residual can be modified by only accounting for the zero order moments.
+is fulfilled.
+Condition \eqref{eq:residualSteady}, which is for example being used in the SU2 code framework \cite{economon2015su2}, measures the change of the solution by a single time iteration.
+Note that to obtain an estimate of the distance to the steady state solution, one has to include the Lipschitz constant of the corresponding fixed point problem.
+Since one is generally interested in low order moments such as the expectation value, the residual \eqref{eq:residualSteady} can be modified by only accounting for the zero order moments.
 

### 2019-12-19 Paper1/results.tex
final changes before arxive?

@@
 All intrusive methods converge to a residual $\varepsilon=6\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=1\cdot 10^{-7}$.
-All computations are performed with with 5 MPI threads.}
+All computations are performed with 5 MPI threads.}
 	\label{fig:L2ErrorSolution}
@@
 	\end{subfigure}
-	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{2-9}$}
+	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{2-9}$.}
 	\label{fig:readosIPMEVar}
@@
 	\end{subfigure}
-	\caption{Relative L$^2$-error with 16 MPI threads for density and energy.}
+	\caption{Relative L$^2$-error with 16 MPI threads for density and energy.
+All intrusive methods are converged to a residual $\varepsilon=6\cdot 10^{-6}$ and all non-intrusive methods are converged to a residual of $\varepsilon=1\cdot 10^{-6}$.}
 	\label{fig:L2ErrorSolution2D}

### 2019-12-19 Paper1/one_shot_ipm.tex
final change before archive!

@@
 Condition \eqref{eq:residualSteady}, which is for example being used in the SU2 code framework \cite{economon2015su2}, measures the change of the solution by a single time iteration.
-Note that to obtain an estimate of the distance to the steady state solution, one has to include the Lipschitz constant of the corresponding fixed point problem.
+Note, that in order to obtain an estimate of the distance to the steady state solution, one has to include the Lipschitz constant of the corresponding fixed point problem.
 Since one is generally interested in low order moments such as the expectation value, the residual \eqref{eq:residualSteady} can be modified by only accounting for the zero order moments.

### 2019-12-23 code/src/thermalpn.cpp
testcase settings

@@ -20,8 +20,8 @@ ThermalPN::ThermalPN( Settings* settings ) : Problem( settings ) {
     _settings->SetExactSolution( false );
     _settings->SetSource( true );
     _settings->SetImplicitSource( false );
-    _suOlson         = false;
     _constitutiveLaw = 2;    // 1 is Su Olson, 2 is constant
+    _testCase        = 1;    // 0 is Su Olson, 1 is test1, 2 is radShock
 
     // physical constants
     _c             = 299792458.0 * 100.0;          // speed of light in [cm/s]
@@ -29,16 +29,28 @@ ThermalPN::ThermalPN( Settings* settings ) : Problem( settings ) {
     _TRef          = 1.0;                          // reference temperature
     _sigma         = 1.0 / 92.6 / 1e-6 / 100.0;    // opacity
     _alpha         = 4.0 * _a;                     // closure relation, can be changed
-    double density = 2.7;
-    _cV            = density * 0.831 * 1e-7;
-    //_cV            = 0.718 * 1e7;      // heat capacity. Air has cV = 0.718*1e7 in [kJ/(kg K)] = [1000 m^2 / s^2 / K]   density * 0.831 * 1e-7
-    double sigmaSB = 5.6704 * 1e-5;    // Stefan Boltzmann constant in [erg/cm^2/s/K^4]
+    double sigmaSB = 5.6704 * 1e-5;                // Stefan Boltzmann constant in [erg/cm^2/s/K^4]
     _a             = 4.0 * sigmaSB / _c;
-    //_epsilon       = 4.0 * _a / _alpha;
+    if( _testCase == 1 ) {
+        double density = 2.7;
+        _cV            = density * 0.831 * 1e-7;    // heat capacity: [kJ/(kg K)] = [1000 m^2 / s^2 / K] therefore density * 0.831 * 1e-7
+        _cV            = density * 0.831 * 1e3;
+    }
+    else if( _testCase == 2 ) {
+        //_cV = 0.718 * 1e7;
+        _cV = 0.718 * 1e-13;    // heat capacity. Air has cV = 0.718*1e7 in [kJ/(kg K)] = [1000 m^2 / s^2 / K]   density * 0.831 * 1e-7
+    }
+
     _epsilon = 1.0 / _sigma;
 
[…]

### 2019-12-23 code/src/thermalpn.h

@@ -14,7 +14,6 @@ class ThermalPN : public Problem
     double _TRef;
     double _sigma;
     double _alpha;
-    bool _suOlson;
     unsigned _constitutiveLaw;
     Matrix _AbsA;
     std::vector<Vector> _xiQuad;
@@ -29,6 +28,8 @@ class ThermalPN : public Problem
     Matrix _Az;
     unsigned _nMoments;    // number of radiative moments
     unsigned _N;
+
+    unsigned _testCase;
 
     // parameter functions for setting up system matrix
     double AParam( int l, int k ) const;

### 2020-01-04 Paper1/introduction.tex
last changes Martin, Ref NACA0012 missing

@@
 Numerical methods for approximating these coefficients can be divided into intrusive and non-intrusive techniques.
-A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g.
+A popular non-intrusive method is the Stochastic Collocation (SC) method, see e.g.
 \cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule.
-Commonly, SC uses sparse grids, since they posses a reduced number of collocation points for multi-dimensional problems.
+Commonly, SC uses sparse grids, since they possess a reduced number of collocation points for multi-dimensional problems.
 Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort.
@@
 
-When studying hyperbolic equations, the moment approximations of various methods such as stochastic-Galerkin \cite{le2004uncertainty,kusch2018filtered}, IPM \cite{kusch2018filtered,poette2019contribution} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
+When studying hyperbolic equations, the moment approximations of various methods such as stochastic-Galerkin \cite{le2004uncertainty,kusch2018filtered}, IPM \cite{kusch2018filtered,poette2019contribution} and Stochastic Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space.
 These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through either the spatial numerical method \cite{offner2017stability} or by filters \cite{kusch2018filtered}.
 Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{wan2006multi,durrwachter2018hyperbolicity}.
-Non-intrusive Monte-Carlo methods \cite{mishra2012multi,mishra2012sparse,mishra2016numerical}, which randomly sample input uncertainties to compute quantities of interest are robust, but suffer from a slow rate of convergence while again lacking the ability to use adaptivity to their full extend.
+Non-intrusive Monte-Carlo methods \cite{mishra2012multi,mishra2012sparse,mishra2016numerical}, which randomly sample input uncertainties to compute quantities of interest are robust, but suffer from a slow rate of convergence while again lacking the ability to use adaptivity to its full extent.
 Discontinuous structures commonly arise on a small portion of the space-time domain.
@@
 In addition to that, the number of moments needed to obtain a certain order with intrusive methods is smaller than the number of quadrature points for SC.
-An additional downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals.
-Consequently, collocation methods typically require a higher number of unknowns than intrusive methods to reach a given accuracy \cite{xiu2009fast,alekseev2011estimation}.
+Furthermore, collocation methods typically require a higher number of unknowns than intrusive methods to reach a given accuracy \cite{xiu2009fast,alekseev2011estimation}.
 Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex and high-dimensional settings.
@@
 
-In this paper, we propose acceleration techniques for intrusive methods and compare them against stochastic-Collocation.
+In this paper, we propose acceleration techniques for intrusive methods and compare them against Stochastic Collocation.
 For steady and unsteady problems, we use adaptivity, for which intrusive methods provide a convenient framework: %Since complex structures in the uncertain domain tend to arise in small portions of the spatial mesh, our aim is to locally increase the accuracy of the stochastic discretization in region that show a complex structure in the random domain, while choosing a low order method in the remainder:
@@
 
-The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.
+The effectiveness of these acceleration ideas are tested by comparing results with Stochastic Collocation for the uncertain NACA test case as well as a bent shock tube problem.
 Our numerical studies show the following main results:
@@
 While facilitating the task of implementing general intrusive methods, this framework reduces the number of operations required to compute numerical fluxes.
-Also, it provides the ability to base the intrusive method on the same deterministic solver as used in the implementation of a black-box fashion stochastic-Collocation code, which allows for, what we believe, a fair comparison between intrusive and non-intrusive methods.
+Also, it provides the ability to base the intrusive method on the same deterministic solver as used in the implementation of a black-box fashion Stochastic Collocation code, which allows for, what we believe, a fair comparison between intrusive and non-intrusive methods.
 The code is publicly available to allow reproducibility \cite{uqcreator}.
@@
 \end{subequations}
-where the solution $\bm u\in\mathbb{R}^m$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^p$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,p$.
+where the solution $\bm u\in\mathbb{R}^m$ depends on time $t\in\mathbb{R}^+$, spatial position $\bm{x}\in D\subseteq \mathbb{R}^d$ as well as a vector of independent random variables $\bm{\xi}\in\Theta\subseteq\mathbb{R}^p$ with given probability density functions $f_{\Xi,i}(\xi_i)$ for $i = 1,\cdots,p$.
 Hence, the probability density function of $\bm{\xi}$ is $f_{\Xi}(\bm\xi):=\prod_{i=1}^p f_{\Xi,i}(\xi_i)$.
@@
 basis functions when defining $|i|:=\sum_{n = 1}^p |i_n|$.
-Commonly, these functions are chosen to be orthonormal polynomials \cite{wiener1938homogeneous} with respect to the probability function, i.e.
+Commonly, these functions are chosen to be orthonormal polynomials \cite{wiener1938homogeneous} with respect to the probability density function, i.e.
 $\langle \varphi_i \varphi_j \rangle =\prod_{n=1}^p\delta_{i_nj_n}$.
@@
 
-The core idea of the stochastic-Collocation method is to compute the moments in the gPC expansion with a quadrature rule.
+The core idea of the Stochastic Collocation method is to compute the moments in the gPC expansion with a quadrature rule.
 Given a set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by

### 2020-01-04 Paper1/main.tex
clickable refs
pipe testcase first draft; figs not final
last changes Martin, Ref NACA0012 missing

--- 2019-11-22 Paper1/main.tex
@@
 \usepackage{array}
+\usepackage[hidelinks]{hyperref}
 
@@
 Non-intrusive methods allow the usage of a given deterministic solver in a black-box manner, while being embarrassingly parallel.
-However, avoiding intrusive modifications of a given solver takes away the ability to use several inherently intrusive numerical acceleration tools.
+On the other hand, intrusive modifications allow for certain acceleration techniques.
 Moreover, intrusive methods are expected to reach a given accuracy with a smaller number of unknowns compared to non-intrusive techniques.
 This effect is amplified in settings with high dimensional uncertainty.
-A downside of intrusive methods is however the need to guarantee hyperbolicity of the resulting moment system.
+A downside of intrusive methods is the need to guarantee hyperbolicity of the resulting moment system.
 In contrast to stochastic-Galerkin (SG), the Intrusive Polynomial Moment (IPM) method is able to maintain hyperbolicity at the cost of solving an optimization problem in every spatial cell and every time step.\\
-In this work, we propose several acceleration techniques for intrusive methods and study their advantages and shortcomings compared to the non-intrusive stochastic-Collocation method.
+In this work, we propose several acceleration techniques for intrusive methods and study their advantages and shortcomings compared to the non-intrusive Stochastic Collocation method.
 When solving steady problems with IPM, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by using concepts from PDE-constrained optimization.

### 2020-01-04 Paper1/one_shot_ipm.tex
last changes Martin, Ref NACA0012 missing

@@
 Assume that the classical IPM iteration is contractive at its fixed point $\bm{\hat u}^*$.
-Then the Jacobi matrix $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$.
+Then the Jacobian $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$.
 \end{theorem}

### 2020-01-04 Paper1/results.tex
last changes Martin, Ref NACA0012 missing

@@
 The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods.
-To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
+To be able to measure the quality of the obtained solutions, we compute a reference solution using Stochastic Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}.
 In the following, we investigate the L$^2$-error of the variance and the expectation value.
@@
 The collocation runtimes, that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D}, are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error.
-Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e.
+Our Stochastic Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e.
 $5,9$ and $17$ quadrature points.
@@
 The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}.
-To determine the error, we used stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule.
+To determine the error, we used Stochastic Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule.
 As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method.

### 2020-01-04 Paper1/summary_outlook.tex
last changes Martin, Ref NACA0012 missing

@@
 Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
-Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, as shown in this work, it can be helpful to intrusively modify the code in order to fully exploit all acceleration potentials.
+Even though the non-intrusive nature of Stochastic Collocation or Monte Carlo methods allows an easy implementation, as shown in this work, it can be helpful to intrusively modify the code in order to fully exploit all acceleration potentials.
 Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods.

### 2020-01-06 code/src/thermalpn.cpp
thermal pn cV bugfix

@@ -25,16 +25,17 @@ ThermalPN::ThermalPN( Settings* settings ) : Problem( settings ) {
 
     // physical constants
     _kB            = 1.38064852e-16;               // Boltzmann's constant [cm^2 g / (s^2 K)]
+    _sigmaSB       = 5.6704 * 1e-5;                // Stefan-Boltzmann constant [erg⋅cm−2⋅s−1⋅K−4]
     _c             = 299792458.0 * 100.0;          // speed of light in [cm/s]
     _a             = 7.5657 * 1e-15;               // radiation constant [erg/(cm^3 K^4)]
     _TRef          = 1.0;                          // reference temperature
-    _sigma         = 1.0 / 92.6 / 1e-6 / 100.0;    // opacity
+    _sigma         = 1.0 / 9.26 / 1e-6 / 100.0;    // opacity old: 1.0 / 92.6 / 1e-6 / 100.0
     _alpha         = 4.0 * _a;                     // closure relation, can be changed
     double sigmaSB = 5.6704 * 1e-5;                // Stefan Boltzmann constant in [erg/cm^2/s/K^4]
     _a             = 4.0 * sigmaSB / _c;
+    double density = 2.7;
     if( _testCase == 1 ) {
-        double density = 2.7;
-        _cV            = density * 0.831 * 1e-7;    // heat capacity: [kJ/(kg K)] = [1000 m^2 / s^2 / K] therefore density * 0.831 * 1e-7
+        _cV = density * 0.831 * 1e7;    // heat capacity: [kJ/(kg K)] = [1000 m^2 / s^2 / K] therefore density * 0.831 * 1e7
         //_cV            = density * 0.831 * 1e3;
     }
     else if( _testCase == 2 ) {
@@ -49,6 +50,7 @@ ThermalPN::ThermalPN( Settings* settings ) : Problem( settings ) {
         if( _testCase == 2 ) _TRef = pow( _cV / _a, 1.0 / 4.0 );    // ensure eTilde = O(1)
     }
     if( _testCase == 1 ) {
+        //_sigma   = 4000.0 * density;
         _epsilon = 1.0 / _sigma;
         _TRef    = 80.0 * 11604.0;
     }
@@ -402,8 +404,8 @@ Matrix ThermalPN::F( const Matrix& u ) {
[…]

### 2020-01-06 code/src/thermalpn.h

@@ -9,6 +9,7 @@ class ThermalPN : public Problem
   private:
     double _epsilon;
     double _kB;
+    double _sigmaSB;
     double _c;
     double _cV;
     double _a;

### 2020-01-06 Paper1/introduction.tex
added NACA Reference

@@
 
-The effectiveness of these acceleration ideas are tested by comparing results with Stochastic Collocation for the uncertain NACA test case as well as a bent shock tube problem.
+The effectiveness of these acceleration ideas are tested by comparing results with Stochastic Collocation for the NACA test case \cite{jacobs1933characteristics} with uncertainties as well as a bent shock tube problem.
 Our numerical studies show the following main results:

### 2020-01-10 Paper1/one_shot_ipm.tex
Error in one-shot Algorithm corrected

@@
 \begin{align}
-&\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
-&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_{j}^{n+1},\bm{\lambda}_{j+1}^{n+1}\right)\label{eq:oneshotIPMmoment}.
+&\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all }j \label{eq:oneshotIPMdual}\\
+&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_{j}^{n+1},\bm{\lambda}_{j+1}^{n+1}\right) \enskip \text{ for all }j \label{eq:oneshotIPMmoment}.
 \end{align}
@@
 \State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
+\EndFor
+\For{$j=1$ to $N_x$}
 \State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$

### 2020-01-11 Paper1/one_shot_ipm.tex
one shot "IPM" added to text

@@
 We call this method One-Shot IPM, since it is inspired by One-Shot optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables.
-Note that the dual variables from the One-Shot iteration are written without a hat to indicate that they are not the exact solution of the dual problem.
+Note that the dual variables from the One-Shot IPM iteration are written without a hat to indicate that they are not the exact solution of the dual problem.
 

### 2020-01-11 Paper1/CoverLetter/main.tex
cover letter

 %% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.
%Version for spanish users, by dgarhdez

\documentclass[11pt,a4paper,roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
\usepackage[english]{babel}

% moderncv themes
\moderncvstyle{classic}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{blue}                              % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
%\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

% personal data
\name{}{Jonas Kusch}
\title{Cover letter}                               % optional, remove / comment the line if not wanted
\address{Karlsruhe Institute of Technology}{}{Karlsruhe, Germany}% optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
%\phone[mobile]{000-000-000-000}                   % optional, remove / comment the line if not wanted
%\phone[fixed]{+2~(345)~678~901}                    % optional, remove / comment the line if not wanted
%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
\email{jonas.kusch@kit.edu}                               % optional, remove / comment the line if not wanted
%\homepage{www.johndoe.com}                         % optional, remove / comment the line if not wanted
%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}
%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
\recipient{~}{}
\date{\today}
\opening{Dear Editors,}
\closing{Yours sincerely, \\ (on behalf of the authors)}
\makelettertitle

I am sending you our research article entitled \textit{"Intrusive acceleration strategies for Uncertainty Quantification for hyperbolic systems of conservation laws"}. Uncertainty Quantification for hyperbolic problems becomes a challenging task, especially when the solution shows shocks in the random space. To enable the use of adaptivity, we use intrusive methods, which provide a set of equations describing the time evolution of the polynomial chaos (PC) coefficients. The Intrusive Polynomial Moment (IPM) method possesses a large number of desirable quantities that are violated by standard intrusive methods such as stochastic-Galerkin. However, the IPM system requires solving a convex optimization problem which computes the entropy variables corresponding to a given set of PC coefficients. This optimization problem needs to be solved in every spatial cell in each time step, yielding high computational costs.

In our paper, we propose different methods to accelerate the IPM algorithm. First, we introduce a framework using adaptivity in the random space, which allows refining the solution in spatial cells that show shocks or complex structures in the uncertain domain. Furthermore, to accelerate the IPM optimization problem for steady problems, we propose to not solve the IPM optimization problem exactly. Thereby, we converge the entropy variables and PC coefficients to their steady state simultaneously, which significantly speeds up the calculation. We can show that the proposed method has local convergence, which is the same convergence result as provided by the standard IPM algorithm. The effectiveness of the proposed methods is shown by studying an uncertain NACA0012 as well as a pipe test case for uncertainties up to dimension three. The results are compared with Stochastic Collocation methods, showing that the use of the proposed acceleration techniques enables intrusive methods to compete with non-intrusive methods.

The presented theoretical findings and numerical methods could have a considerable impact on the uncertainty quantification community. While non-intrusive techniques are gaining popularity, we believe that the proposed adaptive algorithm makes IPM (and intrusive methods in general) an important tool to quantify uncertainties for hyperbolic problems, which often show uncertain shocks in small portions of the spatial domain. A further novelty of this paper is the IPM code, which can be run for two dimensional domains and high-dimensional uncertainties. The code is made publicly available to guarantee reproducibility of the presented results.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal. All authors have approved the manuscript and agree with its submission to the "Methods and Algorithms for Scientific Computing" category of the SIAM Journal on Scientific Computing.

\vspace{0.5cm}

\makeletterclosing

\end{document}

%% end of file `template.tex'.

### 2020-01-14 Paper1/introduction.tex
typo

@@
 \begin{itemize}
-\item Since complex structures in the uncertain domain tend to arise in small portions of the spatial mesh, our aim is to locally increase the accuracy of the stochastic discretization in region that show a complex structure in the random domain, while choosing a low order method in the remainder.
+\item Since complex structures in the uncertain domain tend to arise in small portions of the spatial mesh, our aim is to locally increase the accuracy of the stochastic discretization in regions that show a complex structure in the random domain, while choosing a low order method in the remainder.
 Such an adaptive treatment cannot be realized with non--intrusive methods, since one needs to break up the black-box approach.

### 2020-01-16 Paper1/CoverLetter/main.tex
Paper to SISC

@@
 
-\documentclass[11pt,a4paper,roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
+\documentclass[10pt,a4paper,roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
 \usepackage[english]{babel}

### 2020-01-16 Paper1/main.tex
funding DFG added
Paper to SISC

@@
 
+\journal{SIAM Journal on Scientific Computing}
+
 \begin{document}
@@
 The effectiveness of the proposed adaptations is demonstrated for multi-dimensional uncertainties in fluid dynamics applications, resulting in the observation of requiring a smaller number of unknowns to achieve a given accuracy when using intrusive methods.
-Furthermore, using the proposed acceleration techniques, our implementation reaches a given accuracy faster than stochastic-Collocation.
+Furthermore, using the proposed acceleration techniques, our implementation reaches a given accuracy faster than Stochastic Collocation.
 \end{abstract}
@@
 
+\section*{Acknowledgment} \noindent
+This work was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – FR 2841/6-1.
+
 \newpage

### 2020-01-30 Paper1/SignificanceAndNovelty/SignificanceAndNovelty.tex
submission JCP

--- 2020-01-16 Paper1/CoverLetter/main.tex
@@
 
-\documentclass[10pt,a4paper,roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
-\usepackage[english]{babel}
+\documentclass[10pt,a4paper]{article}      % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
+%\usepackage[english]{babel}
+\pagenumbering{gobble}
+\setlength\parindent{0pt}
 
-% moderncv themes
-\moderncvstyle{classic}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
-\moderncvcolor{blue}                              % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
-%\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
-%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page
-
-% character encoding
-\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
-%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean
-
-% adjust the page margins
-\usepackage[scale=0.75]{geometry}
-%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
-%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks.
-be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...
-
-% personal data
-\name{}{Jonas Kusch}
-\title{Cover letter}                               % optional, remove / comment the line if not wanted
-\address{Karlsruhe Institute of Technology}{}{Karlsruhe, Germany}% optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
-%\phone[mobile]{000-000-000-000}                   % optional, remove / comment the line if not wanted
-%\phone[fixed]{+2~(345)~678~901}                    % optional, remove / comment the line if not wanted
-%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
-\email{jonas.kusch@kit.edu}                               % optional, remove / comment the line if not wanted
-%\homepage{www.johndoe.com}                         % optional, remove / comment the line if not wanted
-%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
-%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
-%\quote{Some quote}                                 % optional, remove / comment the line if not wanted
-
-% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
-%\makeatletter
-%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
-%\makeatother
-%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS
-
-% bibliography with mutiple entries
-%\usepackage{multibib}
-%\newcites{book,misc}{{Books},{Others}}
 %----------------------------------------------------------------------------------
@@
 \begin{document}
-%-----       letter       ---------------------------------------------------------
-% recipient data
-\recipient{~}{}
-\date{\today}
-\opening{Dear Editors,}
-\closing{Yours sincerely, \\ (on behalf of the authors)}
-\makelettertitle
 
-I am sending you our research article entitled \textit{"Intrusive acceleration strategies for Uncertainty Quantification for hyperbolic systems of conservation laws"}.
-Uncertainty Quantification for hyperbolic problems becomes a challenging task, especially when the solution shows shocks in the random space.
-To enable the use of adaptivity, we use intrusive methods, which provide a set of equations describing the time evolution of the polynomial chaos (PC) coefficients.
-The Intrusive Polynomial Moment (IPM) method possesses a large number of desirable quantities that are violated by standard intrusive methods such as stochastic-Galerkin.
-However, the IPM system requires solving a convex optimization problem which computes the entropy variables corresponding to a given set of PC coefficients.
-This optimization problem needs to be solved in every spatial cell in each time step, yielding high computational costs.
-
-In our paper, we propose different methods to accelerate the IPM algorithm.
+{\centering
+\subsection*{Significance and novelty of this paper}}
+In our paper, we propose different methods to accelerate the Intrusive Polynomial Moment (IPM) method, which is an intrusive technique for uncertainty quantification.
 First, we introduce a framework using adaptivity in the random space, which allows refining the solution in spatial cells that show shocks or complex structures in the uncertain domain.
@@
 The effectiveness of the proposed methods is shown by studying an uncertain NACA0012 as well as a pipe test case for uncertainties up to dimension three.
-The results are compared with Stochastic Collocation methods, showing that the use of the proposed acceleration techniques enables intrusive methods to compete with non-intrusive methods.
+The results are compared with stochastic-Galerkin and Stochastic Collocation methods, showing that the use of the proposed acceleration techniques enables intrusive methods to compete with non-intrusive methods.
 
-The presented theoretical findings and numerical methods could have a considerable impact on the uncertainty quantification community.
 While non-intrusive techniques are gaining popularity, we believe that the proposed adaptive algorithm makes IPM (and intrusive methods in general) an important tool to quantify uncertainties for hyperbolic problems, which often show uncertain shocks in small portions of the spatial domain.
-A further novelty of this paper is the IPM code, which can be run for two dimensional domains and high-dimensional uncertainties.
+A further novelty of this paper is the IPM code, which can be run for two-dimensional domains and high-dimensional uncertainties.
 The code is made publicly available to guarantee reproducibility of the presented results.
 
-We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.
-All authors have approved the manuscript and agree with its submission to the "Methods and Algorithms for Scientific Computing" category of the SIAM Journal on Scientific Computing.
-
-\vspace{0.5cm}
-
-\makeletterclosing
-
+%Filters have been used in various spectral methods such as discontinuous Galerkin or P$_N$ in kinetic theory.
+Their main challenge is the need to pick a reasonable filter strength, which mitigates oscillations while maintaining important solution characteristics.
+To overcome this problem, we propose a new filter, which is based on Lasso regression and allows an automated and local choice of the filter strength.
+We test the filtered SG method for Burgers' and the Euler equations.
+Compared to SG, the results are less oscillatory, yielding an improved approximation of expectation value and variance.
 \end{document}

### 2020-01-30 Paper1/SignificanceAndNovelty/highlights.tex
submission JCP

--- 2020-01-30 Paper1/SignificanceAndNovelty/SignificanceAndNovelty.tex
@@
 %\usepackage[english]{babel}
-\pagenumbering{gobble}
-\setlength\parindent{0pt}
 
@@
 
-{\centering
-\subsection*{Significance and novelty of this paper}}
-In our paper, we propose different methods to accelerate the Intrusive Polynomial Moment (IPM) method, which is an intrusive technique for uncertainty quantification.
-First, we introduce a framework using adaptivity in the random space, which allows refining the solution in spatial cells that show shocks or complex structures in the uncertain domain.
-Furthermore, to accelerate the IPM optimization problem for steady problems, we propose to not solve the IPM optimization problem exactly.
-Thereby, we converge the entropy variables and PC coefficients to their steady state simultaneously, which significantly speeds up the calculation.
-We can show that the proposed method has local convergence, which is the same convergence result as provided by the standard IPM algorithm.
-The effectiveness of the proposed methods is shown by studying an uncertain NACA0012 as well as a pipe test case for uncertainties up to dimension three.
-The results are compared with stochastic-Galerkin and Stochastic Collocation methods, showing that the use of the proposed acceleration techniques enables intrusive methods to compete with non-intrusive methods.
-
-While non-intrusive techniques are gaining popularity, we believe that the proposed adaptive algorithm makes IPM (and intrusive methods in general) an important tool to quantify uncertainties for hyperbolic problems, which often show uncertain shocks in small portions of the spatial domain.
-A further novelty of this paper is the IPM code, which can be run for two-dimensional domains and high-dimensional uncertainties.
-The code is made publicly available to guarantee reproducibility of the presented results.
-
-%Filters have been used in various spectral methods such as discontinuous Galerkin or P$_N$ in kinetic theory.
-Their main challenge is the need to pick a reasonable filter strength, which mitigates oscillations while maintaining important solution characteristics.
-To overcome this problem, we propose a new filter, which is based on Lasso regression and allows an automated and local choice of the filter strength.
-We test the filtered SG method for Burgers' and the Euler equations.
-Compared to SG, the results are less oscillatory, yielding an improved approximation of expectation value and variance.
+\subsection*{Highlights}
+- intrusive methods for uncertainty quantification have high numerical costs\\
+- adaptivity and the One-Shot approach can be used to accelerate intrusive methods\\
+- the proposed techniques are studied for an uncertain NACA0012 and pipe testcase with uncertainties up to order three\\
+- using the discussed acceleration techniques allows intrusive methods to compete with collocation methods \\
+- in our numerical experiments, intrusive methods show a faster convergence to steady state and an improved accuracy compared to collocation methods
 \end{document}

### 2020-01-30 Paper1/adaptivity.tex
Error in one-shot Algorithm corrected
submission JCP

@@
 Translating the idea of the so-called discontinuity sensor which has been defined in \cite{persson2006sub} to uncertainty quantification, we define the polynomial approximation at refinement level $\ell$ as
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \bm{\tilde u}_{\ell} := \sum_{|i|\leq M_{\ell}} \bm{\hat{u}}_i \varphi_i.
-\end{align*}
+\end{align*}\end{linenomath*}
 Now the indicator for a moment vector at level $\ell$ is defined as
-\begin{align}\label{eq:errorIndicator}
+\begin{linenomath*}\begin{align}\label{eq:errorIndicator}
 \bm S_{\ell} := \frac{\langle \left(\bm{\tilde u}_{\ell} - \bm{\tilde u}_{\ell-1}\right)^2\rangle}{\langle \bm{\tilde u}_{\ell}^2\rangle},
-\end{align}
+\end{align}\end{linenomath*}
 where divisions and multiplications are performed element-wise.
@@
 $\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times m}\times\mathbb{R}^{N_{\ell}\times m}\to\mathbb{R}^{N_{\ell}\times m}$, given by
-\begin{align}\label{eq:dualIterationFunctionAd}
+\begin{linenomath*}\begin{align}\label{eq:dualIterationFunctionAd}
 \bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{H}_{\ell}^{-1}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
-\end{align}
+\end{align}\end{linenomath*}
 where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $M_\ell$. The Hessian $\bm{H}_{\ell}$ is given by 
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \bm{H}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{s} (\bm{\lambda}^T\bm{\varphi}_{\ell})\otimes\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{T}.
-\end{align*}
+\end{align*}\end{linenomath*}
 An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times m}\times \mathbb{R}^{N_{\ell_2'}\times m}\times \mathbb{R}^{N_{\ell_3'}\times m}\rightarrow \mathbb{R}^{N_{\ell}\times m}$ and given by
-\begin{align}\label{eq:adaptiveFVUpdate}
+\begin{linenomath*}\begin{align}\label{eq:adaptiveFVUpdate}
 \bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= &\langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T \\&- \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
 \nonumber
-\end{align}
+\end{align}\end{linenomath*}
 Hence, the index vector $\bm\ell'\in\mathbb{N}^{3}$ denotes the refinement levels of the stencil cells, which are used to compute the time updated moment vector at level $\ell$.
@@
 \State $\ell_j^{n+1}\leftarrow \max\{\text{DetermineRefinementLevel}\left(\bm{\lambda}_j^{n+1}\right),\ell_l^*\}$
+\EndFor
+\For{$j=1$ to $N_x$}
 \State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$

### 2020-01-30 Paper1/appendix.tex
continued
minor fix appendix
submission JCP

@@
 In the following, we briefly discuss the number of operations needed when precomputing integrals versus the use of a kinetic flux for Burgers' equation. The stochastic Burgers' equation reads
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \partial_t u + \partial_x \frac{u^2}{2} &= 0,\\
 u(t=0,x,\xi) &= u_{IC}(x,\xi).
-\end{align*}
+\end{align*}\end{linenomath*}
 The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials.
 Choosing the SG ansatz \eqref{eq:SGClosure} and testing with the gPC basis functions yields the SG moment system
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \partial_t \hat u_i + \partial_x \frac12\sum_{n,m = 0}^M \hat u_n \hat u_m \langle \varphi_n\varphi_m\varphi_i \rangle = 0.
-\end{align*}
+\end{align*}\end{linenomath*}
 Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{N\times N}$ gives
-\begin{align*}
-\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = 0
-\end{align*}
+\begin{linenomath*}\begin{align*}
+\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = \bm 0
+\end{align*}\end{linenomath*}
 with $F_i(\bm{\hat u}) = \frac12\bm{\hat u}^T\bm C_i\bm{\hat u}$. Note that $\bm{C}_i$ can be computed analytically, hence choosing a Lax-Friedrichs flux
-\begin{align}\label{eq:numFluxAnalytic}
+\begin{linenomath*}\begin{align}\label{eq:numFluxAnalytic}
 G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i
-\end{align}
+\end{align}\end{linenomath*}
 requires no integral evaluations. Recall, that the numerical flux choice made in this work gives
-\begin{align}\label{eq:numericalFluxIPMBurgers}
+\begin{linenomath*}\begin{align}\label{eq:numericalFluxIPMBurgers}
  \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k),
-\end{align}
+\end{align}\end{linenomath*}
 where $\mathcal{U}$ is the SG ansatz \eqref{eq:SGClosure}.
@@
 These solution values are then used to compute the numerical flux
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 G_i(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) &= \sum_{k=1}^Q w_k g(u^{(\ell)}_k,u^{(r)}_k)a_{ki}f_{\Xi}(\xi_k),
-\end{align*}
+\end{align*}\end{linenomath*}
 which again requires $O(N\cdot Q)$ operations, i.e. the costs are $O(N^2)$. The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.
@@
 In the following, we provide details on the implementation of IPM for the 2D Euler equations.
-We for clarity, we denote the momentum by $m_1 := \rho v_1$ and $m_2:=\rho v_2$ and the energy by $E:=\rho e$.
+For ease of presentation, we denote the momentum by $m_1 := \rho v_1$ and $m_2:=\rho v_2$ and the energy by $E:=\rho e$.
 Then, the vector of conserved variables is $\bm u = (\rho,m_1,m_2,E)^T$.
 The entropy used is
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 s(\bm u) = -\rho \ln \left(\rho^{-\gamma} \left(E - \frac{m_1^2 + m_2^2}{2
 \rho}\right)\right).
-\end{align*}
+\end{align*}\end{linenomath*}
 Now the gradient of the entropy $\nabla_{\bm u} s$ has the components
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \frac{\partial s}{\partial \rho} &= -\ln \left(\rho^{-\gamma}\left(E-\frac{m_1^2+m_2^2}{2 \rho }\right)\right)+\frac{m_1^2+m_2^2}{-2 \rho  E+m_1^2+m_2^2}+\gamma, \\
@@
 \frac{\partial s}{\partial E} &=-\frac1\rho\left(E-\frac{m_1^2+m_2^2}{2 \rho }\right).
-\end{align*}
+\end{align*}\end{linenomath*}
 To compute $\bm u_s(\bm\Lambda) = (\nabla_{\bm u}s)^{-1}(\bm \Lambda)$, we set $\bm \Lambda = \nabla_{\bm u}s(\bm u)$ and rearrange with respect to $\bm u$. Let us define
-%\begin{align*}
+%\begin{linenomath*}\begin{align*}
 %\alpha(\bm\Lambda) := \text{exp}\left(-\frac{0.5 \left(-10 \Lambda_1 \Lambda_4+5 \Lambda_2^2+5 \Lambda_3^2+14 \Lambda_4\right)}{\Lambda_4}\right) \cdot (-\Lambda_4)^{\frac{1}{1+\gamma}}.
-%\end{align*}
-\begin{align*}
+%\end{align*}\end{linenomath*}
+\begin{linenomath*}\begin{align*}
 \alpha(\bm\Lambda) := \text{exp}\left(\frac{ \Lambda_2^2 + \Lambda_3^2 - 2\Lambda_1  \Lambda_4 - 2 \Lambda_4  \gamma }{ 2 \Lambda_4(1-\gamma) } \right) \cdot (-\Lambda_4)^{\frac{1}{1-\gamma}}
-\end{align*}
+\end{align*}\end{linenomath*}
 Then the solution ansatz $\bm u_s$ is given by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \rho(\bm\Lambda) &= \alpha(\bm\Lambda),\qquad m_1(\bm\Lambda) = -\frac{\Lambda_2 \alpha(\bm{\Lambda})}{\Lambda_4},\qquad m_2(\bm\Lambda) = -\frac{\Lambda_3 \alpha(\bm{\Lambda})}{\Lambda_4}, \\
 E(\bm\Lambda) &= -\frac{  \alpha(\bm{\Lambda}) ( -\Lambda_2^2 - \Lambda_3^2 + 2\Lambda_4 ) }{ 2 \Lambda_4^2}.
-\end{align*}
+\end{align*}\end{linenomath*}

### 2020-01-30 Paper1/framework.tex
index fix in paper
p to m
correction root
bm j corrected
one more notation mistake in framework
submission JCP

@@
 Omitting initial conditions and assuming a one-dimensional spatial domain, we can write the IPM system  as
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \partial_t \bm{\hat u}+\partial_x \bm{F}(\bm{\hat u}) = \bm{0}
-\end{align*}
+\end{align*}\end{linenomath*}
 with the flux $\bm{F}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$.
@@
 We choose the discrete unknowns which represent the solution to be the spatial averages over each cell at time $t_n$, given by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \bm{\hat u}_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}\bm{\hat u}_i(t_n,x) dx.
-\end{align*}
+\end{align*}\end{linenomath*}
 If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{ij}^n)_{\vert i\vert\leq M}\in\mathbb{R}^{N\times m}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
-\begin{align}\label{eq:IPMDiscretization}
+\begin{linenomath*}\begin{align}\label{eq:IPMDiscretization}
 \bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{\hat u}_{j}^{n},\bm{\hat u}_{j+1}^{n})- \bm{G}(\bm{\hat u}_{j-1}^{n},\bm{\hat u}_{j}^{n})\right)
-\end{align}
+\end{align}\end{linenomath*}
 for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$. Here, the number of spatial cells is denoted by $N_x$ and the number of time steps by $N_t$.
@@
 When a consistent numerical flux $\bm g:\mathbb{R}^m\times\mathbb{R}^m\to\mathbb{R}^m$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the original problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take the numerical flux
-\begin{align*}
+\begin{linenomath*}\begin{align*}
  \bm{\tilde G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T
-\end{align*}
+\end{align*}\end{linenomath*}
 in \eqref{eq:IPMDiscretization}. Commonly, this integral cannot be evaluated analytically and therefore needs to be approximated by a quadrature rule
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \langle h \rangle \approx \langle h \rangle_{Q} := \sum_{k=1}^Q w_k h(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
-\end{align*}
+\end{align*}\end{linenomath*}
 The approximated numerical flux then becomes
-\begin{align}\label{eq:numericalFluxIPM}
+\begin{linenomath*}\begin{align}\label{eq:numericalFluxIPM}
  \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}(\bm{\hat u}_j^n),\mathcal{U}(\bm{\hat u}_{j+1}^n))\bm{\varphi}^T\rangle^T_Q.
-\end{align}
-Note that the numerical flux requires evaluating the ansatz $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^p \to \mathbb{R}^p$,
-\begin{align*}
+\end{align}\end{linenomath*}
+Note that the numerical flux requires evaluating the ansatz $\mathcal{U}(\bm{\hat u}_j^n)$. To simplify notation, we define $\bm{u}_{s}:\mathbb{R}^m \to \mathbb{R}^m$,
+\begin{linenomath*}\begin{align*}
 \bm{u}_{s}(\bm\Lambda):=\left( \nabla_{\bm{u}} s \right)^{-1}(\bm\Lambda),
-\end{align*}
+\end{align*}\end{linenomath*}
 meaning that the IPM ansatz \eqref{eq:ansatz} at cell $j$ in timestep $n$ can be written as
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \mathcal{U}(\bm{\hat u}_j^n) = \bm{u}_{s}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
-\end{align*}
+\end{align*}\end{linenomath*}
 The computation of the dual variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.
 Therefore, to determine the dual variables for a given moment vector $\bm{\hat{u}}$, the cost function
-\begin{align}\label{eq:L}
+\begin{linenomath*}\begin{align}\label{eq:L}
 L(\bm{\lambda};\bm{\hat{u}}) := \langle s_*(\bm{\lambda}^T \bm\varphi)\rangle_Q - \sum_{i\leq M}\bm{\lambda}_i^T \bm{\hat u}_i
-\end{align}
+\end{align}\end{linenomath*}
 needs to be minimized. Hence, one needs to find the root of
-\begin{align*}
-\nabla_{\bm{\lambda_i}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}_i = \langle \bm u_s(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}_i,
-\end{align*}
+\begin{linenomath*}\begin{align*}
+\nabla_{\bm{\lambda}}L(\bm{\lambda};\bm{\hat{u}}) = \langle \nabla s_*(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u} = \langle \bm u_s(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u},
+\end{align*}\end{linenomath*}
 where we used $\nabla s_* \equiv \bm u_s$.
@@
 Newton's method uses the iteration function $\bm{d}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$,
-\begin{align}\label{eq:dualIterationFunction}
+\begin{linenomath*}\begin{align}\label{eq:dualIterationFunction}
 \bm{d}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{H}(\bm{\lambda})^{-1}\cdot\nabla_{\bm{\lambda}}L(\bm{\lambda};\bm{\hat{u}}),
-\end{align}
-where $\bm H\in\mathbb{R}^{N \cdot n\times N\cdot m}$ is the Hessian of \eqref{eq:L}, given by
-\begin{align*}
+\end{align}\end{linenomath*}
+where $\bm H\in\mathbb{R}^{N \cdot m\times N\cdot m}$ is the Hessian of \eqref{eq:L}, given by
+\begin{linenomath*}\begin{align*}
 \bm{H}(\bm{\lambda}) := \langle \nabla \bm{u}_{s} (\bm{\lambda}^T\bm{\varphi})\otimes\bm{\varphi}\bm{\varphi}^T\rangle_Q^{T}.
-\end{align*}
+\end{align*}\end{linenomath*}
 %Inside the Newton update \eqref{eq:dualIterationFunction}, we abuse notation for better readability by making use of
-%\begin{align*}
+%\begin{linenomath*}\begin{align*}
 %\left(\bm{H}^{-1}\cdot\nabla_{\bm{\lambda}}L\right)_{ij} := \sum_{i' = 1}^N\sum_{j' = 1}^m \left(\bm{H}^{-1}\right)_{m(j-1) + i,m(j'-1) + i'}\cdot\nabla_{\bm{\lambda}}L_{i'j'}.
-%\end{align*}
+%\end{align*}\end{linenomath*}
 The function $\bm d$ will in the following be called dual iteration function. Now, the Newton iteration $l$ for spatial cell $j$ is given by
-\begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}_j}).
-\end{align}
+\begin{linenomath*}\begin{align}\label{eq:dualIteration1}
+\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j).
+\end{align}\end{linenomath*}
 The exact dual state is then obtained by computing the fixed point of $\bm{d}$, meaning that one converges the iteration \eqref{eq:dualIteration1}, i.e.
@@
 To obtain a finite number of iterations for the iteration in cell $j$, the stopping criterion 
-\begin{align}\label{eq:tauCrit}
+\begin{linenomath*}\begin{align}\label{eq:tauCrit}
 \sum_{i=0}^m\left\Vert \nabla_{\bm{\lambda_i}}L(\bm{\lambda}_j^{(l)};\bm{\hat{u}}_j^n) \right\Vert < \tau
-\end{align}
+\end{align}\end{linenomath*}
 is used.
@@
 We now write down the entire scheme: To obtain a more compact notation, we define
-\begin{align}\label{eq:momentIterationFunction}
+\begin{linenomath*}\begin{align}\label{eq:momentIterationFunction}
 \bm{c}\left(\bm{\lambda}_{\ell},\bm{\lambda}_c,\bm{\lambda}_r\right):= \langle \bm u_{s}(\bm{\lambda}_c^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_r^T\bm{\varphi}))\bm{\varphi}^T\rangle_Q^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{\ell}^T\bm{\varphi}),\bm u_{s}(\bm{\lambda}_c^T\bm{\varphi}))\bm{\varphi}^T\rangle_Q^T\right).
-\end{align}
+\end{align}\end{linenomath*}
 The moment iteration is then given by
-\begin{align}\label{eq:momentIteration}
+\begin{linenomath*}\begin{align}\label{eq:momentIteration}
 \bm{\hat u}_j^{n+1} = \bm{c}\left(\bm{\hat\lambda}(\bm{\hat u}_{j-1}^n),\bm{\hat\lambda}(\bm{\hat u}_{j}^n),\bm{\hat\lambda}(\bm{\hat u}_{j+1}^n)\right),
-\end{align}
-where the map from the moment vector to the dual variables, i.e. $\bm{\lambda}(\bm{\hat u}_{j}^n)$, is obtained by iterating
-\begin{align}\label{eq:dualIteration}
+\end{align}\end{linenomath*}
+where the map from the moment vector to the dual variables, i.e. $\bm{\hat\lambda}(\bm{\hat u}_{j}^n)$, is obtained by iterating
+\begin{linenomath*}\begin{align}\label{eq:dualIteration}
 \bm{\lambda}_j^{(l+1)} = \bm{d}(\bm{\lambda}_{j}^{(l)};\bm{\hat u}_j^{n}).
-\end{align}
+\end{align}\end{linenomath*}
 until condition \eqref{eq:tauCrit} is fulfilled. This gives Algorithm \ref{alg:IPM}.

### 2020-01-30 Paper1/introduction.tex
submission JCP

@@
 Due to the randomness of the solution, one is interested in determining the expectation value or the variance, i.e.
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \text{E}[\bm{u}] = \langle \bm{u} \rangle,\qquad \text{Var}[\bm{u}] = \langle \left( \bm{u}-\text{E}[\bm{u}]\right)^2\rangle,
-\end{align*}
+\end{align*}\end{linenomath*}
 where we use the bracket operator $\langle \cdot \rangle := \int_{\Theta} \cdot f_{\Xi}(\bm\xi)d\xi_1 \cdots d\xi_p$.
@@
 Note that this yields
-\begin{align}\label{eq:numberBasisFcts}
+\begin{linenomath*}\begin{align}\label{eq:numberBasisFcts}
 N:=\begin{pmatrix}
@@
 \end{pmatrix}
-\end{align}
+\end{align}\end{linenomath*}
 basis functions when defining $|i|:=\sum_{n = 1}^p |i_n|$.
@@
 The generalized polynomial chaos (gPC) expansion \cite{xiu2002wiener} approximates the solution by
-\begin{align}\label{eq:SGClosure}
+\begin{linenomath*}\begin{align}\label{eq:SGClosure}
 \mathcal{U}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \bm{\hat{u}}^T\bm{\varphi}(\bm\xi),
-\end{align}
+\end{align}\end{linenomath*}
 where the deterministic expansion coefficients $\bm{\hat{u}}_i\in\mathbb{R}^m$ are called moments.
@@
 One can also use the ansatz \eqref{eq:SGClosure} to compute the quantities of interest as
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \text{E}[\mathcal{U}(\bm{\hat u})] = \bm{\hat u}_0,\quad \text{Var}[\mathcal{U}(\bm{\hat u})] = \text{E}[\mathcal{U}(\bm{\hat u})^2] - \text{E}[\mathcal{U}(\bm{\hat u})]^2 = \left(\sum_{i = 1}^N \hat{u}_{\ell i}^2\right)_{\ell = 1,\cdots,m}.
-\end{align*}
+\end{align*}\end{linenomath*}
 
@@
 Given a set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \bm{\hat u}_i = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{Q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
-\end{align*} 
+\end{align*}\end{linenomath*} 
 Hence, the moments can be computed by running a given deterministic solver for the original problem at each quadrature point.
@@
 \begin{subequations}\label{eq:SGMomentSystem}
-\begin{align}
+\begin{linenomath*}\begin{align}
 \partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{f}(\mathcal{U}(\bm{\hat u})) \varphi_i\rangle = \bm{0}, \\
 \bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle,
-\end{align}
+\end{align}\end{linenomath*}
 \end{subequations}
@@
 For a given convex entropy $s:\mathbb{R}^m\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, the IPM solution ansatz is given by
-\begin{align}\label{eq:primalProblem}
+\begin{linenomath*}\begin{align}\label{eq:primalProblem}
 \mathcal{U}(\bm{\hat u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } |i| \leq M.
-\end{align}
+\end{align}\end{linenomath*}
 Rewritten in its dual form, \eqref{eq:primalProblem} is transformed into an unconstrained optimization problem.
 Defining the variables $\bm{\lambda}_i\in\mathbb{R}^m$, where $i$ is again a multi index, gives the unconstrained dual problem
-\begin{align}\label{eq:dualProblem}
+\begin{linenomath*}\begin{align}\label{eq:dualProblem}
  \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{N \times m}}
   \left\{\langle s_*(\bm{\lambda}^T \bm\varphi)\rangle - \sum_{|i|\leq M}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
-\end{align}
+\end{align}\end{linenomath*}
 where $s_*:\mathbb{R}^m\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}:=(\bm{\hat{\lambda}}_i)_{|i|\leq M}\in \mathbb{R}^{N \times m}$ are called the dual variables.
 The solution to \eqref{eq:primalProblem} is then given by
-\begin{align}\label{eq:ansatz}
+\begin{linenomath*}\begin{align}\label{eq:ansatz}
  \mathcal{U}(\bm{\hat u}) = \left( \nabla_{\bm{u}} s \right)^{-1}(\bm{\hat{\lambda}}(\bm{\hat u})^T \bm{\varphi}).
-\end{align}
+\end{align}\end{linenomath*}
 When plugging this ansatz into the original equations \eqref{eq:hyperbolicProblem} and projecting the resulting residual to zero again yields the moment system \eqref{eq:SGMomentSystem}, but with the ansatz \eqref{eq:ansatz} instead of \eqref{eq:SGClosure}.

### 2020-01-30 Paper1/one_shot_ipm.tex
submission JCP

@@
 In the following section we only consider steady state problems, i.e. equation \eqref{eq:fulleq} reduces to
-\begin{align}\label{eq:hyperbolicProblemSteady}
+\begin{linenomath*}\begin{align}\label{eq:hyperbolicProblemSteady}
 \nabla\cdot\bm{f}(\bm{u}(\bm{x},\bm{\xi})) = \bm{0} \enskip \text{ in } D
-\end{align}
+\end{align}\end{linenomath*}
 with adequate boundary conditions.
@@
 Similar to the unsteady case \eqref{eq:hyperbolicProblem}, we can again derive a moment system for \eqref{eq:hyperbolicProblemSteady} given by
-\begin{align}\label{eq:MomentSystemSteady}
+\begin{linenomath*}\begin{align}\label{eq:MomentSystemSteady}
 \nabla\cdot\langle\bm{f}(\bm{u}(\bm{x},\bm{\xi}))\bm{\varphi}^T\rangle^T = \bm{0} \enskip \text{ in } D
-\end{align}
+\end{align}\end{linenomath*}
 which is again needed for the construction of intrusive methods.
@@
 Note that now, the time iteration is not performed for a fixed number of time steps $N_t$, but until the condition
-\begin{align}\label{eq:residualSteady}
+\begin{linenomath*}\begin{align}\label{eq:residualSteady}
 \sum_{j = 1}^{N_x} \Delta x_j \Vert \bm{\hat{u}}_j^n - \bm{\hat{u}}_j^{n-1} \Vert \leq \varepsilon
-\end{align}
+\end{align}\end{linenomath*}
 is fulfilled.
@@
 \begin{subequations}\label{eq:oneshotIPM}
-\begin{align}
+\begin{linenomath*}\begin{align}
 &\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all }j \label{eq:oneshotIPMdual}\\
 &\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_{j}^{n+1},\bm{\lambda}_{j+1}^{n+1}\right) \enskip \text{ for all }j \label{eq:oneshotIPMmoment}.
-\end{align}
+\end{align}\end{linenomath*}
 \end{subequations}
@@
 First, to understand what contraction of the classical IPM iteration implies, we rewrite the moment iteration \eqref{eq:momentIteration} of the classical IPM scheme: When defining the update function
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \bm{\tilde c}\left(\bm{\hat{u}}_{\ell},\bm{\hat{u}}_{c},\bm{\hat{u}}_{r}\right):=\bm{c}\left(\bm{\hat{\lambda}}(\bm{\hat{u}}_{\ell}),\bm{\hat{\lambda}}(\bm{\hat{u}}_{c}),\bm{\hat{\lambda}}(\bm{\hat{u}}_{r})\right)
-\end{align*}
+\end{align*}\end{linenomath*}
 we can rewrite the classical moment iteration as
-\begin{align}\label{eq:shortIPMIt}
+\begin{linenomath*}\begin{align}\label{eq:shortIPMIt}
 \bm{\hat u}_j^{n+1} = \bm{\tilde c}\left(\bm{\hat u}_{j-1}^n,\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^n\right).
-\end{align}
+\end{align}\end{linenomath*}
 Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \nabla_{\bm{\hat u}}\bm{\tilde c} = 
@@
     \end{pmatrix},
-\end{align*}
+\end{align*}\end{linenomath*}
 where we define $\bm{\tilde c}_{j}:=\bm{\tilde c}\left(\bm{\hat u}_{j-1}^*,\bm{\hat u}_{j}^*,\bm{\hat u}_{j+1}^*\right)$ for all $j$.
 Now for each term inside the matrix $\nabla_{\bm{\hat u}}\bm{\tilde c}$ we have
-\begin{align}\label{eq:cTildeDer}
+\begin{linenomath*}\begin{align}\label{eq:cTildeDer}
 \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\hat \lambda}_{\ell}}\frac{\partial \bm{\hat \lambda}(\bm{\hat u}_{j-1}^*)}{\partial \bm{\hat u}},\enskip\partial_{\bm{\hat u}_c}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\hat \lambda}_c}\frac{\partial \bm{\hat \lambda}(\bm{\hat u}_j^*)}{\partial \bm{\hat u}},\enskip\partial_{\bm{\hat u}_r}\bm{\tilde c}_{j} = \frac{\partial \bm{c}_j}{\partial \bm{\hat \lambda}_r}\frac{\partial \bm{\hat \lambda}(\bm{\hat u}_{j+1}^*)}{\partial \bm{\hat u}}.
-\end{align}
+\end{align}\end{linenomath*}
 We first wish to understand the structure of the terms $\partial_{\bm{\hat u}} \bm{\hat \lambda}(\bm{\hat u})$. For this, we note that the exact dual variables fulfill
-\begin{align}\label{eq:ulambda}
+\begin{linenomath*}\begin{align}\label{eq:ulambda}
 \bm{\hat u} = \langle \bm{u}_s(\bm{\hat \lambda}^T\bm{\varphi})\bm{\varphi}^T\rangle^T =: \bm{h}(\bm{\hat \lambda}),
-\end{align}
+\end{align}\end{linenomath*}
 which is why we have the mapping $\bm{\hat u}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{\hat u}(\bm{\hat \lambda}) = \bm{h}(\bm{\hat \lambda})$.
 Since the solution of the dual problem for a given moment vector is unique, this mapping is bijective and therefore we have an inverse function
-\begin{align}\label{eq:lambdau}
+\begin{linenomath*}\begin{align}\label{eq:lambdau}
 \bm{\hat \lambda} = \bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda})).
-\end{align}
+\end{align}\end{linenomath*}
 Now we differentiate both sides w.r.t. $\bm{\hat \lambda}$ to get
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \bm{I}_{d} = \frac{\partial \bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda}))}{\partial \bm{\hat u}}\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}.
-\end{align*}
+\end{align*}\end{linenomath*}
 We multiply with the matrix inverse of $\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}$ to get
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \left(\frac{\partial \bm{\hat u}(\bm{\hat \lambda})}{\partial \bm{\hat \lambda}}\right)^{-1} = \frac{\partial \bm{h}^{-1}(\bm{\hat u}(\bm{\hat \lambda}))}{\partial \bm{\hat u}}.
-\end{align*}
+\end{align*}\end{linenomath*}
 Note that on the left-hand-side we have the inverse of a matrix and on the right-hand-side, we have the inverse of a multi-dimensional function.
@@
 $\bm{\hat \lambda}$, one obtains
-\begin{align}\label{eq:dudlambdaex}
+\begin{linenomath*}\begin{align}\label{eq:dudlambdaex}
 \partial_{\bm{\hat u}} \bm{\hat \lambda}(\bm{\hat u}) = \langle \nabla\bm{u}_s(\bm{\hat \lambda}^T\bm{\varphi})\bm{\varphi}\bm{\varphi}^T\rangle^{-T}.
-\end{align}
+\end{align}\end{linenomath*}
 Now we begin to derive the spectrum of the \textit{One-Shot IPM} iteration \eqref{eq:oneshotIPM}.
@@
 To obtain a fixed point iteration, we plug the dual iteration step \eqref{eq:oneshotIPMdual} into the moment iteration \eqref{eq:oneshotIPMmoment} to obtain
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 &\bm{\lambda}_j^{n+1} = \bm{d}(\bm{\lambda}_j^{n},\bm{\hat u}_j^{n}) \enskip \text{ for all j} \\
 &\bm{\hat u}_j^{n+1} =  \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{\hat u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{\hat u}_{j+1}^{n})\right).
-\end{align*}
+\end{align*}\end{linenomath*}
 The Jacobian $\bm{J}\in\mathbb{R}^{2N\cdot N_x \times 2N\cdot N_x}$ has the form
-\begin{align}\label{eq:Jacobian}
+\begin{linenomath*}\begin{align}\label{eq:Jacobian}
 \bm{J} = 
@@
 \end{pmatrix},
-\end{align}
+\end{align}\end{linenomath*}
 where each block has entries for all spatial cells. We start by looking at $\partial_{\bm{\lambda}} \bm{d}$. For the columns belonging to cell $j$, we have
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\bm\lambda)^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\bm\lambda)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right) \\
 &=- \partial_{\bm{\lambda}}\bm{H}(\bm\lambda)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right).
-\end{align*}
+\end{align*}\end{linenomath*}
 Recall that at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
 For the block $\partial_{\bm{\hat u}} \bm{d}$, we get
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\bm\lambda)^{-1},
-\end{align*}
+\end{align*}\end{linenomath*}
 hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$:
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\lambda}} = \bm{0},
-\end{align*}
+\end{align*}\end{linenomath*}
 since we already showed that by the choice of $\bm{H}(\bm\lambda)^{-1}$ the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
@@
 For the last block, we have that
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
-\end{align*}
+\end{align*}\end{linenomath*}
 by the choice of $\bm{H}(\bm\lambda)^{-1}$ as well as \eqref{eq:cTildeDer} and \eqref{eq:dudlambdaex}.
@@
 for all starting points $(\bm{\lambda}^0,\bm{\hat u}^0)\in B_{\delta}(\bm{\lambda}^*,\bm{\hat u}^*)$ we have
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \Vert (\bm{\lambda}^n,\bm{\hat u}^n) - (\bm{\lambda}^*,\bm{\hat u}^*)\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
-\end{align*}
+\end{align*}\end{linenomath*}
 \end{theorem}
@@
 Furthermore, the determinant of $\bm{\tilde{J}}:=\bm{J}-\lambda \bm{I}_d$ is a polynomial of continuous functions, since
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \text{det}(\bm{\tilde J}) = \sum_{\sigma} \text{sgn}(\sigma)\prod_{i = 1}^{2 N_x N} \tilde J_{\sigma(i),i}.
-\end{align*}
+\end{align*}\end{linenomath*}
 Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{\hat u})$.

### 2020-01-30 Paper1/results.tex
submission JCP

@@
 The stochastic Euler equations in two dimensions are given by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \partial_t
@@
 =\bm{0}.
-\end{align*}
+\end{align*}\end{linenomath*}
 These equations determine the time evolution of the conserved variables $(\rho,\rho \bm v, \rho e)$, i.e. density, momentum and energy. A closure for the pressure $p$ is given by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 p = (\gamma-1)\rho\left(e-\frac12(v_1^2+v_2^2)\right).
-\end{align*}
+\end{align*}\end{linenomath*}
 Here, the heat capacity ratio $\gamma$ is chosen to be $1.4$.
@@
 The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
-\end{align*}
+\end{align*}\end{linenomath*}
 Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
-\begin{align*}
+\begin{linenomath*}\begin{align*}
 \frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
-\end{align*}
+\end{align*}\end{linenomath*}
 The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.

### 2020-05-18 Paper1/ResponseToReviewers/Response_to_Reviewer.tex
more review changes

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand*\xoverline[2][0.75]{%
    \sbox{\myboxA}{$\m@th#2$}%
    \setbox\myboxB\null% Phantom box
    \ht\myboxB=\ht\myboxA%
    \dp\myboxB=\dp\myboxA%
    \wd\myboxB=#1\wd\myboxA% Scale phantom
    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
    \addtolength\mylenA{-\the\wd\myboxB}%
    \ifdim\wd\myboxB<\wd\myboxA%
       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
    \else
        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
    \fi}
\makeatother
\usepackage{bm}

\title{Reply to Reviewers}
\author{Martin Frank, Jonas Kusch and Jannick Wolters}
\date{\today}                                           % Activate to display a given date or no date

\setlength{\parindent}{0pt}

\begin{document}
\maketitle

We would like to thank the referee for taking the time to carefully read the article. The suggestions and corrections have been helpful and increase the quality of this article. The document has been edited under consideration of the remarks made in the review.

\vspace{1em}

The attached documents contain a version of the paper where all changes are highlighted in blue. In the following, we address each of the remarks individually:

\begin{enumerate}
\item The paper is globally well written, with all the key arguments and algorithms clearly explained. The reference list is comprehensive.
\item Some notations are non fully rigorous. Typical is line 182 page 5 where $\bm{\hat u}_i$
is the i-th moment (scalar function of x and t), while $\bm{\hat u}_j$ in line 225 page
6 is something else (a vector, the index is the cell number). I recommend
to check all these kinds of formal inaccuracies.
\\ \textbf{Response:} We added a time index to $\bm{\hat u}_j$ which was missing in the original version. Furthermore, we added a comment when introducing the moments at the spacial cells in line TODO in order to avoid confusion.
\item Is algorithm 1 (page 7) similar to the one by Poette? Can you explain the differences?
\\ \textbf{Response:} Algorithm 1 is very similar to the original IPM algorithm by Poette. The only difference is that we are not using the old moments $\bm{\hat u}_j^{n+1}$ in the finite volume update (line 11 in algorithm 1), but we are using the term $\langle \bm u_{s}((\bm{\lambda}_j^n)^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T$. This step ensures that the error from Newton's method does not yield non-physical moments which can lead to a failure of the method, see Section 4.3 in [8].
\item The formulation of Th. 1 is incomplete because $\bm\lambda_*$ is not specified. Are
conditions on $\bm\lambda_*$? It must be explained.
\\ \textbf{Response:} We added an explanation for $\bm\lambda^*$, see TODO. Furthermore, to underline that $\bm\lambda^*$ collects the dual variables at all spatial cells, we now denote $\bm\lambda^*$ as $\bm{\lambda}_{\Delta}^*=(\bm{\lambda}_{1}^*,\cdots,\bm{\lambda}_{N_x}^*)^T$.
\item The super-attractive property of the Newton-Raphson method is well
known, and in my mind, this is exactly what means $\partial_{\lambda}d = 0$. Can you
confirm?
\\ \textbf{Response:} Yes, this is correct. At the fixed point $\partial_{\lambda}d = 0$ which comes from choosing the Newton-Raphson method to iterate the dual variables.
\item Line 357 page 10 is impossible to understand since the right hand side
seems independent of $j$ and $n$. The same for line 364 same page. Actually
it is mandatory that the dependence with respect to $j$ and $n$ are the same
throughout the whole proof.
\\ \textbf{Response:} You are correct. We forgot to include indices and the manuscript lacked a clear notation at some points. We now added indices and define our notation more clearly. 
\item And also my own calculations from (11) yield that $\partial_u d = -H(\lambda)^{-1}\lambda$. Can you check?
\\ \textbf{Response:} We have that
\begin{align*}
\bm d(\bm{\lambda},\bm{\hat u}) =& \bm{\lambda} - \bm{H}^{-1}(\bm\lambda)\left(\langle \bm u_s(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T - \bm{\hat u}\right) \\
=& \bm{\lambda} - \bm{H}^{-1}(\bm\lambda)\langle \bm u_s(\bm{\lambda}^T \bm\varphi)\bm\varphi^T\rangle_Q^T +\bm{H}^{-1}(\bm\lambda) \bm{\hat u}
\end{align*}
Since only the last term $\bm{H}^{-1}(\bm\lambda) \bm{\hat u}$ depends on the moment vector $\bm{\hat u}$, we have
\begin{align*}
\partial_{\bm{\hat u}} d(\bm{\lambda},\bm{\hat u}) = \bm{H}^{-1}(\bm\lambda).
\end{align*}
Note that for the One-Shot method, the moments $\bm{\hat u}$ and dual variables $\bm\lambda$ do not depend on each other.  
\item page 12 in Algo. 3, it would be better to write ”Determine Refinement
Level ” instead of ”DetermineRefinementLevel ”. Please check all similar
notations.
\\ \textbf{Response:} We have changed this in the manuscript.
\item Line 496 page 14, I do not understand the need to specify the temperature $T$ since it does not show up in line 487.
\\ \textbf{Response:} We added a short description on how the defined quantities relate to the conserved variables TODO.
\item I missed the definition of the reference solution $u_{\Delta}$. It must be explicit
(or more visible).
\item Figure 3: the threshold $\varepsilon = 10^{-6}$ is not visible on the plots, as it should
be. Also the intrusive and non intrusives methods must be more explicit:
which ones (number of equations, . . . ).
\item Lines 553/554 page 17 seem weird. What do you call positivity of mass?
Is it total mass? If yes, there is not problem with a conservative scheme.
\item Line 691 in the conclusion (Additionally, the ability to adaptively change
692 the truncation order helps intrusive methods to compete with SC in
terms of computational runtime) are an interesting asset of the methods
presented in this work.
\end{enumerate}

\end{document}

### 2020-05-18 Paper1/framework.tex
more review changes

@@
 for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$. Here, the number of spatial cells is denoted by $N_x$ and the number of time steps by $N_t$.
-The numerical flux is assumed to be consistent, i.e. $\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$.
+The numerical flux is assumed to be consistent, i.e.
+$\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$.
+\comment{The moment vector inside a spatial cell, denoted by $\bm{\hat u}_{j}^{n}$ is not to be confused with the moment of order $i$ in \eqref{eq:SGClosure}, which is a continuous function in time and space.
+Note that whenever the moment vector includes a time index $n$, we are speaking of a discrete quantity.
+Furthermore, we use the index $j\in\mathbb{N}$ to denote spatial cells and the multi-index $i\in\mathbb{N}_0^p$ to denote the moment order.}
 
@@
 \begin{linenomath*}\begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j).
+\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j^{\comment{n}}).
 \end{align}\end{linenomath*}

### 2020-05-18 Paper1/one_shot_ipm.tex
more review changes

@@
 \end{subequations}
-This yields Algorithm \ref{alg:osIPM}.
+This yields Algorithm~\ref{alg:osIPM}.
 \begin{algorithm}[H]
@@
 \begin{theorem}\label{th:Contractive}
-Assume that the classical IPM iteration is contractive at its fixed point $\bm{\hat u}^*$.
-Then the Jacobian $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$.
+\comment{We denote the fixed point of the classical IPM iteration \eqref{eq:momentIteration} by $\bm{\hat u}^*_{\Delta} = (\bm{\hat u}_{1}^*,\cdots,\bm{\hat u}_{N_x}^*)^T$, i.e. we have
+\begin{linenomath*}\begin{align*}
+\bm{\hat u}_j^* = \bm{c}\left(\bm{\hat\lambda}(\bm{\hat u}_{j-1}^*),\bm{\hat\lambda}(\bm{\hat u}_{j}^*),\bm{\hat\lambda}(\bm{\hat u}_{j+1}^*)\right)
+\end{align*}\end{linenomath*}
+for all spatial cells $j$.
+The fixed point for the One-Shot IPM iteration \eqref{eq:oneshotIPM} is denoted by $(\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)$, where $\bm{\lambda}_{\Delta}^*=(\bm{\lambda}_{1}^*,\cdots,\bm{\lambda}_{N_x}^*)^T$.
+The fixed point of the One-Shot iteration fulfills
+\begin{subequations}
+\begin{linenomath*}\begin{align*}
+&\bm{\lambda}_{j}^{*} =  \bm{d}(\bm{\lambda}_j^{*},\bm{u}_j^{*}),\\
+&\bm{u}_j^{*} =  \bm{c}\left(\bm{\lambda}_{j-1}^{*},\bm{\lambda}_{j}^{*},\bm{\lambda}_{j+1}^{*}\right),
+\end{align*}\end{linenomath*}
+\end{subequations}
+for all spatial cells $j$.
+}
+Assume that the classical IPM iteration is contractive at its fixed point.
+Then the Jacobian $\bm{J}$ of the One-Shot IPM iteration \eqref{eq:oneshotIPM} has a spectral radius $\rho(\bm{J})<1$ at the fixed point $(\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)$
 \end{theorem}
@@
 \end{align}\end{linenomath*}
-Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
+Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}_{\comment{\Delta}}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
 \begin{linenomath*}\begin{align*}
@@
 \begin{linenomath*}\begin{align*}
-\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\bm\lambda)^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\bm\lambda)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right) \\
-&=- \partial_{\bm{\lambda}}\bm{H}(\bm\lambda)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}\right).
+\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \comment{\bm{\hat u}_j^n}\right) \\
+&=- \partial_{\bm{\lambda}}\bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \comment{\bm{\hat u}_j^n}\right).
 \end{align*}\end{linenomath*}
-Recall that at the fixed point $(\bm{\lambda}^*,\bm{\hat u}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
+\comment{Recall that if $\bm{\hat u}_j^n$ and $\bm{\lambda}_j^n$ belong to the fixed point $(\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)$,} we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \comment{\bm{\hat u}_j^n}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
 For the block $\partial_{\bm{\hat u}} \bm{d}$, we get
 \begin{linenomath*}\begin{align*}
-\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\bm\lambda)^{-1},
+\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\comment{\bm\lambda_j^n})^{-1},
 \end{align*}\end{linenomath*}
-hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$:
+hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$. \comment{To shorten notation, let us define
 \begin{linenomath*}\begin{align*}
-\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\lambda}} = \bm{0},
+\bm{c}_j := \bm{c}\left(\bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n}),\bm{d}(\bm{\lambda}_{j}^{n},\bm{\hat u}_{j}^{n}),\bm{d}(\bm{\lambda}_{j+1}^{n},\bm{\hat u}_{j+1}^{n})\right).
 \end{align*}\end{linenomath*}
-since we already showed that by the choice of $\bm{H}(\bm\lambda)^{-1}$ the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
+Then we have}
+\begin{linenomath*}\begin{align*}
+\frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\lambda}} = \bm{0},
+\end{align*}\end{linenomath*}
+since we already showed that the term $\partial_{\bm{\lambda}} \bm{d}$ is zero.
 We can show the same result for all spatial cells and all inputs of $\bm{c}$ analogously, hence $\partial_{\bm{\lambda}} \bm{c} = \bm{0}$.
@@
 \begin{linenomath*}\begin{align*}
-\frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
+\frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
 \end{align*}\end{linenomath*}
@@
 \begin{theorem}\label{th:localConvergence}
-With the assumptions from Theorem \ref{th:Contractive}, the One-Shot IPM converges locally, i.e.
-there exists a $\delta>0$ s.t.
-for all starting points $(\bm{\lambda}^0,\bm{\hat u}^0)\in B_{\delta}(\bm{\lambda}^*,\bm{\hat u}^*)$ we have
+With the assumptions from Theorem \ref{th:Contractive}, the One-Shot IPM converges locally.
+I.e.
+\comment{if we denote the solution at iteration $n$ as $(\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)})$}, there exists a $\delta>0$ s.t.
+for all starting points $(\bm{\lambda}_{\comment{\Delta}}^{(0)},\bm{\hat u}_{\comment{\Delta}}^{(0)})\in B_{\delta}(\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)$ we have
 \begin{linenomath*}\begin{align*}
-\Vert (\bm{\lambda}^n,\bm{\hat u}^n) - (\bm{\lambda}^*,\bm{\hat u}^*)\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
+\left\Vert (\bm{\lambda}\comment{_{\Delta}^{(n)}},\bm{\hat u}\comment{_{\Delta}^{(n)}}) - (\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)\right\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
 \end{align*}\end{linenomath*}
@@
 \end{align*}\end{linenomath*}
-Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda},\bm{\hat u})$.
+Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda}_{\comment{\Delta}},\bm{\hat u}_{\comment{\Delta}})$.
 Hence there exists an open ball with radius $\delta$ around the fixed point in which the eigenvalues remain in the interval $(-1,1)$.

### 2020-05-18 Paper1/results.tex
more review changes

@@
 At the airfoil boundary, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal.
-At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of $273.15$ K.
+At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of \comment{$T=273.15$ K.
+Note that with the specific gas constant $R=287.87\frac{J}{kg\cdot K}$ and a given angle of attack $\phi$ this fixes the conserved variables at the far field boundary, since
+\begin{linenomath*}\begin{align*}
+\rho = \frac{p}{R\cdot T}, \quad \rho\bm{v} = \rho\cdot\text{Ma}\cdot \sqrt{\gamma R T} \begin{pmatrix}
+\cos\phi \\ 
+\sin\phi
+\end{pmatrix}, \quad \rho e = \frac{p}{\gamma - 1} + \frac{1}{2}\rho (v_1^2 + v_2^2).
+\end{align*}\end{linenomath*}}
 Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees, i.e.

### 2020-05-19 Paper1/ResponseToReviewers/Response_to_Reviewer.tex
review done?

@@
 \title{Reply to Reviewers}
-\author{Martin Frank, Jonas Kusch and Jannick Wolters}
+\author{Jonas Kusch, Jannick Wolters and Martin Frank}
 \date{\today}                                           % Activate to display a given date or no date
@@
 \begin{align*}
-\partial_{\bm{\hat u}} d(\bm{\lambda},\bm{\hat u}) = \bm{H}^{-1}(\bm\lambda).
+\partial_{\bm{\hat u}} \bm d(\bm{\lambda},\bm{\hat u}) = \bm{H}^{-1}(\bm\lambda).
 \end{align*}
@@
 (or more visible).
-\item Figure 3: the threshold $\varepsilon = 10^{-6}$ is not visible on the plots, as it should
-be. Also the intrusive and non intrusives methods must be more explicit:
-which ones (number of equations, . . . ).
+\\ \textbf{Response:} We compute the reference solution with a finely resolved collocation method.
+The one-dimensional testcase uses 100 Gauss-Legendre quadrature points (as mentioned in line 506 in the original manuscript), the two-dimensional testcase uses $50^2$ quadrature points (as mentioned in line 620 in the original manuscript) and the three-dimensional testcase uses $50^3$ quadrature points (as mentioned in line 661 in the original manuscript).
+You are correct that this information gets lost in the text which is why we added a short description whenever we plot reference solutions (which we do in Figures 1, 7 and 8).
+\item Figure 3: the threshold $\varepsilon = 10^{-6}$ is not visible on the plots, as it should be.
+Also the intrusive and non intrusives methods must be more explicit: which ones (number of equations, .
+.
+.
+).
+\\ \textbf{Response:} We added the names of intrusive and non-intrusive methods and include references to the corresponding algorithms.
+The residual cannot be depicted in the plot itself, since the y-axis is the relative error, not the residual.
+We added references to the relative error as well as to the residual equations to avoid confusion.
 \item Lines 553/554 page 17 seem weird. What do you call positivity of mass?
 Is it total mass? If yes, there is not problem with a conservative scheme.
+\\ \textbf{Response:} We meant to say density, which we changed in the manuscript.
+The total mass (here the expected value of the total mass) is conserved by the scheme, however the solution at individual values for $\xi$ might yield negative densities, since we are using polynomials to span the solution in $\xi$.
 \item Line 691 in the conclusion (Additionally, the ability to adaptively change

### 2020-05-19 Paper1/results.tex
review done?

@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.
+\comment{The reference solution has been computed with stochastic-Collocation using $100$ Gauss-Legendre quadrature points.}}
 	\label{fig:referenceSolution}
@@
 Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
-\begin{linenomath*}\begin{align*}
+\begin{linenomath*}\begin{align}\label{eq:relErrorEVar}
 \frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
-\end{align*}\end{linenomath*}
+\end{align}\end{linenomath*}
 The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.
@@
 	\end{subfigure}
-	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-All intrusive methods converge to a residual $\varepsilon=6\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=1\cdot 10^{-7}$.
-All computations are performed with 5 MPI threads.}
+	\caption{Comparison of the relative L$^2$-error \comment{\eqref{eq:relErrorEVar}} for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
+All intrusive methods \comment{(i.e.
+SG, IPM, osIPM, adosIPM, readosIPM) are converged} to a residual $\varepsilon=6\cdot 10^{-6}$ \comment{(see equation \eqref{eq:residualSteady})}, whereas all non-intrusive methods \comment{(i.e.
+SC) are converged} to a residual $\varepsilon=1\cdot 10^{-7}$ \comment{in order to obtain an error which remains constant in time.
+SG and IPM are computed with Algorithm~\ref{alg:IPM} (where SG uses a quadratic entropy) adapted for steady problems.
+For osIPM we used Algorithm~\ref{alg:osIPM}, for adosIPM and readosIPM we use Algorithm~\ref{alg:adosIPM} with and without refinement retardation.} All computations are performed with 5 MPI threads.}
 	\label{fig:L2ErrorSolution}
@@
 
-First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration.
-The standard SG method however preserves positivity of mass, energy and pressure.
-The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance.
+First, let us mention that the adaptive SG method fails, since it yields negative \comment{gas} densities during the iteration.
+The standard SG method \comment{without adaptivity} however preserves positivity of \comment{density}, energy and pressure \comment{in our numerical experiments.
+Note that in general, stochastic-Galerkin does not guarantee positivity of these quantities, since the polynomial approximation does not guarantee any solution bounds.} The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance.
 When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
@@
 	\caption{Reference solution (\subref{fig:referenceSolutionsub2DE}) E$[\rho]$ and (\subref{fig:referenceSolutionsub2DVar}) Var$[\rho]$ and with logarithmic scaled (\subref{fig:referenceSolutionsub2DVarLog}) Var$[\rho]$.
-readosIPM$_{1-9}$ solution for (\subref{fig:reados2DE}) E$[\rho]$ and (\subref{fig:reados2DVar}) Var$[\rho]$, with the resulting (\subref{fig:readosIPMEVar2DRefinement}) refinement levels.}
+\comment{The reference solution has been computed with stochastic-Collocation using $50^2$ Gauss-Legendre quadrature points.} readosIPM$_{1-9}$ solution for (\subref{fig:reados2DE}) E$[\rho]$ and (\subref{fig:reados2DVar}) Var$[\rho]$, with the resulting (\subref{fig:readosIPMEVar2DRefinement}) refinement levels.}
 	\label{fig:refAndreadosIPMEVar2D}
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step.
+\comment{The reference solution has been computed with stochastic-Collocation using $50^3$ Gauss-Legendre quadrature points.}}
 	\label{fig:solution3D}

### 2020-05-19 Paper1/ResponseToReviewers/Response_to_Reviewer.tex
linenumbers added

@@
 \\ \textbf{Response:} We added a time index to $\bm{\hat u}_j$ which was missing in the original version.
-Furthermore, we added a comment when introducing the moments at the spacial cells in line TODO in order to avoid confusion.
+Furthermore, we added a comment when introducing the moments at the spacial cells in line 189 in order to avoid confusion.
 \item Is algorithm 1 (page 7) similar to the one by Poette? Can you explain the differences?
@@
 conditions on $\bm\lambda_*$? It must be explained.
-\\ \textbf{Response:} We added an explanation for $\bm\lambda^*$, see TODO.
+\\ \textbf{Response:} We added an explanation for $\bm\lambda^*$, see line 312.
 Furthermore, to underline that $\bm\lambda^*$ collects the dual variables at all spatial cells, we now denote $\bm\lambda^*$ as $\bm{\lambda}_{\Delta}^*=(\bm{\lambda}_{1}^*,\cdots,\bm{\lambda}_{N_x}^*)^T$.
@@
 \item Line 496 page 14, I do not understand the need to specify the temperature $T$ since it does not show up in line 487.
-\\ \textbf{Response:} We added a short description on how the defined quantities relate to the conserved variables TODO.
+\\ \textbf{Response:} We added a short description on how the defined quantities relate to the conserved variables in line 517.
 \item I missed the definition of the reference solution $u_{\Delta}$. It must be explicit

### 2020-05-25 code/src/euler.cpp
continued hyperbolicitylimiter

@@ -99,9 +99,9 @@ Vector Euler::IC( const Vector& x, const Vector& xi ) {
     double gamma = 1.4;
 
     double rhoL = 1.0;
-    double rhoR = 0.1;
+    double rhoR = 0.125;
     double pL   = 1.0;
-    double pR   = 0.125;
+    double pR   = 0.1;
     double uL   = 0.0;
     double uR   = 0.0;
 
@@ -116,7 +116,7 @@ Vector Euler::IC( const Vector& x, const Vector& xi ) {
         y[0]                  = rhoL;
         y[1]                  = rhoL * uL;
         double kineticEnergyL = 0.5 * rhoL * pow( uL, 2 );
-        double innerEnergyL   = ( pL / ( rhoL * ( gamma - 1 ) ) ) * rhoL;
+        double innerEnergyL   = pL / ( gamma - 1.0 );
         y[2]                  = kineticEnergyL + innerEnergyL;
     }
     else {
@@ -146,8 +146,8 @@ Matrix Euler::ExactSolution( double t, const Matrix& x, const Vector& xi ) const
     double rho_l      = 1.0;
     double P_l        = 1.0;
     double u_l        = 0.0;
-    double rho_r      = 0.1;
-    double P_r        = 0.125;
+    double rho_r      = 0.125;
+    double P_r        = 0.1;
     double u_r        = 0.0;
 
     if( !sodShockTube ) {

### 2020-05-25 code/src/hyperbolicitylimiter.cpp

@@ -3,24 +3,41 @@
 
 HyperbolicityLimiter::HyperbolicityLimiter( Settings* settings )
     : Closure( settings ), _gamma( _settings->GetGamma() ), _lambda( _settings->GetFilterStrength() ) {
-    _alpha = 1.0;
+    _alpha          = 1.0;
+    double epsilonM = std::numeric_limits<double>::denorm_min();
+    _c              = log( epsilonM );
     if( _lambda < 0 ) _lambda = 0.0;
     unsigned nMoments = _settings->GetNMoments();
     _filterFunction   = Vector( _settings->GetNTotal(), 1.0 );
-    for( unsigned s = 0; s < _settings->GetNStates(); ++s ) {
-        for( unsigned i = 0; i < _settings->GetNTotal(); ++i ) {
-            for( unsigned l = 0; l < _settings->GetNDimXi(); ++l ) {
-                // if( _settings->GetDistributionType( l ) == DistributionType::D_LEGENDRE ) n = 0;
-                // if( _settings->GetDistributionType( l ) == DistributionType::D_HERMITE ) n = 1;
-                unsigned index =
-                    unsigned( ( i - i % unsigned( std::pow( nMoments + 1, l ) ) ) / unsigned( std::pow( nMoments + 1, l ) ) ) % ( nMoments + 1 );
-                _filterFunction[i] *= 1.0 / ( 1.0 + _lambda * pow( index, 2 ) * pow( index + 1, 2 ) );
-            }
-        }
+
+    try {
+        auto file = cpptoml::parse_file( _settings->GetInputFile() );
+
+        auto problem = file->get_table( "moment_system" );
+        _filterOrder = problem->get_as<double>( "filterOrder" ).value_or( 1 );
+    } catch( const cpptoml::parse_exception& e ) {
+        _log->error( "[SplineFilter] Failed to parse {0}: {1}", _settings->GetInputFile(), e.what() );
+        exit( EXIT_FAILURE );
     }
+
+    std::cout << "-------------" << std::endl;
+    for( unsigned i = 0; i < _settings->GetNTotal(); ++i ) {
+        for( unsigned l = 0; l < _settings->GetNDimXi(); ++l ) {
+            // if( _settings->GetDistributionType( l ) == DistributionType::D_LEGENDRE ) n = 0;
[…]

### 2020-05-25 code/src/hyperbolicitylimiter.h

@@ -9,7 +9,10 @@ class HyperbolicityLimiter : public Closure
     HyperbolicityLimiter() = delete;
     double _gamma;
     Vector _filterFunction;
+    unsigned _filterOrder;    // order of the filter
     double _lambda;
+    double _c;    // machine precision constant
+    double FilterFunction( double eta ) const;
 
   public:
     HyperbolicityLimiter( Settings* settings );

### 2020-05-25 Paper1/ResponseToReviewers/CRediT.tex
Revised Submission JCP

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand*\xoverline[2][0.75]{%
    \sbox{\myboxA}{$\m@th#2$}%
    \setbox\myboxB\null% Phantom box
    \ht\myboxB=\ht\myboxA%
    \dp\myboxB=\dp\myboxA%
    \wd\myboxB=#1\wd\myboxA% Scale phantom
    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
    \addtolength\mylenA{-\the\wd\myboxB}%
    \ifdim\wd\myboxB<\wd\myboxA%
       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
    \else
        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
    \fi}
\makeatother
\usepackage{bm}

\title{Reply to Reviewers}
\author{Jonas Kusch, Jannick Wolters and Martin Frank}
\date{\today}                                           % Activate to display a given date or no date

\setlength{\parindent}{0pt}

\begin{document}
\textbf{Jonas Kusch:} Conceptualization, Software, Validation, Formal analysis, Investigation, Writing - Original Draft, Writing - Review \& Editing \\
\textbf{Jannick Wolters:} Software, Validation, Investigation,  Data Curation, Writing - Original Draft, Visualization \\
\textbf{Martin Frank:} Methodology, Writing - Original Draft, Supervision, Funding acquisition  
\end{document}

### 2020-07-03 Paper1/SignificanceAndNovelty/highlights.tex
Paper version elsevier

@@
 - adaptivity and the One-Shot approach can be used to accelerate intrusive methods\\
-- the proposed techniques are studied for an uncertain NACA0012 and pipe testcase with uncertainties up to order three\\
-- using the discussed acceleration techniques allows intrusive methods to compete with collocation methods \\
-- in our numerical experiments, intrusive methods show a faster convergence to steady state and an improved accuracy compared to collocation methods
+- proposed techniques are studied for uncertainties up to order three\\
+- acceleration techniques allow intrusive methods to compete with collocation methods
 \end{document}

### 2020-07-03 Paper1/framework.tex
Paper version elsevier

--- 2020-01-30 Paper1/framework.tex
@@
 for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$. Here, the number of spatial cells is denoted by $N_x$ and the number of time steps by $N_t$.
-The numerical flux is assumed to be consistent, i.e. $\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$.
+The numerical flux is assumed to be consistent, i.e.
+$\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$.
+The moment vector inside a spatial cell, denoted by $\bm{\hat u}_{j}^{n}$ is not to be confused with the moment of order $i$ in \eqref{eq:SGClosure}, which is a continuous function in time and space.
+Note that whenever the moment vector includes a time index $n$, we are speaking of a discrete quantity.
+Furthermore, we use the index $j\in\mathbb{N}$ to denote spatial cells and the multi-index $i\in\mathbb{N}_0^p$ to denote the moment order.
 
@@
 \begin{linenomath*}\begin{align}\label{eq:dualIteration1}
-\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j).
+\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j^{n}).
 \end{align}\end{linenomath*}

### 2020-07-03 Paper1/one_shot_ipm.tex
Paper version elsevier

@@
 \begin{theorem}\label{th:Contractive}
-\comment{We denote the fixed point of the classical IPM iteration \eqref{eq:momentIteration} by $\bm{\hat u}^*_{\Delta} = (\bm{\hat u}_{1}^*,\cdots,\bm{\hat u}_{N_x}^*)^T$, i.e. we have
+We denote the fixed point of the classical IPM iteration \eqref{eq:momentIteration} by $\bm{\hat u}^*_{\Delta} = (\bm{\hat u}_{1}^*,\cdots,\bm{\hat u}_{N_x}^*)^T$, i.e. we have
 \begin{linenomath*}\begin{align*}
@@
 for all spatial cells $j$.
-}
 Assume that the classical IPM iteration is contractive at its fixed point.
@@
 \end{align}\end{linenomath*}
-Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}_{\comment{\Delta}}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
+Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}_{\Delta}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by
 \begin{linenomath*}\begin{align*}
@@
 \begin{linenomath*}\begin{align*}
-\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \comment{\bm{\hat u}_j^n}\right) \\
-&=- \partial_{\bm{\lambda}}\bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \comment{\bm{\hat u}_j^n}\right).
+\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\bm\lambda_j^n)^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\bm\lambda_j^n)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}_j^n\right) \\
+&=- \partial_{\bm{\lambda}}\bm{H}(\bm\lambda_j^n)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}_j^n\right).
 \end{align*}\end{linenomath*}
-\comment{Recall that if $\bm{\hat u}_j^n$ and $\bm{\lambda}_j^n$ belong to the fixed point $(\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)$,} we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \comment{\bm{\hat u}_j^n}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
+Recall that if $\bm{\hat u}_j^n$ and $\bm{\lambda}_j^n$ belong to the fixed point $(\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}_j^n$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$.
 For the block $\partial_{\bm{\hat u}} \bm{d}$, we get
 \begin{linenomath*}\begin{align*}
-\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\comment{\bm\lambda_j^n})^{-1},
+\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\bm\lambda_j^n)^{-1},
 \end{align*}\end{linenomath*}
-hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$. \comment{To shorten notation, let us define
+hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$. To shorten notation, let us define
 \begin{linenomath*}\begin{align*}
@@
 \end{align*}\end{linenomath*}
-Then we have}
+Then we have
 \begin{linenomath*}\begin{align*}
@@
 \begin{linenomath*}\begin{align*}
-\frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
+\frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j
 \end{align*}\end{linenomath*}
@@
 I.e.
-\comment{if we denote the solution at iteration $n$ as $(\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)})$}, there exists a $\delta>0$ s.t.
-for all starting points $(\bm{\lambda}_{\comment{\Delta}}^{(0)},\bm{\hat u}_{\comment{\Delta}}^{(0)})\in B_{\delta}(\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)$ we have
+if we denote the solution at iteration $n$ as $(\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)})$, there exists a $\delta>0$ s.t.
+for all starting points $(\bm{\lambda}_{\Delta}^{(0)},\bm{\hat u}_{\Delta}^{(0)})\in B_{\delta}(\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)$ we have
 \begin{linenomath*}\begin{align*}
-\left\Vert (\bm{\lambda}\comment{_{\Delta}^{(n)}},\bm{\hat u}\comment{_{\Delta}^{(n)}}) - (\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)\right\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
+\left\Vert (\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)}) - (\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)\right\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.
 \end{align*}\end{linenomath*}
@@
 \end{align*}\end{linenomath*}
-Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda}_{\comment{\Delta}},\bm{\hat u}_{\comment{\Delta}})$.
+Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda}_{\Delta},\bm{\hat u}_{\Delta})$.
 Hence there exists an open ball with radius $\delta$ around the fixed point in which the eigenvalues remain in the interval $(-1,1)$.

### 2020-07-03 Paper1/results.tex
Paper version elsevier

--- 2020-05-18 Paper1/results.tex
@@
 At the airfoil boundary, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal.
-At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of \comment{$T=273.15$ K.
+At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101\;325$ Pa and a temperature of $T=273.15$ K.
 Note that with the specific gas constant $R=287.87\frac{J}{kg\cdot K}$ and a given angle of attack $\phi$ this fixes the conserved variables at the far field boundary, since
@@
 \end{pmatrix}, \quad \rho e = \frac{p}{\gamma - 1} + \frac{1}{2}\rho (v_1^2 + v_2^2).
-\end{align*}\end{linenomath*}}
+\end{align*}\end{linenomath*}
 Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees, i.e.
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.
+The reference solution has been computed with stochastic-Collocation using $100$ Gauss-Legendre quadrature points.}
 	\label{fig:referenceSolution}
@@
 	\end{subfigure}
-	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
-All intrusive methods converge to a residual $\varepsilon=6\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=1\cdot 10^{-7}$.
+	\caption{Comparison of the relative L$^2$-error \eqref{eq:relErrorEVar} for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row).
+All intrusive methods (i.e.
+SG, IPM, osIPM, adosIPM, readosIPM) are converged to a residual $\varepsilon=6\cdot 10^{-6}$ (see equation \eqref{eq:residualSteady}), whereas all non-intrusive methods (i.e.
+SC) are converged to a residual $\varepsilon=1\cdot 10^{-7}$ in order to obtain an error which remains constant in time.
+SG and IPM are computed with Algorithm~\ref{alg:IPM} (where SG uses a quadratic entropy) adapted for steady problems.
+For osIPM we used Algorithm~\ref{alg:osIPM}, for adosIPM and readosIPM we use Algorithm~\ref{alg:adosIPM} with and without refinement retardation.
 All computations are performed with 5 MPI threads.}
@@
 
-First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration.
-The standard SG method however preserves positivity of mass, energy and pressure.
+First, let us mention that the adaptive SG method fails, since it yields negative gas densities during the iteration.
+The standard SG method without adaptivity however preserves positivity of density, energy and pressure in our numerical experiments.
+Note that in general, stochastic-Galerkin does not guarantee positivity of these quantities, since the polynomial approximation does not guarantee any solution bounds.
 The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance.
@@
 	\caption{Reference solution (\subref{fig:referenceSolutionsub2DE}) E$[\rho]$ and (\subref{fig:referenceSolutionsub2DVar}) Var$[\rho]$ and with logarithmic scaled (\subref{fig:referenceSolutionsub2DVarLog}) Var$[\rho]$.
+The reference solution has been computed with stochastic-Collocation using $50^2$ Gauss-Legendre quadrature points.
 readosIPM$_{1-9}$ solution for (\subref{fig:reados2DE}) E$[\rho]$ and (\subref{fig:reados2DVar}) Var$[\rho]$, with the resulting (\subref{fig:readosIPMEVar2DRefinement}) refinement levels.}
@@
 	\end{subfigure}
-	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step}
+	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step.
+The reference solution has been computed with stochastic-Collocation using $50^3$ Gauss-Legendre quadrature points.}
 	\label{fig:solution3D}
